### 2012-10-22 Research training portfolio.tex
Started working on internship portfolio

\documentclass[12pt]{article}
\title{Research training portfolio}
\author{Wessel Stoop}

\begin{document}
\maketitle

\section{Introduction}

\section{What I have done}

\subsection{Project structure}
\subsection{Problems encountered}
\subsection{Results}

\section{What I have learned}
Although I did have some experience with both programming and language technology, and even with some software co-developed by my supervisor, this internship turned out to be a adventures full of novelty: no single semester in my Radboud University career had such a high skills-acquired/day-ratio as this one.

\subsection{Working with Linux and SSH}
The novelty already started with something as non-trivial as the choice of operating system. As a Microsoft Windows user, I was surprised to find out all machine learning software I needed was developed and best used on Linux (see the next section). I only had a few hours of experience with Ubuntu, so I had to invest several hours into installing and getting more familiar with it. Most time went into learning the terminal commands - its GUI turned out to be almost identical to Windows'. Windows of course has a command prompt as well, but it is rarely used. Ubuntu's commands on the other hand are essential for having full control. The first hours of the internship were therefore spent figuring out basic actions like \emph{ls} (showing folders and files in a directory) and \emph{mv} (renaming and moving files). This knowledge later turned out to be incredibly handy when I needed to control external servers over an SSH connection, as the Linux machines can be controller with commands only. Here again, it took some time to figure out how to do this properly.

% Maar.... nu kan ik het heel goed?
% Voordelen Linux?

\subsection{Working with machine learning software}
Fowlt uses the machine learning software TiMBL, TiMBLserver and WOPR.

% Namen ontwikkelaars.
% With the help of my supervisors

\subsection{Working with GitHub}
Working with multiple people on the same software can cause quite a lot of practical problems. Version control system Git solves these problems by keeping track of which version everyone of working on, providing tools to easily merge the work when done, and to remember older versions of the project, among other things. Although I had used Git for the latter function in personal hobby projects, I'd never worked together with other people on programming tasks. During this internship I learned (1) this can be done easily with GitHub and (2) how to handle GitHub. GItHub is a website that offers open source project a free, central and always reachable place for a Git repository, and also provides tools to look into this repo and its history online.

\subsection{Working with CLAM, setting up a NLP webservice}

\subsection{General}

\section{Research paper}

\subsection{Introduction}

\subsection{Method}

\subsection{Results}

\subsection{Conclusion}

\section{General onclusion}

\end{document}

### 2012-10-23 Research training portfolio.tex
Worked on 'What I have learned', changed font

@@
 \documentclass[12pt]{article}
+
 \title{Research training portfolio}
 \author{Wessel Stoop}
+
+\renewcommand{\familydefault}{\sfdefault}
 
@@
 \section{What I have learned}
-Although I did have some experience with both programming and language technology, and even with some software co-developed by my supervisor, this internship turned out to be a adventures full of novelty: no single semester in my Radboud University career had such a high skills-acquired/day-ratio as this one.
+Although had some experience with both programming and language technology, and even with some software co-developed by my supervisor, this internship turned out to be an adventure full of novelty: no single semester in my Radboud University career had such a high skills-acquired/day-ratio as this one.
+In the following sections, I give a non-exhaustive overview of what I learned while working on Fowlt.
 
@@
 I only had a few hours of experience with Ubuntu, so I had to invest several hours into installing and getting more familiar with it.
-Most time went into learning the terminal commands - its GUI turned out to be almost identical to Windows'.
-Windows of course has a command prompt as well, but it is rarely used.
-Ubuntu's commands on the other hand are essential for having full control.
+Ubuntu's GUI turned out to be almost identical to Windows', so most time went into learning the terminal commands.
+Unlike the commands for Windows' command prompt, Linux commands are essential for having full control.
 The first hours of the internship were therefore spent figuring out basic actions like \emph{ls} (showing folders and files in a directory) and \emph{mv} (renaming and moving files).
-This knowledge later turned out to be incredibly handy when I needed to control external servers over an SSH connection, as the Linux machines can be controller with commands only.
-Here again, it took some time to figure out how to do this properly.
+This knowledge was also essential later, when I needed to control external servers over an SSH connection, as the Linux machines can be controlled with commands only.
+Here again, it took some time to figure out how to make the servers do what I wanted - transferring a python script from my machine to a server, for instance.
+Fortunately, working with Linux became routine quickly.
 
-% Maar.... nu kan ik het heel goed?
 % Voordelen Linux?
@@
 Fowlt uses the machine learning software TiMBL, TiMBLserver and WOPR.
+TiMBL is consists of several memory-based learning algorithms, among which IB1-IG, an implementation of k-nearest neighbor classification with feature weighting suitable for symbolic feature spaces, and IGTree, a decision-tree approximation of IB1-IG.
+TiMBLserver adds server functionality to TiMBL.
+WOPR is a wrapper around the k-nearest neighbor classifier in TiMBL, offering word prediction and language modeling functionalities.
+Trained on a text corpus, WOPR can predict missing words, report perplexities at the word level and the text level, and generate spelling correction hypotheses.
+For Fowlt, only WOPR's spelling modus is used, of course.
+\\\indent
+Unfortunately, all three programs have very little documentation, and documentation that does exist often assumes background knowledge.
+With the help of my supervisor and Maarten van Gompel I learned how to (1) how to make language models by training the algorithms on large amounts of texts, (2) how to use these models to predict new words (and in my case, find spelling errors) and (3) how to test the model's accuracy.
+During my steps, I've repeated these three steps numerous times.
+\\\indent
+Besides TiMBL, TiMBLserver and WOPR, I also learned to use Ucto, software to tokenize texts so TiMBL and WOPR can handle them, and the PyNLPl Python library, which can read and create FoLiA-XML.
+FoLiA-XML is the XML-format used by Fowlt's output.
 
-% Namen ontwikkelaars.
 % With the help of my supervisors
@@
 During this internship I learned (1) this can be done easily with GitHub and (2) how to handle GitHub.
-GItHub is a website that offers open source project a free, central and always reachable place for a Git repository, and also provides tools to look into this repo and its history online.
+GitHub is a website that offers open source project a free, central and always reachable place for a Git repository, and also provides tools to look into this repo and its history online.
 
@@
 \subsection{General}
+
+% Python stuff
+% C compilen
 

### 2012-10-23 Research training portfolio.tex
Added intro

@@
 \section{Introduction}
+Fowlt is the English version of the Dutch spelling checker Valkuil.net.
+Both Valkuil and Fowlt are unlike most well-known spell checkers (like the one in Microsoft Word): whereas these spelling checkers mostly try to find errors by comparing all words to a built-in dictionary and flag the word as an error if they can't find a match, Valkuil and Fowlt are 'context-based'.
+This entails that they look at the words around every word.
+If, based on this context, another word was expected, and the algorithm is really certain about it, it is flagged as an error.
+This means, for example, that Fowlt is able to replace the incorrect "there' in "there really nice people" into "they're" simply because "there" usually isn't followed by "really nice people", while "they're" is.
+\\\indent
+To be able to make these kinds of correction suggestion, Fowlt makes use of language models.
+These models are made with the help of machine learning software TiMBL and WOPR and very large datasets.
+However, this also means that if the context of a particular error is very different from what the model has seen in the training data, it won't be able to correct the error.
+This means users should be aware of the fact that, although Fowlt can much kinds of errors than regular spell checkers, it might still miss a few.
+This is particularly true because we told Fowlt to only flag a word as an error if the algorithms are really sure about it.
 
@@
 
-% With the help of my supervisors
-
 \subsection{Working with GitHub}
@@
 Although I had used Git for the latter function in personal hobby projects, I'd never worked together with other people on programming tasks.
-During this internship I learned (1) this can be done easily with GitHub and (2) how to handle GitHub.
-GitHub is a website that offers open source project a free, central and always reachable place for a Git repository, and also provides tools to look into this repo and its history online.
+During this internship I learned (1) this can be done easily by using Git in combination with GitHub and (2) how to handle GitHub.
+GitHub is a website that offers open source projects a free, central and always reachable place for a Git repository, and also provides tools to look into this repo and its history online.
 

### 2012-10-23 Research training portfolio.tex
Added project structure

@@
 \section{What I have done}
+The best way to get a good overview of what I have done is probably to look at Fowlt's internal structure, so that's what I present in the next section.
+The first part of my internship basically consisted of adding these modules one by one.
+To get a better idea of what parts took more time to 'get right', I also added a section about problems I encountered.
 
 \subsection{Project structure}
+Fowlt exist of multiple little spelling checkers, specialized in doing one thing.
+These spelling checkers are called 'modules'.
+Fowlt consists of four context-free modules (relatively similar to normal spelling checkers' functionality) and, as I write this, sixteen context-based modules.
+This number will probably be different when you read this, though, as new context-based modules are created easily, and can be removed if it turns out they're not accurate enough.
+
+\subsubsection{The processchain}
+In Fowlt's main script, fowlt\_processchain.py, all modules are started up, and their outputs are collected.
+Making changes in this script (based on the one for Valkuil, written by Maarten van Gompel) to make sure the new English based modules worked correctly, is one of the main things I did during my internship.
+
+\subsubsection{ErrorlistChecker (context-free)}
+Martin Reynaert provided us with a large list of common typos and their corrections.
+This module, written by me, checks for every word in the input whether it's in the errorlist, and adds the correction when it is.
+
+\subsubsection{LexiconChecker (context-free)}
+Google provided us with an enormous list of word frequencies on the internet in 2008.
+The LexiconChecker checks for every word in the input whether there's a word on the lexiconchecker that is very similar to it, but is much more frequent.
+
+\subsubsection{RunonChecker (context-free)}
+The RunonChecker also uses Google's list, but uses it check if any spaces where forgotten in the input.
+This is done by looking if splitting up long words produces two words which both are much more frequent that the original word.
+
+\subsubsection{SplitChecker (context-free)}
+The SplitChecker is the exact opposite of the RunonChecker: instead of looking whether any spaces are forgotten, it checks whether any spaces have to be added.
+This is done by looking whether each combination of two words, when combined, produces a word which is much more frequent than the two original words.
+
+\subsubsection{WOPRChecker (context-based)}
+The WOPRChecker contacts WOPR, running on a server on localhost or elsewhere.
+It gives it the context of each word, but not the words themselves.
+If WOPR predicts another word with much certainty, this word is added as a correction.
+
+\subsubsection{The ConfusibleCheckers (context-based)}
+The ConfusibleCheckers work similar to the WOPRChecker, with the only difference that they are trained on one specific language problem, instead of on language in general.
+As I write this, there are fifteen active ConfusibleCheckers, all focussing on their own expertise: lose-loose, it's-its, you're-your, than-then, who-which-that, whether-weather, effect-affect, lie-lay, they're-their-there, don't-doesn't,to-too-two, advice-advise, any-some, less-fewer, practice-practise and chose-choose.
+\\\indent
+A ConfusibleChecker trained on the difference between 'then' and 'than' for example filters out these two words and their contexts, sends these contexts to a Timblserver running somewhere.
+This Timblserver then says which of the two it would predict on the basis of the context.
+If this prediction is different from the actual word and the server is very sure about this, the word is flagged as an error.
+Training new ConfusibleModules has been a large part of my internship.
+
 \subsection{Problems encountered}
@@
 
-\subsection{Working with CLAM, setting up a NLP webservice}
+\subsection{Working with CLAM, setting up an NLP webservice}
 
@@
 
-\section{General onclusion}
+\section{General conclusion}
 

### 2012-10-23 Research training portfolio.tex
Added problems

@@
 \subsection{Problems encountered}
+
+\subsubsection{The original confusiblemodule could only handle two confusibles}
+Most confusibles consist of two words only, but English has two that consist of three: they're-there-their and to-too-two.
+The original c-code could not handle confusibles with more than two words, so while the original plan was to either use Valkuil's c-code or rewrite it in Python, the quickest solution here was me diving into the c-code of my supervisor and add support for three.
+
+\subsubsection{Problems with Ucto}
+Although Ucto includes a configuration file for English, it didn't produce the correct output right out of the box.
+For example, for the It's-ItsModule to work, it is essential that "it's" is recognized as one word, which is not the case in original English configuration file.
+After a lot of tweaking of the regular expressions, I now have a tokenizer that produces output which can be used by all modules.
+
+\subsubsection{The ConfusibleModules weren't good enough}
+I didn't only train the ConfusibleModules, I also tested them with a program written by my supervisor.
+This program takes 10\% from the training data and saves it in a seperate file, trains TiMBL on the remaining 90\%, and then returns the percentage of correct predictions the resulting language can make about the 10\%.
+The results differed for the various confusibles, but the models were correct roughly 85-90\% of the time.
+This may seem a lot, but for a spellingchecker to be useful it's essential that a model can have an accuracy of over 99\%.
+Following the advice of my supervisor, I did two things to accomplish this: (1) I retrained TiMBL on the balanced training sets, which entails that there is an even amount of examples for each word the confusible consists of, and (2) I made the ConfusibleChecker only add a correction when it is very sure.
+Especially the latter solution increases the accuracy dramatically, but at the price of finding a lot less errors.
+
 \subsection{Results}

### 2012-10-24 Research training portfolio.tex
Revision of whole text

@@
 Both Valkuil and Fowlt are unlike most well-known spell checkers (like the one in Microsoft Word): whereas these spelling checkers mostly try to find errors by comparing all words to a built-in dictionary and flag the word as an error if they can't find a match, Valkuil and Fowlt are 'context-based'.
-This entails that they look at the words around every word.
-If, based on this context, another word was expected, and the algorithm is really certain about it, it is flagged as an error.
+This entails that they also take into account the context, which is the words around every word.
+If, for a particular word, they expected another word based on the context, and the algorithm is really certain about it, it is flagged as an error.
 This means, for example, that Fowlt is able to replace the incorrect "there' in "there really nice people" into "they're" simply because "there" usually isn't followed by "really nice people", while "they're" is.
 \\\indent
-To be able to make these kinds of correction suggestion, Fowlt makes use of language models.
-These models are made with the help of machine learning software TiMBL and WOPR and very large datasets.
+To be able to make these kinds of correction suggestions, Fowlt makes use of language models.
+These models are created by giving lots of texts to the machine learning software TiMBL and WOPR.
+It's on the basis of these texts the model knows "there" mostly is not followed by "really nice people".
 However, this also means that if the context of a particular error is very different from what the model has seen in the training data, it won't be able to correct the error.
-This means users should be aware of the fact that, although Fowlt can much kinds of errors than regular spell checkers, it might still miss a few.
-This is particularly true because we told Fowlt to only flag a word as an error if the algorithms are really sure about it.
+This means users should be aware of the fact that, although Fowlt can recognize more kinds of errors than regular spell checkers, it might still miss a few.
+This is particularly true because Fowlt has been taught to only flag a word as an error if the algorithms are really sure about it.
 
@@
 The best way to get a good overview of what I have done is probably to look at Fowlt's internal structure, so that's what I present in the next section.
-The first part of my internship basically consisted of adding these modules one by one.
+The first part of my internship basically consisted of adding the modules described below, one by one.
 To get a better idea of what parts took more time to 'get right', I also added a section about problems I encountered.
@@
 These spelling checkers are called 'modules'.
-Fowlt consists of four context-free modules (relatively similar to normal spelling checkers' functionality) and, as I write this, sixteen context-based modules.
+Fowlt consists of four context-free modules (relatively similar to normal spelling checkers functionality) and, as I write this, sixteen context-based modules.
 This number will probably be different when you read this, though, as new context-based modules are created easily, and can be removed if it turns out they're not accurate enough.
@@
 Google provided us with an enormous list of word frequencies on the internet in 2008.
-The LexiconChecker checks for every word in the input whether there's a word on the lexiconchecker that is very similar to it, but is much more frequent.
+The LexiconChecker checks for every word in the input whether there's a word on the frequencylist that is very similar to it, but much more frequent.
 
 \subsubsection{RunonChecker (context-free)}
-The RunonChecker also uses Google's list, but uses it check if any spaces where forgotten in the input.
-This is done by looking if splitting up long words produces two words which both are much more frequent that the original word.
+The RunonChecker also uses Google's list, but uses it to check if any spaces were forgotten in the input.
+This is done by looking whether splitting up long words produces two words which both are much more frequent that the original word.
 
@@
 The SplitChecker is the exact opposite of the RunonChecker: instead of looking whether any spaces are forgotten, it checks whether any spaces have to be added.
-This is done by looking whether each combination of two words, when combined, produces a word which is much more frequent than the two original words.
+This is done by looking whether each combination of two words produces a word which is much more frequent than the two original words.
 
@@
 I didn't only train the ConfusibleModules, I also tested them with a program written by my supervisor.
-This program takes 10\% from the training data and saves it in a seperate file, trains TiMBL on the remaining 90\%, and then returns the percentage of correct predictions the resulting language can make about the 10\%.
+This program takes 10\% from the training data and saves it in a seperate file, trains TiMBL on the remaining 90\%, and then returns the percentage of correct predictions the resulting model can make about the 10\%.
 The results differed for the various confusibles, but the models were correct roughly 85-90\% of the time.
@@
 The first hours of the internship were therefore spent figuring out basic actions like \emph{ls} (showing folders and files in a directory) and \emph{mv} (renaming and moving files).
-This knowledge was also essential later, when I needed to control external servers over an SSH connection, as the Linux machines can be controlled with commands only.
+This knowledge was also essential later, when I needed to control external servers over an SSH connection, as these Linux machines can be controlled with commands only.
 Here again, it took some time to figure out how to make the servers do what I wanted - transferring a python script from my machine to a server, for instance.
@@
 Fowlt uses the machine learning software TiMBL, TiMBLserver and WOPR.
-TiMBL is consists of several memory-based learning algorithms, among which IB1-IG, an implementation of k-nearest neighbor classification with feature weighting suitable for symbolic feature spaces, and IGTree, a decision-tree approximation of IB1-IG.
+TiMBL consists of several memory-based learning algorithms, among which IB1-IG, an implementation of k-nearest neighbor classification with feature weighting suitable for symbolic feature spaces, and IGTree, a decision-tree approximation of IB1-IG.
 TiMBLserver adds server functionality to TiMBL.
@@
 Unfortunately, all three programs have very little documentation, and documentation that does exist often assumes background knowledge.
-With the help of my supervisor and Maarten van Gompel I learned how to (1) how to make language models by training the algorithms on large amounts of texts, (2) how to use these models to predict new words (and in my case, find spelling errors) and (3) how to test the model's accuracy.
-During my steps, I've repeated these three steps numerous times.
+With the help of my supervisor and Maarten van Gompel I learned how to (1) make language models by training the algorithms on large amounts of texts, (2) use these models to predict new words (and in my case, find spelling errors) and (3) test the model's accuracy.
+During my internship, I've repeated these three steps numerous times.
 \\\indent
-Besides TiMBL, TiMBLserver and WOPR, I also learned to use Ucto, software to tokenize texts so TiMBL and WOPR can handle them, and the PyNLPl Python library, which can read and create FoLiA-XML.
+Besides TiMBL, TiMBLserver and WOPR, I also learned to use Ucto, software to tokenize texts so TiMBL and WOPR can handle them, and the PyNLPl Python library, which can read and create FoLiA-XML (among other things).
 FoLiA-XML is the XML-format used by Fowlt's output.

### 2012-10-25 Research paper.tex
Added info table, added research paper outline

\documentclass[12pt]{article}

\title{Research paper}
\author{Wessel Stoop}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}
\maketitle

\section{Introduction}
\section{The goal of the project}
\section{Scientific importance and relevance for society}
\section{Methodology}
\section{Theoretical framework}
\section{Achievements so far}
\section{Main outstanding questions}
\section{Strengths and weaknesses}
\section{Possibilities for future research}
\section{Conclusion}
\end{document}

### 2012-11-03 Research paper.tex
Added introduction, added ideas under various headers

@@
 \title{Research paper}
-\author{Wessel Stoop}
+\author{Wessel Stoop, s0808709}
+
+\usepackage{covington}
 
@@
 \section{Introduction}
-\section{The goal of the project}
+
+In this paper, I will give a detailed overview of the research project Colibri.
+Colibri, carried out a the Radboud University Nijmegen by Maarten van Gompel and Antal van den Bosch, investigates the identification and extraction of constructions in natural language, and the usage of such constructions in Machine Translation.
+These constructions can be a various lengths, and, importantly, can also contain one or more gaps.
+The constructions are not identified on the basis of explicit 'human' knowledge about language and grammar ('linguistic theory'), but are distilled from large amounts of text ('text corpora') with the help of context-sensitive machine-learning techniques.
+
+\section{Project description}
+
+%Stukje over dat het software is, dus tekst over wat de software inhoudt is tegelijkertijd een beschrijving van wat het softwarepakket inhoudt
+
+\subsection{Pattern extraction}
+
+Colibri, 'Constructions as Linguistics Bridges', investigates constructions (or patterns), as its name suggests.
+A construction can be any combination of multiple words (a so-called 'n-gram') in any natural language that in some way forms an entity.
+An example is 'on the basis of' in the following sentence:
+
+\begin{examples}
+\item On the basis of these ideas, software can be developed.
+\end{examples}
+
+Importantly, constructions can also have one or more gaps (so-called 'skip-grams'). '\_\_ have been \_\_' in the following example, is a construction with two such gaps.
+
+\begin{examples}
+\item I have been working
+\end{examples}
+
+The first gap is usually filled with a noun phrase, the second one verb phrase.
+Of course, not every possible combination of words is a construction; what makes some combinations special?
+The exact nature of linguistic constructions has been the subject of many publications and even an entire linguistic framework (construction grammar).
+For this project however, it suffices to say that constructions emerge because some combinations of words are more frequent than others.
+
+% Drie vormen freq: raw frequency, context frequency, multilingual alignmetn
+% Blabla, want freq is waar mach learning gebriuk van maakt
+
+\subsection{Pattern querying} Interactively query generated models for specific patterns.
+\subsection{Pattern/corpus comparison} Match patterns generated from one corpus to another corpus.
+\subsection{Graph computation \& visualisation} Compute graph relations between pattern models and visualise graphs
+\subsection{Alignment algorithms} Algorithms to extract and align patterns accross languages; resulting in translation pairs for constructions.
+\subsection{Machine Translation Decoding} Reassembles translated fragments into one coherent sentence in the target-language; seeking the best (statistically most probably) solution in a vast search space of possible search hypotheses.
+\subsection{Memory-based Machine Translation} Machine learning algorithms can be used to consider source-side context in translations. Classifiers can be built and directly used by the decoder.
+\subsection{Experiment Framework} Colibri comes with an extensive experiment framework for conducting Machine Translation experiments.\
+
+\subsection{Hypotheses} %Of moet dit bij main outstanding questions?
+
+Constructions can be found efficiently in corpus data
+Graph-based relations can be used to constrain to good constructions
+Constructions can be aligned without resorting to word alignments as a basis
+In MT, constructions (i.e. possibly with gaps) result in better translation than mere consecutive phrases
+
 \section{Scientific importance and relevance for society}
+
+% Hier dingen zeggen over hoe er eerst vertaald werd. Of is dat theoretical framework?
+% Ook lingusitic relevance: Abstracting fully lexicalised constructions
+%Finding semantic subclasses in constructions: from time-expression to time-expression"
+%Collapsing constructions with word disjunctions (\he/she/it") or part-of-speech tags
+% Correlations with experimental findings
+% Switch tasks, cloze tests, reaction times, ...
+
 \section{Methodology}
+
+% C++, Python
+
 \section{Theoretical framework}

### 2012-11-04 Research paper.tex
More fancy intro

@@
 
-In this paper, I will give a detailed overview of the research project Colibri.
-Colibri, carried out a the Radboud University Nijmegen by Maarten van Gompel and Antal van den Bosch, investigates the identification and extraction of constructions in natural language, and the usage of such constructions in Machine Translation.
+When one of Starfleet's vessels meets an alien ship in areas where no man has gone before, they can (almost) always communicate with these aliens (almost) instantly; they speak to them in English, and these aliens answer in English.
+This is possible because, according to the Star Trek lore, the protagonists make use of a so-called Universal Translator.
+This Universal Translator can (1) recognize speech in every language, (2) translate what it has 'heard' to any other language, (3) make its results audible with speech synthesis and (4) learn itself new languages when needed.
+Unfortunately, like many technologies in Star Trek, this is far from reality: there have been various attempts at the first three tasks in particular, but no software currently in existence can do this as perfect as the Universal Translator.
+\\\indent
+In this paper, I will give a detailed overview of the research project Colibri, which basically is an attempt to improve the technology needed for the second task: automatically translating text from one language to another.
+More specifically, the project investigates the identification and extraction of constructions in natural language, and how these can be used in Machine Translation.
 These constructions can be a various lengths, and, importantly, can also contain one or more gaps.
@@
 
+\section{Theoretical framework}
+
 \section{Scientific importance and relevance for society}
@@
 
-\section{Theoretical framework}
 \section{Achievements so far}

### 2012-11-06 Research paper.tex
RP: wat structuur en flarden toegevoegd
RP: revision, start with classical MT. RTP: start with evaluation

@@
 \usepackage{covington}
+\usepackage{graphicx}
 
@@
 This Universal Translator can (1) recognize speech in every language, (2) translate what it has 'heard' to any other language, (3) make its results audible with speech synthesis and (4) learn itself new languages when needed.
-Unfortunately, like many technologies in Star Trek, this is far from reality: there have been various attempts at the first three tasks in particular, but no software currently in existence can do this as perfect as the Universal Translator.
+Unfortunately, like many technologies in Star Trek, this is far from reality: there is lots of active research at the first three tasks in particular, but no software currently in existence can do this as perfect as the Universal Translator.
 \\\indent
@@
 More specifically, the project investigates the identification and extraction of constructions in natural language, and how these can be used in Machine Translation.
-These constructions can be a various lengths, and, importantly, can also contain one or more gaps.
+These constructions can be of various lengths, and, importantly, can also contain one or more gaps.
 The constructions are not identified on the basis of explicit 'human' knowledge about language and grammar ('linguistic theory'), but are distilled from large amounts of text ('text corpora') with the help of context-sensitive machine-learning techniques.
+The Colibri project is carried out by Maarten van Gompel and supervised by Antal van den Bosch.
 
@@
 
-%Stukje over dat het software is, dus tekst over wat de software inhoudt is tegelijkertijd een beschrijving van wat het softwarepakket inhoudt
+Besides an attempt to discover more about the possibilities for machine translation technology, Colibri is also an attempt to actually develop this technology.
+The following project overview therefore necessarily also is an overview of the software.
 
@@
 
-The first gap is usually filled with a noun phrase, the second one verb phrase.
+The first gap is usually filled with a noun phrase, the second one with a verb phrase.
 Of course, not every possible combination of words is a construction; what makes some combinations special?
@@
 
-\section{Theoretical framework}
+\section{Theoretical background and framework}
+
+\subsection{Classical Machine Translation}
+
+Classical Machine Translation is best characterized by the use of human created rules.
+That is, authors of translation algorithms tried to include their knowledge about the languages and translation.
+There have been four main approaches, summarized in the Vauquois triangle: \\
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.8\textwidth]{vauquois.png}
+\caption{The Vauquois triangle}
+\label{fig:vauquois}
+\end{figure}
+
+The horizontal arrows represent the approaches, I'll go through them from bottom to top:
+
+\begin{itemize}
+\item The direct approach does very little source language analysis, and could thus be viewed as the automatical version of looking up every word in the source language in a dictionary and replacing it with the corresponding word in the target language.
+This often doesn't yield very good results: if I for example would have translated this paper from my mother tongue Dutch to English word for word, 'would we now not real a good result have'.
+It should be noted, however, that direct approach can be a little more intelligent than that: %wwat dan?
+\item The syntactic approach
+\item The semantic approach
+\item All of the approaches so far have in common that they require a seperate rule set for each language pair; no or very few rules included for translations from English to German can also be used for translations from English to French.
+The Interlingua approach is an attempt to solve this problem, by pairing up every language with one general language, the Interlingua.
+This means a translation from English to German would entail a translation from English to the Interlingua (using the ruleset English $\rightarrow$ Interlingua) and then a translation from the Interlingua to German (using the ruleset Interlingua $\rightarrow$ German).
+When translating from English to French, the same two steps would be taken, and the same ruleset English $\rightarrow$ Interlingua can be used.
+This Interlingua is a formal representation.
+\end{itemize}
+
+% Voorbeelden?
+
+\subsection{Statistical Machine Translation}
 
@@
 
-% Hier dingen zeggen over hoe er eerst vertaald werd. Of is dat theoretical framework?
 % Ook lingusitic relevance: Abstracting fully lexicalised constructions
@@
 \section{Conclusion}
+
+\section{References}
+
+% Documentatie Colibri
+% Maartens thesis
+
 \end{document}

### 2012-11-06 Research training portfolio.tex
Added info table, added research paper outline
Added student number
RP: revision, start with classical MT. RTP: start with evaluation

@@
 \maketitle
+
+\begin{table}[h]
+\begin{tabular}{ll}
+Name&Wessel Stoop\\
+Student numbers&s0808709 (Nijmegen), u1249664 (Tilburg)\\
+Project&Fowlt, a context-based spelling checker\\
+Supervisor&Antal van den Bosch\\
+Period&September - December 2012\\
+\end{tabular}
+\end{table}
 
@@
 
+% Hier moet ergens een beetje beschreven worden hoe zo'n Timblmodule werkt.
+
+\subsubsection{Evaluation}
+
+In this section, I describe the various evaluation measures I used.
+This is necessary because evaluation of spelling corrector has not been very uniform in the spelling checker literature so far; whereas various scholars only measured how many errors their algorithm can detect in an error list and called it accuracy (e.g.
+Agirre et al., 1998; Brill \& Moore, 2000; Toutanova \& Moore, 2002; Van Delden et al., 2004), others use or argue for the precision-recall terminology (Reynaert, 2008; Pollock \& Zamora, 1984, although they use other terms for the two concepts) or argue against it (Starlander \& Popescu-Belis, 2002).\\\indent
+
+Although most of the evaluation measures used can be useful and informative, they only tell part of the story.
+I think it is important to recognize that spelling correction is a multi-step process, and each step comes with its own measures.
+Using a confusible corrector as part of a spelling corrector, there are three steps: word prediction, error detection and error correction.
+I'll go through these steps and their evaluation measures one by one.
+
+\paragraph{Word prediction}
+
+\begin{itemize}
+\item Accuracy.
+Although word prediction may be the most complicated part of the confusible corrector, its evaluation is actually the simplest.
+Calculating rediction accuracy entails removing all confusibles from a text, and see in how many of the cases the language model can correctly predict which word was removed.
+This percentage can be increased by only counting predictions that are above a certain threshold.
+Results with multiple thresholds (i.e.
+to what extent does the prediction improve if the threshold is raised) can be visualized easily with a graph, as shown in the result section.
+\end{itemize}
+
+\paragraph{Error detection}
+\begin{itemize}
+\item The confusion matrix.
+\item Recall.
+\item Precision.
+\item Accuracy.
+\end{itemize}
+
+\paragraph{Error correction} Some text.
+
+\paragraph{Problems with precision and recall.} Some text.
+
 \subsection{Results}
@@
 
+\section{References}
+% Ene paper Antal
+% roc101
+% (e.g. Agirre et al., 1998; Brill & Moore, 2000; Toutanova & Moore, 2002; Van Delden et al., 2004)
+% (Reynaert, 2008; Pollock & Zamora, 1984, although they use other terminology)
+% (Starlander & Popescu-Belis, 2002)
+
 \end{document}

### 2012-11-06 Research paper.tex
Reworded, reexplained some things

@@
 Colibri, 'Constructions as Linguistics Bridges', investigates constructions (or patterns), as its name suggests.
-A construction can be any combination of multiple words (a so-called 'n-gram') in any natural language that in some way forms an entity.
+A construction can be any group of consecutive words (a so-called 'n-gram') in any natural language that in some way forms an entity.
 An example is 'on the basis of' in the following sentence:
@@
 
-Importantly, constructions can also have one or more gaps (so-called 'skip-grams'). '\_\_ have been \_\_' in the following example, is a construction with two such gaps.
+Importantly, constructions can also have one or more gaps (so-called 'skip-grams'). 'from \_\_ point of view' in the following example, is a construction with one such gap.
 
 \begin{examples}
-\item I have been working
+\item I understand things better when I look at them from his point of view.
 \end{examples}
 
-The first gap is usually filled with a noun phrase, the second one with a verb phrase.
-Of course, not every possible combination of words is a construction; what makes some combinations special?
+The gap is filled here is filled with 'his', but the construction 'from \_\_ point of view' can also be filled with many other words ('my', 'the', 'another', 'Obama's', etc.).
+This shows the advantage of skip-grams over n-grams for machine translation: whereas n-grams would only have been able to capture the construction 'point of view', because the word directly before 'point of view' is variable, skip-grams are also able to recognize that the word 'for' is part of the construction.\\\indent
+Of course, not every possible group of consecutive words is a construction; what makes some groups special?
 The exact nature of linguistic constructions has been the subject of many publications and even an entire linguistic framework (construction grammar).

### 2012-11-06 Research training portfolio.tex
RT: more about evaluation, busy with references in latex

@@
 \author{Wessel Stoop}
+
+\usepackage{graphicx}
+\usepackage{natbib}
 
@@
 This is necessary because evaluation of spelling corrector has not been very uniform in the spelling checker literature so far; whereas various scholars only measured how many errors their algorithm can detect in an error list and called it accuracy (e.g.
-Agirre et al., 1998; Brill \& Moore, 2000; Toutanova \& Moore, 2002; Van Delden et al., 2004), others use or argue for the precision-recall terminology (Reynaert, 2008; Pollock \& Zamora, 1984, although they use other terms for the two concepts) or argue against it (Starlander \& Popescu-Belis, 2002).\\\indent
+Agirre et al., 1998; Brill \& Moore, 2000; Toutanova \& Moore, 2002; \citet{tm02,vandelden04}), others use or argue for the precision-recall terminology (Reynaert, 2008; Pollock \& Zamora, 1984, although they use other terms for the two concepts) or argue against it \citep{sp02}.\\\indent
 
@@
 \begin{itemize}
-\item Accuracy.
+\item Prediction accuracy.
 Although word prediction may be the most complicated part of the confusible corrector, its evaluation is actually the simplest.
-Calculating rediction accuracy entails removing all confusibles from a text, and see in how many of the cases the language model can correctly predict which word was removed.
+Calculating prediction accuracy entails removing all confusibles from a text, and see in how many of the cases the language model can correctly predict which word was removed.
 This percentage can be increased by only counting predictions that are above a certain threshold.
 Results with multiple thresholds (i.e.
-to what extent does the prediction improve if the threshold is raised) can be visualized easily with a graph, as shown in the result section.
+to what extent does the prediction accuracy improve if the threshold is raised) can be visualized with a graph, as shown in the result section.
 \end{itemize}
@@
 \paragraph{Error detection}
+
 \begin{itemize}
 \item The confusion matrix.
-\item Recall.
-\item Precision.
-\item Accuracy.
+From this step onwards, we concentrate on spelling errors.
+This means that our input no longer is errorless text (like in the previous step).
+We also interpret the results of the prediction differently now: whereas incorrectly predicted words were considered errors of the system in the previous task, lowering the accuracy, they will now be considered corrections.
+This means that the system can do two things right (correct incorrect words in the text and don't correct correct words), and two things wrong (correct correct words and don't correct incorrect words).
+This can be summarized in a confusion matrix (taken from \citet{fawcett04}):
+
+\begin{figure}[htb]
+\centering
+\includegraphics[width=0.8\textwidth]{confusion_matrix.png}
+\caption{The confusion matrix}
+\label{fig:confusion}
+\end{figure}
+
+\item Detection accuracy. Detection accuracy only indicates how many mistakes the confusible detector makes, not what kind of mistakes. It can be calculated with the following formula:
+
+%formula
+
+It is similar to prediction accuracy in that it indicates how well the sytem does in general, but keep in mind that they're measuring two completely different concepts.
+Although a system with a good prediction accuracy also has a good detection accuracy, it might be possible that a system very good at predicting confusibles in errorless texts behaves differently in texts that do contain errors.
+
+\item Detection recall.
+\item Detection precision.
 \end{itemize}
@@
 \paragraph{Problems with precision and recall.} Some text.
+
+Waarom geen AUC?
 
@@
 
-\section{References}
+\begin{thebibliography}{99}
+
+\bibitem[Agirre et al., 1998]{agirre98}
+Agirre et al., 1998
+\bibitem[Brill \& Moore, 2000]{bm00}
+Brill \& Moore, 1998
+\bibitem[Van Delden et al., 2004]{vandelden04}
+Van Delden et al, 2004.
+\bibitem[Fawcett, 2004]{fawcett04}
+Fawcett, 2004.
+\bibitem[Pollock \& Zamora, 1984]{pz84}
+Pollock \& Zamora, 1984
+\bibitem[Reynaert, 2008]{reynaert08}
+Reyneart, 2008.
+\bibitem[Starlander \& Popescu-Belis, 2002]{sp02}
+Starlander \& Popescu-Belis, 2002
+\bibitem[Toutanova \& Moore, 2002]{tm02}
+Toutanova \& Moore, 2002.
+
+\end{thebibliography}
+
 % Ene paper Antal
 % roc101
-% (e.g. Agirre et al., 1998; Brill & Moore, 2000; Toutanova & Moore, 2002; Van Delden et al., 2004)
-% (Reynaert, 2008; Pollock & Zamora, 1984, although they use other terminology)
-% (Starlander & Popescu-Belis, 2002)
 

### 2012-11-08 Research training portfolio.tex
RTP: added abstract and more evaluation

@@
 \renewcommand{\familydefault}{\sfdefault}
+
+\let\stdsection\section
+\renewcommand\section{\newpage\stdsection}
 
@@
 
+\subsection*{Abstract}
+
+For many confusibles, like '/emph{you're} versus /emph{your}', one option is much more frequent than the other.
+This poses a problem when one tries to create a context-sensitive memory-based confusible corrector, as having more data for one option will make it more likely that the confusible corrector will predict that option.
+In some cases this can be solved by balancing the training data so that it has exactly the same amount of examles for each option, but in other cases this means a very large part of the data have to be thrown away, making the corrector even worse.
+In this paper I investigate %ja, wat eigenlijk?
+
 \subsection{Introduction}
+
+Dit is een confusible.
+Zo kan je ze herkennen.
+Dit hebben andere mensen gedaan.
+Dit zijn de confusibles waarnaar ik ga kijken.
+Dit zijn de verdelingen in mijn testset...
+kijk ze zijn scheef.
+Ik kan ze niet altijd weggooien, want dat wordt het nog erger.
 
@@
 In this section, I describe the various evaluation measures I used.
-This is necessary because evaluation of spelling corrector has not been very uniform in the spelling checker literature so far; whereas various scholars only measured how many errors their algorithm can detect in an error list and called it accuracy (e.g.
-Agirre et al., 1998; Brill \& Moore, 2000; Toutanova \& Moore, 2002; \citet{tm02,vandelden04}), others use or argue for the precision-recall terminology (Reynaert, 2008; Pollock \& Zamora, 1984, although they use other terms for the two concepts) or argue against it \citep{sp02}.\\\indent
+This is necessary because evaluation of spelling correctors has not been very uniform in the spelling checker literature so far; whereas various scholars only measured how many errors their algorithm can detect in an error list and called it accuracy (e.g.
+\citealp{agirre98, bm00, tm02,vandelden04}), others use or argue for the precision-recall terminology (\citealp{reynaert08,pz84}, although they use other terms for the two concepts) or argue against it \citep{sp02}.\\\indent
 
-Although most of the evaluation measures used can be useful and informative, they only tell part of the story.
+Although most of the evaluation measures used in the literature can be useful and informative, they only tell part of the story.
 I think it is important to recognize that spelling correction is a multi-step process, and each step comes with its own measures.
-Using a confusible corrector as part of a spelling corrector, there are three steps: word prediction, error detection and error correction.
+When using a confusible module as part of a spelling corrector, there are three steps: word prediction, error detection and error correction.
 I'll go through these steps and their evaluation measures one by one.
@@
 \item Prediction accuracy.
-Although word prediction may be the most complicated part of the confusible corrector, its evaluation is actually the simplest.
+Although word prediction may be the most complicated part of the confusible module, its evaluation is actually the simplest.
 Calculating prediction accuracy entails removing all confusibles from a text, and see in how many of the cases the language model can correctly predict which word was removed.
@@
 Results with multiple thresholds (i.e.
-to what extent does the prediction accuracy improve if the threshold is raised) can be visualized with a graph, as shown in the result section.
+to what extent does the prediction accuracy improve if the threshold is raised?) can be visualized with a graph, as shown in the result section.
 \end{itemize}
@@
 From this step onwards, we concentrate on spelling errors.
-This means that our input no longer is errorless text (like in the previous step).
+This means that our input no longer is errorless text, like in the previous step.
 We also interpret the results of the prediction differently now: whereas incorrectly predicted words were considered errors of the system in the previous task, lowering the accuracy, they will now be considered corrections.
@@
 
+When reporting the results of a confusible module, the names of the various cells will be replaced by numbers.
+These numbers are the data needed to calculate various evaluation measures, like the three below.
+Beside these evaluation measures, I will always give the confusion matrix itself, so readers can calculate other measures if desired.
+
 \item Detection accuracy. Detection accuracy only indicates how many mistakes the confusible detector makes, not what kind of mistakes. It can be calculated with the following formula:
@@
 
-It is similar to prediction accuracy in that it indicates how well the sytem does in general, but keep in mind that they're measuring two completely different concepts.
-Although a system with a good prediction accuracy also has a good detection accuracy, it might be possible that a system very good at predicting confusibles in errorless texts behaves differently in texts that do contain errors.
+It is similar to prediction accuracy in that it indicates how well the sytem does in general, but keep in mind that prediction and detection accuracy measure two completely different concepts.
+Although a system with a good prediction accuracy is very likely to also have a good detection accuracy, it might be possible that a system very good at predicting confusibles in errorless texts behaves differently in texts that do contain errors.
 
@@
 
-\paragraph{Error correction} Some text.
+\paragraph{Error correction}
+
+\begin{itemize}
+\item Confusion matrix.
+\item Detection accuracy.
+\item Detection recall.
+\item Detection precision.
+\end{itemize}
 

### 2012-11-10 Research paper.tex
RT: more about evaluation, busy with references in latex
RP: Completed rule-based approach, started on others

@@
 
-\section{Theoretical background and framework}
+\section{Theoretical background / framework}
 
-\subsection{Classical Machine Translation}
+In this section I will try to give a short overview of the various approaches in machine translation, with special attention to the example-based approach, as that is the approach used in this project.
+This section is mainly based on %Maartens thesis.
 
-Classical Machine Translation is best characterized by the use of human created rules.
+\subsection{The rule-based approach}
+
+The rule-based approach, also known as Classical Machine Translation, is best characterized by the use of human created rules.
 That is, authors of translation algorithms tried to include their knowledge about the languages and translation.
-There have been four main approaches, summarized in the Vauquois triangle: \\
+The rule-based approache can be subdivided into four subapproaches, summarized in the Vauquois triangle: \\
 
@@
 
-The horizontal arrows represent the approaches, I'll go through them from bottom to top:
+The horizontal arrows represent the approaches. I'll go through them from bottom to top:
 
 \begin{itemize}
-\item The direct approach does very little source language analysis, and could thus be viewed as the automatical version of looking up every word in the source language in a dictionary and replacing it with the corresponding word in the target language.
+\item \emph{The direct approach} does very little source language analysis, and could thus be viewed as the automatical version of looking up every word in the source language in a dictionary and replacing it with the corresponding word in the target language.
 This often doesn't yield very good results: if I for example would have translated this paper from my mother tongue Dutch to English word for word, 'would we now not real a good result have'.
-It should be noted, however, that direct approach can be a little more intelligent than that: %wwat dan?
-\item The syntactic approach
-\item The semantic approach
-\item All of the approaches so far have in common that they require a seperate rule set for each language pair; no or very few rules included for translations from English to German can also be used for translations from English to French.
-The Interlingua approach is an attempt to solve this problem, by pairing up every language with one general language, the Interlingua.
+It should be noted, however, that the direct approach mostly is more complicated than this; for example, in some cases morphological analysis is needed before it can be looked up in a dictionary, and more morphological analysis is needed for the translation word to fit in the sentence.
+Word reordering rules can also be included.
+\item \emph{The syntactic approach} tries to find out the syntactic structure of the source sentence before it starts translating, using a parser.
+This structure is then translated to the target language, and filled with translated words.
+\item \emph{The semantic approach} goes even one step further and also tries to grasp the meaning of a sentence, roughly.
+For a sentence like /emph{Mummy kisses daddy} this for example entails that mummy is the agent, and daddy the patient of the sentence.
+After having identified these semantic roles, the algorithm would then look how this meaning is expressed in the target language, find the approriate syntactic structure for that, and fill it with the translated words.
+\item \emph{The interlingua approach}.
+All of the approaches so far have in common that they require a seperate rule set for each language pair; no or very few rules included for translations from English to German can also be used for translations from English to French.
+The Interlingua approach is an attempt to solve this problem, by pairing up every language with a formal representation of meaning, the Interlingua.
 This means a translation from English to German would entail a translation from English to the Interlingua (using the ruleset English $\rightarrow$ Interlingua) and then a translation from the Interlingua to German (using the ruleset Interlingua $\rightarrow$ German).
 When translating from English to French, the same two steps would be taken, and the same ruleset English $\rightarrow$ Interlingua can be used.
-This Interlingua is a formal representation.
 \end{itemize}
 
-% Voorbeelden?
+An advantage of the rule-based approach is that one has a lot of control over the output.
+In theory, if something has gone wrong, all one has to do is find out which rule or dictionary item caused the problem and improve it.
+In practice, however, it is almost impossible to have a dictionary and rule system good enough for more complicated sentences, and things like rule interactions and ambiguity make everything even more complicated.
+The next two approaches therefore replace hand-crafted rules with automatically generated ones.
 
-\subsection{Statistical Machine Translation}
+% Voorbeelden? References?
+
+\subsection{The statistical approach}
+
+Unlike the ruled-based approach, the statistical approach creates its own 'knowledge', using bilingual text corpora.
+With this knowledge, the algorithm is able to generate various translation hypotheses, from which the best one is chosen with various statistical measures.
+I'll explain these stages in more detail in the following subsections.
+
+\subsubsection{Corpus alignment}
+
+Having a bilingual corpus does not mean one .
+
+\subsubsection{Hypothesis generation}
+
+\subsubsection{Hypothesis evaluation}
+
+\paragraph{The translation model}
+
+\paragraph{The language model}
+
+\subsection{The example-based approach}
+
+In this section, I'll give an overview of the example-based approach.
+I'll focus on the memory-based version of the example-based approach in particular, because that is the version the Colibri project uses.
+Other versions of example-based machine translation might vary slightly.
+\\\indent
+The memory-based approach is in many ways similar to the statistical approach - it too generates translations on the basis of large bilingual corpora, for instance.
+The main difference, however, is that it is context-sensitive; the idea is that the system will find another translation for \emph{left} in the sentence \emph{The toilets are at your left hand} than \emph{left} in the sentence \emph{Elvis has left the building}.
 

### 2012-11-11 Research paper.tex
RP: completed early version of statistical and example based

@@
 \usepackage{graphicx}
+\usepackage{natbib}
 
@@
 The Colibri project is carried out by Maarten van Gompel and supervised by Antal van den Bosch.
+
+%Beschrijving van ik ga doen.
 
@@
 With this knowledge, the algorithm is able to generate various translation hypotheses, from which the best one is chosen with various statistical measures.
-I'll explain these stages in more detail in the following subsections.
+I will explain these stages in more detail:
 
-\subsubsection{Corpus alignment}
+\begin{itemize}
+\item \emph{Corpus alignment.} Only having a bilingual corpus is not enough: for such corpus to be useful one also needs to know which parts of the sentence in language A correspond to which part of the sentence B.
+This is not only complicated by different word orders, but also by concepts that in language A take more or less words than in language B.
+A well-known word alignment algorithm is the Expectation Maximisation algorithm \citep{dempster77}.
 
-Having a bilingual corpus does not mean one .
+%GIZA++?
 
-\subsubsection{Hypothesis generation}
+\item \emph{Hypothesis generation.} We now have some sort of automatically generated dictionary: for each word in language A, we know which words in language B are often used for it.
+On the basis of this knowledge, we can generate various possible translations.
 
-\subsubsection{Hypothesis evaluation}
+\item \emph{Hypothesis evaluation.} Which one of these possible translation is the best one is determined on the basis of two measures: faithfulness and fluency (which can also be called grammaticality).
+These measures often are in conflict: a solution to get an optimally fluent translation would be to always return the same well-formed sentence, but that would not be faithful to the source language at all, while a very faithful translation, for example translating a sentence word by word, is usually not considered very fluent.
+Faithfulness and fluency are measured by the translation model and the language model respectively:
 
-\paragraph{The translation model}
+\begin{itemize}
+\item \emph{The translation model.} The translation model returns a probability of source sentence given the target sentence; it thus looks back whether what has been created fits the original.
+This probability is calculated on the basis of statistical word alignment algorithms like the IBM models \citep{brown93}.
 
-\paragraph{The language model}
+\item \emph{The language model.} The language model predicts to what extent the generated sentence is expected for the target language; in other words, is this sentence a well-formed sentence or not?
+Because the unlimited number of ways words can be combined into sentences, it is unlikely the sentence under investigation will actually be found in a corpus, so the language model tries to do this prediction on the basis of how often smaller parts of the sentence (n-grams) occur in a corpus.
+
+\end{itemize}
+
+\end{itemize}
 
@@
 The main difference, however, is that it is context-sensitive; the idea is that the system will find another translation for \emph{left} in the sentence \emph{The toilets are at your left hand} than \emph{left} in the sentence \emph{Elvis has left the building}.
+\\\indent
+In the memory-based approach, this context-sensitivy is achieved with the help of machine learning: by showing a so-called classifier lots of examples of something, and telling it in which class it belongs, it is able to predict the class of new examples.
+For example, we can give a classifier the external characteristics of 10.000 humans (for example length, type of cloths, length of hair), and also tell it whether these humans are male or female.
+If we then ask for the sex of a relatively short human, wearing a dress, with long hair, it will probably be able to tell us this is a woman.
+While this example only has two classes, it is in principle possible to use an unlimited number of classes.
+This is exactly what the memory-based approach does: on the basis of a particular word and its context (for example, its direct neighbours), a classifier in machine translation returns as a class the most likely n-gram.
+Thus, instead of human characteristics the classifiers are trained on words and their context from the source language, and instead of sex it uses n-grams from the target language as classes.
+\\\indent
+There are two types of classifaction methods: \emph{eager learning} and \emph{lazy learning}.
+Whereas eager learning algorithms try to creating a small, abstract, general model, filtering out infrequent cases, lazy learning algorithms take into account everything they encounter.
+According to \citet{dvdb05}, for natural language processing the lazy learning approach works best, because infrequent cases are an important part of the knowledge needed.
+\\\indent
+Like I did with statistical machine translation, I will now give an overview of the memory-based machine translation process, based on \citet{vdbb09}.
+For this example, I will describe a translator that uses trigrams, but it should be noted that the Colibri system uses n-grams of various lengths.
+
+\begin{itemize}
+\item \emph{Local phase.} In the description of \citet{vdbb09}, the local phase entails both training the classifier and actual classification, but I think it is important to make a sharp distinction here: whereas the training is done only once and \emph{before} actual translation, and could therefore be seen as 'preparing' the translation software for its task, the classification is done every time the user asks for a translation.
+Training consists of (1) aligning the bilingual corpus, using the word alignment algorithms described in the previous sections, (2) decomposing the corpus into trigrams, and (3) giving the results to the classifier as training material.
+The trigrams of the source language constitute the examples, the trigrams of the target language are the classes.
+Classification means (1) decomposing the source text into trigrams and (2) classifying each trigram.
+We end up with a list of trigrams in the target language.
+
+\item \emph{Global phase.} In the global phase we try to merge the trigrams into one text.
+The part of the translation software responsible for this is called the \emph{decoder}.
+There might be a lot of overlap in the resulting trigram (for example: '\emph{\_ I love}', '\emph{I love you}' and '\emph{love you \_}'), making the decoder's task relatively easy.
+Such overlap might even give hints for the word order of the target language.
+In many cases, unfortunately, such overlap does not exist or is even misleading.
+As with statistical translation, here too we can pick the best translation hypothesis by using a language model and a translation model.
+
+% Constraint Satisfaction Inference
+
+\end{itemize}
+
+%Moet ik hier iets zeggen over phrase based?
 
@@
 
-\section{References}
+\begin{thebibliography}{99}
 
+\bibitem[Van den Bosch \& Berk, 2009]{vdbb09}
+Van den Bosch \& Berk, 2009
+
+\bibitem[Brown et al., 1993]{brown93}
+Brown et al., 1993
+
+\bibitem[Daelemans \& van den Bosch, 2005]{dvdb05}
+Daelemans \& van den Bosch, 2005
+
+\bibitem[Dempster et al., 1997]{dempster77}
+Dempster et al., 1997
+
+\end{thebibliography}
 % Documentatie Colibri

### 2012-11-12 Research training portfolio.tex
RTP: Extended abstract, started with intro

@@
 
+\usepackage{covington}
 \usepackage{graphicx}
@@
 
-For many confusibles, like '/emph{you're} versus /emph{your}', one option is much more frequent than the other.
+For many confusibles, like "\emph{you're} versus \emph{your}", one option is much more frequent than the other.
 This poses a problem when one tries to create a context-sensitive memory-based confusible corrector, as having more data for one option will make it more likely that the confusible corrector will predict that option.
 In some cases this can be solved by balancing the training data so that it has exactly the same amount of examles for each option, but in other cases this means a very large part of the data have to be thrown away, making the corrector even worse.
-In this paper I investigate %ja, wat eigenlijk?
+In this paper I investigate in which cases it is better to use a balanced training set, and in which cases it is better to train the modules on the larger, unbalanced set.
+I do this by building x confusible correctors and evaluating their results.
+
+%hoeveel?
 
@@
 
+This paper results from my attempts to solve a rather practical issue while building Fowlt, an online spell checker freely available on the internet.
+Fowlt, just like its Dutch equivalent Valkuil, is a context-sensitive spell checker.
+This context-sensitivity makes it possible to detect errors like the following:
+
+\begin{examples}
+
+\item There really nice people. \emph{Intended: they're.}
+\item I'd like too see you two. \emph{Intended: to and too.}
+\item I don't want to loose you. \emph{Intended: lose.}
+
+\end{examples}
+
+These errors, in which one word is confused with another (often homophone) word, are called \emph{confusible errors}.
+For the most common confusible errors, Fowlt contains a specialized modules.
+Unlike typos, confusible errors cannot be detected by simply looking at the text word for word, because the error is in the relation of a particular word with the other words in the sentence.
+Therefore, the confusible modules detect them by looking at the context of the word, and see if that differs from the contexts they have seen in the past for that word.
+If it, on the basis of the context, would predict another word in that position, and is really certain about it, it flags the word as an error.
+This means, for example, that Fowlt is able to replace the incorrect "there" in the last example into "they're", simply because "there" usually isn't followed by "really nice people", while "they're" is.
+\\\indent
+This knowledge was not inserted manually, but was created with the help of Machine Learning software TiMBL.
+A more detailed account of its workings will be given in the next section.
+For now it is important to understand that TiMBL extracts this knowledge from large text corpora.
+From these corpora, all examples of the confusible have been extracted together with their content.
+The confusible corrector's results improve when TiMBL has seen more training material to base its decision on.\\\indent
+Fowlt contains language models, and thus specialized confusible correctors, for the following confusibles (among others):
+
+Dit is Fowlt.
 Dit is een confusible.

### 2012-11-13 Research paper.tex
RP: Swapped background and project description
Added CLIN-abstract, RP: added sentence about phrase-based mt + references

@@
 The Colibri project is carried out by Maarten van Gompel and supervised by Antal van den Bosch.
+\\\indent
+In section 2, I'll give a short overview of what has been achieved in machine translation, and in section 3 the current project will be described.
+This order was chosen to show that Colibri builds on previous research directly.
 
-%Beschrijving van ik ga doen.
-
-\section{Project description}
-
-Besides an attempt to discover more about the possibilities for machine translation technology, Colibri is also an attempt to actually develop this technology.
-The following project overview therefore necessarily also is an overview of the software.
-
-\subsection{Pattern extraction}
-
-Colibri, 'Constructions as Linguistics Bridges', investigates constructions (or patterns), as its name suggests.
-A construction can be any group of consecutive words (a so-called 'n-gram') in any natural language that in some way forms an entity.
-An example is 'on the basis of' in the following sentence:
-
-\begin{examples}
-\item On the basis of these ideas, software can be developed.
-\end{examples}
-
-Importantly, constructions can also have one or more gaps (so-called 'skip-grams'). 'from \_\_ point of view' in the following example, is a construction with one such gap.
-
-\begin{examples}
-\item I understand things better when I look at them from his point of view.
-\end{examples}
-
-The gap is filled here is filled with 'his', but the construction 'from \_\_ point of view' can also be filled with many other words ('my', 'the', 'another', 'Obama's', etc.).
-This shows the advantage of skip-grams over n-grams for machine translation: whereas n-grams would only have been able to capture the construction 'point of view', because the word directly before 'point of view' is variable, skip-grams are also able to recognize that the word 'for' is part of the construction.\\\indent
-Of course, not every possible group of consecutive words is a construction; what makes some groups special?
-The exact nature of linguistic constructions has been the subject of many publications and even an entire linguistic framework (construction grammar).
-For this project however, it suffices to say that constructions emerge because some combinations of words are more frequent than others.
-
-% Drie vormen freq: raw frequency, context frequency, multilingual alignmetn
-% Blabla, want freq is waar mach learning gebriuk van maakt
-
-\subsection{Pattern querying} Interactively query generated models for specific patterns.
-\subsection{Pattern/corpus comparison} Match patterns generated from one corpus to another corpus.
-\subsection{Graph computation \& visualisation} Compute graph relations between pattern models and visualise graphs
-\subsection{Alignment algorithms} Algorithms to extract and align patterns accross languages; resulting in translation pairs for constructions.
-\subsection{Machine Translation Decoding} Reassembles translated fragments into one coherent sentence in the target-language; seeking the best (statistically most probably) solution in a vast search space of possible search hypotheses.
-\subsection{Memory-based Machine Translation} Machine learning algorithms can be used to consider source-side context in translations. Classifiers can be built and directly used by the decoder.
-\subsection{Experiment Framework} Colibri comes with an extensive experiment framework for conducting Machine Translation experiments.\
-
-\subsection{Hypotheses} %Of moet dit bij main outstanding questions?
-
-Constructions can be found efficiently in corpus data
-Graph-based relations can be used to constrain to good constructions
-Constructions can be aligned without resorting to word alignments as a basis
-In MT, constructions (i.e. possibly with gaps) result in better translation than mere consecutive phrases
+%Rest van de beschrijving.
 
@@
 In this section I will try to give a short overview of the various approaches in machine translation, with special attention to the example-based approach, as that is the approach used in this project.
-This section is mainly based on %Maartens thesis.
+This section is mainly based on \citet{vangompel09}
 
@@
 
-%Moet ik hier iets zeggen over phrase based?
+On a final note, it should be added that although I focussed word-based translation (as described in \citet{stroppa07} and \citet{vdbb09}) in the previous examples,  \citet{vangompel09}, \citet{vangompelea09} and \citet{vangompel11} showed that even better results can be achieved using phrases of variable length.
+Investigating the advantage of multi-word translation units is one of the main goals of the Colibri project, as will be explained in the next section.
+
+\section{Project description}
+
+Besides an attempt to discover more about the possibilities for machine translation technology, Colibri is also an attempt to actually develop this technology.
+The following project overview therefore necessarily also is an overview of the software.
+
+\subsection{Pattern extraction}
+
+Colibri, 'Constructions as Linguistics Bridges', investigates constructions (or patterns), as its name suggests.
+A construction can be any group of consecutive words (a so-called 'n-gram') in any natural language that in some way forms an entity.
+An example is 'on the basis of' in the following sentence:
+
+\begin{examples}
+\item On the basis of these ideas, software can be developed.
+\end{examples}
+
+Importantly, constructions can also have one or more gaps (so-called 'skip-grams'). 'from \_\_ point of view' in the following example, is a construction with one such gap.
+
+\begin{examples}
+\item I understand things better when I look at them from his point of view.
+\end{examples}
+
+The gap is filled here is filled with 'his', but the construction 'from \_\_ point of view' can also be filled with many other words ('my', 'the', 'another', 'Obama's', etc.).
+This shows the advantage of skip-grams over n-grams for machine translation: whereas n-grams would only have been able to capture the construction 'point of view', because the word directly before 'point of view' is variable, skip-grams are also able to recognize that the word 'for' is part of the construction.\\\indent
+Of course, not every possible group of consecutive words is a construction; what makes some groups special?
+The exact nature of linguistic constructions has been the subject of many publications and even an entire linguistic framework (construction grammar).
+For this project however, it suffices to say that constructions emerge because some combinations of words are more frequent than others.
+
+% Drie vormen freq: raw frequency, context frequency, multilingual alignmetn
+% Blabla, want freq is waar mach learning gebriuk van maakt
+
+\subsection{Pattern querying} Interactively query generated models for specific patterns.
+\subsection{Pattern/corpus comparison} Match patterns generated from one corpus to another corpus.
+\subsection{Graph computation \& visualisation} Compute graph relations between pattern models and visualise graphs
+\subsection{Alignment algorithms} Algorithms to extract and align patterns accross languages; resulting in translation pairs for constructions.
+\subsection{Machine Translation Decoding} Reassembles translated fragments into one coherent sentence in the target-language; seeking the best (statistically most probably) solution in a vast search space of possible search hypotheses.
+\subsection{Memory-based Machine Translation} Machine learning algorithms can be used to consider source-side context in translations. Classifiers can be built and directly used by the decoder.
+\subsection{Experiment Framework} Colibri comes with an extensive experiment framework for conducting Machine Translation experiments.\
+
+\subsection{Hypotheses} %Of moet dit bij main outstanding questions?
+
+Constructions can be found efficiently in corpus data
+Graph-based relations can be used to constrain to good constructions
+Constructions can be aligned without resorting to word alignments as a basis
+In MT, constructions (i.e. possibly with gaps) result in better translation than mere consecutive phrases
 
@@
 
-\bibitem[Van den Bosch \& Berk, 2009]{vdbb09}
-Van den Bosch \& Berk, 2009
+\bibitem[Van den Bosch \& Berck, 2009]{vdbb09}
+Van den Bosch, A., \& Berck, P. (2009). Memory-based machine translation and language
+modeling. The Prague Bulletin of Mathematical Linguistics, 91, 17\-26.
 
@@
 
+\bibitem[Van Gompel, 2009]{vangompel09}
+Van Gompel, M. Phrase-based memory-based machine translation. Unpublished
+master's thesis, Tilburg University.
+
+\bibitem[Van Gompel et al., 2009]{vangompelea09}
+Van Gompel, M., Van den Bosch, A., \& Berck, P. (2009). Extending memory-based
+machine translation to phrases. In M. Forcada \& A. Way (Eds.), Proceedings of the
+third workshop on example-based machine translation (pp. 79\-86). Dublin, Ireland.
+
+\bibitem[Van Gompel et al., 2011]{vangompel11}
+Van Gompel, M., Van den Bosch, A., \& Berck, P. (2011). Extending memory-based
+machine translation to phrases. In T. Markus, P. Monachesi, \& E. Westerhout (Eds.),
+Computational linguistics in the netherlands 2010: Selected papers from the twentieth
+clin meeting (pp. 45\-58). Utrecht, the Netherlands: LOT.
+
+\bibitem[Stroppa et al., 2007]{stroppa07}
+Stroppa, N., Van den Bosch, A., \& Way, A. (2007). Exploiting source similarity for SMT
+using context-informed features. In A. Way \& B. Gawronska (Eds.), Proceedings of the
+11th international conference on theoretical issues in machine translation (tmi 2007)
+(pp. 231\-240). Skövde, Sweden.
+
 \end{thebibliography}

### 2012-11-13 abstract.tex
Added CLIN-abstract, RP: added sentence about phrase-based mt + references

\documentclass[12pt]{article}

\title{The inside of an English context-sensitive spell checker}
\author{Wessel Stoop, Antal van den Bosch\\ \emph{Radboud University Nijmegen}}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}
\maketitle

In this talk, we explain the workings of fowlt.net, an online spell checker freely available on the internet. Fowlt, just like it's Dutch equivalent Valkuil, is a context-sensitive spell checker, using memory based learning software Timbl and word prediction software WOPR; the general idea is that a certain word is flagged as an error if our language models expected another word there and are very certain about it. This makes detection of errors like the one in the second sentence of this abstract possible. Fowlt consist of multiple modules, all approaching the task of detecting and correcting errors from a different viewpoint. We will explain these modules, how they work together, and discuss the practical problems we encountered while preparing them for their task.

\end{document}

### 2012-11-13 Research training portfolio.tex
RTP: added formulas, F-score header
RPT: reviewed some things, added more introduction

@@
 This poses a problem when one tries to create a context-sensitive memory-based confusible corrector, as having more data for one option will make it more likely that the confusible corrector will predict that option.
-In some cases this can be solved by balancing the training data so that it has exactly the same amount of examles for each option, but in other cases this means a very large part of the data have to be thrown away, making the corrector even worse.
+In some cases this can be solved by balancing the training data so that it has exactly the same amount of examples for each option, but in other cases this means a very large part of the data have to be thrown away, making the corrector even worse.
 In this paper I investigate in which cases it is better to use a balanced training set, and in which cases it is better to train the modules on the larger, unbalanced set.
@@
 These errors, in which one word is confused with another (often homophone) word, are called \emph{confusible errors}.
-For the most common confusible errors, Fowlt contains a specialized modules.
+For most common confusible errors, Fowlt contains specialized modules.
 Unlike typos, confusible errors cannot be detected by simply looking at the text word for word, because the error is in the relation of a particular word with the other words in the sentence.
@@
 For now it is important to understand that TiMBL extracts this knowledge from large text corpora.
-From these corpora, all examples of the confusible have been extracted together with their content.
-The confusible corrector's results improve when TiMBL has seen more training material to base its decision on.\\\indent
+From these corpora, all examples of the confusible have been extracted together with their context.
+The confusible corrector's results improve when TiMBL has seen more training material to base its language models on.\\\indent
 Fowlt contains language models, and thus specialized confusible correctors, for the following confusibles (among others):
 
-Dit is Fowlt.
-Dit is een confusible.
-Zo kan je ze herkennen.
-Dit hebben andere mensen gedaan.
-Dit zijn de confusibles waarnaar ik ga kijken.
-Dit zijn de verdelingen in mijn testset...
-kijk ze zijn scheef.
-Ik kan ze niet altijd weggooien, want dat wordt het nog erger.
+What the confusible modules within Fowlt for each of these confusibles do is removing the confusible words and predicting which word was removed on the basis of the context.
+If they predicted the word incorrectly, BUT are very certain about it, they assume the original word was wrong and flag it as an error.
+\\\indent
+Ideally, a language model is created on the basis of a balanced training set, i.e.
+a training set that consists of roughly the same amount of examples for each confusible option.
+If we would use a training set with much more contexts for option A than for option B, which may bias the model to option A.
+For example, if the word 'cookie' can occur perfectly both in the context of option A and option B, when seeing 'cookie' in a particular context the confusible corrector should not be able to make a prediction on the basis of this word alone.
+However, if the language model is based on an unbalanced set it will predict option A, simply because A occured together with 'cookie' more often.
+\\\indent
 
@@
 
-%formula
+\[
+\frac{tp+tn}{tp+tn+fp+fn}
+\]
 
@@
 
-\item Detection recall.
-\item Detection precision.
+\item Detection recall and detection precision.
+
+\[
+\frac{tp}{tp+fn}
+\]
+
+\[
+\frac{tp}{tp+fp}
+\]
+
+\item Detection F-score.
+
+\[
+2 * \frac{precision * recall}{precision + recall}
+\]
+
 \end{itemize}
@@
 \item Confusion matrix.
-\item Detection accuracy.
-\item Detection recall.
-\item Detection precision.
+\item Correction accuracy.
+\item Correction recall and correction precision.
+\item Correction F-score.
 \end{itemize}
@@
 % Ene paper Antal
-% roc101
 

### 2012-11-14 Research paper.tex
RP: worked on project description, added headers for criticism, revised some text

@@
 This Universal Translator can (1) recognize speech in every language, (2) translate what it has 'heard' to any other language, (3) make its results audible with speech synthesis and (4) learn itself new languages when needed.
-Unfortunately, like many technologies in Star Trek, this is far from reality: there is lots of active research at the first three tasks in particular, but no software currently in existence can do this as perfect as the Universal Translator.
+Unfortunately, like many technologies in Star Trek, this is far from reality: there is lots of active research at the first three tasks in particular (and even some of the fourth, see \citet{biemann11}), but no software currently in existence can do this as perfect as the Universal Translator.
 \\\indent
@@
 The memory-based approach is in many ways similar to the statistical approach - it too generates translations on the basis of large bilingual corpora, for instance.
-The main difference, however, is that it is context-sensitive; the idea is that the system will find another translation for \emph{left} in the sentence \emph{The toilets are at your left hand} than \emph{left} in the sentence \emph{Elvis has left the building}.
+The main difference, however, is that it is context-sensitive; the idea is that the system will find another translation for \emph{left} in the sentence \emph{The toilets are at your left hand} than for \emph{left} in the sentence \emph{Elvis has left the building}.
 \\\indent
-In the memory-based approach, this context-sensitivy is achieved with the help of machine learning: by showing a so-called classifier lots of examples of something, and telling it in which class it belongs, it is able to predict the class of new examples.
-For example, we can give a classifier the external characteristics of 10.000 humans (for example length, type of cloths, length of hair), and also tell it whether these humans are male or female.
-If we then ask for the sex of a relatively short human, wearing a dress, with long hair, it will probably be able to tell us this is a woman.
+In the memory-based approach, this context-sensitivy is achieved with the help of machine learning: by showing a so-called classifier lots of examples of something, and telling it to which class it belongs, the classifier is able to predict the class of new examples.
+For example, we can give a classifier the external characteristics of 10.000 humans (length, type of cloths, length of hair, etc.), and also tell it whether these humans are male or female.
+If we then ask for the sex of a relatively short human, wearing a dress, with long hair, the classifier will probably be able to tell us this is a woman.
 While this example only has two classes, it is in principle possible to use an unlimited number of classes.
-This is exactly what the memory-based approach does: on the basis of a particular word and its context (for example, its direct neighbours), a classifier in machine translation returns as a class the most likely n-gram.
-Thus, instead of human characteristics the classifiers are trained on words and their context from the source language, and instead of sex it uses n-grams from the target language as classes.
+This is exactly what the memory-based approach to machine translation does: on the basis of a particular word and its context (for example, its direct neighbours), a classifier in machine translation returns as a class the most likely n-gram.
+Thus, instead of on human characteristics the classifiers are trained on words and their context from the source language, and instead of sex it uses n-grams from the target language as classes.
 \\\indent
 There are two types of classifaction methods: \emph{eager learning} and \emph{lazy learning}.
-Whereas eager learning algorithms try to creating a small, abstract, general model, filtering out infrequent cases, lazy learning algorithms take into account everything they encounter.
+Whereas eager learning algorithms try to create a small, abstract, general model, filtering out infrequent cases, lazy learning algorithms take into account everything they encounter.
 According to \citet{dvdb05}, for natural language processing the lazy learning approach works best, because infrequent cases are an important part of the knowledge needed.
@@
 Like I did with statistical machine translation, I will now give an overview of the memory-based machine translation process, based on \citet{vdbb09}.
-For this example, I will describe a translator that uses trigrams, but it should be noted that the Colibri system uses n-grams of various lengths.
+For this example, I will describe a translator that uses trigrams, but all kinds of n-grams could be used here, even n-grams of variable length.
+In fact, \citet{vangompel09}, \citet{vangompelea09} and \citet{vangompel11} showed that better results can be achieved using phrases of variable length; investigating the nature of multi-word units in translation therefore is one of the main goals of the Colibri project.
 
@@
 \item \emph{Local phase.} In the description of \citet{vdbb09}, the local phase entails both training the classifier and actual classification, but I think it is important to make a sharp distinction here: whereas the training is done only once and \emph{before} actual translation, and could therefore be seen as 'preparing' the translation software for its task, the classification is done every time the user asks for a translation.
-Training consists of (1) aligning the bilingual corpus, using the word alignment algorithms described in the previous sections, (2) decomposing the corpus into trigrams, and (3) giving the results to the classifier as training material.
+Training consists of (1) aligning the bilingual corpus, using the word alignment algorithms described in the previous sections, (2) decomposing the complete corpus into trigrams, and (3) giving the results to the classifier as training material.
 The trigrams of the source language constitute the examples, the trigrams of the target language are the classes.
@@
 We end up with a list of trigrams in the target language.
+
+%vdb b 09 ook echt lezen
 
@@
 
-On a final note, it should be added that although I focussed word-based translation (as described in \citet{stroppa07} and \citet{vdbb09}) in the previous examples,  \citet{vangompel09}, \citet{vangompelea09} and \citet{vangompel11} showed that even better results can be achieved using phrases of variable length.
-Investigating the advantage of multi-word translation units is one of the main goals of the Colibri project, as will be explained in the next section.
-
 \section{Project description}
 
-Besides an attempt to discover more about the possibilities for machine translation technology, Colibri is also an attempt to actually develop this technology.
-The following project overview therefore necessarily also is an overview of the software.
+In this section, I will describe the main idea behind the project (subsection 1), and how this is operationalized (section 2).
+Where possible, I will add information of the current status of the project.
 
-\subsection{Pattern extraction}
+\subsection{The building blocks: constructions}
 
 Colibri, 'Constructions as Linguistics Bridges', investigates constructions (or patterns), as its name suggests.
-A construction can be any group of consecutive words (a so-called 'n-gram') in any natural language that in some way forms an entity.
+A construction can be any group of consecutive words in any natural language that in some way forms an entity.
 An example is 'on the basis of' in the following sentence:
@@
 
-Importantly, constructions can also have one or more gaps (so-called 'skip-grams'). 'from \_\_ point of view' in the following example, is a construction with one such gap.
+Note that a simple trigram approach couldn't have discoverd this construction.
+Importantly, constructions can also have one or more gaps (so-called 'skip-grams').
+'From \_\_ point of view' in the following example, is a construction with one such gap.
 
@@
 Of course, not every possible group of consecutive words is a construction; what makes some groups special?
-The exact nature of linguistic constructions has been the subject of many publications and even an entire linguistic framework (construction grammar).
-For this project however, it suffices to say that constructions emerge because some combinations of words are more frequent than others.
+The exact nature of linguistic constructions has been the subject of many lignuistic publications and even an entire linguistic framework (construction grammar).
+For this project, however, it suffices to say that constructions emerge because some combinations of words are more frequent than others.
 
+% Referentie construction grammar
 % Drie vormen freq: raw frequency, context frequency, multilingual alignmetn
@@
 
-\subsection{Pattern querying} Interactively query generated models for specific patterns.
-\subsection{Pattern/corpus comparison} Match patterns generated from one corpus to another corpus.
-\subsection{Graph computation \& visualisation} Compute graph relations between pattern models and visualise graphs
-\subsection{Alignment algorithms} Algorithms to extract and align patterns accross languages; resulting in translation pairs for constructions.
-\subsection{Machine Translation Decoding} Reassembles translated fragments into one coherent sentence in the target-language; seeking the best (statistically most probably) solution in a vast search space of possible search hypotheses.
-\subsection{Memory-based Machine Translation} Machine learning algorithms can be used to consider source-side context in translations. Classifiers can be built and directly used by the decoder.
-\subsection{Experiment Framework} Colibri comes with an extensive experiment framework for conducting Machine Translation experiments.\
+\subsection{The builder: machine translation}
+Besides an attempt to discover more about the possibilities for machine translation technology, Colibri is also an attempt to actually develop this technology.
+A project overview should be complete without an overview of what the software developed.
+Colibri's functionality can be split in four parts: (1) identifying and extracting constructions in a language, (2) identifying aligned constructions between language-pairs, (3) memory-based classification using construction experts and (4) reassembling constructions into a coherent target sentence.
+If we look at these parts int he light of the light of the memory-based machine translation process as described in the previous section, the first three parts would constitute the local phase, while part 4 would constitute the global phase.
+I will discuss the four parts in more detail consecutively:
 
+\subsubsection{Identifying and extracting constructions in a language}
+When using large corpora, the process of discovering simple n-grams might already take quite a lot of time; adding variability in lenght and one or multiple gaps to the n-grams makes the process even more - in fact, many times more - complex.
+However, as I write this, this problem has been solved by using two techniques: (1) by optimizing how save the various tokens and the n-grams and skipgrams they are part of in memory, and (2) by using an iterative counting algorithm to limit storage.\\\indent
+Besides identifying the constructions, various frequency statistics can be given. These frequencies can then be used to create graphs which display the relations between the various constructions.
+
+%Hoe is dit probleem precies opgelost?
+
+\subsubsection{Identifying aligned constructions between language-pairs}
+The classical approach for aligning constructions from a bilingual corpus is to first align the individual words, and then align the construction themselves \citep{koehn03}.
+As part of the Colibri project, it was attempted to skip the first step, and try to align the construction directly.
+To achieve this, it was simply measured how often the constructions occur together, among other things.
+Because at this stage the decoder was not yet ready (see subsection 4), the decoder from translation system MOSES was used.
+This showed that the approach using constructions for alignment could not provide results as good as the approach using words can; therefore, it was decided to go back to the classical method.
+It should be noted that, because the MOSES decoder was used, the influence of skipgrams on alignment could not be measured.
+However, because they increase complexity, they are not expected to improve the alignments.\\\indent
+
+\subsubsection{Memory-based classification using construction experts}
+\subsubsection{Reassembling constructions into a coherent target sentence}
 \subsection{Hypotheses} %Of moet dit bij main outstanding questions?
@@
 \section{Strengths and weaknesses}
+
+\subsection{Weaknesses}
+
+\subsubsection{Machine translation and the status of translators}
+
+Machine translation is far from perfect, and although this project is an interesting attempt to bring the technology closer to the 'perfect'-end, it will almost certainly not result in an application which is able to replace human translators.
+However, when non-perfect software like this is made available to the public, ignorant users might however treat it as such.
+A similar critisism can be heard in the field of spell checkers:
+
+%Overzicht kritiek, zie Wikipedia
+
+Since the release of Google Translate, various examples can be found on the internet of websites and other texts translated with this service.
+
+\subsubsection{Translation as an art}
+
 \section{Possibilities for future research}
@@
 \begin{thebibliography}{99}
+
+\bibitem[Biemann, 2011]{biemann11}
+Biemann, C. (2011). Structure discovery in natural language. New York / Dordrecht: Springer.
 
@@
 
+\bibitem[Koehn et al, 2003]{koehn03}
+Koehn et al, 2003
+
 \bibitem[Stroppa et al., 2007]{stroppa07}

### 2012-11-15 Research training portfolio.tex
RTP: worked on prec and recall

@@
 \item Detection recall and detection precision. 
+Precision and recall are concepts well-known within the field of Information Retrieval; precision reflects how much of the retrieved things were relevant, while recall reflects how much of the the relevant things were retrieved.
+Often there is a inverse relationship between precision and recall: if you change an algorithm so that its recall increases, its precision decreases, and the other way around.
+In case of spelling checkers, this means you either have a system that catches a lot of errors, but also has a lot of false alarms, or a system that only finds true errors, but also misses a lot; you can't have the best of both.
 
@@
 \item Detection F-score. 
+Although you mostly cannot have an optimal precision \emph{and} recall at the same time, you can of course try to have them both as high as possible.
+Calculating the mean does not help here, because the mean can be relatively high if one of the two is high, but what you want to know is whether \emph{both} are high.
 

### 2012-11-17 Research paper.tex
RP: Added a more strenghts and weaknesses, completed project description, moved hypotheses

@@
 Besides an attempt to discover more about the possibilities for machine translation technology, Colibri is also an attempt to actually develop this technology.
-A project overview should be complete without an overview of what the software developed.
+A project overview therefore should not be complete without an overview of the software's functionality.
 Colibri's functionality can be split in four parts: (1) identifying and extracting constructions in a language, (2) identifying aligned constructions between language-pairs, (3) memory-based classification using construction experts and (4) reassembling constructions into a coherent target sentence.
@@
 When using large corpora, the process of discovering simple n-grams might already take quite a lot of time; adding variability in lenght and one or multiple gaps to the n-grams makes the process even more - in fact, many times more - complex.
-However, as I write this, this problem has been solved by using two techniques: (1) by optimizing how save the various tokens and the n-grams and skipgrams they are part of in memory, and (2) by using an iterative counting algorithm to limit storage.\\\indent
-Besides identifying the constructions, various frequency statistics can be given. These frequencies can then be used to create graphs which display the relations between the various constructions.
+However, as I write this, this problem has been solved by using two techniques: (1) by optimizing how the various tokens and the n-grams and skipgrams they are part of are saved in memory, and (2) by using an iterative counting algorithm to limit storage.
+Besides identifying the constructions, various frequency statistics can be given.
+These frequencies can then be used to create graphs which display the relations between the various constructions.
 
@@
 The classical approach for aligning constructions from a bilingual corpus is to first align the individual words, and then align the construction themselves \citep{koehn03}.
-As part of the Colibri project, it was attempted to skip the first step, and try to align the construction directly.
-To achieve this, it was simply measured how often the constructions occur together, among other things.
-Because at this stage the decoder was not yet ready (see subsection 4), the decoder from translation system MOSES was used.
-This showed that the approach using constructions for alignment could not provide results as good as the approach using words can; therefore, it was decided to go back to the classical method.
+As part of the Colibri project, it was attempted to skip the first step, and to try to align the constructions directly.
+To achieve this, you can for example simply measured how often the constructions found in both languages occur together.\\\indent
+Because at this stage the decoder was not yet ready (see subsection 4), the decoder from translation system MOSES was used to evaluate the results.
+These results showed that the approach using constructions for alignment could not provide results as good as the approach using words; therefore, it was decided to go back to the classical method.
 It should be noted that, because the MOSES decoder was used, the influence of skipgrams on alignment could not be measured.
 However, because they increase complexity, they are not expected to improve the alignments.\\\indent
+Although first aligning the words and after that aligning the construction is a good way to do this task, this does not solve the problems with the skipgrams: including skipgrams in the alignment process has negative effects on the results.
+To fix this, a algorithm specialized for skipgrams was developed, which indeed was able to find good translations.
+Unfortunately, when investigating the effects in a full translation setting, it turned out the skipgrams do not seem to improve the translation quality, because the decoder often picks the translation hypotheses consisting of sentences formed without skipgrams.
 
 \subsubsection{Memory-based classification using construction experts}
+As may have become clear from the previous sections, the Colibri project uses machine learning techniques to get from the source language construction to the target language constructions.
+More specifically, Colibri uses little expert classifiers, trained for only one construction, inspired by the effectivity of this approach in word sense ambiguation \citep{vangompel10}.
+The results of this approach are compared to the results of a combined classifier approach and statistical machine translation.
+So far, the experiments have not yet shown an improvement in translation quality, but the researchers are hopeful in achieving this in the future.
+They also want to add support for skipgrams.
+
 \subsubsection{Reassembling constructions into a coherent target sentence}
-\subsection{Hypotheses} %Of moet dit bij main outstanding questions?
-
-Constructions can be found efficiently in corpus data
-Graph-based relations can be used to constrain to good constructions
-Constructions can be aligned without resorting to word alignments as a basis
-In MT, constructions (i.e. possibly with gaps) result in better translation than mere consecutive phrases
+A decoder was developed from scratch, following state-of-the-art techniques as described by \citet{koehn03}. On top of that, support of skip-grams was added.
 
@@
 
-\section{Achievements so far}
 \section{Main outstanding questions}
@@
 
+\subsection{Strenghts}
+
+\subsubsection{Easily extensible to other languages}
+
+When describing the rule-based approaches to machine translation, I pointed out that for most approaches one needs a specific ruleset for each language pair.
+The Interlingua is an attempt to solve this problem by adding a language in between, so that only a ruleset for every language-Interlingua pair is needed.
+Statistical and memory-based approaches solve this problem in an even more elegant way: the only thing needed are computational power, time, and a lot of data.
+As digital texts are available in large quantities for the more important languages of the world, adding more languages to Colibri should not be that problematic.
+And also less important languages, even languages of which the grammar has not yet been investigated, can be added easily, as long as there is enough data to train the classifiers (but also see the first two weaknesses).
+Obviously, the opposite is true for the rule-based approach.
+
+\subsubsection{Improving communication between the world's communities}
+
 \subsection{Weaknesses}
 
-\subsubsection{Machine translation and the status of translators}
+\subsubsection{Memory based learning and the quantity/quality of the data}
 
-Machine translation is far from perfect, and although this project is an interesting attempt to bring the technology closer to the 'perfect'-end, it will almost certainly not result in an application which is able to replace human translators.
-However, when non-perfect software like this is made available to the public, ignorant users might however treat it as such.
+%Lack of control
+
+\subsubsection{Existension possibilities to non-European languages?}
+
+\subsubsection{Machine translation and the societal status of translators}
+
+I'd like to close this section with a more general critique of machine translation.
+Machine translation is far from perfect.
+Colibri of course is an interesting attempt to bring the technology closer to the 'perfect'-end, it will almost certainly not result in an application which is able to replace human translators.
+However, when non-perfect software like this is made available to the public, ignorant users might treat it as if it was perfect software, thereby turning off their own intelligence, as it were.
 A similar critisism can be heard in the field of spell checkers:
@@
 
-Since the release of Google Translate, various examples can be found on the internet of websites and other texts translated with this service.
-
-\subsubsection{Translation as an art}
+And although giving a detailed analysis of the validity of these arguments is beyond the scope of this paper, I'd like to point out that something similar already seems to be happening: since the release of Google Translate, various examples of websites, street signs, restaurant menus and other texts translated by the service can be found all over the internet.
+The authors of the text, apparently not skilled in the target language, have relied on machine translation only and didn't check the quality translation in any other way, and have put this incorrect (and often funny) language in a place for the world to see.\\\indent
+An to make the problem even more complicated, there probably is no such thing as a 'correct' translation, in contrast to what the previous paragraph may have suggested.
+That is, if you ask n translators to translate a long text in language A to language B, n different translations will result.
+Translation is a problem for which no clearly defined answer exists.
+Instead, what one considers a good translation is largely a matter of taste, beliefs, cultural preference, etc., making translation much closer to an art.
+Viewing translation as something a computer can do might be a huge underestimation of the task, and giving the ignorant user this impression could be argued to be very misleading.
 
@@
 
+\bibitem[Van Gompel, 2010]{vangompel10}
+Van Gompel, M. (2010). Uvt-wsd1: A cross-lingual word sense disambiguation system.
+In Semeval '10: Proceedings of the 5th international workshop on semantic evaluation
+(pp. 238\-241). Morristown, NJ, USA: Association for Computational Linguistics.
+
 \bibitem[Van Gompel et al., 2011]{vangompel11}
@@
 \end{document}
+
+%\subsection{Hypotheses}
+
+%Constructions can be found efficiently in corpus data
+%Graph-based relations can be used to constrain to good constructions
+%Constructions can be aligned without resorting to word alignments as a basis
+%In MT, constructions (i.e. possibly with gaps) result in better translation than mere consecutive phrases

### 2012-11-18 Research paper.tex
Added methodology, completed strengths and weaknesses

@@
 
-% C++, Python
+Because of Colibri's practical nature, its methodology forms the core of the project and has been explained in detail in the previous sections, making this section largely superfluous.
+However, this might be the right place to note that Colibri's unusual methodology also makes its time planning quite different from more conventional research projects: whereas other projects would go through stages like (1) reading the previous literature, (2) building a theoretical model, (3) testing the model, for example with experiments and (4) refining the model on the basis of the results, Colibri takes an entirely different approach.
+The first year has been devoted to implementing ideas developed previously, resulting in a fully functional machine translation system.
+With this technology in place, the years after that can be used refining (and possibly extending) this technology, as well as evaluating its results.
 
 \section{Main outstanding questions}
+
 \section{Strengths and weaknesses}
@@
 
-\subsubsection{Easily extensible to other languages}
+\subsubsection{The example-based approach makes the system easily extensible to other languages}
 
 When describing the rule-based approaches to machine translation, I pointed out that for most approaches one needs a specific ruleset for each language pair.
-The Interlingua is an attempt to solve this problem by adding a language in between, so that only a ruleset for every language-Interlingua pair is needed.
-Statistical and memory-based approaches solve this problem in an even more elegant way: the only thing needed are computational power, time, and a lot of data.
-As digital texts are available in large quantities for the more important languages of the world, adding more languages to Colibri should not be that problematic.
-And also less important languages, even languages of which the grammar has not yet been investigated, can be added easily, as long as there is enough data to train the classifiers (but also see the first two weaknesses).
+The Interlingua approach is an attempt to solve this problem by adding a language in between, so that only a ruleset for every language-Interlingua pair is needed.
+Statistical and memory-based approaches solve this problem in an even more elegant way: the only things needed are computational power, time, and a lot of data.
+As digital translated texts are available in large quantities for the more important languages of the world, adding more languages to Colibri should not be that problematic.
+But even less important languages can be added easily, as long as there is enough data to train the classifiers (but see the first two weaknesses).
 Obviously, the opposite is true for the rule-based approach.
 
-\subsubsection{Improving communication between the world's communities}
+\subsubsection{Machines translation improves communication between the world's communities}
 
@@
 
-\subsubsection{Memory based learning and the quantity/quality of the data}
+\subsubsection{Memory based learning depends on quantity and quality of the data}
 
-%Lack of control
+Memory-based learning takes the rule-creating process out of the hands of the authors.
+This can be a good thing, because it means translation systems can be made without any knowledge of the source or the target language, but it also means the authors loses control over what the system produces.
+That is, if the system makes errors, solving this problem is not as simple as changing how the system works.
+\\\indent
+Instead, the task of the creator now is to provide enough bilingual texts for training - preferably as much as possible - as this improves the change of finding a matching n-gram.
+And importantly, to avoid errors, this data should be high quality.
+One incorrect translation, if not found anywhere else in the corpus, can already end up in the translation results.
+For example, when translating \emph{Will Justin Bieber ever hit puberty} from English to Vietnamese with Google Translate, which is also based on bilingual corpora, the result is \emph{Justin Bieber se bao gio den tuoi day thi}, which means as much as 'Justin Bieber will never hit puberty'.
 
-\subsubsection{Existension possibilities to non-European languages?}
+\subsubsection{It is unclear how well the memory-based approach works for non-European languages}
 
-\subsubsection{Machine translation and the societal status of translators}
+Memory-based learning generated a translation on the basis of context.
+This works well for languages in which word order is largely fixed, like English: for these languages, if you change the order (and thus the context) of a particular word, you also change the meaning.
+There are languages, however, what happens for languages with free word order; so a word can have roughly the same meaning on various places in the sentence, and thus with various contexts.
+In other words, in these languages the main source of the knowledge of memory-based learning isn't fully available.
+
+% Uitzoeken
+
+\subsubsection{Machine translation is not good for common view on translation}
 
@@
 However, when non-perfect software like this is made available to the public, ignorant users might treat it as if it was perfect software, thereby turning off their own intelligence, as it were.
-A similar critisism can be heard in the field of spell checkers:
-
-%Overzicht kritiek, zie Wikipedia
-
-And although giving a detailed analysis of the validity of these arguments is beyond the scope of this paper, I'd like to point out that something similar already seems to be happening: since the release of Google Translate, various examples of websites, street signs, restaurant menus and other texts translated by the service can be found all over the internet.
-The authors of the text, apparently not skilled in the target language, have relied on machine translation only and didn't check the quality translation in any other way, and have put this incorrect (and often funny) language in a place for the world to see.\\\indent
-An to make the problem even more complicated, there probably is no such thing as a 'correct' translation, in contrast to what the previous paragraph may have suggested.
-That is, if you ask n translators to translate a long text in language A to language B, n different translations will result.
+A similar critisism can be heard in the field of spell checkers: the invention of automated  spell checkers would have caused a general trend in relying on the technology too much, an regarding work as finished without even proofreading it.
+\citet{galletta05} for example show that people deliver texts with more false-positive and false-negative errors in it when a spell checker was present, and that this even holds for people with good verbal skills.
+\\\indent
+And although giving a detailed analysis of the validity of these arguments is beyond the scope of this paper, I'd like to point out that something similar already seems to be happening for machine translation: since the release of Google Translate, various examples of websites, street signs, restaurant menus and other texts incorrectly translated by the service can be found all over the internet.
+The authors of the text, apparently not skilled in the target language, have relied on machine translation only and didn't check the quality translation in any other way.\\\indent
+And to make the problem even more complicated, there probably is no such thing as a 'correct' translation, in contrast to what the previous paragraphs may have suggested.
+That is, if you ask \emph{n} translators to translate a long text from language A to language B, \emph{n} different translations will be produced.
 Translation is a problem for which no clearly defined answer exists.
@@
 Dempster et al., 1997
+
+\bibitem[Galletta et al., 2005]{galletta05}
+Galletta, D. F., Durcikova, A., Everard, A. \& B. M. Jones. Does spelling checker software need warning label? Communications of the ACM, 48 (7), 82 - 86.
 
@@
 % Documentatie Colibri
-% Maartens thesis
 

### 2012-11-19 Research paper.tex
RP: corrected some sentences

@@
 That is, authors of translation algorithms tried to include their knowledge about the languages and translation.
-The rule-based approache can be subdivided into four subapproaches, summarized in the Vauquois triangle: \\
+The rule-based approach can be subdivided into four subapproaches, summarized in the Vauquois triangle: \\
 
@@
 This means a translation from English to German would entail a translation from English to the Interlingua (using the ruleset English $\rightarrow$ Interlingua) and then a translation from the Interlingua to German (using the ruleset Interlingua $\rightarrow$ German).
-When translating from English to French, the same two steps would be taken, and the same ruleset English $\rightarrow$ Interlingua can be used.
+When translating from English to French, the same two steps would be taken, and the same ruleset English $\rightarrow$ Interlingua can be re-used.
 \end{itemize}
@@
 Unlike the ruled-based approach, the statistical approach creates its own 'knowledge', using bilingual text corpora.
-With this knowledge, the algorithm is able to generate various translation hypotheses, from which the best one is chosen with various statistical measures.
+With this knowledge, the algorithm is able to generate various translation hypotheses, from which the best one is chosen with the help of various statistical measures.
 I will explain these stages in more detail:
@@
 \begin{itemize}
-\item \emph{Corpus alignment.} Only having a bilingual corpus is not enough: for such corpus to be useful one also needs to know which parts of the sentence in language A correspond to which part of the sentence B.
+\item \emph{Corpus alignment.} Only having a bilingual corpus is not enough: for such a corpus to be useful one also needs to know which parts of the sentence in language A correspond to which part of the sentence B.
 This is not only complicated by different word orders, but also by concepts that in language A take more or less words than in language B.
-A well-known word alignment algorithm is the Expectation Maximisation algorithm \citep{dempster77}.
+A well-known word alignment algorithm that tries to solve this problem is the Expectation Maximisation algorithm \citep{dempster77}.
 
@@
 
-\item \emph{Hypothesis generation.} We now have some sort of automatically generated dictionary: for each word in language A, we know which words in language B are often used for it.
+\item \emph{Hypothesis generation.} We now have some sort of automatically generated dictionary: for each word in language A, we know which word in language B is often used for it.
 On the basis of this knowledge, we can generate various possible translations.
@@
 \begin{itemize}
-\item \emph{The translation model.} The translation model returns a probability of source sentence given the target sentence; it thus looks back whether what has been created fits the original.
+\item \emph{The translation model.} The translation model returns a probability of the source sentence given the target sentence; it thus looks back whether what has been created fits the original.
 This probability is calculated on the basis of statistical word alignment algorithms like the IBM models \citep{brown93}.
@@
 
-In this section, I'll give an overview of the example-based approach.
-I'll focus on the memory-based version of the example-based approach in particular, because that is the version the Colibri project uses.
+For my overview of the example-based approach, I'll focus on the memory-based version of the example-based approach in particular, because that is the version the Colibri project uses.
 Other versions of example-based machine translation might vary slightly.
@@
 The memory-based approach is in many ways similar to the statistical approach - it too generates translations on the basis of large bilingual corpora, for instance.
-The main difference, however, is that it is context-sensitive; the idea is that the system will find another translation for \emph{left} in the sentence \emph{The toilets are at your left hand} than for \emph{left} in the sentence \emph{Elvis has left the building}.
+The main difference, however, is that it is context-sensitive; the idea is that the system will find another translation for \emph{left} in the sentence \emph{The toilets are at your left hand} than for \emph{left} in the sentence \emph{Elvis has left the building} simply because they occur in different contexts.
 \\\indent
@@
 The part of the translation software responsible for this is called the \emph{decoder}.
-There might be a lot of overlap in the resulting trigram (for example: '\emph{\_ I love}', '\emph{I love you}' and '\emph{love you \_}'), making the decoder's task relatively easy.
+There might be a lot of overlap in the resulting trigrams (for example, when the trigrams would be '\emph{\_ I love}', '\emph{I love you}' and '\emph{love you \_}'), making the decoder's task relatively easy.
 Such overlap might even give hints for the word order of the target language.

### 2012-11-20 Research paper.tex
RP: more revisions, more structure, RTP: more work on intro and method

@@
 For example, we can give a classifier the external characteristics of 10.000 humans (length, type of cloths, length of hair, etc.), and also tell it whether these humans are male or female.
-If we then ask for the sex of a relatively short human, wearing a dress, with long hair, the classifier will probably be able to tell us this is a woman.
+If we then ask for the sex of a relatively short human, wearing a dress, with long hair, the classifier will probably be able to tell us this is a woman, because most of the examples with similar features also were women.
+It thus is frequency that plays a central role here.\\\indent
 While this example only has two classes, it is in principle possible to use an unlimited number of classes.
@@
 Colibri, 'Constructions as Linguistics Bridges', investigates constructions (or patterns), as its name suggests.
-A construction can be any group of consecutive words in any natural language that in some way forms an entity.
+A construction can be any group of consecutive words  that in some way forms an entity, in any natural language.
 An example is 'on the basis of' in the following sentence:
@@
 
-Note that a simple trigram approach couldn't have discoverd this construction.
-Importantly, constructions can also have one or more gaps (so-called 'skip-grams').
+Note that a simple trigram approach could not have covered this construction.
+Importantly, constructions can also have one or more gaps (so-called 'skipgrams').
 'From \_\_ point of view' in the following example, is a construction with one such gap.
@@
 The gap is filled here is filled with 'his', but the construction 'from \_\_ point of view' can also be filled with many other words ('my', 'the', 'another', 'Obama's', etc.).
-This shows the advantage of skip-grams over n-grams for machine translation: whereas n-grams would only have been able to capture the construction 'point of view', because the word directly before 'point of view' is variable, skip-grams are also able to recognize that the word 'for' is part of the construction.\\\indent
+This shows the advantage of skipgrams over n-grams for machine translation: whereas n-grams would only have been able to capture the construction 'point of view', because the word directly before 'point of view' is variable, skipgrams are also able to recognize that the word 'for' is part of the construction.
+Frequent examples of the skipgrams (I expect the variant with 'my' to be very frequent, for instance) of course \emph{would} be recognized by the n-gram approach.\\\indent
 Of course, not every possible group of consecutive words is a construction; what makes some groups special?
@@
 For this project, however, it suffices to say that constructions emerge because some combinations of words are more frequent than others.
+As we've seen in the previous section, taking into account frequency in the training material is how classifiers are able to do their job.
 
@@
 % Drie vormen freq: raw frequency, context frequency, multilingual alignmetn
-% Blabla, want freq is waar mach learning gebriuk van maakt
 
@@
 Besides an attempt to discover more about the possibilities for machine translation technology, Colibri is also an attempt to actually develop this technology.
-A project overview therefore should not be complete without an overview of the software's functionality.
+A project overview therefore cannot be complete without an overview of the software's functionality.
 Colibri's functionality can be split in four parts: (1) identifying and extracting constructions in a language, (2) identifying aligned constructions between language-pairs, (3) memory-based classification using construction experts and (4) reassembling constructions into a coherent target sentence.
-If we look at these parts int he light of the light of the memory-based machine translation process as described in the previous section, the first three parts would constitute the local phase, while part 4 would constitute the global phase.
+If we look at these parts in the light of the light of the memory-based machine translation process as described in the previous section, the first three parts would constitute the local phase, while part 4 would constitute the global phase.
 I will discuss the four parts in more detail consecutively:
@@
 \subsubsection{Identifying aligned constructions between language-pairs}
-The classical approach for aligning constructions from a bilingual corpus is to first align the individual words, and then align the construction themselves \citep{koehn03}.
-As part of the Colibri project, it was attempted to skip the first step, and to try to align the constructions directly.
-To achieve this, you can for example simply measured how often the constructions found in both languages occur together.\\\indent
+The classical approach for aligning constructions from a bilingual corpus is to first align the individual words, and then align the constructions themselves \citep{koehn03}.
+It was attempted to skip the first step, and to try to align the constructions directly.
+To achieve this, you can for example simply measured how often the constructions found in both languages occur together (that is, in both the source and the target language).\\\indent
 Because at this stage the decoder was not yet ready (see subsection 4), the decoder from translation system MOSES was used to evaluate the results.
@@
 However, because they increase complexity, they are not expected to improve the alignments.\\\indent
-Although first aligning the words and after that aligning the construction is a good way to do this task, this does not solve the problems with the skipgrams: including skipgrams in the alignment process has negative effects on the results.
-To fix this, a algorithm specialized for skipgrams was developed, which indeed was able to find good translations.
-Unfortunately, when investigating the effects in a full translation setting, it turned out the skipgrams do not seem to improve the translation quality, because the decoder often picks the translation hypotheses consisting of sentences formed without skipgrams.
+When support for skipgrams was added to the implementation of the classical approach, it turned out that including skipgrams in the alignment process had negative effects on the results.
+To fix this, a algorithm specialized for aligning skipgrams was developed, which indeed was able to find good translations.
+Unfortunately, when investigating its effects in a full translation setting, it turned out the skipgrams do not seem to improve the translation quality, because the decoder mostly picks the translation hypotheses consisting of sentences formed without skipgrams.
 
@@
 So far, the experiments have not yet shown an improvement in translation quality, but the researchers are hopeful in achieving this in the future.
-They also want to add support for skipgrams.
+They also want to add support for skipgrams in this stage.
 
 \subsubsection{Reassembling constructions into a coherent target sentence}
-A decoder was developed from scratch, following state-of-the-art techniques as described by \citet{koehn03}. On top of that, support of skip-grams was added.
+A decoder was developed from scratch, following state-of-the-art techniques as described by \citet{koehn03}. On top of that, support of skipgrams was added.
 
-\section{Scientific importance and relevance for society}
+\section{Relevance}
+
+\subsection{For science}
 
@@
 
+\subseciton{For society}
+
 \section{Methodology}
@@
 However, this might be the right place to note that Colibri's unusual methodology also makes its time planning quite different from more conventional research projects: whereas other projects would go through stages like (1) reading the previous literature, (2) building a theoretical model, (3) testing the model, for example with experiments and (4) refining the model on the basis of the results, Colibri takes an entirely different approach.
-The first year has been devoted to implementing ideas developed previously, resulting in a fully functional machine translation system.
+The first year has almost completely been devoted to implementing ideas developed previously, resulting in a fully functional machine translation system.
 With this technology in place, the years after that can be used refining (and possibly extending) this technology, as well as evaluating its results.
@@
 
+In this section, I will try to give an overview of the project's strong and weak points.
+Some points concern this particular research project, other concern the memory based approach, and others zoom out even further and concern machine translation in general.
+
 \subsection{Strenghts}
+
+\subsubsection{The theory is tested immediately}
 
@@
 
+%Bruggetje naar relevance for society
+
 \subsection{Weaknesses}
@@
 Instead, the task of the creator now is to provide enough bilingual texts for training - preferably as much as possible - as this improves the change of finding a matching n-gram.
-And importantly, to avoid errors, this data should be high quality.
+And importantly, to avoid errors, this data should be of high quality.
 One incorrect translation, if not found anywhere else in the corpus, can already end up in the translation results.
@@
 Memory-based learning generated a translation on the basis of context.
-This works well for languages in which word order is largely fixed, like English: for these languages, if you change the order (and thus the context) of a particular word, you also change the meaning.
-There are languages, however, what happens for languages with free word order; so a word can have roughly the same meaning on various places in the sentence, and thus with various contexts.
+This works well for languages in which word order is largely fixed, like English: for these languages, if you change the position of a particular word in the sentence (and thus the context), you also change the meaning of that sentence.
+There are languages, however, with relatively free word order; so a word can have roughly the same meaning on various places in the sentence, and thus with various contexts.
 In other words, in these languages the main source of the knowledge of memory-based learning isn't fully available.
@@
 
-\subsubsection{Machine translation is not good for common view on translation}
+\subsubsection{Machine translation is not good for the common view on translation}
 
-I'd like to close this section with a more general critique of machine translation.
 Machine translation is far from perfect.
-Colibri of course is an interesting attempt to bring the technology closer to the 'perfect'-end, it will almost certainly not result in an application which is able to replace human translators.
+Colibri of course is an interesting attempt to bring the technology closer to the 'perfect' end of the scale, but it will almost certainly not result in an application which is able to replace human translators.
 However, when non-perfect software like this is made available to the public, ignorant users might treat it as if it was perfect software, thereby turning off their own intelligence, as it were.
-A similar critisism can be heard in the field of spell checkers: the invention of automated  spell checkers would have caused a general trend in relying on the technology too much, an regarding work as finished without even proofreading it.
+A similar critisism can be heard in the field of spell checkers: the invention of automated  spell checkers would have caused a general trend in relying on the technology too much, and regarding work as finished without even proofreading it.
 \citet{galletta05} for example show that people deliver texts with more false-positive and false-negative errors in it when a spell checker was present, and that this even holds for people with good verbal skills.
 \\\indent
-And although giving a detailed analysis of the validity of these arguments is beyond the scope of this paper, I'd like to point out that something similar already seems to be happening for machine translation: since the release of Google Translate, various examples of websites, street signs, restaurant menus and other texts incorrectly translated by the service can be found all over the internet.
+Although giving a detailed analysis of these viewpoint is beyond the scope of this paper, I'd like to point out that something similar already seems to be happening for machine translation: since the release of Google Translate, various examples of websites, street signs, restaurant menus and other texts incorrectly translated by the service can be found all over the internet.
 The authors of the text, apparently not skilled in the target language, have relied on machine translation only and didn't check the quality translation in any other way.\\\indent

### 2012-11-20 Research training portfolio.tex
RP: more revisions, more structure, RTP: more work on intro and method

@@
 a training set that consists of roughly the same amount of examples for each confusible option.
-If we would use a training set with much more contexts for option A than for option B, which may bias the model to option A.
-For example, if the word 'cookie' can occur perfectly both in the context of option A and option B, when seeing 'cookie' in a particular context the confusible corrector should not be able to make a prediction on the basis of this word alone.
-However, if the language model is based on an unbalanced set it will predict option A, simply because A occured together with 'cookie' more often.
+If we would use a training set with much more contexts for option A than for option B, we may bias the model to option A.
+For example, if the word 'cookie' has an equal chance to occur both in the context of option A and option B, the confusible corrector should not be able to make a prediction on the basis of this word alone.
+However, if the language model is based on an unbalanced set it will predict option A, simply because A occurred together with 'cookie' more often.
 \\\indent
+A solution to this could of course be to always offer balanced training sets.
+In practice, this means removing part of the data: if 100 examples of option A, and 500 examples of option B are available, 400 examples of B will have to be removed.
+The more skewed the distribution is, the more data have to be removed, which can lead to situations where less than 1 percent of the training material can be used.
+As the results of the confusible correctors improve if they have seen more data, it is very likely that removing so much data deteriorates the results, rather than improving them.
+In the remainder of this paper I will therefore investigate in which cases it is better to use a balanced training set, and in which cases it is better to train the modules on the larger, unbalanced set.
+As will be shown, the exact behavior of the confusible corrector differs per module, so it is hard to draw an exact boundary, but ...
+
+% Hier even opletten of wat ikz eg wel klopt met mijn uiteindelijke resultaten
+% UIteindelijke resultaten toevoegen
 
@@
 
-% Hier moet ergens een beetje beschreven worden hoe zo'n Timblmodule werkt.
+\subsubsection{The confusible correctors}
+To investigate the differences between confusible correctors trained on balanced and unbalanced training sets, I will use the confusibles listed in the previous sections.
+For each of these confusibles, a confusible corrector is trained to predict which option is likely to occur in which context.
+To be able to recognize contexts, the correctors use \emph{Memory Based Learning}, which means that it saves all contexts previously seen in memory.
+When it is given a new context, it generates a list of previously seen contexts which are relatively similar.
+For each of the contexts of this list, the confusible ...
+This approach is called the k-nearest neighbour algorithm.
 
-\subsubsection{Evaluation}
+The classifier algorithm used using IGTree, a decision-tree approximation of IB1-IG.
+
+% Hier ref toevoegen
+
+\subsubsection{Evaluation of the confusible correctors}
 

### 2012-11-21 Research paper.tex
Added another strength

@@
 
-\subsubsection{The theory is tested immediately}
+\subsubsection{The theory is tested immediately and completely}
+
+An idea is very different from something that is really useful or true.
+Many fields of science get stuck somewhere in this transition.
+In some fields scholars only write down their ideas in human language, not fully formulating how the idea should work, and thus never discovering there might be flaws in the idea somewhere.
+In other fields scholars can make the idea formal, but have no way to test the ideas in any other way than trusting on intuitions.
+In again other fields one can test it, for example by means of experiments, but it remains unclear how the results should be interpreted and to what extent they generalize to the whole population.
+All this is not a problem for the Colibri project.
+Claims are made (for example 'skipgrams improve machine translation'), formalized ('this is how they should work within in a machine translation system'), but also implemented and fully tested (i.e.
+in a fully functional machine learning system).
+In other words, except maybe for the evaluation measures, there is little reason to doubt the results of this project.
 

### 2012-11-23 Research paper.tex
RP: added references

@@
 A well-known word alignment algorithm that tries to solve this problem is the Expectation Maximisation algorithm \citep{dempster77}.
+It should be added that after aligning words, phrased could be formed and aligned.
+\citet{koehn03} show this provides better results than word-based translation.
 
@@
 \\\indent
-Like I did with statistical machine translation, I will now give an overview of the memory-based machine translation process, based on \citet{vdbb09}.
+Like I did with statistical machine translation, I will now give an overview of the memory-based machine translation process, adapted from \citet{vdbb09}.
 For this example, I will describe a translator that uses trigrams, but all kinds of n-grams could be used here, even n-grams of variable length.
@@
 \begin{itemize}
-\item \emph{Local phase.} In the description of \citet{vdbb09}, the local phase entails both training the classifier and actual classification, but I think it is important to make a sharp distinction here: whereas the training is done only once and \emph{before} actual translation, and could therefore be seen as 'preparing' the translation software for its task, the classification is done every time the user asks for a translation.
+\item \emph{Training.} Training is done only once and \emph{before} actual translation, and could therefore be seen as 'preparing' the translation software for its task, the classification is done every time the user asks for a translation.
 Training consists of (1) aligning the bilingual corpus, using the word alignment algorithms described in the previous sections, (2) decomposing the complete corpus into trigrams, and (3) giving the results to the classifier as training material.
 The trigrams of the source language constitute the examples, the trigrams of the target language are the classes.
-Classification means (1) decomposing the source text into trigrams and (2) classifying each trigram.
-We end up with a list of trigrams in the target language.
 
-%vdb b 09 ook echt lezen
+\item \emph{Local classification.} Classification means (1) decomposing the source text into trigrams and (2) classifying each trigram, based on the language model created in the training phase.
+The output are the classes, i.e.
+the trigrams, the classifiers associates with each source trigram.
 
-\item \emph{Global phase.} In the global phase we try to merge the trigrams into one text.
+\item \emph{Global search.} In the global phase we try to merge the trigrams into one text.
 The part of the translation software responsible for this is called the \emph{decoder}.
@@
 
-\subseciton{For society}
+\subsection{For society}
 

### 2012-11-29 Research training portfolio.tex
Added info about the website, the slider, and the confidence formulas needed for it

@@
 This number will probably be different when you read this, though, as new context-based modules are created easily, and can be removed if it turns out they're not accurate enough.
+Although it is possible to run Fowlt as a stand-alone application, most people will use it as a web application.
+I will devote a subsection to this 'web shell' around Fowlt as well.
+
+%Plaatje toevoegen
 
 \subsubsection{The processchain}
-In Fowlt's main script, fowlt\_processchain.py, all modules are started up, and their outputs are collected.
-Making changes in this script (based on the one for Valkuil, written by Maarten van Gompel) to make sure the new English based modules worked correctly, is one of the main things I did during my internship.
+In Fowlt's main script, fowlt\_processchain.py, all modules are started up, their outputs are collected and are all saved in xml form.
+Making changes in this script (based on the ones for Valkuil, written by Maarten van Gompel) to make sure the new English based modules worked correctly, is one of the main things I did during my internship.
 
@@
 
+\subsubsection{Fowlt as a web application}
+%Picture reference
+Fowlt can also (and, probably, will mostly) be used as a website.
+In this web variant, Fowlt works exactly as described above, but some extra mechanisms are built around to make sure it works and looks nice in the browser.
+The transition from a standalone application to a website is a two-step process, as is exemplified by figure...
+.
+Firstly, Fowlt is made available online webservice using the software CLAM.
+Among other things, CLAM adds the possibility to use Fowlt as a RESTful service.
+However, ordinary users don't want to send HTTP requests or read xml files when they want to texts corrected.
+This is solved in the second step, which is the Fowlt.net interface.
+In this interface, users can paste their texts, and see the mistakes flagged in red in these texts themselves.
+The interface even adds functionality to correct any errors.
+Behind the screens, the interface communicates of course with the CLAM webservice.
+\\\indent
+%Section reference
+Most of the functionality described above was already working for Valkuil, and so only had to be adapted to English to be function within Fowlt.
+One feature that I created for Fowlt that Valkuil doesn't have (yet), is an error tolerance slider.
+With this slider, users can change on the fly how confident Fowlt has to be about an error before it displays it.
+The implementation of this slider was, however, not without problems.
+See section for more information on that.
+
 \subsection{Problems encountered}
@@
 
-\subsubsection{The ConfusibleModules weren't good enough}
-I didn't only train the ConfusibleModules, I also tested them with a program written by my supervisor.
+%Deze paragraaf is twijfelachtig
+%\subsubsection{The ConfusibleModules weren't good enough}
+%I didn't only train the ConfusibleModules, I also tested them with a program written by my supervisor.
 This program takes 10\% from the training data and saves it in a seperate file, trains TiMBL on the remaining 90\%, and then returns the percentage of correct predictions the resulting model can make about the 10\%.
@@
 
-\subsection{Results}
+\subsubsection{Most modules didn't provide confidences measures}
+Unlike the confusible modules, the WOPR module and the other modules do not return how confident they are a particular error is indeed an error.
+This was a problem for the error tolerance slider, because for it to work it needs have confidence values for all errors, not just the confusibles.
+This not only entailed going back to the C-code of the original modules, but also to come up with formulas to represent exactly how confident the modules are.
+I would like to use this final subsection to explain in detail the formulas I have chosen:
+
+\paragraph{Lexicon Checker}
+
+\[
+1 - \frac{wf}{cf/ft}
+\]
+
+The lexicon checker gives a correction \emph{c} for a word \emph{w} if \emph{c} is very similar to \emph{w}, but \emph{frequency threshold} times more frequent.
+As I write this, the frequency threshold (ft) used for the lexicon checker is 10.000.
+This confidence formula reflects how many times the correction frequency (cf) is 10.000 times more frequent than the frequency of the original word (wf).
+Thus, if the correction is only 10.000 times more frequent, the result will be 0, if it is 20.000 more frequent, it will be 0.5, etc.
+The larger the difference between the original word and the correction, the closer to 1 the result will be.
+
+\begin{table}[h]
+\begin{tabular}{|l|lll|}
+\hline
+&word frequency (wf)&correction frequency (cf)&frequency threshold (ft)\\ 
+\hline
+anticcipate&1*&2.710.230&10.000\\
+newwspaper&876&15.200.566&10.000\\
+\hline
+\end{tabular}
+\end{table}
+
+anticcipate $\rightarrow$ anticipate
+
+\[
+1 - \frac{1}{2.710.230/10.000} = 0.996
+\]
+
+newwspaper $\rightarrow$ newspaper
+
+\[
+1 - \frac{876}{15.200.566/10.000} = 0.43
+\]
+
+* If the word frequency is lower than 1, it is assumed to be 1.
+
+\paragraph{Runon Checker}
+
+\[
+1 - \frac{1}{(\frac{(lf+rf)/2}{t})}
+\]
+
+The left fraction (lf) and the right fraction (rf) of a word represent how often the left and the right part of that word are part of a bigger word.
+If for these two parts the number is above a threshold (3000 as I write this), the Runon Checker suggests the two parts are probably words on their own.
+This confidence formula reflects how high these fractions are.
+It first calculates the mean fraction and relates this to the threshold (t).
+The result is then transformed into a number between 1 and 0.
+
+\begin{table}[h]
+\begin{tabular}{|l|lll|}
+\hline
+&left fraction (lf)&right fraction (rf)&threshold (t)\\ 
+\hline
+fightscrime&37.274,75&94.063,234&3000\\
+unexpectedhorror&81.098,078&55.204,098&3000\\
+\hline
+\end{tabular}
+\end{table}
+
+fightscrime $\rightarrow$ fights crime
+
+\[
+1 - \frac{1}{(\frac{(37.274,75+94.063,234)/2}{3000})} = 0.954
+\]
+
+unexpectedhorror $\rightarrow$ unexpected horror
+
+\[
+1 - \frac{1}{(\frac{(81.098,078+55.204,098)/2}{3000})} = 0.956
+\]
+
+\paragraph{Split Checker}
+
+\[
+1 - \frac{1}{(\frac{cf}{((fpw+fcw)/fr)/2})}
+\]
+
+The Split Checker combines every word with the word before it, and checks if the frequency of the result is higher than the frequencies of both original words.
+Before this comparison is done, the frequencies of the original words are divided by the frequency ratio (2 as I write this), so the module will respond more often.
+This confidence formula reflects the difference between the mean frequency of both original words (fwp and fcw) and the frequency of the new word (cf).
+It first calculates this mean, and also takes the frequency ratio (fr) into account.
+This mean is then compared to the frequency of the mean of the combination, and the result is then transformed into a number between 1 and 0.
+
+\begin{table}[h]\footnotesize
+\begin{tabular}{|l|lll|}
+\hline
+&combination frequency (cf)&previous word frequency (pwf)&current word frequency (cwf)\\ 
+\hline
+unexp ected&1054275&2124&7223\\
+flabberg asted&97389&0&1913\\
+\hline
+\end{tabular}
+\end{table}
+
+unexp ected $\rightarrow$ unexpected
+
+\[
+1 - \frac{1}{(\frac{1054275}{((2124+7223)/2)/2})} = 0.982
+\]
+
+flabberg asted $\rightarrow$ flabbergasted
+
+\[
+1 - \frac{1}{(\frac{97389}{((0+1913)/2)/2})} = 0.995
+\]
 

### 2012-11-29 Research training portfolio.tex
RTP: added section what I learned about setting up a webservice

@@
 
-\subsection{Working with CLAM, setting up an NLP webservice}
+\subsection{Working with Django and CLAM, setting up an NLP webservice}
+Although I do have experience with web development, everything I had to do to get Fowlt online was completely new to me.
+Whereas I was used to web programming language PHP, Fowlt works with Django, a Python library for web development.
+Fortunately, the Django documentation turned out to be very good and clear, so I got the hang of it quickly.
+In fact, Django turned out to be so elegant and easy that I might use it in future projects as well.
+The fact that most code needed could be copied from Valkuil made this part extra easy.
+CLAM, which turns an NLP application into a webservice, also turned out to be easy to use, and has a good documentation.
+I had Fowlt working with CLAM within an hour.
+\\\indent
+However, the real problems with setting up an NLP webservice were not related to Django or CLAM, but to all kinds of unexpected behavior of the university's servers.
+For example, the user groups on Applejack were not set up correctly when I started, prohibiting me to reach some locations, both Spitfire and Applejack reset file permissions when files are updated, and the server software caches server settings, even after update the configuration file.
+In retrospect, the problems were minor and could be solved easily (either by me or by an administrator), but it took me quite a lot of time figure out what the problem was and how to solve it.
 

### 2012-12-04 Research training portfolio.tex
RTP: worked on intro, added refs

@@
+%newgjhg
 \documentclass[12pt]{article}
@@
 This poses a problem when one tries to create a context-sensitive memory-based confusible corrector, as having more data for one option will make it more likely that the confusible corrector will predict that option.
-In some cases this can be solved by balancing the training data so that it has exactly the same amount of examples for each option, but in other cases this means a very large part of the data have to be thrown away, making the corrector even worse.
-In this paper I investigate in which cases it is better to use a balanced training set, and in which cases it is better to train the modules on the larger, unbalanced set.
-I do this by building x confusible correctors and evaluating their results.
+In the literature it has been suggested that balancing the training data, so that it has exactly the same amount of examples for each option, might improve results of classifiers in general.
+In this paper I investigate whether this is true for classifiers in a spelling correction environment.
 
 %hoeveel?
+% UIteindelijke resultaten toevoegen aan het abstract
 
@@
 A more detailed account of its workings will be given in the next section.
-For now it is important to understand that TiMBL extracts this knowledge from large text corpora.
+For now, it is important to understand that TiMBL extracts this knowledge from large text corpora.
 From these corpora, all examples of the confusible have been extracted together with their context.
 The confusible corrector's results improve when TiMBL has seen more training material to base its language models on.\\\indent
-Fowlt contains language models, and thus specialized confusible correctors, for the following confusibles (among others):
+Fowlt contains language models, and thus specialized confusible correctors for numerous well-known mistakes in English. In this paper, I will investigate the following:
+
+\begin{itemize}
+\item \emph{advice} versus \emph{advise}
+\item \emph{lose} versus \emph{loose}
+\item \emph{practice} versus \emph{practise}
+\item \emph{than} versus \emph{then}
+\item \emph{weather} versus \emph{whether}
+\item \emph{you're} versus \emph{your}
+\end{itemize}
 
@@
 \\\indent
-A solution to this could of course be to always offer balanced training sets.
+A solution to this could of course be to always offer balanced training sets \citep{provost, gh09}.
+MEER TOEVOEGEN UIT DE LITERATUUR.
 In practice, this means removing part of the data: if 100 examples of option A, and 500 examples of option B are available, 400 examples of B will have to be removed.
@@
 As the results of the confusible correctors improve if they have seen more data, it is very likely that removing so much data deteriorates the results, rather than improving them.
-In the remainder of this paper I will therefore investigate in which cases it is better to use a balanced training set, and in which cases it is better to train the modules on the larger, unbalanced set.
-As will be shown, the exact behavior of the confusible corrector differs per module, so it is hard to draw an exact boundary, but ...
+In the remainder of this paper I will therefore investigate whether it is better to use a balanced training set or to to train the modules on the larger, unbalanced set.
 
+% Meer literatuur toevoegen
 % Hier even opletten of wat ikz eg wel klopt met mijn uiteindelijke resultaten
-% UIteindelijke resultaten toevoegen
 
@@
 Fawcett, 2004.
+\bibitem[García \& Herrera, 2009]{gh09}
+García \& Herrera, 2009. Evolutionary Undersampling for Classification with Imbalanced Datasets: Proposals and Taxonomy.
 \bibitem[Pollock \& Zamora, 1984]{pz84}
 Pollock \& Zamora, 1984
+\bibitem[Provost, 2000]{provost}
+Provost, F (2000). Machine Learning from Imbalanced Data Sets 101. Extended abstract. 
+%jaar is nep
 \bibitem[Reynaert, 2008]{reynaert08}

### 2012-12-04 Research paper.tex
Some little changes on the basis of Jurafsky & Martin, added Goldberg

@@
 In this section I will try to give a short overview of the various approaches in machine translation, with special attention to the example-based approach, as that is the approach used in this project.
-This section is mainly based on \citet{vangompel09}
+This section is mainly based on \citet{jm09} and \citet{vangompel09}.
 
@@
 All of the approaches so far have in common that they require a seperate rule set for each language pair; no or very few rules included for translations from English to German can also be used for translations from English to French.
+This means that for every time a new language is added to the system, the amount of rules that have to be added grows exponentially.
 The Interlingua approach is an attempt to solve this problem, by pairing up every language with a formal representation of meaning, the Interlingua.
@@
 
+I should add that translation system focussing on one such approach are rare.
+As \citet{jm09} put it: 'Real systems tend to involve combinations of elements from these three architectures; thus, each is best thought of as a point in an algorithmic design space rather than as an actual algorithm.' The Systran system (\citealp{hs92,senellartea01}) is an example of this.\\\indent
 An advantage of the rule-based approach is that one has a lot of control over the output.
@@
 The next two approaches therefore replace hand-crafted rules with automatically generated ones.
-
-% Voorbeelden? References?
 
@@
 This is not only complicated by different word orders, but also by concepts that in language A take more or less words than in language B.
-A well-known word alignment algorithm that tries to solve this problem is the Expectation Maximisation algorithm \citep{dempster77}.
 It should be added that after aligning words, phrased could be formed and aligned.
 \citet{koehn03} show this provides better results than word-based translation.
-
-%GIZA++?
 
@@
 Of course, not every possible group of consecutive words is a construction; what makes some groups special?
-The exact nature of linguistic constructions has been the subject of many lignuistic publications and even an entire linguistic framework (construction grammar).
+The exact nature of linguistic constructions has been the subject of many linguistic publications and even an entire group of linguistic frameworks (construction grammar, see \citet{goldberg95} for a well-known example).
 For this project, however, it suffices to say that constructions emerge because some combinations of words are more frequent than others.
@@
 
-% Referentie construction grammar
 % Drie vormen freq: raw frequency, context frequency, multilingual alignmetn
@@
 
+\bibitem[Goldberg, 1995]{goldberg95}
+Goldberg, A (1995). Constructions: A Construction Grammar Approach to Argument Structure. Chicago: University of Chicago Press.
+
 \bibitem[Van Gompel, 2009]{vangompel09}
@@
 
+\bibitem[Hutchins \& Somers, 1992]{hs92}
+Hutchins \& Somers, 1992.
+
+\bibitem[Jurafsky \& Martin, 2009]{jm09}
+Jurafsky, D.
+\& J.
+H.
+Martin (2000).
+Speech and language processing: an introduction to natural language processing, computational linguistics, and speech recognition.
+Second edition.
+New Jersey: Pearson Education.
+
 \bibitem[Koehn et al, 2003]{koehn03}
 Koehn et al, 2003
+
+\bibitem[Senellart et al., 2001]{senellartea01}
+Senellart et al., 2001
 

### 2012-12-06 Research training portfolio.tex
RTP: rewrote part of the intro, worked on method

@@
-%newgjhg
 \documentclass[12pt]{article}
@@
 These errors, in which one word is confused with another (often homophone) word, are called \emph{confusible errors}.
-For most common confusible errors, Fowlt contains specialized modules.
-Unlike typos, confusible errors cannot be detected by simply looking at the text word for word, because the error is in the relation of a particular word with the other words in the sentence.
-Therefore, the confusible modules detect them by looking at the context of the word, and see if that differs from the contexts they have seen in the past for that word.
+For most common confusible errors, Fowlt contains specialized modules, which I will call \emph{confusible correctors} in the remainder of this paper.
+These confusible correctors cannot detect the errors they're supposed to detect by simply looking at the text word for word, because the error is in the relation of a particular word with the other words in the sentence.
+Therefore, they detect them by looking at the context of the word, and see if that differs from the contexts they have seen in the past for that word.
 If it, on the basis of the context, would predict another word in that position, and is really certain about it, it flags the word as an error.
 This means, for example, that Fowlt is able to replace the incorrect "there" in the last example into "they're", simply because "there" usually isn't followed by "really nice people", while "they're" is.
-\\\indent
-This knowledge was not inserted manually, but was created with the help of Machine Learning software TiMBL.
-A more detailed account of its workings will be given in the next section.
-For now, it is important to understand that TiMBL extracts this knowledge from large text corpora.
-From these corpora, all examples of the confusible have been extracted together with their context.
-The confusible corrector's results improve when TiMBL has seen more training material to base its language models on.\\\indent
-Fowlt contains language models, and thus specialized confusible correctors for numerous well-known mistakes in English. In this paper, I will investigate the following:
+Fowlt includes many confusible correctors like this, but for this paper I will limit myself to six examples.
+These are the confusibles they are specialized in:
 
@@
 
-What the confusible modules within Fowlt for each of these confusibles do is removing the confusible words and predicting which word was removed on the basis of the context.
-If they predicted the word incorrectly, BUT are very certain about it, they assume the original word was wrong and flag it as an error.
+All confusible correctors in Fowlt consist of two parts: \emph{TiMBL}, which is the same for every corrector, and a \emph{language model}, which is based on the specific confusible error the corrector was created for.
+TiMBL is machine learning software that looks at the contexts of possible, and decides whether it would have expected that word on the basis of that context.
+The language model is the 'knowledge' TiMBL uses to make this decision.
+This knowledge was not inserted manually, but was created automatically by looking at large text corpora (so-called \emph{training material}).
+Importantly, the confusible corrector's results improve when the language is based on more training material.
+So the more 'training material', the larger and more complete the language model, which in turn increases the likelihood that TiMBL will make the correct decision.
 \\\indent
@@
 \\\indent
-A solution to this could of course be to always offer balanced training sets \citep{provost, gh09}.
-MEER TOEVOEGEN UIT DE LITERATUUR.
-In practice, this means removing part of the data: if 100 examples of option A, and 500 examples of option B are available, 400 examples of B will have to be removed.
-The more skewed the distribution is, the more data have to be removed, which can lead to situations where less than 1 percent of the training material can be used.
-As the results of the confusible correctors improve if they have seen more data, it is very likely that removing so much data deteriorates the results, rather than improving them.
-In the remainder of this paper I will therefore investigate whether it is better to use a balanced training set or to to train the modules on the larger, unbalanced set.
+A solution to this could of course be to always offer balanced training sets.
+A balanced training set can be achieved in two ways: oversampling and undersampling \citep{provost, gh09}.
+Oversampling means that extra data are somehow created for the less frequent option.
+Because of the unnatural situations this creates, oversampling probably is not desirable in the spell checking domain (and possibly for NLP in general).
+Undersampling means removing examples for the more frequent option: if 100 examples of option A, and 500 examples of option B are available, 400 examples of B will have to be removed.
+The more skewed the distribution is, the more data have to be removed, which can lead to situations where only a very small part of the originally collected training material can be used.
+For example, the distribution of the \emph{practice} versus \emph{practise} confusible is so unbalanced that 85.57 percent of the data had to be removed to create a balanced trainig set.
+\\\indent
+This leaves us with a dilemma: on the one hand, we want to throw away data to create balanced training set, so we can improve the results, and on the other hand, we want to keep as much data as possible, so we can improve the results.
+This paper is an attempt to solve this dilemma.
+This is done by creating two language models for each of the confusibles mentioned above - one on the basis of a balanced training set, and one of the basis of an unbalanced one - and evaluating their results.
 
@@
 \subsection{Method}
+This section is subdivided into three parts.
+In section \ref{timbl}, I will explain the workings of the software behind Fowlt's confusible correctors: TiMBL.
+This is important, because the results reported here might not hold for the confusible correctors of other spell checkers.
+Section \ref{langmod} contains information about the language models created with TiMBl.
+Section \ref{eval}, finally, explains how the models were evaluated.
 
-\subsubsection{The confusible correctors}
+\subsubsection{TiMBL} \label{timbl}
 To investigate the differences between confusible correctors trained on balanced and unbalanced training sets, I will use the confusibles listed in the previous sections.
@@
 
-% Hier ref toevoegen
+\subsubsection{The language models} \label{langmod}
 
-\subsubsection{Evaluation of the confusible correctors}
+\subsubsection{Evaluation procedures} \label{eval}
 
@@
 I think it is important to recognize that spelling correction is a multi-step process, and each step comes with its own measures.
-When using a confusible module as part of a spelling corrector, there are three steps: word prediction, error detection and error correction.
+When using a confusible corrector as part of a spelling corrector, there are three steps: word prediction, error detection and error correction.
 I'll go through these steps and their evaluation measures one by one.
@@
 
-\section{General conclusion}
+\section{General conclusion}%Beter uitleggen hoe Fowlt werkt, eenduidige terminologie hanteren
 
@@
 \end{document}
+
+% TODO
+% Eenduidige terminologie handteren: context-sensitive, modulenamen, 'confusiblecorrectors'
+% Ref TiMBL invoegen

### 2012-12-11 Research training portfolio.tex
RTP: added refs to introduction, worked on method

@@
 \maketitle
-
-\begin{table}[h]
-\begin{tabular}{ll}
-Name&Wessel Stoop\\
-Student numbers&s0808709 (Nijmegen), u1249664 (Tilburg)\\
-Project&Fowlt, a context-based spelling checker\\
-Supervisor&Antal van den Bosch\\
-Period&September - December 2012\\
-\end{tabular}
-\end{table}
 
@@
 Therefore, they detect them by looking at the context of the word, and see if that differs from the contexts they have seen in the past for that word.
-If it, on the basis of the context, would predict another word in that position, and is really certain about it, it flags the word as an error.
+If it, on the basis of that context, would predict another word in that position, and is really certain about it, it flags the word as an error.
 This means, for example, that Fowlt is able to replace the incorrect "there" in the last example into "they're", simply because "there" usually isn't followed by "really nice people", while "they're" is.
-Fowlt includes many confusible correctors like this, but for this paper I will limit myself to six examples.
+Fowlt has many confusible correctors like this, but for this paper I will limit myself to six examples.
 These are the confusibles they are specialized in:
@@
 All confusible correctors in Fowlt consist of two parts: \emph{TiMBL}, which is the same for every corrector, and a \emph{language model}, which is based on the specific confusible error the corrector was created for.
-TiMBL is machine learning software that looks at the contexts of possible, and decides whether it would have expected that word on the basis of that context.
+TiMBL is machine learning software that looks at the contexts of possible confusible errors, and decides whether it would have expected that word on the basis of that context.
 The language model is the 'knowledge' TiMBL uses to make this decision.
-This knowledge was not inserted manually, but was created automatically by looking at large text corpora (so-called \emph{training material}).
-Importantly, the confusible corrector's results improve when the language is based on more training material.
-So the more 'training material', the larger and more complete the language model, which in turn increases the likelihood that TiMBL will make the correct decision.
+This knowledge was not created manually, but automatically by looking at large text corpora (so-called \emph{training material}).
+The more 'training material', the larger and more complete the language model, which in turn increases the likelihood that TiMBL will make the correct decision.
 \\\indent
-Ideally, a language model is created on the basis of a balanced training set, i.e.
-a training set that consists of roughly the same amount of examples for each confusible option.
+But just using lots of data isn't the only thing you can do to create a good language model.
+Ideally, a language model is also created on the basis of a \emph{balanced} training set, i.e.
+a training set that consists of roughly the same amount of examples for each confusible option (see \citealt{hg07} for an overview of literature on this problem).
 If we would use a training set with much more contexts for option A than for option B, we may bias the model to option A.
@@
 A solution to this could of course be to always offer balanced training sets.
-A balanced training set can be achieved in two ways: oversampling and undersampling \citep{provost, gh09}.
+A balanced training set can be achieved in two ways: oversampling and undersampling \citep{provost, gh09, vhkn07, hg07}.
 Oversampling means that extra data are somehow created for the less frequent option.
-Because of the unnatural situations this creates, oversampling probably is not desirable in the spell checking domain (and possibly for NLP in general).
+Because of the unnatural situations this creates, oversampling is probably not desirable in the spell checking domain (and possibly for NLP in general).
 Undersampling means removing examples for the more frequent option: if 100 examples of option A, and 500 examples of option B are available, 400 examples of B will have to be removed.
 The more skewed the distribution is, the more data have to be removed, which can lead to situations where only a very small part of the originally collected training material can be used.
-For example, the distribution of the \emph{practice} versus \emph{practise} confusible is so unbalanced that 85.57 percent of the data had to be removed to create a balanced trainig set.
+For example, the distribution of the \emph{practice} versus \emph{practise} confusible is so unbalanced that 85.57 percent of the data had to be removed to create a balanced training set.
 \\\indent
@@
 
-% Meer literatuur toevoegen
 % Hier even opletten of wat ikz eg wel klopt met mijn uiteindelijke resultaten
@@
 This is important, because the results reported here might not hold for the confusible correctors of other spell checkers.
-Section \ref{langmod} contains information about the language models created with TiMBl.
+Section \ref{langmod} contains information about the language models  TiMBl uses.
 Section \ref{eval}, finally, explains how the models were evaluated.
@@
 \subsubsection{TiMBL} \label{timbl}
-To investigate the differences between confusible correctors trained on balanced and unbalanced training sets, I will use the confusibles listed in the previous sections.
-For each of these confusibles, a confusible corrector is trained to predict which option is likely to occur in which context.
-To be able to recognize contexts, the correctors use \emph{Memory Based Learning}, which means that it saves all contexts previously seen in memory.
-When it is given a new context, it generates a list of previously seen contexts which are relatively similar.
-For each of the contexts of this list, the confusible ...
+TiMBL compares contexts.
+For example, when it is given the task to decide whether "your" in "I think your really nice" is a mistake or not, it compares "I think \_ really nice" to the contexts it has seen in the training material.
+These training contexts are divided into two groups: the contexts for the "your" option and the contexts for the "you're" contexts.
+The context "I think \_ really nice" is more similar to the contexts in the "you're" group, so TiMBL will predict that "you're" should be used here.
+Because this doesn't match with the original word, Fowlt will flag this word as an error.
 This approach is called the k-nearest neighbour algorithm.
-
-The classifier algorithm used using IGTree, a decision-tree approximation of IB1-IG.
+The specific implementation of this algorithm used for Fowlt is IGTree, a decision-tree approximation of IB1-IG \citep{dvdbw97}.
 
 \subsubsection{The language models} \label{langmod}
+
+The six confusibles used for this research were chosen because they vary in their unbalancedness. Here I summarize some descriptive statistics about them:
+
+\begin{table}[h] \footnotesize
+\begin{tabular}{|l|l|p{3cm}|l|p{3cm}|}
+\hline
+Confusible&Distribution&Examples in unbalanced data set&Examples in balanced set&Percentage of data lost in balanced set\\
+\hline
+\emph{advice} versus \emph{advise}&82.8 - 17.2&12480&4094&67.2\\
+\emph{lose} versus \emph{loose}&67.7 - 32.2&9570&6178&35.44\\
+\emph{practice} versus \emph{practise}&92.82 - 7.18&18407&2638&85.67\\
+\emph{than} versus \emph{then}&46.97 - 53.03&302913&284546&6.06\\
+\emph{weather} versus \emph{whether}&13.66 - 86.34&41747&11400&72.69\\
+\emph{you're} versus \emph{your}&21.8 - 78.2&176801&77070&56.41\\
+\hline
+\end{tabular}
+\end{table}
 
@@
 When using a confusible corrector as part of a spelling corrector, there are three steps: word prediction, error detection and error correction.
-I'll go through these steps and their evaluation measures one by one.
+I'll go through these steps and their evaluation measures one by one in the next three paragraphs.
 
-\paragraph{Word prediction}
+\paragraph{Step 1: Word prediction}
 
@@
 \item Prediction accuracy.
-Although word prediction may be the most complicated part of the confusible module, its evaluation is actually the simplest.
-Calculating prediction accuracy entails removing all confusibles from a text, and see in how many of the cases the language model can correctly predict which word was removed.
-This percentage can be increased by only counting predictions that are above a certain threshold.
-Results with multiple thresholds (i.e.
-to what extent does the prediction accuracy improve if the threshold is raised?) can be visualized with a graph, as shown in the result section.
+Although word prediction may be the most complicated part of the confusible correction, its evaluation is actually the simplest.
+Calculating prediction accuracy simply entails removing all confusibles from a text, and see in how many of the cases the language model can correctly predict which word was removed.
+This percentage can be increased by only taking into account predictions that are above a certain threshold.
 \end{itemize}
 
-\paragraph{Error detection}
+\paragraph{Step 2: Error detection}
 
 \begin{itemize}
-\item The confusion matrix.
+\item The detection confusion matrix.
 From this step onwards, we concentrate on spelling errors.
@@
 These numbers are the data needed to calculate various evaluation measures, like the three below.
-Beside these evaluation measures, I will always give the confusion matrix itself, so readers can calculate other measures if desired.
+Beside the evaluation measures, I will always give the confusion matrix itself, so readers can calculate other measures if desired.
 
-\item Detection accuracy. Detection accuracy only indicates how many mistakes the confusible detector makes, not what kind of mistakes. It can be calculated with the following formula:
+\item Detection accuracy.
+Detection accuracy only indicates how many mistakes the confusible detector makes, not what kind of mistakes.
+It can be calculated from the confusion matrix with the following formula:
 
@@
 It is similar to prediction accuracy in that it indicates how well the sytem does in general, but keep in mind that prediction and detection accuracy measure two completely different concepts.
-Although a system with a good prediction accuracy is very likely to also have a good detection accuracy, it might be possible that a system very good at predicting confusibles in errorless texts behaves differently in texts that do contain errors.
+Although a system with a good prediction accuracy is likely to also have a good detection accuracy, it might be possible that a system very good at predicting confusibles in errorless texts behaves differently in texts that do contain errors.
 
@@
 In case of spelling checkers, this means you either have a system that catches a lot of errors, but also has a lot of false alarms, or a system that only finds true errors, but also misses a lot; you can't have the best of both.
+\\
+
+Precision
 
@@
 \]
+
+Recall
 
@@
 Calculating the mean does not help here, because the mean can be relatively high if one of the two is high, but what you want to know is whether \emph{both} are high.
+This is fixed by the F-score, calculated with the following formula.
 
@@
 
+Because this F-score 'summarizes' both precision and recall, it comes close to a measure that indicates how well the system does with it is supposed to do, and that is of course what we are interested in for this research question.
+For this reason, I will leave precision and recall out, and focus on the F-scores.
+
 \end{itemize}
 
-\paragraph{Error correction}
+\paragraph{Step 3: Error correction.} Detecting an error is of course something different from correctly correcting it.
+If correct corrections is what the user is interested in, it is possible to only count as true positives the errors correctly corrected, while the incorrectly corrected errors are counted as false positives.
+From the resulting correction confusion matrix, again the measures correction accuracy, correction recall, correction precision and correction F-score can be calculated.
+\\\indent
+For this paper, the correction measures will always be identical to the detecion measures, and will therefore not be mentioned seperately.
+This is because only two-option confusibles are investigated, which means that the correct correction is always 'the other option' once an error is detected correctly.
+This of course does not hold for confusibles with three or more options, like \emph{to} versus \emph{too} versus \emph{two} or \emph{they're} versus \emph{there} versus \emph{their}.
 
-\begin{itemize}
-\item Confusion matrix.
-\item Correction accuracy.
-\item Correction recall and correction precision.
-\item Correction F-score.
-\end{itemize}
+\paragraph{Multiple thresholds} Results with multiple thresholds (i.e.
+to what extent does the prediction accuracy improve if the threshold is raised?) can be visualized with a graph, as shown in the appendix section.
 
@@
 Brill \& Moore, 1998
+\bibitem[Daelemans, Van den Bosch \& Weijters, 1997]{dvdbw97}
+Daelemans, W., A. Van den Bosch, \& A. Weijters, 1997. IGTree: using trees for compression and classification in lazy learning algorithms. Artificial Intelligence Review, 11:407–423.
 \bibitem[Van Delden et al., 2004]{vandelden04}
@@
 Fawcett, 2004.
-\bibitem[García \& Herrera, 2009]{gh09}
-García \& Herrera, 2009. Evolutionary Undersampling for Classification with Imbalanced Datasets: Proposals and Taxonomy.
+\bibitem[Garcia \& Herrera, 2009]{gh09}
+Garcia \& Herrera, 2009. Evolutionary Undersampling for Classification with Imbalanced Datasets: Proposals and Taxonomy.
+\bibitem[He \& Garcia, 2007]{hg07}
+He \& Garcia, 2007. Learning from imbalanced data. IEEE TRANSACTIONS ON KNOWLEDGE AND DATA ENGINEERING, VOL. 21, NO. 9.
+\bibitem[Van Hulse, Khoshgoftaar \& Napolitano, 2007]{vhkn07}
+Van Hulse, Khoshgoftaar \& Napolitano, 2007. Experimental Perspectives on Learning from Imbalanced Data. Proceedings of the 24 th International Conference on Machine Learning.
 \bibitem[Pollock \& Zamora, 1984]{pz84}

### 2012-12-12 Research training portfolio.tex
RTP: worked on method

@@
 In the literature it has been suggested that balancing the training data, so that it has exactly the same amount of examples for each option, might improve results of classifiers in general.
-In this paper I investigate whether this is true for classifiers in a spelling correction environment.
-
-%hoeveel?
-% UIteindelijke resultaten toevoegen aan het abstract
+In this paper I show this does not hold for classifiers in a spelling correction environment.
 
@@
 This is important, because the results reported here might not hold for the confusible correctors of other spell checkers.
-Section \ref{langmod} contains information about the language models  TiMBl uses.
+Section \ref{langmod} contains information about the language models  TiMBL used.
 Section \ref{eval}, finally, explains how the models were evaluated.
@@
 
-The six confusibles used for this research were chosen because they vary in their unbalancedness. Here I summarize some descriptive statistics about them:
+The language models are the actual decision trees TiMBL uses to makes its decision.
+These models were created by collecting all relevant examples and their contexts from the British National Corpus (ref), and randomly removing 10\% (which were used for evaluation) of them.
+And example consists of the confusible word, three words to its left, and three words to its right.
+The algorithm used to create the models was also part of the TiMBL software package.\\\indent
+As described above, language models were made for six confusibles. These confusibles were chosen because they vary in their unbalancedness. Here I summarize some descriptive statistics about them:
+
+%ref
 
@@
 
-In this section, I describe the various evaluation measures I used.
+The remaining 10\% of the collected data were used to evaluate the confusible correctors.
+In this section, I describe the various evaluation measures I used to do this.
 This is necessary because evaluation of spelling correctors has not been very uniform in the spelling checker literature so far; whereas various scholars only measured how many errors their algorithm can detect in an error list and called it accuracy (e.g.
 \citealp{agirre98, bm00, tm02,vandelden04}), others use or argue for the precision-recall terminology (\citealp{reynaert08,pz84}, although they use other terms for the two concepts) or argue against it \citep{sp02}.\\\indent
-
 Although most of the evaluation measures used in the literature can be useful and informative, they only tell part of the story.
@@
 I'll go through these steps and their evaluation measures one by one in the next three paragraphs.
+The paragraphs after that highlight other concerns when evaluating spelling correctors.
 
@@
 \paragraph{Step 3: Error correction.} Detecting an error is of course something different from correctly correcting it.
-If correct corrections is what the user is interested in, it is possible to only count as true positives the errors correctly corrected, while the incorrectly corrected errors are counted as false positives.
-From the resulting correction confusion matrix, again the measures correction accuracy, correction recall, correction precision and correction F-score can be calculated.
+That is, when a spelling checker corrects 'i like you alot' into 'i like you allot' it might have detected the error, but replaced it with the wrong word.
+If approriate corrections is what the user is interested in, it is possible to only count as true positives the errors correctly corrected, while the incorrectly corrected errors are counted as false positives.
+From the resulting correction confusion matrix the measures correction accuracy, correction recall, correction precision and correction F-score can then be calculated again.
 \\\indent
 For this paper, the correction measures will always be identical to the detecion measures, and will therefore not be mentioned seperately.
-This is because only two-option confusibles are investigated, which means that the correct correction is always 'the other option' once an error is detected correctly.
+This is because only two-option confusibles are investigated, which means that the approriate correction is always 'the other option' once an error is detected correctly.
 This of course does not hold for confusibles with three or more options, like \emph{to} versus \emph{too} versus \emph{two} or \emph{they're} versus \emph{there} versus \emph{their}.
 
-\paragraph{Multiple thresholds} Results with multiple thresholds (i.e.
-to what extent does the prediction accuracy improve if the threshold is raised?) can be visualized with a graph, as shown in the appendix section.
+\paragraph{Simulating errors.} In the description of these measures, I have assumed the existence of a text corpus full of annotated natural errors, so we can easily see which errors were correctly identified, and which were missed.
+Such a corpus unfortunately does not exist, and because of time and money constraints I was unable to create one myself.
+To solve this, such a corpus was created artificially by adding errors to 10\% of the examples used for evaluation.
+I would like to stress, however, that this is a dangerous approach: I risk adding very unnatural errors, making the evaluation results different from real usage results, and thus less useful.
+\\\indent
+Fortunately, our focus on confusibles largely eliminates this problem: we don't only already know which words are often spelled wrong, we also know \emph{how} language users spell them wrong (unlike, for example, typos).
+A problem that still remains, though, is the fact that some errors are not made as often as others.
+For example, "your" might be used instead of "you're" much more than vice versa; ref for example claims that confusible errors are largely based on frequencies.
+Research into the occurence frequencies for all confusible errors would be needed to be able to simulate errors in a more natural way.
+
+%ref
+%Dominiek Sandra toevoegen
+
+\paragraph{Thresholds.} As mentioned in the section about prediction accuracy, predictions come with a confidence, and it is possible to only take into account predictions above a particular threshold.
+This decision has an enormous influence on the rest of the confusible correction process, as in the detection phase every prediction below the threshold means a change either from a true positive into a false negative or from a false positive into a true negative.
+Because a higher threshold means the corrector has to be more confident to flag a word as an error, raising the threshold essentially means lowering the amount of false positives at the cost of lowering the amount of true positives; in other words, increasing the precision at the cost of recall.
+To find out at which threshold there is an optimal precision-recall ratio, the F-score is calculated for multiple thresholds, as can be seen in the results section.
 

### 2012-12-13 Research training portfolio.tex
RTP: Added results, conclusion, rewrote lots of little things.

@@
 \usepackage{natbib}
+\usepackage{float}
 
@@
 
+\begin{table}[h]
+\begin{tabular}{ll}
+Name&Wessel Stoop\\
+Student numbers&s0808709 (Nijmegen), u1249664 (Tilburg)\\
+Project&Fowlt, a context-based spelling checker\\
+Supervisor&Antal van den Bosch\\
+Period&September - December 2012\\
+\end{tabular}
+\end{table}
+
 \section{Introduction}
-Fowlt is the English version of the Dutch spelling checker Valkuil.net.
-Both Valkuil and Fowlt are unlike most well-known spell checkers (like the one in Microsoft Word): whereas these spelling checkers mostly try to find errors by comparing all words to a built-in dictionary and flag the word as an error if they can't find a match, Valkuil and Fowlt are 'context-based'.
+Fowlt is the English version of the Dutch spelling corrector Valkuil.net.
+Both Valkuil and Fowlt are unlike most well-known spelling correctors (like the one in Microsoft Word): whereas these spelling correctors mostly try to find errors by comparing all words to a built-in dictionary and flag the word as an error if they can't find a match, Valkuil and Fowlt are 'context-based'.
 This entails that they also take into account the context, which is the words around every word.
@@
 However, this also means that if the context of a particular error is very different from what the model has seen in the training data, it won't be able to correct the error.
-This means users should be aware of the fact that, although Fowlt can recognize more kinds of errors than regular spell checkers, it might still miss a few.
+This means users should be aware of the fact that, although Fowlt can recognize more kinds of errors than regular spelling correctors, it might still miss a few.
 This is particularly true because Fowlt has been taught to only flag a word as an error if the algorithms are really sure about it.
@@
 \subsection{Project structure}
-Fowlt exist of multiple little spelling checkers, specialized in doing one thing.
-These spelling checkers are called 'modules'.
-Fowlt consists of four context-free modules (relatively similar to normal spelling checkers functionality) and, as I write this, sixteen context-based modules.
+Fowlt exist of multiple little spelling correctors, specialized in doing one thing.
+These spelling correctors are called 'modules'.
+Fowlt consists of four context-free modules (relatively similar to normal spelling correctors functionality) and, as I write this, sixteen context-based modules.
 This number will probably be different when you read this, though, as new context-based modules are created easily, and can be removed if it turns out they're not accurate enough.
@@
 The implementation of this slider was, however, not without problems.
-See section for more information on that.
+See section \ref{formulas} more information on that.
 
@@
 
-\subsubsection{Most modules didn't provide confidences measures}
-Unlike the confusible modules, the WOPR module and the other modules do not return how confident they are a particular error is indeed an error.
-This was a problem for the error tolerance slider, because for it to work it needs have confidence values for all errors, not just the confusibles.
-This not only entailed going back to the C-code of the original modules, but also to come up with formulas to represent exactly how confident the modules are.
+\subsubsection{WOPR didn't behave as expected}
+WOPR wasn't expected to give a lot of problems: it works well in Valkuil, so simply retraining it on English data should have been enough.
+Unfortunately, this turned out not be the case: although WOPR was trained with exactly the same settings, and was used in exactly the same way, it always returned that I didn't discover any errors.
+As I write this, this problem still isn't fully solved, but we did discover three things:
+
+\begin{itemize}
+
+\item The problem is not in the communication between Fowlt and WOPR, as WOPR starts to respond when you tell it it can also give suggestion larger than Levenshtein Distance 1.
+\item The problem also isn't in the training data, as WOPR does find errors when it is used from the command line (instead of while communicating with Fowlt in server mode).
+\item WOPR finds more errors when you turn off the setting \emph{min\_ratio:100}.
+This setting makes sure its suggestion is 100 times more frequent than the original word.
+It is completely unclear why turning this setting on and off makes such a big difference, as it is always on for Valkuil.
+
+\end{itemize}
+
+\subsubsection{Most modules didn't provide confidences measures} \label{formulas}
+Implementing the confidence slider described above was a hard and labor-intensive process.
+Not only did I have to create a working slider in Javascript that behaved like expected in all browsers, this slider also had to look like designer Monica Hajek had envisioned it, and the position of the handle had to be linked to both the errors marked in the text and the errors in the menu (directly above the slider).
+Before this could work, however, the slider needed to be able to extract the confidence for each error from Fowlt's output xml.
+This information about confidence originally wasn't there either, so I had to tell the process chain to ask each module for this information, and add it to the right place in the output.
+All modules subsequently had to be changed so they actually did return their confidence for each error.
+In other words, imlementing the slider was a long journey that made me revisit all layers of Fowlt's architecture.\\\indent
+The biggest problem in this process, however, is that most modules not only didn't return some confidence measure for each error, they actually didn't calculate one.
+This not only meant I had to go back to the C-code of the original modules, but also that I had to come up with formulas to represent exactly how confident the modules are.
 I would like to use this final subsection to explain in detail the formulas I have chosen:
@@
 
-This paper results from my attempts to solve a rather practical issue while building Fowlt, an online spell checker freely available on the internet.
-Fowlt, just like its Dutch equivalent Valkuil, is a context-sensitive spell checker.
+This paper results from my attempts to solve a rather practical issue while building Fowlt, an online spelling corrector freely available on the internet.
+Fowlt, just like its Dutch equivalent Valkuil, is a context-sensitive spelling corrector.
 This context-sensitivity makes it possible to detect errors like the following:
@@
 Oversampling means that extra data are somehow created for the less frequent option.
-Because of the unnatural situations this creates, oversampling is probably not desirable in the spell checking domain (and possibly for NLP in general).
+Because of the unnatural situations this creates, oversampling is probably not desirable in the spelling correction domain (and possibly for NLP in general).
 Undersampling means removing examples for the more frequent option: if 100 examples of option A, and 500 examples of option B are available, 400 examples of B will have to be removed.
@@
 This paper is an attempt to solve this dilemma.
-This is done by creating two language models for each of the confusibles mentioned above - one on the basis of a balanced training set, and one of the basis of an unbalanced one - and evaluating their results.
+The main research question thus is:
+
+\begin{examples}
+
+\item Does throwing away training data to obtain a balanced data set improve the results of confusible correctors?
+
+\end{examples}
+
+This is investigated by creating two language models for each of the confusibles mentioned above - one on the basis of a balanced training set, and one of the basis of an unbalanced one - and evaluating their results.
 
@@
 In section \ref{timbl}, I will explain the workings of the software behind Fowlt's confusible correctors: TiMBL.
-This is important, because the results reported here might not hold for the confusible correctors of other spell checkers.
+This is important, because the results reported here might not hold for the confusible correctors of other spelling correctors.
 Section \ref{langmod} contains information about the language models  TiMBL used.
@@
 A problem that still remains, though, is the fact that some errors are not made as often as others.
-For example, "your" might be used instead of "you're" much more than vice versa; ref for example claims that confusible errors are largely based on frequencies.
+For example, "your" might be used instead of "you're" much more than vice versa; \citet{sfd99} for example claim that confusible errors are largely based on frequencies.
 Research into the occurence frequencies for all confusible errors would be needed to be able to simulate errors in a more natural way.
@@
 %ref
-%Dominiek Sandra toevoegen
 
@@
 
-\paragraph{Problems with precision and recall.} Some text.
-
-Waarom geen AUC?
+\paragraph{Problems with precision and recall.} Although I limit myself to precision, recall, and the F-scores calculated from them, these measures are not without problems.
+As can be seen from the formulas given above, for example, they completely ignore the 'true negative' box in the confusion matrix.
+As \citet{fawcett04} shows, this can have quite extreme implications when the class distribution changes, and proposes ROC curves instead; this entails by pitting the true positive rate against the false positive rate.
+However, for spelling correction this measure is not useful: because the amount of negatives (correct words) usually is many times larger than the amount of positives (incorrect words), this always leads to false positive rates that are near-zero.
+This in turns causes the Area Under the Curve is almost completely determined by the recall, which is of course only part of what we're interested in.
 
@@
 
+In this section I present the F-scores of all confusible correctors, calculated for 21 different thresholds.
+Because all examples were assigned to either the training set and the test set randomly (as was explained in the Methods section, 90\% of the data were used to train the correctors, while 10\% was used to calculated these measures), and also because the errors added to the test set were added randomly, these results are not exactly reproducable: a confusible corrector trained on other examples and tested on other errors might of course have other results.
+Nevertheless, a quick test showed that the F-scores are relatively stable: for all correctors, when the procedure was done again, the F-scores were less than 0.01 higher or lower, which suggest that the results presented here are good reflections of the correctors' actual performance.\\\indent
+Prediction accuracy for multiple thresholds and confusion matrices were calculated as well. These can be found in the appendix.
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{fscore_adviceadvise.png}
+\caption{The F-scores of corrector for the \emph{advice} versus \emph{advise}, trained on both a balanced and an inbalanced data set.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{fscore_loseloose.png}
+\caption{The F-scores of corrector for the \emph{lose} versus \emph{loose}, trained on both a balanced and an inbalanced data set.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{fscore_practicepractise.png}
+\caption{The F-scores of corrector for the \emph{practice} versus \emph{practise}, trained on both a balanced and an inbalanced data set.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{fscore_thanthen.png}
+\caption{The F-scores of corrector for the \emph{than} versus \emph{then}, trained on both a balanced and an inbalanced data set.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{fscore_weatherwhether.png}
+\caption{The F-scores of corrector for the \emph{weather} versus \emph{whether}, trained on both a balanced and an inbalanced data set.}
+\end{figure}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{fscore_you'reyour.png}
+\caption{The F-scores of corrector for the \emph{you're} versus \emph{your}, trained on both a balanced and an inbalanced data set.}
+\end{figure}
+
+With respect to our research question, two important facts can be distilled from these data:
+
+\begin{itemize}
+
+\item The corrector trained on the balanced data set (the blue line in the graph) always performs similar or worse than the unbalanced set.
+Only for the \emph{lose} versus \emph{loose} corrector we see that the balanced set provides the best results.
+However, the difference between the F-scores of the balanced and unbalanced set in less than 0.01, which means that this better performance can attributed to chance.
+\item How much better the corrector trained on the unbalanced data set actually is seems to be related to how much data was lost.
+For example, for the balanced set for the \emph{than} versus \emph{then} corrector only 6.06\% of the data had to be thrown away, and we indeed see that the correctors based on the balanced and unbalanced perform similar.
+For the \emph{practice} versus \emph{practise} corrector, on the other hand, 85.67\% of the data had to be thrown away, and here was see that the corrector trained on the large, unbalanced data set performs much better.
+Spearman's rho\footnote{Spearman's rho was chosen over Pearson's r, because an early plot of the data indicated that the correlation was not linear.} indicates that there is indeed a strong positive correlation between difference in F-score and proportion of data lost ($r_s(4)=.812, p=.05$).
+This means that the more data that are lost, the more the corrector based on the unbalanced set will outperform the corrector based on the balanced data set.
+
+% stat toevoegen
+
+\end{itemize}
+
 \subsection{Conclusion}
+In this study, six confusible correctors were trained on both a balanced and an inbalanced data set.
+It was found that the correctors using the imbalanced training data always performed similar and mostly better than the balanced training data.
+A strong correlation was found between how much data had to be removed to create the balanced data set and to what extent the corrector based on the imbalanced training data outperformed its balanced counterpart.
+This suggests that the dataset being large is more important than the dataset being balanced.
+These findings form a contrast with suggestions in the literature that imbalanced training data should be avoided.
 
-\section{General conclusion}%Beter uitleggen hoe Fowlt werkt, eenduidige terminologie hanteren
+\section{General conclusion}
+
+\section{Appendix}
 
@@
 Reyneart, 2008.
+\bibitem[Sandra et al., 1999]{sfd99}
+Sandra, D., Frisson, S.,Daems, F., 1999. Why simple verb forms can be so difficult to spell: the influence of homophone frequency and distance in Dutch. Brain Lang. 1999 Jun 1-15;68(1-2):277-83.
 \bibitem[Starlander \& Popescu-Belis, 2002]{sp02}
@@
 % TODO
-% Eenduidige terminologie handteren: context-sensitive, modulenamen, 'confusiblecorrectors'
-% Ref TiMBL invoegen
+% Eenduidige terminologie hanteren:  modulenamen
+% Appendix

### 2012-12-18 Research training portfolio.tex
RTP: added full references

@@
 Ideally, a language model is also created on the basis of a \emph{balanced} training set, i.e.
-a training set that consists of roughly the same amount of examples for each confusible option (see \citealt{hg07} for an overview of literature on this problem).
+a training set that consists of roughly the same amount of examples for each confusible option (see \citealt{hg09} for an overview of literature on this problem).
 If we would use a training set with much more contexts for option A than for option B, we may bias the model to option A.
@@
 A solution to this could of course be to always offer balanced training sets.
-A balanced training set can be achieved in two ways: oversampling and undersampling \citep{provost, gh09, vhkn07, hg07}.
+A balanced training set can be achieved in two ways: oversampling and undersampling \citep{provost, gh09, vhkn07, hg09}.
 Oversampling means that extra data are somehow created for the less frequent option.
@@
 \bibitem[Agirre et al., 1998]{agirre98}
-Agirre et al., 1998
+Agirre, E., K. Gojenola, K. Sarasola, \& A. Voutilainen (1998). Towards a Single Proposal in Spelling Correction. \emph{COLING-ACL ’98},  22-28.
 \bibitem[Brill \& Moore, 2000]{bm00}
-Brill \& Moore, 1998
+Brill, E. \& R. C. Moore (2000). An improved error model for noisy channel spelling correction. In \emph{Proceedings of the 38th Annual Meeting of the ACL}, 286-293.
 \bibitem[Daelemans, Van den Bosch \& Weijters, 1997]{dvdbw97}
-Daelemans, W., A. Van den Bosch, \& A. Weijters, 1997. IGTree: using trees for compression and classification in lazy learning algorithms. Artificial Intelligence Review, 11:407–423.
+Daelemans, W., A. van den Bosch, \& A. Weijters (1997). IGTree: using trees for compression and classification in lazy learning algorithms. \emph{Artificial Intelligence Review, 11}, 407–423.
 \bibitem[Van Delden et al., 2004]{vandelden04}
-Van Delden et al, 2004.
+Van Delden, S., D.
+B.
+Bracewell, \& F.
+Gomez (2004).
+Supervised and unsupervised automatic spelling correction algorithms.
+In D.
+Zhang, E.
+Gregoire \& D.
+DeGroot (eds), \emph{Proceedings of the 2004 IEEE International Conference on Information Reuse and Integration}, 530–535, Las Vegas.
 \bibitem[Fawcett, 2004]{fawcett04}
-Fawcett, 2004.
+Fawcett, T. (2004). \emph{ROC Graphs: Notes and Practical Considerations for Researchers.} Available online at binf.gmu.edu/mmasso/ROC101.pdf.
 \bibitem[Garcia \& Herrera, 2009]{gh09}
-Garcia \& Herrera, 2009. Evolutionary Undersampling for Classification with Imbalanced Datasets: Proposals and Taxonomy.
-\bibitem[He \& Garcia, 2007]{hg07}
-He \& Garcia, 2007. Learning from imbalanced data. IEEE TRANSACTIONS ON KNOWLEDGE AND DATA ENGINEERING, VOL. 21, NO. 9.
+Garcia, S. \& F. Herrera (2009). Evolutionary Undersampling for Classification with Imbalanced Datasets: Proposals and Taxonomy. \emph{Evolutionary Computation 17}, 275-306.
+\bibitem[He \& Garcia, 2009]{hg09}
+He, H. \& E. Garcia (2009). Learning from imbalanced data. \emph{IEEE Transactions on knowledge and data engineering, 21(9)}, 1263-1281.
 \bibitem[Van Hulse, Khoshgoftaar \& Napolitano, 2007]{vhkn07}
-Van Hulse, Khoshgoftaar \& Napolitano, 2007. Experimental Perspectives on Learning from Imbalanced Data. Proceedings of the 24 th International Conference on Machine Learning.
+Van Hulse, J. T. M. Khoshgoftaar \& A. Napolitano (2007). Experimental Perspectives on Learning from Imbalanced Data. \emph{Proceedings of the 24 th International Conference on Machine Learning.}
 \bibitem[Pollock \& Zamora, 1984]{pz84}
-Pollock \& Zamora, 1984
-\bibitem[Provost, 2000]{provost}
-Provost, F (2000). Machine Learning from Imbalanced Data Sets 101. Extended abstract. 
-%jaar is nep
+Pollock J. J. \& A. Zamora (1984). Automatic spelling correction in scientific and scholarly text. \emph{Commun. ACM, 27(4)}, 358-368.
+\bibitem[Provost, ?]{provost}
+Provost, F. (year unkown). \emph{Machine Learning from Imbalanced Data Sets 101.} Extended abstract.
 \bibitem[Reynaert, 2008]{reynaert08}
-Reyneart, 2008.
+Reyneart, M.
+(2008).
+All, and only, the errors: More complete and consistent spelling and OCR-error correction evaluation.
+\emph{Proceedings of the Sixth International Language Resources and Evaluation (LREC’08), Marrakech, Morocco}.
 \bibitem[Sandra et al., 1999]{sfd99}
-Sandra, D., Frisson, S.,Daems, F., 1999. Why simple verb forms can be so difficult to spell: the influence of homophone frequency and distance in Dutch. Brain Lang. 1999 Jun 1-15;68(1-2):277-83.
+Sandra, D., Frisson, S.,Daems, F., (1999). Why simple verb forms can be so difficult to spell: the influence of homophone frequency and distance in Dutch. \emph{Brain Language, 68(1-2),} 277-283.
 \bibitem[Starlander \& Popescu-Belis, 2002]{sp02}
-Starlander \& Popescu-Belis, 2002
+Starlander, M.
+\& A.
+Popescu-Belis (2002).
+Corpusbased evaluation of a French spelling and grammar checker.
+In \emph{LREC 2002: Third International Conference on language resources and evaluation, volume 1}, 268–274.
+Paris: ELRA, European Language Resources.
 \bibitem[Toutanova \& Moore, 2002]{tm02}
-Toutanova \& Moore, 2002.
+Toutanova, K. \& R. C. Moore (2002). Pronunciation modeling for improved spelling correction. In \emph{Proceedings of the 40th Annual Meeting of the ACL}, 144–151.
 

### 2012-12-18 Research training portfolio.tex
RTP: added pictures and bibliography

@@
 
-%Plaatje toevoegen
+\begin{figure}[H]
+\centering
+\includegraphics[width=1\textwidth]{fowlt_chain.png}
+\caption{Fowlt's architecture}
+\label{fig:chain}
+\end{figure}
 
@@
 
-\subsubsection{ErrorlistChecker (context-free)}
+\subsubsection{Errorlist module (context-free)}
 Martin Reynaert provided us with a large list of common typos and their corrections.
@@
 
-\subsubsection{LexiconChecker (context-free)}
+\subsubsection{Lexicon module (context-free)}
 Google provided us with an enormous list of word frequencies on the internet in 2008.
@@
 
-\subsubsection{RunonChecker (context-free)}
+\subsubsection{Runon module (context-free)}
 The RunonChecker also uses Google's list, but uses it to check if any spaces were forgotten in the input.
@@
 
-\subsubsection{SplitChecker (context-free)}
+\subsubsection{Split module (context-free)}
 The SplitChecker is the exact opposite of the RunonChecker: instead of looking whether any spaces are forgotten, it checks whether any spaces have to be added.
@@
 
-\subsubsection{WOPRChecker (context-based)}
+\subsubsection{WOPR module(context-based)}
 The WOPRChecker contacts WOPR, running on a server on localhost or elsewhere.
@@
 
-\subsubsection{The ConfusibleCheckers (context-based)}
+\subsubsection{The Confusible modules (context-based)}
 The ConfusibleCheckers work similar to the WOPRChecker, with the only difference that they are trained on one specific language problem, instead of on language in general.
@@
 \subsubsection{Fowlt as a web application}
-%Picture reference
 Fowlt can also (and, probably, will mostly) be used as a website.
 In this web variant, Fowlt works exactly as described above, but some extra mechanisms are built around to make sure it works and looks nice in the browser.
-The transition from a standalone application to a website is a two-step process, as is exemplified by figure...
-.
+The transition from a standalone application to a website is a two-step process, as is exemplified by Figure \ref{fig:chain}.
 Firstly, Fowlt is made available online webservice using the software CLAM.
@@
 Fortunately, working with Linux became routine quickly.
-
-% Voordelen Linux?
 
@@
 
-% stat toevoegen
-
 \end{itemize}
@@
 \section{Appendix}
+
+\subsection{Advice versus advice corrector}
+
+\begin{table}[H] \footnotesize
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+&&P&N\\
+\hline
+balanced&Y&0&0\\
+&N&0&0\\
+\hline
+\hline
+unbalanced&Y&0&0\\
+&N&0&0\\
+\hline
+\end{tabular}
+\caption{The confusion matrices for the corrector for the \emph{advice} versus \emph{advise}, trained on both a balanced and an inbalanced data set.}
+\end{table}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{accuracy_adviceadvise.png}
+\caption{The accuracies of corrector for the \emph{advice} versus \emph{advise}, trained on both a balanced and an inbalanced data set.}
+\end{figure}
+
+\subsection{Lose versus loose corrector}
+
+\begin{table}[H] \footnotesize
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+&&P&N\\
+\hline
+balanced&Y&0&0\\
+&N&0&0\\
+\hline
+\hline
+unbalanced&Y&0&0\\
+&N&0&0\\
+\hline
+\end{tabular}
+\caption{The confusion matrices for the corrector for the \emph{lose} versus \emph{loose}, trained on both a balanced and an inbalanced data set.}
+\end{table}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{accuracy_loseloose.png}
+\caption{The accuracies of corrector for the \emph{lose} versus \emph{loose}, trained on both a balanced and an inbalanced data set.}
+\end{figure}
+
+\subsection{Practice versus practise corrector}
+
+\begin{table}[H] \footnotesize
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+&&P&N\\
+\hline
+balanced&Y&0&0\\
+&N&0&0\\
+\hline
+\hline
+unbalanced&Y&0&0\\
+&N&0&0\\
+\hline
+\end{tabular}
+\caption{The confusion matrices for the corrector for the \emph{practice} versus \emph{practise}, trained on both a balanced and an inbalanced data set.}
+\end{table}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{accuracy_practicepractise.png}
+\caption{The accuracies of corrector for the \emph{practice} versus \emph{practise}, trained on both a balanced and an inbalanced data set.}
+\end{figure}
+
+\subsection{Than versus then corrector}
+
+\begin{table}[H] \footnotesize
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+&&P&N\\
+\hline
+balanced&Y&0&0\\
+&N&0&0\\
+\hline
+\hline
+unbalanced&Y&0&0\\
+&N&0&0\\
+\hline
+\end{tabular}
+\caption{The confusion matrices for the corrector for the \emph{than} versus \emph{then}, trained on both a balanced and an inbalanced data set.}
+\end{table}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{accuracy_thanthen.png}
+\caption{The accuracies of corrector for the \emph{than} versus \emph{then}, trained on both a balanced and an inbalanced data set.}
+\end{figure}
+
+\subsection{Weather versus whether corrector}
+
+\begin{table}[H] \footnotesize
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+&&P&N\\
+\hline
+balanced&Y&0&0\\
+&N&0&0\\
+\hline
+\hline
+unbalanced&Y&0&0\\
+&N&0&0\\
+\hline
+\end{tabular}
+\caption{The confusion matrices for the corrector for the \emph{weather} versus \emph{whether}, trained on both a balanced and an inbalanced data set.}
+\end{table}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{accuracy_weatherwhether.png}
+\caption{The accuracies of corrector for the \emph{weather} versus \emph{whether}, trained on both a balanced and an inbalanced data set.}
+\end{figure}
+
+\subsection{You're versus your corrector}
+
+\begin{table}[H] \footnotesize
+\centering
+\begin{tabular}{|c|c|c|c|}
+\hline
+&&P&N\\
+\hline
+balanced&Y&0&0\\
+&N&0&0\\
+\hline
+\hline
+unbalanced&Y&0&0\\
+&N&0&0\\
+\hline
+\end{tabular}
+\caption{The confusion matrices for the corrector for the \emph{you're} versus \emph{your}, trained on both a balanced and an inbalanced data set.}
+\end{table}
+
+\begin{figure}[H]
+\centering
+\includegraphics[width=0.8\textwidth]{accuracy_you'reyour.png}
+\caption{The accuracies of corrector for the \emph{you're} versus \emph{your}, trained on both a balanced and an inbalanced data set.}
+\end{figure}
 
@@
 \end{document}
-
-% TODO
-% Eenduidige terminologie hanteren:  modulenamen
-% Appendix

### 2012-12-18 Research training portfolio.tex
RTP: revised the non-research parts, added general conclusion

@@
 Student numbers&s0808709 (Nijmegen), u1249664 (Tilburg)\\
-Project&Fowlt, a context-based spelling checker\\
+Project&Fowlt, a context-sensitive spelling checker\\
 Supervisor&Antal van den Bosch\\
@@
 Fowlt is the English version of the Dutch spelling corrector Valkuil.net.
-Both Valkuil and Fowlt are unlike most well-known spelling correctors (like the one in Microsoft Word): whereas these spelling correctors mostly try to find errors by comparing all words to a built-in dictionary and flag the word as an error if they can't find a match, Valkuil and Fowlt are 'context-based'.
+Both Valkuil and Fowlt are unlike most well-known spelling correctors (like the one in Microsoft Word): whereas these spelling correctors mostly try to find errors by comparing all words to a built-in dictionary and flag the word as an error if they can't find a match, Valkuil and Fowlt are 'context-sensitive'.
 This entails that they also take into account the context, which is the words around every word.
@@
 This means users should be aware of the fact that, although Fowlt can recognize more kinds of errors than regular spelling correctors, it might still miss a few.
-This is particularly true because Fowlt has been taught to only flag a word as an error if the algorithms are really sure about it.
+\\\indent
+Although most of software was already there when I started, this project has been a long journey, with many problems.
+In this portfolio, I hope to give the reader an idea of what I did, what I learned, and what went wrong.
+On top of that, I did a little study on Fowlt's confusible checkers.
+I report this study in section 4.
+\\\indent
 
@@
 \subsection{Project structure}
-Fowlt exist of multiple little spelling correctors, specialized in doing one thing.
+Fowlt exists of multiple little spelling correctors, specialized in doing one thing.
 These spelling correctors are called 'modules'.
-Fowlt consists of four context-free modules (relatively similar to normal spelling correctors functionality) and, as I write this, sixteen context-based modules.
-This number will probably be different when you read this, though, as new context-based modules are created easily, and can be removed if it turns out they're not accurate enough.
+Fowlt consists of four context-free modules (relatively similar to normal spelling correctors functionality) and, as I write this, sixteen context-sensitive modules.
+This number will probably be different when you read this, though, as new context-sensitive modules are created easily, and can be removed if it turns out they're not accurate enough.
 Although it is possible to run Fowlt as a stand-alone application, most people will use it as a web application.
@@
 In Fowlt's main script, fowlt\_processchain.py, all modules are started up, their outputs are collected and are all saved in xml form.
-Making changes in this script (based on the ones for Valkuil, written by Maarten van Gompel) to make sure the new English based modules worked correctly, is one of the main things I did during my internship.
+Making changes in this script (based on the one for Valkuil, written by Maarten van Gompel) to make sure the new English based modules worked correctly, is one of the main things I did during my internship.
 
@@
 \subsubsection{Runon module (context-free)}
-The RunonChecker also uses Google's list, but uses it to check if any spaces were forgotten in the input.
+The Runon module also uses Google's list, but uses it to check if any spaces were forgotten in the input.
 This is done by looking whether splitting up long words produces two words which both are much more frequent that the original word.
@@
 \subsubsection{Split module (context-free)}
-The SplitChecker is the exact opposite of the RunonChecker: instead of looking whether any spaces are forgotten, it checks whether any spaces have to be added.
+The Split module is the exact opposite of the RunonChecker: instead of looking whether any spaces are forgotten, it checks whether any spaces have to be added.
 This is done by looking whether each combination of two words produces a word which is much more frequent than the two original words.
 
-\subsubsection{WOPR module(context-based)}
-The WOPRChecker contacts WOPR, running on a server on localhost or elsewhere.
+\subsubsection{WOPR module(context-sensitive)}
+The WOPR module contacts WOPR, running on a server on localhost or elsewhere.
 It gives it the context of each word, but not the words themselves.
@@
 
-\subsubsection{The Confusible modules (context-based)}
-The ConfusibleCheckers work similar to the WOPRChecker, with the only difference that they are trained on one specific language problem, instead of on language in general.
-As I write this, there are fifteen active ConfusibleCheckers, all focussing on their own expertise: lose-loose, it's-its, you're-your, than-then, who-which-that, whether-weather, effect-affect, lie-lay, they're-their-there, don't-doesn't,to-too-two, advice-advise, any-some, less-fewer, practice-practise and chose-choose.
+\subsubsection{The Confusible modules (context-sensitive)}
+The confusible modules work similar to the WOPR module, with the only difference that they are trained on one specific language problem, instead of on language in general.
+As I write this, there are fifteen active confusible modules, all focussing on their own expertise: lose-loose, it's-its, you're-your, than-then, who-which-that, whether-weather, effect-affect, lie-lay, they're-their-there, don't-doesn't,to-too-two, advice-advise, any-some, less-fewer, practice-practise and chose-choose.
 \\\indent
-A ConfusibleChecker trained on the difference between 'then' and 'than' for example filters out these two words and their contexts, sends these contexts to a Timblserver running somewhere.
+A confusible module trained on the difference between 'then' and 'than' for example filters out these two words and their contexts, sends these contexts to a Timblserver running somewhere.
 This Timblserver then says which of the two it would predict on the basis of the context.
 If this prediction is different from the actual word and the server is very sure about this, the word is flagged as an error.
-Training new ConfusibleModules has been a large part of my internship.
 
@@
 Fowlt can also (and, probably, will mostly) be used as a website.
-In this web variant, Fowlt works exactly as described above, but some extra mechanisms are built around to make sure it works and looks nice in the browser.
-The transition from a standalone application to a website is a two-step process, as is exemplified by Figure \ref{fig:chain}.
+In this web variant, Fowlt works exactly as described above, but some extra mechanisms are built around it to make sure it works and looks nice in the browser.
+The transition from a standalone application to a website is a two-step process.
 Firstly, Fowlt is made available online webservice using the software CLAM.
 Among other things, CLAM adds the possibility to use Fowlt as a RESTful service.
-However, ordinary users don't want to send HTTP requests or read xml files when they want to texts corrected.
+However, ordinary users don't want to send HTTP requests or read xml files when they want their texts corrected.
 This is solved in the second step, which is the Fowlt.net interface.
@@
 The interface even adds functionality to correct any errors.
-Behind the screens, the interface communicates of course with the CLAM webservice.
+Behind the screens, the interface communicates with the CLAM webservice.
 \\\indent
-%Section reference
 Most of the functionality described above was already working for Valkuil, and so only had to be adapted to English to be function within Fowlt.
@@
 
-\subsubsection{The original confusiblemodule could only handle two confusibles}
+\subsubsection{The original confusible module could only handle two confusibles}
 Most confusibles consist of two words only, but English has two that consist of three: they're-there-their and to-too-two.
@@
 WOPR wasn't expected to give a lot of problems: it works well in Valkuil, so simply retraining it on English data should have been enough.
-Unfortunately, this turned out not be the case: although WOPR was trained with exactly the same settings, and was used in exactly the same way, it always returned that I didn't discover any errors.
+Unfortunately, this turned out not be the case: although WOPR was trained with exactly the same settings, and was used in exactly the same way, it always returned that it didn't discover any errors.
 As I write this, this problem still isn't fully solved, but we did discover three things:
@@
 
-\item The problem is not in the communication between Fowlt and WOPR, as WOPR starts to respond when you tell it it can also give suggestion larger than Levenshtein Distance 1.
+\item The problem is not in the communication between Fowlt and WOPR, as WOPR starts to respond when you tell it it can also give suggestions that have a larger Levenshtein Distance than 1.
 \item The problem also isn't in the training data, as WOPR does find errors when it is used from the command line (instead of while communicating with Fowlt in server mode).
@@
 
+\subsection{All browsers and operating systems treat css differently}
+Fowlt's web interface, designed by Monica Hajek, is very similar to Valkuil's interface.
+I thought that simply changing the logo and some colors would be all interface-related things I had to do.
+I couldn't have been more wrong: Monica wanted every GUI element to be on exactly the right place, pixel-perfect.
+This meant that every single element I had added (and even some elements I hadn't added) had to be repositioned.\\\indent
+Having used css several times before, I thought this entailed changing some values and adding some things to Fowlt's stylesheet.
+Unfortunately, this turned out to be a very hard, frustrating and time-consuming task, as no two browsers behave exactly the same.
+This means that having all positioning correct in one browsers means having all positioning horribly wrong in another.
+And to make things even more complicated, things differ per operating system as well: although you can use Chrome on Windows, Mac and Linux, Fowlt looks very different on each of them.\\\indent
+After many frustrated mornings, I decided to give up on this problem, and instead focus on the most used browser-operating system combination, which is Windows+Chrome.
+Monica provided me with a normalizer style sheet, which claims to eliminate browsers differences and indeed fixes some problems, but at the moment the problem is far from solved.
+
 \subsubsection{Most modules didn't provide confidences measures} \label{formulas}
 Implementing the confidence slider described above was a hard and labor-intensive process.
-Not only did I have to create a working slider in Javascript that behaved like expected in all browsers, this slider also had to look like designer Monica Hajek had envisioned it, and the position of the handle had to be linked to both the errors marked in the text and the errors in the menu (directly above the slider).
+Not only did I have to create a working slider in Javascript that behaved like expected in all browsers, this slider also had to look like Monica had envisioned it, and the position of the handle had to be linked to both the errors marked in the text and the errors in the menu (directly above the slider).
 Before this could work, however, the slider needed to be able to extract the confidence for each error from Fowlt's output xml.
@@
 However, the real problems with setting up an NLP webservice were not related to Django or CLAM, but to all kinds of unexpected behavior of the university's servers.
-For example, the user groups on Applejack were not set up correctly when I started, prohibiting me to reach some locations, both Spitfire and Applejack reset file permissions when files are updated, and the server software caches server settings, even after update the configuration file.
-In retrospect, the problems were minor and could be solved easily (either by me or by an administrator), but it took me quite a lot of time figure out what the problem was and how to solve it.
+For example, the user groups on Applejack were not set up correctly when I started, prohibiting me to reach some locations, both Spitfire and Applejack reset file permissions when files are updated, and the server software caches server settings, even after updating the configuration file.
+In retrospect, the problems were minor and could be solved easily (either by me or by an administrator), but it took me days to figure out what the problems were and how to solve them.
 
@@
 
-% Python stuff
-% C compilen
+Besides these general themes, there were several similar learning moments per day.
+Things that I learned include how to use the Google Calendar, the concept IRC (of which I had only a very basic understanding), the fact that bash shell aliases make your life a lot easier, how to make Latex show pictures exactly where you want them, what Python list comprehensions and lambda functions really are, and how to compile C code (and that it should be compiled in the first place).
 
@@
 \section{General conclusion}
+In this research training portfolio, I hope to have shown that creating a spelling corrector isn't something that can be done in a day, even if most of the code is already available.
+I didn't only change what was there to work with English, I also had to solve several problems related to processing the English language, extend the software on some places (for example, with the confidence slider), and did a little study on the confusible correctors' training data.
+While doing this, I had to get familiar with new concepts (SSH, GitHub, MBL) on an almost daily basis.\\\indent
+Perhaps suprisingly, the fact that I have created something which could (hopefully) be useful to hundreds, if not thousands or millions of people around the globe, and the fact that I have gathered lots of new and super useful knowledge, are not the most valuable things I got out of this internship.
+The most valuable thing is that I rediscovered language technology as something I love, and that I now know what I want to spend my future doing.
 

### 2012-12-20 Research training portfolio.tex
RTP: Final revisions

@@
 For many confusibles, like "\emph{you're} versus \emph{your}", one option is much more frequent than the other.
-This poses a problem when one tries to create a context-sensitive memory-based confusible corrector, as having more data for one option will make it more likely that the confusible corrector will predict that option.
+This poses a problem when one tries to use a memory-based classifier as a context-sensitive confusible corrector, as having more data for one option might make it more likely that the confusible corrector will incorrectly predict that option.
 In the literature it has been suggested that balancing the training data, so that it has exactly the same amount of examples for each option, might improve results of classifiers in general.
@@
 Therefore, they detect them by looking at the context of the word, and see if that differs from the contexts they have seen in the past for that word.
-If it, on the basis of that context, would predict another word in that position, and is really certain about it, it flags the word as an error.
-This means, for example, that Fowlt is able to replace the incorrect "there" in the last example into "they're", simply because "there" usually isn't followed by "really nice people", while "they're" is.
+If the confusible correctors, on the basis of the context, would predict another word in that position, and are really certain about it, it flags the word as an error.
+This means, for example, that Fowlt is able to replace the incorrect "there" in "there really nice people" into "they're", simply because "there" usually isn't followed by "really nice people", while "they're" is.
 Fowlt has many confusible correctors like this, but for this paper I will limit myself to six examples.
@@
 Ideally, a language model is also created on the basis of a \emph{balanced} training set, i.e.
-a training set that consists of roughly the same amount of examples for each confusible option (see \citealt{hg09} for an overview of literature on this problem).
+a training set that consists of roughly the same amount of examples for each confusible option (see \citealt{hg09} for an overview of literature on this topic).
 If we would use a training set with much more contexts for option A than for option B, we may bias the model to option A.
@@
 \\\indent
-This leaves us with a dilemma: on the one hand, we want to throw away data to create balanced training set, so we can improve the results, and on the other hand, we want to keep as much data as possible, so we can improve the results.
+This leaves us with a dilemma: on the one hand, we want to throw away data to create balanced training set, so we can improve the results, and on the other hand, we want to keep as much data as possible, also to improve the results.
 This paper is an attempt to solve this dilemma.
@@
 This approach is called the k-nearest neighbour algorithm.
-The specific implementation of this algorithm used for Fowlt is IGTree, a decision-tree approximation of IB1-IG \citep{dvdbw97}.
+The specific implementation of this algorithm used for Fowlt is IGTree, a decision-tree approximation of IB1-IG \citep*{dvdbw97}.
 
@@
 The language models are the actual decision trees TiMBL uses to makes its decision.
-These models were created by collecting all relevant examples and their contexts from the British National Corpus (ref), and randomly removing 10\% (which were used for evaluation) of them.
+These models were created by collecting all relevant examples and their contexts from the British National Corpus, and randomly removing 10\% of them (which were used for evaluation).
 And example consists of the confusible word, three words to its left, and three words to its right.
@@
 As described above, language models were made for six confusibles. These confusibles were chosen because they vary in their unbalancedness. Here I summarize some descriptive statistics about them:
-
-%ref
 
@@
 This means that the system can do two things right (correct incorrect words in the text and don't correct correct words), and two things wrong (correct correct words and don't correct incorrect words).
-This can be summarized in a confusion matrix (taken from \citet{fawcett04}):
+This can be summarized in a confusion matrix, as shown in figure \ref{fig:confusion} (taken from \citet{fawcett04}).
 
@@
 
-It is similar to prediction accuracy in that it indicates how well the sytem does in general, but keep in mind that prediction and detection accuracy measure two completely different concepts.
+It is similar to prediction accuracy in that it indicates how well the system does in general, but keep in mind that prediction and detection accuracy measure two completely different concepts.
 Although a system with a good prediction accuracy is likely to also have a good detection accuracy, it might be possible that a system very good at predicting confusibles in errorless texts behaves differently in texts that do contain errors.
@@
 Precision and recall are concepts well-known within the field of Information Retrieval; precision reflects how much of the retrieved things were relevant, while recall reflects how much of the the relevant things were retrieved.
-Often there is a inverse relationship between precision and recall: if you change an algorithm so that its recall increases, its precision decreases, and the other way around.
+Often there is a inverse relationship between precision and recall: if an algorithm os changed so that its recall increases, its precision decreases, and the other way around.
 In case of spelling checkers, this means you either have a system that catches a lot of errors, but also has a lot of false alarms, or a system that only finds true errors, but also misses a lot; you can't have the best of both.
-\\
+They are calculated as follows:\\
 
@@
 Although you mostly cannot have an optimal precision \emph{and} recall at the same time, you can of course try to have them both as high as possible.
-Calculating the mean does not help here, because the mean can be relatively high if one of the two is high, but what you want to know is whether \emph{both} are high.
-This is fixed by the F-score, calculated with the following formula.
+Calculating the mean of the two does not help here, because the mean can be relatively high if one of the two is high.
+What is relevant is whether \emph{both} are high.
+This is fixed by the F-score, calculated with the following formula:
 
@@
 To solve this, such a corpus was created artificially by adding errors to 10\% of the examples used for evaluation.
-I would like to stress, however, that this is a dangerous approach: I risk adding very unnatural errors, making the evaluation results different from real usage results, and thus less useful.
+I would like to stress, however, that this is a dangerous approach: by adding artifical errors one risks adding very unnatural errors, making the evaluation results different from real usage results, and thus less useful.
 \\\indent
@@
 
-%ref
-
 \paragraph{Thresholds.} As mentioned in the section about prediction accuracy, predictions come with a confidence, and it is possible to only take into account predictions above a particular threshold.
@@
 Because a higher threshold means the corrector has to be more confident to flag a word as an error, raising the threshold essentially means lowering the amount of false positives at the cost of lowering the amount of true positives; in other words, increasing the precision at the cost of recall.
-To find out at which threshold there is an optimal precision-recall ratio, the F-score is calculated for multiple thresholds, as can be seen in the results section.
+To find out at which threshold there is an optimal precision-recall ratio, each confusible corrector's F-score was calculated for multiple thresholds, as can be seen in the results section.
 
@@
 As can be seen from the formulas given above, for example, they completely ignore the 'true negative' box in the confusion matrix.
-As \citet{fawcett04} shows, this can have quite extreme implications when the class distribution changes, and proposes ROC curves instead; this entails by pitting the true positive rate against the false positive rate.
+As \citet{fawcett04} shows, this can have quite extreme implications when the class distribution changes, and proposes ROC curves instead; this entails pitting the true positive rate against the false positive rate.
 However, for spelling correction this measure is not useful: because the amount of negatives (correct words) usually is many times larger than the amount of positives (incorrect words), this always leads to false positive rates that are near-zero.
-This in turns causes the Area Under the Curve is almost completely determined by the recall, which is of course only part of what we're interested in.
+This in turn causes the Area Under the Curve to be almost completely determined by the recall, which is of course only part of what we're interested in.
 
@@
 
-In this section I present the F-scores of all confusible correctors, calculated for 21 different thresholds.
+In this section I present the F-scores of the 6 confusible correctors, calculated for 21 different thresholds.
 Because all examples were assigned to either the training set and the test set randomly (as was explained in the Methods section, 90\% of the data were used to train the correctors, while 10\% was used to calculated these measures), and also because the errors added to the test set were added randomly, these results are not exactly reproducable: a confusible corrector trained on other examples and tested on other errors might of course have other results.
-Nevertheless, a quick test showed that the F-scores are relatively stable: for all correctors, when the procedure was done again, the F-scores were less than 0.01 higher or lower, which suggest that the results presented here are good reflections of the correctors' actual performance.\\\indent
+Nevertheless, a quick test showed that the F-scores are relatively stable: for all correctors, when the procedure was done again, the F-scores were less than 0.01 higher or lower, which suggests that the results presented here are good reflections of the correctors' actual performance.\\\indent
 Prediction accuracy for multiple thresholds and confusion matrices were calculated as well. These can be found in the appendix.
@@
 
-\item The corrector trained on the balanced data set (the blue line in the graph) always performs similar or worse than the unbalanced set.
+\item The corrector trained on the balanced data set (the blue line in the graphs) always performs similar or worse than the unbalanced set.
 Only for the \emph{lose} versus \emph{loose} corrector we see that the balanced set provides the best results.
@@
 I didn't only change what was there to work with English, I also had to solve several problems related to processing the English language, extend the software on some places (for example, with the confidence slider), and did a little study on the confusible correctors' training data.
-While doing this, I had to get familiar with new concepts (SSH, GitHub, MBL) on an almost daily basis.\\\indent
+While doing this, I also had to get familiar with new concepts (SSH, GitHub, MBL) on an almost daily basis.\\\indent
 Perhaps suprisingly, the fact that I have created something which could (hopefully) be useful to hundreds, if not thousands or millions of people around the globe, and the fact that I have gathered lots of new and super useful knowledge, are not the most valuable things I got out of this internship.
@@
 \hline
-balanced&Y&0&0\\
-&N&0&0\\
+balanced&Y&33&44\\
+&N&8&324\\
 \hline
 \hline
-unbalanced&Y&0&0\\
-&N&0&0\\
+unbalanced&Y&123&44\\
+&N&2&1079\\
 \hline
@@
 \includegraphics[width=0.8\textwidth]{accuracy_adviceadvise.png}
-\caption{The accuracies of corrector for the \emph{advice} versus \emph{advise}, trained on both a balanced and an inbalanced data set.}
+\caption{The prediction accuracies of corrector for the \emph{advice} versus \emph{advise}, trained on both a balanced and an inbalanced data set.}
 \end{figure}
@@
 \hline
-balanced&Y&0&0\\
-&N&0&0\\
+balanced&Y&55&54\\
+&N&7&502\\
 \hline
 \hline
-unbalanced&Y&0&0\\
-&N&0&0\\
+unbalanced&Y&85&75\\
+&N&11&786\\
 \hline
@@
 \includegraphics[width=0.8\textwidth]{accuracy_loseloose.png}
-\caption{The accuracies of corrector for the \emph{lose} versus \emph{loose}, trained on both a balanced and an inbalanced data set.}
+\caption{The prediction accuracies of corrector for the \emph{lose} versus \emph{loose}, trained on both a balanced and an inbalanced data set.}
 \end{figure}
@@
 \hline
-balanced&Y&0&0\\
-&N&0&0\\
+balanced&Y&22&35\\
+&N&4&202\\
 \hline
 \hline
-unbalanced&Y&0&0\\
-&N&0&0\\
+unbalanced&Y&177&50\\
+&N&7&1606\\
 \hline
@@
 \includegraphics[width=0.8\textwidth]{accuracy_practicepractise.png}
-\caption{The accuracies of corrector for the \emph{practice} versus \emph{practise}, trained on both a balanced and an inbalanced data set.}
+\caption{The prediction accuracies of corrector for the \emph{practice} versus \emph{practise}, trained on both a balanced and an inbalanced data set.}
 \end{figure}
@@
 \hline
-balanced&Y&0&0\\
-&N&0&0\\
+balanced&Y&2636&1821\\
+&N&210&23788\\
 \hline
 \hline
-unbalanced&Y&0&0\\
-&N&0&0\\
+unbalanced&Y&2814&1959\\
+&N&215&25301\\
 \hline
@@
 \includegraphics[width=0.8\textwidth]{accuracy_thanthen.png}
-\caption{The accuracies of corrector for the \emph{than} versus \emph{then}, trained on both a balanced and an inbalanced data set.}
+\caption{The prediction accuracies of corrector for the \emph{than} versus \emph{then}, trained on both a balanced and an inbalanced data set.}
 \end{figure}
@@
 \hline
-balanced&Y&0&0\\
-&N&0&0\\
+balanced&Y&102&77\\
+&N&12&949\\
 \hline
 \hline
-unbalanced&Y&0&0\\
-&N&0&0\\
+unbalanced&Y&404&92\\
+&N&14&3665\\
 \hline
@@
 \includegraphics[width=0.8\textwidth]{accuracy_weatherwhether.png}
-\caption{The accuracies of corrector for the \emph{weather} versus \emph{whether}, trained on both a balanced and an inbalanced data set.}
+\caption{The prediction accuracies of corrector for the \emph{weather} versus \emph{whether}, trained on both a balanced and an inbalanced data set.}
 \end{figure}
@@
 \hline
-balanced&Y&0&0\\
-&N&0&0\\
+balanced&Y&691&698\\
+&N&80&6237\\
 \hline
 \hline
-unbalanced&Y&0&0\\
-&N&0&0\\
+unbalanced&Y&1634&1192\\
+&N&134&14718\\
 \hline
@@
 \includegraphics[width=0.8\textwidth]{accuracy_you'reyour.png}
-\caption{The accuracies of corrector for the \emph{you're} versus \emph{your}, trained on both a balanced and an inbalanced data set.}
+\caption{The prediction accuracies of corrector for the \emph{you're} versus \emph{your}, trained on both a balanced and an inbalanced data set.}
 \end{figure}
@@
 Brill, E. \& R. C. Moore (2000). An improved error model for noisy channel spelling correction. In \emph{Proceedings of the 38th Annual Meeting of the ACL}, 286-293.
-\bibitem[Daelemans, Van den Bosch \& Weijters, 1997]{dvdbw97}
+\bibitem[Daelemans, Van den Bosch \& Weijters(1997)]{dvdbw97}
 Daelemans, W., A. van den Bosch, \& A. Weijters (1997). IGTree: using trees for compression and classification in lazy learning algorithms. \emph{Artificial Intelligence Review, 11}, 407–423.
@@
 He, H. \& E. Garcia (2009). Learning from imbalanced data. \emph{IEEE Transactions on knowledge and data engineering, 21(9)}, 1263-1281.
-\bibitem[Van Hulse, Khoshgoftaar \& Napolitano, 2007]{vhkn07}
+\bibitem[Van Hulse, Khoshgoftaar \& Napolitano(2007)]{vhkn07}
 Van Hulse, J. T. M. Khoshgoftaar \& A. Napolitano (2007). Experimental Perspectives on Learning from Imbalanced Data. \emph{Proceedings of the 24 th International Conference on Machine Learning.}

### 2012-12-20 Research training portfolio.tex
RTP: fixed typos, RP: deleted comments

@@
 Google provided us with an enormous list of word frequencies on the internet in 2008.
-The LexiconChecker checks for every word in the input whether there's a word on the frequencylist that is very similar to it, but much more frequent.
+The LexiconChecker checks for every word in the input whether there's a word on the frequency list that is very similar to it, but much more frequent.
 
@@
 All modules subsequently had to be changed so they actually did return their confidence for each error.
-In other words, imlementing the slider was a long journey that made me revisit all layers of Fowlt's architecture.\\\indent
+In other words, implementing the slider was a long journey that made me revisit all layers of Fowlt's architecture.\\\indent
 The biggest problem in this process, however, is that most modules not only didn't return some confidence measure for each error, they actually didn't calculate one.
@@
 That is, when a spelling checker corrects 'i like you alot' into 'i like you allot' it might have detected the error, but replaced it with the wrong word.
-If approriate corrections is what the user is interested in, it is possible to only count as true positives the errors correctly corrected, while the incorrectly corrected errors are counted as false positives.
+If appropriate corrections is what the user is interested in, it is possible to only count as true positives the errors correctly corrected, while the incorrectly corrected errors are counted as false positives.
 From the resulting correction confusion matrix the measures correction accuracy, correction recall, correction precision and correction F-score can then be calculated again.
 \\\indent
-For this paper, the correction measures will always be identical to the detecion measures, and will therefore not be mentioned seperately.
-This is because only two-option confusibles are investigated, which means that the approriate correction is always 'the other option' once an error is detected correctly.
+For this paper, the correction measures will always be identical to the detecion measures, and will therefore not be mentioned separately.
+This is because only two-option confusibles are investigated, which means that the appropriate correction is always 'the other option' once an error is detected correctly.
 This of course does not hold for confusibles with three or more options, like \emph{to} versus \emph{too} versus \emph{two} or \emph{they're} versus \emph{there} versus \emph{their}.
@@
 To solve this, such a corpus was created artificially by adding errors to 10\% of the examples used for evaluation.
-I would like to stress, however, that this is a dangerous approach: by adding artifical errors one risks adding very unnatural errors, making the evaluation results different from real usage results, and thus less useful.
+I would like to stress, however, that this is a dangerous approach: by adding artificial errors one risks adding very unnatural errors, making the evaluation results different from real usage results, and thus less useful.
 \\\indent
@@
 For example, "your" might be used instead of "you're" much more than vice versa; \citet{sfd99} for example claim that confusible errors are largely based on frequencies.
-Research into the occurence frequencies for all confusible errors would be needed to be able to simulate errors in a more natural way.
+Research into the occurrence frequencies for all confusible errors would be needed to be able to simulate errors in a more natural way.
 
@@
 While doing this, I also had to get familiar with new concepts (SSH, GitHub, MBL) on an almost daily basis.\\\indent
-Perhaps suprisingly, the fact that I have created something which could (hopefully) be useful to hundreds, if not thousands or millions of people around the globe, and the fact that I have gathered lots of new and super useful knowledge, are not the most valuable things I got out of this internship.
+Perhaps surprisingly, the fact that I have created something which could (hopefully) be useful to hundreds, if not thousands or millions of people around the globe, and the fact that I have gathered lots of new and super useful knowledge, are not the most valuable things I got out of this internship.
 The most valuable thing is that I rediscovered language technology as something I love, and that I now know what I want to spend my future doing.
@@
 \bibitem[Provost, ?]{provost}
-Provost, F. (year unkown). \emph{Machine Learning from Imbalanced Data Sets 101.} Extended abstract.
+Provost, F. (year unknown). \emph{Machine Learning from Imbalanced Data Sets 101.} Extended abstract.
 \bibitem[Reynaert, 2008]{reynaert08}

### 2012-12-20 Research paper.tex
RTP: fixed typos, RP: deleted comments
RP: worked on indentifying skipgrams techniques and mbt with free word orders

@@
 
-% Constraint Satisfaction Inference
-
 \end{itemize}
@@
 
-% Drie vormen freq: raw frequency, context frequency, multilingual alignmetn
-
 \subsection{The builder: machine translation}
@@
 When using large corpora, the process of discovering simple n-grams might already take quite a lot of time; adding variability in lenght and one or multiple gaps to the n-grams makes the process even more - in fact, many times more - complex.
-However, as I write this, this problem has been solved by using two techniques: (1) by optimizing how the various tokens and the n-grams and skipgrams they are part of are saved in memory, and (2) by using an iterative counting algorithm to limit storage.
-Besides identifying the constructions, various frequency statistics can be given.
-These frequencies can then be used to create graphs which display the relations between the various constructions.
+To techniques are used to reduce the processing load:
 
-%Hoe is dit probleem precies opgelost?
+\begin{itemize}
+\item Optimization of how the various the n-grams and skipgrams are saved in memory. By reserving the labels that don't take a lot of memory for the most frequent grams, a lot of memory can be saved.
+\item For trigrams and larger, taking into account the frequencies of the parts (in case of a trigram, two two-grams).
+If these parts do not exist or are very infrequent, there is no use in looking for the whole - if even the parts don't exist, the whole won't exist for sure.
+\end{itemize}
 
@@
 
-\subsubsection{It is unclear how well the memory-based approach works for non-European languages}
+\subsubsection{It is unclear how well the memory-based approach works for languages with a more free word order}
 
@@
 This works well for languages in which word order is largely fixed, like English: for these languages, if you change the position of a particular word in the sentence (and thus the context), you also change the meaning of that sentence.
-There are languages, however, with relatively free word order; so a word can have roughly the same meaning on various places in the sentence, and thus with various contexts.
-In other words, in these languages the main source of the knowledge of memory-based learning isn't fully available.
-
-% Uitzoeken
+There are languages, however, with relatively free word order; Hungarian, Russian and Finish, to name a few.
+In these languages, changing the position and thus the context of a particular word or phrase doesn't (really) change the meaning of the sentence.
+In other words, in these languages the main source of the knowledge of memory-based learning isn't fully available, which may very well mean that memory-based translation may not be useful for these languages.
+To what extent this really is the case is completely unclear.
+\\\indent
 

### 2012-12-21 Research paper.tex
RP: revisions based on interview

@@
 \\\indent
-The memory-based approach is in many ways similar to the statistical approach - it too generates translations on the basis of large bilingual corpora, for instance.
+The memory-based approach is in many ways similar to the statistical approach - it could even easily be argued to be a part of it.
+For example, the memory based approach also generates translations on the basis of large bilingual corpora, and also works with a decoder to mold the various translation pieces found into a sentence.
 The main difference, however, is that it is context-sensitive; the idea is that the system will find another translation for \emph{left} in the sentence \emph{The toilets are at your left hand} than for \emph{left} in the sentence \emph{Elvis has left the building} simply because they occur in different contexts.
@@
 \begin{itemize}
-\item \emph{Training.} Training is done only once and \emph{before} actual translation, and could therefore be seen as 'preparing' the translation software for its task, the classification is done every time the user asks for a translation.
-Training consists of (1) aligning the bilingual corpus, using the word alignment algorithms described in the previous sections, (2) decomposing the complete corpus into trigrams, and (3) giving the results to the classifier as training material.
+\item \emph{Training.} Training is done only once and \emph{before} actual translation, and could therefore be seen as 'preparing' the translation software for its task.
+It consists of (1) aligning the bilingual corpus, using the word alignment algorithms described in the previous sections, (2) decomposing the complete corpus into trigrams, and (3) creating a language model on the basis of this.
 The trigrams of the source language constitute the examples, the trigrams of the target language are the classes.
@@
 
+Colibri is an attempt to improve the quality of machine translation.
+This is done by building a translation system using state-of-the-art techniques, then implementing new concepts, and investigating whether they improve translation quality or not.
+Two concepts investigated so far are the use of constructions instead of phrases as translation units, and the use of context while translation these units.
+In section \ref{constructions}, I will describe what constructions are and why they might improve translation, in section \ref{builder} I describe the actual software that uses these constructions \emph{and} their context.
+
 In this section, I will describe the main idea behind the project (subsection 1), and how this is operationalized (section 2).
@@
 
-\subsection{The building blocks: constructions}
+\subsection{The building blocks: constructions} \label{constructions}
 
-Colibri, 'Constructions as Linguistics Bridges', investigates constructions (or patterns), as its name suggests.
+Colibri, 'Constructions as Linguistics Bridges', investigates constructions, as its name suggests.
 A construction can be any group of consecutive words that in some way forms an entity, in any natural language.
@@
 
-Note that a simple trigram approach could not have covered this construction.
-Importantly, constructions can also have one or more gaps (so-called 'skipgrams').
-'From \_\_ point of view' in the following example, is a construction with one such gap.
+Importantly, constructions can also have one or more gaps (so-called 'skipgrams'). 'From \_\_ point of view' in the following example, is a construction with one such gap.
 
@@
 This shows the advantage of skipgrams over n-grams for machine translation: whereas n-grams would only have been able to capture the construction 'point of view', because the word directly before 'point of view' is variable, skipgrams are also able to recognize that the word 'for' is part of the construction.
-Frequent examples of the skipgrams (I expect the variant with 'my' to be very frequent, for instance) of course \emph{would} be recognized by the n-gram approach.\\\indent
+Frequent examples of the skipgrams (I expect the variant with 'my' to be very frequent, for instance) of course \emph{could} be recognized by the n-gram approach.\\\indent
 Of course, not every possible group of consecutive words is a construction; what makes some groups special?
@@
 
-\subsection{The builder: machine translation}
+\subsection{The builder: machine translation} \label{builder}
 Besides an attempt to discover more about the possibilities for machine translation technology, Colibri is also an attempt to actually develop this technology.
@@
 Colibri's functionality can be split in four parts: (1) identifying and extracting constructions in a language, (2) identifying aligned constructions between language-pairs, (3) memory-based classification using construction experts and (4) reassembling constructions into a coherent target sentence.
-If we look at these parts in the light of the light of the memory-based machine translation process as described in the previous section, the first three parts would constitute the local phase, while part 4 would constitute the global phase.
 I will discuss the four parts in more detail consecutively:
@@
 When using large corpora, the process of discovering simple n-grams might already take quite a lot of time; adding variability in lenght and one or multiple gaps to the n-grams makes the process even more - in fact, many times more - complex.
-To techniques are used to reduce the processing load:
+To reduce the processing load, two techniques are used:
 
@@
 \item For trigrams and larger, taking into account the frequencies of the parts (in case of a trigram, two two-grams).
-If these parts do not exist or are very infrequent, there is no use in looking for the whole - if even the parts don't exist, the whole won't exist for sure.
+If these parts do not exist or are very infrequent, there is no use in looking for the whole - if even the parts don't exist, the whole won't exist for sure and doesnt need to be investigated any further.
 \end{itemize}
@@
 It was attempted to skip the first step, and to try to align the constructions directly.
-To achieve this, you can for example simply measured how often the constructions found in both languages occur together (that is, in both the source and the target language).\\\indent
+To achieve this, one can for example simply measure how often the constructions found in both languages occur together (that is, in both the source and the target language).
 Because at this stage the decoder was not yet ready (see subsection 4), the decoder from translation system MOSES was used to evaluate the results.
@@
 To fix this, a algorithm specialized for aligning skipgrams was developed, which indeed was able to find good translations.
-Unfortunately, when investigating its effects in a full translation setting, it turned out the skipgrams do not seem to improve the translation quality, because the decoder mostly picks the translation hypotheses consisting of sentences formed without skipgrams.
+So Colibri now aligns constructions by aligning the individual words first and constructions later, and uses a seperate algorithm to align skipgrams.
 
@@
 The results of this approach are compared to the results of a combined classifier approach and statistical machine translation.
+
+\subsubsection{Reassembling constructions into a coherent target sentence}
+A decoder was developed from scratch, following state-of-the-art techniques as described by \citet{koehn03}. Support of skipgrams, of course not available in standard decoders, was added.
+
+% Integreren
+\subsection{Results so far}
+Unfortunately, when investigating its effects in a full translation setting, it turned out the skipgrams do not seem to improve the translation quality, because the decoder mostly picks the translation hypotheses consisting of sentences formed without skipgrams.
+
 So far, the experiments have not yet shown an improvement in translation quality, but the researchers are hopeful in achieving this in the future.
@@
 
-\subsubsection{Reassembling constructions into a coherent target sentence}
-A decoder was developed from scratch, following state-of-the-art techniques as described by \citet{koehn03}. On top of that, support of skipgrams was added.
+\section{Main outstanding questions and future research}
+
+\section{Methodology}
+
+Because of Colibri's practical nature, its methodology forms the core of the project and has been explained in detail in the previous sections, making this section largely superfluous.
+However, this might be the right place to note that Colibri's unusual methodology also makes its time planning quite different from more conventional research projects: whereas other projects would go through stages like (1) reading the previous literature, (2) building a theoretical model, (3) testing the model, for example with experiments and (4) refining the model on the basis of the results, Colibri takes an entirely different approach.
+The first year has almost completely been devoted to implementing ideas developed previously, resulting in a fully functional machine translation system.
+With this technology in place, the years after that can be used refining (and possibly extending) this technology, as well as evaluating its results.
 
@@
 \subsection{For society}
-
-\section{Methodology}
-
-Because of Colibri's practical nature, its methodology forms the core of the project and has been explained in detail in the previous sections, making this section largely superfluous.
-However, this might be the right place to note that Colibri's unusual methodology also makes its time planning quite different from more conventional research projects: whereas other projects would go through stages like (1) reading the previous literature, (2) building a theoretical model, (3) testing the model, for example with experiments and (4) refining the model on the basis of the results, Colibri takes an entirely different approach.
-The first year has almost completely been devoted to implementing ideas developed previously, resulting in a fully functional machine translation system.
-With this technology in place, the years after that can be used refining (and possibly extending) this technology, as well as evaluating its results.
-
-\section{Main outstanding questions}
 
@@
 
-\section{Possibilities for future research}
 \section{Conclusion}

### 2012-12-21 Research paper.tex
Worked on current results section

@@
 
-\section{Project description}
+\section{Project description} \label{Description}
 
@@
 In section \ref{constructions}, I will describe what constructions are and why they might improve translation, in section \ref{builder} I describe the actual software that uses these constructions \emph{and} their context.
-
-In this section, I will describe the main idea behind the project (subsection 1), and how this is operationalized (section 2).
-Where possible, I will add information of the current status of the project.
 
@@
 
-% Integreren
-\subsection{Results so far}
-Unfortunately, when investigating its effects in a full translation setting, it turned out the skipgrams do not seem to improve the translation quality, because the decoder mostly picks the translation hypotheses consisting of sentences formed without skipgrams.
-
-So far, the experiments have not yet shown an improvement in translation quality, but the researchers are hopeful in achieving this in the future.
-They also want to add support for skipgrams in this stage.
-
-\section{Main outstanding questions and future research}
-
 \section{Methodology}
@@
 However, this might be the right place to note that Colibri's unusual methodology also makes its time planning quite different from more conventional research projects: whereas other projects would go through stages like (1) reading the previous literature, (2) building a theoretical model, (3) testing the model, for example with experiments and (4) refining the model on the basis of the results, Colibri takes an entirely different approach.
-The first year has almost completely been devoted to implementing ideas developed previously, resulting in a fully functional machine translation system.
+The first year has almost completely been devoted to implementing ideas, resulting in a fully functional machine translation system.
 With this technology in place, the years after that can be used refining (and possibly extending) this technology, as well as evaluating its results.
+
+\section{Current status and future research}
+
+At the moment, the functional machine translation system described in section \ref{builder} is mostly finished.
+It is mainly based on the open source translation system MOSES (ref), and indeed manages to produce similar results.
+However, as was explained in section \ref{Description}, Colibri is an attempt to improve the quality of machine translation by adding skipgrams and memory-based learning.
+Unfortunately it has turned out that both do not add much to the translation quality: memory-based learning only improves the results a little bit, and the skipgrams are mostly ignored by the decoder.
+Apparently, translation hypotheses with skipgrams are more likely to be good translations in most of the cases.
+\\\indent
+This means Colibri's goal, improving translation quality, has not yet been achieved.
+It is possible that larger improvements in translation quality \emph{can} be achieved by changing things like corpora or even languages used, but if this is not the case, new possibilities for improvements have to be investigated.
+One such possibility is to include a system that finds the correct translations of words on the basis of particular keywords in a sentences.
+This system should for example be able to translate the English \emph{bank} into Dutch \emph{oever} ('river bank') when the words like \emph{boat} are mentioned, but into \emph{bank} (financial institution) when \emph{money} is mentioned.
+
+%ref
 

### 2012-12-22 Research paper.tex
RP: Added societal relevance

@@
 
-% Ook lingusitic relevance: Abstracting fully lexicalised constructions
-%Finding semantic subclasses in constructions: from time-expression to time-expression"
-%Collapsing constructions with word disjunctions (\he/she/it") or part-of-speech tags
-% Correlations with experimental findings
-% Switch tasks, cloze tests, reaction times, ...
+Because Colibri's goals are not reached at the moment and there currently is no certainty about whether this will happen, Colibri's scientific relevance for now is relatively small; as described in the previous section, the already existing open source machine translation system MOSES has similar results.
+I should add, however, that the current state the project is in seems to be a fertile ground for breakthroughs in the future.
 
-\subsection{For society}
+\subsection{For society} \label{soc}
+
+Assuming such a breakthrough will indeed come, the relevance for society is a big one: by improving the quality of machine translation, one basically improves (1) to what extent someone can get information from a document written in a human language he or she does not master, and (2) to what extent people without a common language can communicate.
+As for the first point, a lot of the roughly 7000 languages in the world also have a written form.
+This means that, unless you know all of these languages, part of all information available can only be reached by translation.
+This can of course be done a translator, but that's a time-consuming and labour-intensive job, which forces us to concentrate on translating only the more important documents to the more important languages.
+With good machine translation, we can easily make available \emph{all} information available to everyone.
+The second point is perhaps even more important: if for two people there isn't a language they both speak, they have very little ways to communicate.
+In practice this means whole communities of people are excluded from the rest of the world, simply because they do not know the languages of the rest of the world.
+In highly educated societies this is of course less of a problem, as people can easily learn a new language as long as they are willing to do some effort, but foreign language education is not available in large parts of the world.
+If it manages to get good enough, machine translation might save the day here: by building 'linguistic bridges' between the world's communities, it creates the possibility for everyone on the world to communicate with everyone on the world.
+This means one is no longer limited to the community he or she was born, which may increase one's chances for a happy and succesful life seriously.
 
@@
 \subsection{Strenghts}
+
+Besides the general strength described in section \ref{soc}, Colibri has (at least) two more:
 
@@
 Obviously, the opposite is true for the rule-based approach.
-
-\subsubsection{Machines translation improves communication between the world's communities}
-
-%Bruggetje naar relevance for society
 

### 2012-12-26 Paper doen.tex
Added paper doen

\documentclass[12pt]{article}

\title{\emph{Doen}-support in Dutch}
\author{Wessel Stoop, s0808709}

\usepackage{covington}
\usepackage{graphicx}
\usepackage{natbib}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}
\maketitle

\section{Introduction}
In the Dutch main clause, there are two verb positions: the \emph{verb second position} and the \emph{verb final} position. The verb second position is used for the inflected verb, which means that it is always filled; the verb final position is only used when there is more than one verb. This is exemplified in example 1 and 2.

\begin{examples}
\item Ik eet koekjes.
\item Ik heb koekjes gegeten.
\end{examples}

In standard Dutch, no such thing as do-support is known. For example, subject–auxiliary inversion is possible with all verbs to form questions.

\begin{examples}
\item Eet ik koekjes?
\end{examples}

Interestingly, however, in some non-standard, informal varieties of Dutch, something very similar to do-support can be heard. Speakers replace the inflected verb in verb second position with an inflected form of \emph{doen} 'do', and place the meaningful verb in verb final position. The \emph{doen}-variant of example 1 would be:

\begin{examples}
\item Ik doe koekjes eten.
\end{examples}

\citet{g83} notes this construction with \emph{doen} is not popular with language users - or at least the ones aware of the construction: 'language users consider [it] unacceptable, and such sentences are mostly associated with childish language. [It is a] frequent phenomenon in spoken language, contested on the basis of non-linguistic arguments or prescriptive norms' (\citealp{g83}:58, translation WS).\\\indent
This dislike seems to be shared by some linguists, however. \citet[p. 121]{d94} for example says: 'experienced language users do not have trouble with larger formal complexity; they don't need the ease of paraphrasing. It is because of this [\emph{doen}-construction] is devious' (translation WS), and \citet[p. 153]{n62} claims the construction is mainly limited to children and women.\\\indent
Interestingly, the construction has been claimed to be part of dialectal language use several times: it would be part of the dialects of Groningen \citep{tl53}, Twente \citep{n62} and Heerlen \citep{c94}, dialects that have in common that they are all spoken in the east of the Netherlands. In this paper, however, I will show that \emph{doen}-support is not limited to dialectal language use, but can be heard in (the eastern variant of) Standard Dutch as well. Furthermore, I will try to explain what triggers its use. For this, I will use both a regression model and a model based on memory based learning.\\\indent
The remainder of this paper consists of 4 sections. In section \ref{data}, I will elaborate on the collection and the nature of the dataset used. Section \ref{regression} consists of an overview of factors that could possibly have an influence on whether periphrastic \emph{doen} is used or not, and to what extent these factors indeed can explain the data. In section \ref{mbl}, I will explain the working of Memory Based Learning software TiMBL, and discuss its results on predicting in which cases periphastic \emph{doen} is used. The paper will be concluded in section \ref{conc}.

\section{The dataset} \label{data}

\section{The regression model} \label{regression}

\subsection{Factors included}

\paragraph{Phrasal verbs.}

\paragraph{The given before new-principle.}

\paragraph{Person.}

\paragraph{The imperative mood.}

\paragraph{Habitual aspect.}

\subsection{Results}

\section{The Memory Based Learning model} \label{mbl}

\subsection{Introduction}

Like regression models, the goal of Memory Based Learning models is to predict the value of a dependent variable (in our case, whether periphrastic \emph{doen} is used or not) for new examples. The way in which MBL tries to do this precition is radically different, however. Regression models make predictions on the basis of abstractions. With every factor we include in the model, we assume another abstraction; for example, by including the factor person, we assume the existence of an abstract category person in language.\\\indent
Memory based learning doesn't do any of these assumptions. Instead of relying on human annotations for a dataset, it tries to make predictions on the basis of the data themselves. An MBL model thus is created by giving an algorithm as many examples as possible, and telling it to which class each example belongs. When given a new example, this example is compared to all classes seen in the training data. For the training examples that are most similar (that is, contain a lot of the same lexical items), it looks up the classes. The class that most of these training examples have is the class the MBL model predicts for the new example.

\subsection{Method}

\subsection{Results}

\section{Conclusion} \label{conc}

% Gloss

\begin{thebibliography}{99}
\bibitem[Duinhoven(1994)]{d94}
Duinhoven, A.M. (1994), Het hulpwerkwoord doen heeft afgedaan. Forum der Letteren 35, 2: 110-131.
\bibitem[Cornips(1994)]{c94}
Cornips, L. (1994). De hardnekkige vooroordelen over de regionale doen+infinitief-constructie. Forum der Letteren. Jaargang 1994.
\bibitem[Giesbers(1983)]{g83}
Giesbers, H. (1983-1984), Doe jij lief spelen? Notities over het perifrastisch doen. Mededelingen van de Nijmeegse Centrale voor Dialect- en Naamkunde 19, 57-64.
\bibitem[ter Laan(1953)]{tl53}
Laan, K. ter (1953), Proeve van een Groninger spraakkunst. Winschoten: van der Veen.
\bibitem[Nuijtens(1962)]{n62}
Nuijtens, E. (1962). De tweetalige mens. Assen: Van Gorcum \& Comp.
\end{thebibliography}

\end{document}

### 2012-12-29 Paper doen.tex
Doen: added glosses, improved references, small revisions, RP: prefinal revisions

@@
 
-\title{\emph{Doen}-support in Dutch}
+\title{Periphrastic \emph{doen} in Dutch}
 \author{Wessel Stoop, s0808709}
@@
 \usepackage{natbib}
+\usepackage{gb4e}
 
 \renewcommand{\familydefault}{\sfdefault}
+\let\eachwordtwo=\sf
 
@@
 
-\begin{examples}
-\item Ik eet koekjes.
-\item Ik heb koekjes gegeten.
-\end{examples}
+\begin{exe}
+\ex \gll Ik \textbf{eet} koekjes\\
+ 'I \textbf{ate} cookies'\\
+
+\ex \gll Ik \textbf{heb} koekjes \textbf{gegeten}\\
+I \textbf{have} cookies \textbf{eaten}\\
+\trans 'I have eaten cookies'
+\end{exe}
 
@@
 
-\begin{examples}
-\item Eet ik koekjes?
-\end{examples}
+\begin{exe}
+\ex \gll Eet ik koekjes? \\
+Eat I cookies? \\
+\trans 'Did I eat cookies?
+\end{exe}
 
-Interestingly, however, in some non-standard, informal varieties of Dutch, something very similar to do-support can be heard.
+Interestingly, however, in some non-standard, informal varieties of Dutch, something very similar to do-support can be found.
 Speakers replace the inflected verb in verb second position with an inflected form of \emph{doen} 'do', and place the meaningful verb in verb final position.
@@
 
-\begin{examples}
-\item Ik doe koekjes eten.
-\end{examples}
+\begin{exe}
+\ex \gll Ik doe koekjes eten \\
+I do cookies eat\\
+\trans 'I eat cookies'
+\end{exe}
 
-\citet{g83} notes this construction with \emph{doen} is not popular with language users - or at least the ones aware of the construction: 'language users consider [it] unacceptable, and such sentences are mostly associated with childish language.
-[It is a] frequent phenomenon in spoken language, contested on the basis of non-linguistic arguments or prescriptive norms' (\citealp{g83}:58, translation WS).\\\indent
+\citet{g83} notes this construction with \emph{doen}, often called \emph{periphrastic doen} is not popular with language users - or at least the ones aware of the construction: 'language users consider [it] unacceptable, and such sentences are mostly associated with childish language.
+[It is a] frequent phenomenon in spoken language, contested on the basis of non-linguistic arguments or prescriptive norms' (\citealp[p.
+58]{g83}, translation WS).\\\indent
 This dislike seems to be shared by some linguists, however.
@@
 121]{d94} for example says: 'experienced language users do not have trouble with larger formal complexity; they don't need the ease of paraphrasing.
-It is because of this [\emph{doen}-construction] is devious' (translation WS), and \citet[p.
+It is because of this [periphrastic \emph{doen}] is devious' (translation WS), and \citet[p.
 153]{n62} claims the construction is mainly limited to children and women.\\\indent
-Interestingly, the construction has been claimed to be part of dialectal language use several times: it would be part of the dialects of Groningen \citep{tl53}, Twente \citep{n62} and Heerlen \citep{c94}, dialects that have in common that they are all spoken in the east of the Netherlands.
-In this paper, however, I will show that \emph{doen}-support is not limited to dialectal language use, but can be heard in (the eastern variant of) Standard Dutch as well.
+Interestingly, the construction has also been claimed to be part of dialectal language use: it would be part of the dialects of Groningen \citep{tl53}, Twente \citep{n62} and Heerlen \citep{c94}, dialects that have in common that they are all spoken in the east of the Netherlands.\\\indent
+In this paper, I will show that \emph{doen}-support is not limited to dialectal language or the language of children or women,  but seems to be part of (the eastern variant of) Standard Dutch.
 Furthermore, I will try to explain what triggers its use.
@@
 Like regression models, the goal of Memory Based Learning models is to predict the value of a dependent variable (in our case, whether periphrastic \emph{doen} is used or not) for new examples.
-The way in which MBL tries to do this precition is radically different, however.
+The way in which MBL tries to do this prediction is radically different, however.
 Regression models make predictions on the basis of abstractions.
@@
 An MBL model thus is created by giving an algorithm as many examples as possible, and telling it to which class each example belongs.
-When given a new example, this example is compared to all classes seen in the training data.
+When given a new example, this example is compared to all examples seen in the training data.
 For the training examples that are most similar (that is, contain a lot of the same lexical items), it looks up the classes.
@@
 
-% Gloss
-
 \begin{thebibliography}{99}
 \bibitem[Duinhoven(1994)]{d94}
-Duinhoven, A.M. (1994), Het hulpwerkwoord doen heeft afgedaan. Forum der Letteren 35, 2: 110-131.
+Duinhoven, A. M. (1994), Het hulpwerkwoord doen heeft afgedaan. \emph{Forum der Letteren 35}, 2: 110-131.
 \bibitem[Cornips(1994)]{c94}
-Cornips, L. (1994). De hardnekkige vooroordelen over de regionale doen+infinitief-constructie. Forum der Letteren. Jaargang 1994.
+Cornips, L. (1994). De hardnekkige vooroordelen over de regionale doen+infinitief-constructie. \emph{Forum der Letteren. Jaargang 1994.}
 \bibitem[Giesbers(1983)]{g83}
-Giesbers, H. (1983-1984), Doe jij lief spelen? Notities over het perifrastisch doen. Mededelingen van de Nijmeegse Centrale voor Dialect- en Naamkunde 19, 57-64.
+Giesbers, H. (1983-1984), Doe jij lief spelen? Notities over het perifrastisch doen. \emph{Mededelingen van de Nijmeegse Centrale voor Dialect- en Naamkunde 19}, 57-64.
 \bibitem[ter Laan(1953)]{tl53}
-Laan, K. ter (1953), Proeve van een Groninger spraakkunst. Winschoten: van der Veen.
+Laan, K. ter (1953). \emph{Proeve van een Groninger spraakkunst}. Winschoten: van der Veen.
 \bibitem[Nuijtens(1962)]{n62}
-Nuijtens, E. (1962). De tweetalige mens. Assen: Van Gorcum \& Comp.
+Nuijtens, E. (1962). \emph{De tweetalige mens}. Assen: Van Gorcum \& Comp.
 \end{thebibliography}

### 2012-12-29 Research paper.tex
Doen: added glosses, improved references, small revisions, RP: prefinal revisions

@@
 In this paper, I will give a detailed overview of the research project Colibri, which basically is an attempt to improve the technology needed for the second task: automatically translating text from one language to another.
-More specifically, the project investigates the identification and extraction of constructions in natural language, and how these can be used in Machine Translation.
-These constructions can be of various lengths, and, importantly, can also contain one or more gaps.
-The constructions are not identified on the basis of explicit 'human' knowledge about language and grammar ('linguistic theory'), but are distilled from large amounts of text ('text corpora') with the help of context-sensitive machine-learning techniques.
+More specifically, the project investigates the effects of various translation units; for example, whether making use of constructions with one or more gaps or context improves translation quality.
+Importantly, the translation are not done on the basis of explicit 'human' knowledge about language and grammar ('linguistic theory'), but are distilled from large amounts of text ('text corpora').
 The Colibri project is carried out by Maarten van Gompel and supervised by Antal van den Bosch.
@@
 This order was chosen to show that Colibri builds on previous research directly.
-
-%Rest van de beschrijving.
+I give more information on the methodology in section 4, and describe the current and future station section 5.
+Section 6 elaborates on Colibri's relevance, both for science and society, and section 7 describes some strengths and weaknesses of the project.
+The paper is concluded in section 8.
 
@@
 \item \emph{The semantic approach} goes even one step further and also tries to grasp the meaning of a sentence, roughly.
-For a sentence like /emph{Mummy kisses daddy} this for example entails that mummy is the agent, and daddy the patient of the sentence.
+A sentence like \emph{Mummy kisses daddy} for example entails that mummy is the agent, and daddy the patient of the sentence.
 After having identified these semantic roles, the algorithm would then look how this meaning is expressed in the target language, find the approriate syntactic structure for that, and fill it with the translated words.
@@
 This means a translation from English to German would entail a translation from English to the Interlingua (using the ruleset English $\rightarrow$ Interlingua) and then a translation from the Interlingua to German (using the ruleset Interlingua $\rightarrow$ German).
-When translating from English to French, the same two steps would be taken, and the same ruleset English $\rightarrow$ Interlingua can be re-used.
+When translating from English to French, the same two steps would be taken, and the same ruleset English $\rightarrow$ Interlingua can be used.
 \end{itemize}
@@
 I should add that translation system focussing on one such approach are rare.
-As \citet{jm09} put it: 'Real systems tend to involve combinations of elements from these three architectures; thus, each is best thought of as a point in an algorithmic design space rather than as an actual algorithm.' The Systran system (\citealp{hs92,senellartea01}) is an example of this.\\\indent
+As \citet{jm09} put it: 'Real systems tend to involve combinations of elements from these three architectures; thus, each is best thought of as a point in an algorithmic design space rather than as an actual algorithm.' The Systran system (\citealp{hs92,senellartea01}) is an example of this.\\\indent %TODO example of what
 An advantage of the rule-based approach is that one has a lot of control over the output.
@@
 \begin{itemize}
-\item \emph{Corpus alignment.} Only having a bilingual corpus is not enough: for such a corpus to be useful one also needs to know which parts of the sentence in language A correspond to which part of the sentence B.
+\item \emph{Corpus alignment.} Only having a bilingual corpus is not enough: for such a corpus to be useful one also needs to know which parts of the sentence in language A correspond to which parts of the sentence in language B.
 This is not only complicated by different word orders, but also by concepts that in language A take more or less words than in language B.
-It should be added that after aligning words, phrased could be formed and aligned.
+After aligning words, it is possible to form phrases and align these on the basis of the word-alignment info.
 \citet{koehn03} show this provides better results than word-based translation.
@@
 On the basis of this knowledge, we can generate various possible translations.
+How many and which hypotheses are generated differs per approach; you could for example limit the generated hypotheses on the basis of the evaluation of previously generated hypotheses.
 
-\item \emph{Hypothesis evaluation.} Which one of these possible translation is the best one is determined on the basis of two measures: faithfulness and fluency (which can also be called grammaticality).
-These measures often are in conflict: a solution to get an optimally fluent translation would be to always return the same well-formed sentence, but that would not be faithful to the source language at all, while a very faithful translation, for example translating a sentence word by word, is usually not considered very fluent.
+\item \emph{Hypothesis evaluation.} Which one of these possible translations is the best one is determined on the basis of two measures: faithfulness and fluency (which could also be called grammaticality).
+These measures often are in conflict: a solution to get an optimally fluent translation would be to always return the same well-formed sentence, but that would not be faithful to the source sentence at all, while a very faithful translation, for example translating a sentence word by word, is usually not considered very fluent.
 Faithfulness and fluency are measured by the translation model and the language model respectively:
@@
 \end{itemize}
+
+The algorithm that performs the last two steps, hypothesis generation and evaluation, is called a \emph{decoder}.
 
@@
 This is done by building a translation system using state-of-the-art techniques, then implementing new concepts, and investigating whether they improve translation quality or not.
-Two concepts investigated so far are the use of constructions instead of phrases as translation units, and the use of context while translation these units.
+Two concepts investigated so far are the use of constructions instead of phrases as translation units, and the use of context while translating these units.
 In section \ref{constructions}, I will describe what constructions are and why they might improve translation, in section \ref{builder} I describe the actual software that uses these constructions \emph{and} their context.
@@
 
-The gap is filled here is filled with 'his', but the construction 'from \_\_ point of view' can also be filled with many other words ('my', 'the', 'another', 'Obama's', etc.).
+The gap here is filled with 'his', but the construction 'from \_\_ point of view' can also be filled with many other words ('my', 'the', 'another', 'Obama's', etc.).
 This shows the advantage of skipgrams over n-grams for machine translation: whereas n-grams would only have been able to capture the construction 'point of view', because the word directly before 'point of view' is variable, skipgrams are also able to recognize that the word 'for' is part of the construction.
@@
 \item Optimization of how the various the n-grams and skipgrams are saved in memory. By reserving the labels that don't take a lot of memory for the most frequent grams, a lot of memory can be saved.
-\item For trigrams and larger, taking into account the frequencies of the parts (in case of a trigram, two two-grams).
+\item For trigrams and larger, taking into account the frequencies of the parts.
 If these parts do not exist or are very infrequent, there is no use in looking for the whole - if even the parts don't exist, the whole won't exist for sure and doesnt need to be investigated any further.
@@
 \subsubsection{Identifying aligned constructions between language-pairs}
-The classical approach for aligning constructions from a bilingual corpus is to first align the individual words, and then align the constructions themselves \citep{koehn03}.
-It was attempted to skip the first step, and to try to align the constructions directly.
+Once the constructions in both parts of a bilingual corpus are identified, these constructions need to be linked.
+Colibri aligns constructions by aligning the individual words first and constructions later, and uses a separate algorithm to align skipgrams.\\\indent
+It should be noted that this approach is the result of various investigations.
+For example, instead of the classical approach of aligning constructions by aligning the individual words first, as described by \citep{koehn03}, it was attempted to try to align the constructions directly.
 To achieve this, one can for example simply measure how often the constructions found in both languages occur together (that is, in both the source and the target language).
-Because at this stage the decoder was not yet ready (see subsection 4), the decoder from translation system MOSES was used to evaluate the results.
+Because at this stage the decoder was not yet ready (see subsection 4), the decoder from translation system MOSES \citep{koehnea07} was used to evaluate the results.
 These results showed that the approach using constructions for alignment could not provide results as good as the approach using words; therefore, it was decided to go back to the classical method.
@@
 To fix this, a algorithm specialized for aligning skipgrams was developed, which indeed was able to find good translations.
-So Colibri now aligns constructions by aligning the individual words first and constructions later, and uses a seperate algorithm to align skipgrams.
 
@@
 \subsubsection{Reassembling constructions into a coherent target sentence}
-A decoder was developed from scratch, following state-of-the-art techniques as described by \citet{koehn03}. Support of skipgrams, of course not available in standard decoders, was added.
+A decoder was developed from scratch, following state-of-the-art techniques as described by \citet{koehn03}. Support for skipgrams, of course not available in standard decoders, was added.
 
@@
 At the moment, the functional machine translation system described in section \ref{builder} is mostly finished.
-It is mainly based on the open source translation system MOSES (ref), and indeed manages to produce similar results.
+It is mainly based on the open source translation system MOSES, and indeed manages to produce similar results.
 However, as was explained in section \ref{Description}, Colibri is an attempt to improve the quality of machine translation by adding skipgrams and memory-based learning.
 Unfortunately it has turned out that both do not add much to the translation quality: memory-based learning only improves the results a little bit, and the skipgrams are mostly ignored by the decoder.
-Apparently, translation hypotheses with skipgrams are more likely to be good translations in most of the cases.
+Apparently, translation hypotheses without skipgrams are more likely to be good translations in most of the cases.
 \\\indent
@@
 This system should for example be able to translate the English \emph{bank} into Dutch \emph{oever} ('river bank') when the words like \emph{boat} are mentioned, but into \emph{bank} (financial institution) when \emph{money} is mentioned.
-
-%ref
 
@@
 In highly educated societies this is of course less of a problem, as people can easily learn a new language as long as they are willing to do some effort, but foreign language education is not available in large parts of the world.
-If it manages to get good enough, machine translation might save the day here: by building 'linguistic bridges' between the world's communities, it creates the possibility for everyone on the world to communicate with everyone on the world.
-This means one is no longer limited to the community he or she was born, which may increase one's chances for a happy and succesful life seriously.
+If it manages to get good enough, machine translation might save the day here: by building 'linguistic bridges' between the world's communities, it creates the possibility for everyone on the globe to communicate with everyone on the globe.
+This means one is no longer limited to the community he or she was born, which may increase one's chances for a happy and succesful life significantly.
 
@@
 All this is not a problem for the Colibri project.
-Claims are made (for example 'skipgrams improve machine translation'), formalized ('this is how they should work within in a machine translation system'), but also implemented and fully tested (i.e.
-in a fully functional machine learning system).
+Claims are made (for example 'skipgrams improve machine translation'), formalized ('this is how they should work within in a machine translation system'), but also implemented and tested in a fully functional machine learning system.
 In other words, except maybe for the evaluation measures, there is little reason to doubt the results of this project.
@@
 Statistical and memory-based approaches solve this problem in an even more elegant way: the only things needed are computational power, time, and a lot of data.
-As digital translated texts are available in large quantities for the more important languages of the world, adding more languages to Colibri should not be that problematic.
+As digital translated texts are available in large quantities for the more important languages of the world, adding more languages to Colibri should be relatively easy.
 But even less important languages can be added easily, as long as there is enough data to train the classifiers (but see the first two weaknesses).
@@
 \\\indent
-Although giving a detailed analysis of these viewpoint is beyond the scope of this paper, I'd like to point out that something similar already seems to be happening for machine translation: since the release of Google Translate, various examples of websites, street signs, restaurant menus and other texts incorrectly translated by the service can be found all over the internet.
+Although giving a detailed analysis of this viewpoint is beyond the scope of this paper, I'd like to point out that something similar already seems to be happening for machine translation: since the release of Google Translate, various examples of websites, street signs, restaurant menus and other texts incorrectly translated by the service can be found on the internet.
 The authors of the text, apparently not skilled in the target language, have relied on machine translation only and didn't check the quality translation in any other way.\\\indent
@@
 \section{Conclusion}
+In this paper, I hope to have given a good impression of the research project Colibri.
+Various approaches to machine translation were described, among which memory-based learning, the approach Colibri uses.
+After that Colibri itself was discussed in more detail, from various viewpoints.
+We saw how the project is organized, and that skipgrams and context-sensitivity were investigated as possible innovations to machine translation, but also how building a practical application has importantant advantages for doing research, and what improving machine translation might mean for society.
 
@@
 \bibitem[Biemann, 2011]{biemann11}
-Biemann, C. (2011). Structure discovery in natural language. New York / Dordrecht: Springer.
+Biemann, C. (2011). \emph{Structure discovery in natural language.} New York / Dordrecht: Springer.
 
 \bibitem[Van den Bosch \& Berck, 2009]{vdbb09}
-Van den Bosch, A., \& Berck, P. (2009). Memory-based machine translation and language
-modeling. The Prague Bulletin of Mathematical Linguistics, 91, 17\-26.
+Van den Bosch, A., \& P. Berck (2009). Memory-based machine translation and language modeling. \emph{The Prague Bulletin of Mathematical Linguistics}, 91, 17\-26.
 
 \bibitem[Brown et al., 1993]{brown93}
-Brown et al., 1993
+Brown, P. F., V. J. D. Pietra, S. A. D. Pietra \& R. L. Mercer (1993). The mathematics of statistical machine translation: parameter estimation. \emph{Computational Linguistics}, 19(2), 263–311.
 
 \bibitem[Daelemans \& van den Bosch, 2005]{dvdb05}
-Daelemans \& van den Bosch, 2005
+Daelemans, W. \& A. van den Bosch (2005). \emph{Memory-based language processing}. Cambridge, UK: Cambridge University Press.
 
 \bibitem[Dempster et al., 1997]{dempster77}
-Dempster et al., 1997
+Dempster, A.
+P., N.
+M.
+Laird, \& D.
+B.
+Rubin (1977).
+Maximum likelihood from incomplete data via the em algorithm.
+\emph{Journal of the Royal Statistical Society}.
+Series B (Methodological), 39(1), 1–38.
 
 \bibitem[Galletta et al., 2005]{galletta05}
-Galletta, D. F., Durcikova, A., Everard, A. \& B. M. Jones. Does spelling checker software need warning label? Communications of the ACM, 48 (7), 82 - 86.
+Galletta, D. F., A. Durcikova, A. Everard \& B. M. Jones. Does spelling checker software need warning label? \emph{Communications of the ACM}, 48(7), 82-86.
 
 \bibitem[Goldberg, 1995]{goldberg95}
-Goldberg, A (1995). Constructions: A Construction Grammar Approach to Argument Structure. Chicago: University of Chicago Press.
+Goldberg, A (1995). \emph{Constructions: A Construction Grammar Approach to Argument Structure.} Chicago: University of Chicago Press.
 
 \bibitem[Van Gompel, 2009]{vangompel09}
-Van Gompel, M. Phrase-based memory-based machine translation. Unpublished
+Van Gompel, M. \emph{Phrase-based memory-based machine translation.} Unpublished
 master's thesis, Tilburg University.
@@
 \bibitem[Van Gompel et al., 2009]{vangompelea09}
-Van Gompel, M., Van den Bosch, A., \& Berck, P. (2009). Extending memory-based
-machine translation to phrases. In M. Forcada \& A. Way (Eds.), Proceedings of the
-third workshop on example-based machine translation (pp. 79\-86). Dublin, Ireland.
+Van Gompel, M., A.
+Van den Bosch, \& P.
+Berck (2009).
+Extending memory-based machine translation to phrases.
+In M.
+Forcada \& A.
+Way (Eds.), \emph{Proceedings of the third workshop on example-based machine translation} (pp.
+79\-86).
+Dublin, Ireland.
 
 \bibitem[Van Gompel, 2010]{vangompel10}
-Van Gompel, M. (2010). Uvt-wsd1: A cross-lingual word sense disambiguation system.
-In Semeval '10: Proceedings of the 5th international workshop on semantic evaluation
-(pp. 238\-241). Morristown, NJ, USA: Association for Computational Linguistics.
+Van Gompel, M.
+(2010).
+Uvt-wsd1: A cross-lingual word sense disambiguation system.
+In \emph{Semeval '10: Proceedings of the 5th international workshop on semantic evaluation} (pp.
+238-241).
+Morristown, NJ, USA: Association for Computational Linguistics.
 
 \bibitem[Van Gompel et al., 2011]{vangompel11}
-Van Gompel, M., Van den Bosch, A., \& Berck, P. (2011). Extending memory-based
-machine translation to phrases. In T. Markus, P. Monachesi, \& E. Westerhout (Eds.),
-Computational linguistics in the netherlands 2010: Selected papers from the twentieth
-clin meeting (pp. 45\-58). Utrecht, the Netherlands: LOT.
+Van Gompel, M., A. Van den Bosch, \& P. Berck (2011). Extending memory-based
+machine translation to phrases.
+In T.
+Markus, P.
+Monachesi, \& E.
+Westerhout (Eds.), \emph{Computational linguistics in the netherlands 2010: Selected papers from the twentieth clin meeting} (pp.
+45-58).
+Utrecht, the Netherlands: LOT.
 
@@
 Martin (2000).
-Speech and language processing: an introduction to natural language processing, computational linguistics, and speech recognition.
-Second edition.
+\emph{Speech and language processing: an introduction to natural language processing, computational linguistics, and speech recognition.} Second edition.
 New Jersey: Pearson Education.
@@
 \bibitem[Koehn et al, 2003]{koehn03}
-Koehn et al, 2003
+Koehn, P., F.-J. Och, \& D. Marcu (2003). Statistical phrase-based translation. In \emph{Proceedings of hlt-naacl 2003} (pp. 48-54). Edmonton, Canada.
+
+\bibitem[Koehn et al, 2007]{koehnea07}
+Koehn, P., H.
+Hoang, A.
+Birch, C.
+Callison-Burch, M.
+Federico, N.
+Bertoldi, et al.
+(2007).
+Moses: Open source toolkit for statistical machine translation.
+In \emph{ACL.
+The Association for Computer Linguistics.}
 
@@
 \bibitem[Stroppa et al., 2007]{stroppa07}
-Stroppa, N., Van den Bosch, A., \& Way, A. (2007). Exploiting source similarity for SMT
-using context-informed features. In A. Way \& B. Gawronska (Eds.), Proceedings of the
-11th international conference on theoretical issues in machine translation (tmi 2007)
-(pp. 231\-240). Skövde, Sweden.
+Stroppa, N., A.
+Van den Bosch, \& A.
+Way, (2007).
+Exploiting source similarity for SMT using context-informed features.
+In A.
+Way \& B.
+Gawronska (Eds.), \emph{Proceedings of the 11th international conference on theoretical issues in machine translation (tmi 2007)} (pp.
+231\-240).
+Skövde, Sweden.
 
@@
 \end{document}
-
-%\subsection{Hypotheses}
-
-%Constructions can be found efficiently in corpus data
-%Graph-based relations can be used to constrain to good constructions
-%Constructions can be aligned without resorting to word alignments as a basis
-%In MT, constructions (i.e. possibly with gaps) result in better translation than mere consecutive phrases

### 2013-01-02 Research paper.tex
Final things for first draft.

@@
 I should add that translation system focussing on one such approach are rare.
-As \citet{jm09} put it: 'Real systems tend to involve combinations of elements from these three architectures; thus, each is best thought of as a point in an algorithmic design space rather than as an actual algorithm.' The Systran system (\citealp{hs92,senellartea01}) is an example of this.\\\indent %TODO example of what
+As \citet{jm09} put it: 'Real systems tend to involve combinations of elements from these three architectures; thus, each is best thought of as a point in an algorithmic design space rather than as an actual algorithm.' The Systran system \citep{hs92,senellartea01}, for example combines the first three approaches.\\\indent
 An advantage of the rule-based approach is that one has a lot of control over the output.
@@
 
+\bibitem[Van Gompel, 2012]{vangompel12}
+Van Gompel, M. (2012). \emph{Colibri documentation}. Available online at http://proycon.github.com/colibri/doc/.
+
 \bibitem[Hutchins \& Somers, 1992]{hs92}
-Hutchins \& Somers, 1992.
+Hutchins, W. J. \& H. L. Somers (1992). \emph{An introduction to Machine Translation.} Academic Press.
 
@@
 \bibitem[Senellart et al., 2001]{senellartea01}
-Senellart et al., 2001
+Senellart, J., P. Dienes \& T. Varadi (2001). New generation SYSTRAN translation system. In emph{MT Summit 8}.
 
@@
 \end{thebibliography}
-% Documentatie Colibri
 

### 2013-01-05 Research paper.tex
Revision on the basis of Maartens comments

@@
 More specifically, the project investigates the effects of various translation units; for example, whether making use of constructions with one or more gaps or context improves translation quality.
-Importantly, the translation are not done on the basis of explicit 'human' knowledge about language and grammar ('linguistic theory'), but are distilled from large amounts of text ('text corpora').
+Importantly, the translation is not done on the basis of explicit 'human' knowledge about language and grammar ('linguistic theory'), but are distilled from large amounts of text ('text corpora').
 The Colibri project is carried out by Maarten van Gompel and supervised by Antal van den Bosch.
 \\\indent
-In section 2, I'll give a short overview of what has been achieved in machine translation, and in section 3 the current project will be described.
+In section 2, I will give a short overview of what has been achieved in machine translation, and in section 3 the current project will be described.
 This order was chosen to show that Colibri builds on previous research directly.
@@
 
-The horizontal arrows represent the approaches. I'll go through them from bottom to top:
+The horizontal arrows represent the approaches. I will go through them from bottom to top:
 
 \begin{itemize}
-\item \emph{The direct approach} does very little source language analysis, and could thus be viewed as the automatical version of looking up every word in the source language in a dictionary and replacing it with the corresponding word in the target language.
-This often doesn't yield very good results: if I for example would have translated this paper from my mother tongue Dutch to English word for word, 'would we now not real a good result have'.
+\item \emph{The direct approach} does very little source language analysis, and could thus be viewed as the automated version of looking up every word in the source language in a dictionary and replacing it with the corresponding word in the target language.
+This often does not yield very good results: if I for example would have translated this paper from my mother tongue Dutch to English word for word, 'would we now not real a good result have'.
 It should be noted, however, that the direct approach mostly is more complicated than this; for example, in some cases morphological analysis is needed before it can be looked up in a dictionary, and more morphological analysis is needed for the translation word to fit in the sentence.
 Word reordering rules can also be included.
-\item \emph{The syntactic approach} tries to find out the syntactic structure of the source sentence before it starts translating, using a parser.
+\item \emph{The syntactic approach} tries to discover the syntactic structure of the source sentence before it starts translating, using a parser.
 This structure is then translated to the target language, and filled with translated words.
@@
 \item \emph{The interlingua approach}.
-All of the approaches so far have in common that they require a seperate rule set for each language pair; no or very few rules included for translations from English to German can also be used for translations from English to French.
+All of the approaches so far have in common that they require a separate rule set for each language pair; no or very few rules included for translations from English to German can also be used for translations from English to French.
 This means that for every time a new language is added to the system, the amount of rules that have to be added grows exponentially.
@@
 
-I should add that translation system focussing on one such approach are rare.
+I should add that translation systems focussing on one approach are rare.
 As \citet{jm09} put it: 'Real systems tend to involve combinations of elements from these three architectures; thus, each is best thought of as a point in an algorithmic design space rather than as an actual algorithm.' The Systran system \citep{hs92,senellartea01}, for example combines the first three approaches.\\\indent
@@
 
-Unlike the ruled-based approach, the statistical approach creates its own 'knowledge', using bilingual text corpora.
-With this knowledge, the algorithm is able to generate various translation hypotheses, from which the best one is chosen with the help of various statistical measures.
+Unlike the rule-based approach, the statistical approach creates its own 'knowledge', using bilingual text corpora.
+This knowledge is probabilistic: instead of using fixed rules, statistical machine translation systems only know how likely a particular translation is on the basis of the data it has seen before.
+On the basis of these data, various translation hypotheses generated, from which the best one is chosen according to various statistical measures.
 I will explain these stages in more detail:
@@
 \begin{itemize}
-\item \emph{Corpus alignment.} Only having a bilingual corpus is not enough: for such a corpus to be useful one also needs to know which parts of the sentence in language A correspond to which parts of the sentence in language B.
-This is not only complicated by different word orders, but also by concepts that in language A take more or less words than in language B.
-After aligning words, it is possible to form phrases and align these on the basis of the word-alignment info.
+\item \emph{Corpus alignment.} Only having a bilingual corpus is not enough: for such a corpus to be useful one also needs to know which parts of the text in language A correspond to which parts of the text in language B.
+This goes for both the sentence level (which sentence in language A corresponds to which sentence in language B), and the word level (which word in sentence X in language A corresponds to which word in sentence X in language B).
+Aligning words and sentences is complicated by the fact that sometimes language A uses more or less sentences/words than language B to express the same concept.
+After aligning words, it is possible to extract aligned phrases on the basis of word-alignment information.
 \citet{koehn03} show this provides better results than word-based translation.
 
-\item \emph{Hypothesis generation.} We now have some sort of automatically generated dictionary: for each word in language A, we know which word in language B is often used for it.
-On the basis of this knowledge, we can generate various possible translations.
-How many and which hypotheses are generated differs per approach; you could for example limit the generated hypotheses on the basis of the evaluation of previously generated hypotheses.
+\item \emph{Hypothesis generation.} We now have some sort of automatically generated dictionary for each word or phrase in language A, we know which word or phrase in language B is often used for it, and with what probability.
+In phrase-based machine translation, this dictionary is called the \emph{phrase translation table}.
+On the basis of this dictionary or phrase translation table, we can generate various possible translations.
+\\\indent
 
@@
 \item \emph{The language model.} The language model predicts to what extent the generated sentence is expected for the target language; in other words, is this sentence a well-formed sentence or not?
-Because the unlimited number of ways words can be combined into sentences, it is unlikely the sentence under investigation will actually be found in a corpus, so the language model tries to do this prediction on the basis of how often smaller parts of the sentence (n-grams) occur in a corpus.
+Because the unlimited number of ways words can be combined into sentences, it is unlikely the sentence under investigation will actually be found in a corpus, so the language model tries to do this prediction on the basis of how often smaller parts of the sentence (n-grams) occur in a corpus in sequence.
 
@@
 The algorithm that performs the last two steps, hypothesis generation and evaluation, is called a \emph{decoder}.
+I should add that, in practice, not all possible hypotheses will be generated and evaluated, because that would take too much time and memory.
+Instead, the decoder's task could be viewed as search problem: we have an enormous pile of possible hypotheses of which one is this best, and we want to find that one hypothesis as quickly and efficiently as possible.
+To achieve this, only a small set of hypotheses is generated and evaluated immediately.
+New hypotheses are only generated on the basis of the outcomes, elimating search directions which are very unlikely to produce good results.
+In other words, in actual translation systems hypothesis generation and evaluation are mostly intertwined.
 
@@
 
-For my overview of the example-based approach, I'll focus on the memory-based version of the example-based approach in particular, because that is the version the Colibri project uses.
+For my overview of the example-based approach, I will focus on the memory-based version of the example-based approach in particular, because that is the version the Colibri project uses.
 Other versions of example-based machine translation might vary slightly.
 \\\indent
-The memory-based approach is in many ways similar to the statistical approach - it could even easily be argued to be a part of it.
+The memory-based approach is in many ways similar to the statistical approach - it could even easily be argued to be a part or an extension of it.
 For example, the memory based approach also generates translations on the basis of large bilingual corpora, and also works with a decoder to mold the various translation pieces found into a sentence.
-The main difference, however, is that it is context-sensitive; the idea is that the system will find another translation for \emph{left} in the sentence \emph{The toilets are at your left hand} than for \emph{left} in the sentence \emph{Elvis has left the building} simply because they occur in different contexts.
+The main difference, however, is that it takes the context of the source language into account; the idea is that the system will find another translation for \emph{left} in the sentence \emph{The toilets are at your left hand} than for \emph{left} in the sentence \emph{Elvis has left the building} simply because they occur in different contexts.
 \\\indent
@@
 It thus is frequency that plays a central role here.\\\indent
-While this example only has two classes, it is in principle possible to use an unlimited number of classes.
-This is exactly what the memory-based approach to machine translation does: on the basis of a particular word and its context (for example, its direct neighbours), a classifier in machine translation returns as a class the most likely n-gram.
+While this example only has two classes, it is in principle possible to use a very large number of classes.
+This is exactly what the memory-based approach to machine translation does: on the basis of a particular word and its context (for example, its direct neighbours), a classifier in machine translation returns as a class the most likely translation.
 Thus, instead of on human characteristics the classifiers are trained on words and their context from the source language, and instead of sex it uses n-grams from the target language as classes.
 \\\indent
-There are two types of classifaction methods: \emph{eager learning} and \emph{lazy learning}.
+There are two types of classification methods: \emph{eager learning} and \emph{lazy learning}.
 Whereas eager learning algorithms try to create a small, abstract, general model, filtering out infrequent cases, lazy learning algorithms take into account everything they encounter.
@@
 \item \emph{Training.} Training is done only once and \emph{before} actual translation, and could therefore be seen as 'preparing' the translation software for its task.
-It consists of (1) aligning the bilingual corpus, using the word alignment algorithms described in the previous sections, (2) decomposing the complete corpus into trigrams, and (3) creating a language model on the basis of this.
+It consists of (1) aligning the bilingual corpus, using the word alignment algorithms described in the previous sections, (2) decomposing the complete corpus into trigrams, and (3) creating a translation model on the basis of this.
 The trigrams of the source language constitute the examples, the trigrams of the target language are the classes.
 
-\item \emph{Local classification.} Classification means (1) decomposing the source text into trigrams and (2) classifying each trigram, based on the language model created in the training phase.
+\item \emph{Local classification.} Classification means (1) decomposing the source text into trigrams and (2) classifying each trigram, based on the translation model created in the training phase.
 The output are the classes, i.e.
@@
 \end{itemize}
+
+Please note again the similarities with the statistical approach: like in the \emph{hypothesis generation} phase, (parts of) hypotheses are generated during \emph{local classification}, and like the \emph{hypothesis evaluation phase}, hypotheses are evaluated during \emph{global search}.
 
@@
 The gap here is filled with 'his', but the construction 'from \_\_ point of view' can also be filled with many other words ('my', 'the', 'another', 'Obama's', etc.).
-This shows the advantage of skipgrams over n-grams for machine translation: whereas n-grams would only have been able to capture the construction 'point of view', because the word directly before 'point of view' is variable, skipgrams are also able to recognize that the word 'for' is part of the construction.
+This shows the appeal of skipgrams over n-grams for machine translation: whereas n-grams would only have been able to capture the construction 'point of view', because the word directly before 'point of view' is variable, skipgrams are also able to recognize that the word 'for' is part of the construction.
 Frequent examples of the skipgrams (I expect the variant with 'my' to be very frequent, for instance) of course \emph{could} be recognized by the n-gram approach.\\\indent
@@
 For this project, however, it suffices to say that constructions emerge because some combinations of words are more frequent than others.
-As we've seen in the previous section, taking into account frequency in the training material is how classifiers are able to do their job.
+As we have seen in the previous section, taking into account frequency in the training material is how classifiers are able to do their job.
 
@@
 \subsubsection{Identifying and extracting constructions in a language}
-When using large corpora, the process of discovering simple n-grams might already take quite a lot of time; adding variability in lenght and one or multiple gaps to the n-grams makes the process even more - in fact, many times more - complex.
-To reduce the processing load, two techniques are used:
+When using large corpora, the process of discovering simple n-grams might already take quite a lot of time and memory; adding variability in length and one or multiple gaps to the n-grams makes the process even more - in fact, many times more - complex.
+To reduce the processing load and resource consumption, two techniques are used:
 
 \begin{itemize}
-\item Optimization of how the various the n-grams and skipgrams are saved in memory. By reserving the labels that don't take a lot of memory for the most frequent grams, a lot of memory can be saved.
-\item For trigrams and larger, taking into account the frequencies of the parts.
-If these parts do not exist or are very infrequent, there is no use in looking for the whole - if even the parts don't exist, the whole won't exist for sure and doesnt need to be investigated any further.
+\item Optimization of how the various the n-grams and skipgrams are saved in memory. By reserving the labels that do not take a lot of memory for the most frequent grams, a lot of memory can be saved.
+\item For trigrams and larger n-grams, taking into account the frequencies of the parts.
+If these parts do not exist or are very infrequent, there is no use in looking for the whole - if even the parts do not exist, the whole will not exist for sure and doesnt need to be investigated any further.
 \end{itemize}
@@
 Because at this stage the decoder was not yet ready (see subsection 4), the decoder from translation system MOSES \citep{koehnea07} was used to evaluate the results.
-These results showed that the approach using constructions for alignment could not provide results as good as the approach using words; therefore, it was decided to go back to the classical method.
+These results showed that the approach using constructions directly for alignment could not provide results as good as the approach mediated through word alignments; therefore, it was decided to go back to the classical method.
 It should be noted that, because the MOSES decoder was used, the influence of skipgrams on alignment could not be measured.
@@
 When support for skipgrams was added to the implementation of the classical approach, it turned out that including skipgrams in the alignment process had negative effects on the results.
-To fix this, a algorithm specialized for aligning skipgrams was developed, which indeed was able to find good translations.
+To fix this, an algorithm specialized for aligning skipgrams was developed, which indeed was able to find good translations.
 
 \subsubsection{Memory-based classification using construction experts}
-As may have become clear from the previous sections, the Colibri project uses machine learning techniques to get from the source language construction to the target language constructions.
-More specifically, Colibri uses little expert classifiers, trained for only one construction, inspired by the effectivity of this approach in word sense ambiguation \citep{vangompel10}.
+As may have become clear from the previous sections, the Colibri project investigates the use of machine learning techniques to get from the source language construction to the target language constructions.
+More specifically, Colibri uses specialized classifiers, trained for only one construction, inspired by the effectivity of this approach in word sense ambiguation \citep{vangompel10}.
 The results of this approach are compared to the results of a combined classifier approach and statistical machine translation.
@@
 \subsubsection{Reassembling constructions into a coherent target sentence}
-A decoder was developed from scratch, following state-of-the-art techniques as described by \citet{koehn03}. Support for skipgrams, of course not available in standard decoders, was added.
+A decoder was developed from scratch, following state-of-the-art techniques as described by \citet{koehn03}.
+Support for skipgrams and machine learning, of course not available in standard decoders, was added.
 
@@
 Because of Colibri's practical nature, its methodology forms the core of the project and has been explained in detail in the previous sections, making this section largely superfluous.
-However, this might be the right place to note that Colibri's unusual methodology also makes its time planning quite different from more conventional research projects: whereas other projects would go through stages like (1) reading the previous literature, (2) building a theoretical model, (3) testing the model, for example with experiments and (4) refining the model on the basis of the results, Colibri takes an entirely different approach.
-The first year has almost completely been devoted to implementing ideas, resulting in a fully functional machine translation system.
-With this technology in place, the years after that can be used refining (and possibly extending) this technology, as well as evaluating its results.
+The only thing I would like to give some attention is that Colibri's time planning is similar to that of other research projects:
+
+\begin{enumerate}
+\item reading the previous literature
+\item building a theoretical model
+\item testing the model, for example with experiments
+\item refining the model on the basis of the results
+\end{enumerate}
+
+However, what is actually done during these stages is radically different from many other fields of science.
+For example, whereas stage 2 in many research projects will result in a paragraph of text, a table, a flowchart, a formula, or something similar, Colibri's stage 2 actually results in a fully functional translation system.
+And whereas stage 3 in many fields of science entails designing some kind of experiment that will hopefully provide evidence for the model developed during stage 2 (and then finding people to participate in that experiment), Colibri's stage 3 consists of writing a script to evaluate the results of the translation system, using various evaluation measures.
 
@@
 At the moment, the functional machine translation system described in section \ref{builder} is mostly finished.
-It is mainly based on the open source translation system MOSES, and indeed manages to produce similar results.
+It is mainly inspired upon the open source translation system MOSES, and indeed manages to produce similar results.
 However, as was explained in section \ref{Description}, Colibri is an attempt to improve the quality of machine translation by adding skipgrams and memory-based learning.
@@
 This means that, unless you know all of these languages, part of all information available can only be reached by translation.
-This can of course be done a translator, but that's a time-consuming and labour-intensive job, which forces us to concentrate on translating only the more important documents to the more important languages.
+This can of course be done a translator, but that is a time-consuming and labour-intensive job, which forces us to concentrate on translating only the more important documents to the more important languages.
 With good machine translation, we can easily make available \emph{all} information available to everyone.
-The second point is perhaps even more important: if for two people there isn't a language they both speak, they have very little ways to communicate.
+The second point is perhaps even more important: if for two people there is not a language they both speak, they have very few ways to communicate.
 In practice this means whole communities of people are excluded from the rest of the world, simply because they do not know the languages of the rest of the world.
@@
 Claims are made (for example 'skipgrams improve machine translation'), formalized ('this is how they should work within in a machine translation system'), but also implemented and tested in a fully functional machine learning system.
-In other words, except maybe for the evaluation measures, there is little reason to doubt the results of this project.
+In other words, except maybe for the evaluation measures, there is little reason to doubt the results of this project, because of its highly emperical nature.
 
@@
 When describing the rule-based approaches to machine translation, I pointed out that for most approaches one needs a specific ruleset for each language pair.
-The Interlingua approach is an attempt to solve this problem by adding a language in between, so that only a ruleset for every language-Interlingua pair is needed.
+The Interlingua approach is an attempt to solve this problem by adding a language in between, so that only a ruleset for every language $\rightarrow$ Interlingua and Interlingua $\rightarrow$ language pair is needed.
 Statistical and memory-based approaches solve this problem in an even more elegant way: the only things needed are computational power, time, and a lot of data.
 As digital translated texts are available in large quantities for the more important languages of the world, adding more languages to Colibri should be relatively easy.
-But even less important languages can be added easily, as long as there is enough data to train the classifiers (but see the first two weaknesses).
+But even less important languages can be added easily, as long as there is enough data to train the translation model (but see the first two weaknesses).
 Obviously, the opposite is true for the rule-based approach.
@@
 
-Memory-based learning takes the rule-creating process out of the hands of the authors.
+Statistical and Memory-based machine translation takes the rule-creating process out of the hands of the authors.
 This can be a good thing, because it means translation systems can be made without any knowledge of the source or the target language, but it also means the authors loses control over what the system produces.
@@
 Instead, the task of the creator now is to provide enough bilingual texts for training - preferably as much as possible - as this improves the change of finding a matching n-gram.
-And importantly, to avoid errors, this data should be of high quality.
+These data are simply not available for all languages in the world.
+And not only should there be enough bilingual texts, they should also be of high quality.
 One incorrect translation, if not found anywhere else in the corpus, can already end up in the translation results.
@@
 
-Memory-based learning generated a translation on the basis of context.
+Memory-based learning generates a translation on the basis of source-context.
 This works well for languages in which word order is largely fixed, like English: for these languages, if you change the position of a particular word in the sentence (and thus the context), you also change the meaning of that sentence.
-There are languages, however, with relatively free word order; Hungarian, Russian and Finish, to name a few.
-In these languages, changing the position and thus the context of a particular word or phrase doesn't (really) change the meaning of the sentence.
-In other words, in these languages the main source of the knowledge of memory-based learning isn't fully available, which may very well mean that memory-based translation may not be useful for these languages.
-To what extent this really is the case is completely unclear.
+There are languages, however, with relatively free word order; Hungarian, Russian and Finnish, to name a few.
+In these languages, changing the position and thus the context of a particular word or phrase does not (really) change the meaning of the sentence.
+In other words, in these languages the main source of the knowledge of memory-based learning is not fully available, which may very well mean that memory-based translation may not be useful for these languages.
+On the other hand, in many languages with free word order there is something like a canonical word order (so, although the order is free, it is not random), and almost all languages with free word order use case instead.
+Case could be really useful to memory-based learning, because having a word in a particular case in the context might be a very good clue which possible translation is the best.
+To what extent free word order is a problem or case is helpful is completely unclear.
 \\\indent
@@
 However, when non-perfect software like this is made available to the public, ignorant users might treat it as if it was perfect software, thereby turning off their own intelligence, as it were.
-A similar critisism can be heard in the field of spell checkers: the invention of automated  spell checkers would have caused a general trend in relying on the technology too much, and regarding work as finished without even proofreading it.
+A similar criticism can be heard in the field of spell checkers: the invention of automated  spell checkers would have caused a general trend in relying on the technology too much, and regarding work as finished without even proofreading it.
 \citet{galletta05} for example show that people deliver texts with more false-positive and false-negative errors in it when a spell checker was present, and that this even holds for people with good verbal skills.
 \\\indent
-Although giving a detailed analysis of this viewpoint is beyond the scope of this paper, I'd like to point out that something similar already seems to be happening for machine translation: since the release of Google Translate, various examples of websites, street signs, restaurant menus and other texts incorrectly translated by the service can be found on the internet.
-The authors of the text, apparently not skilled in the target language, have relied on machine translation only and didn't check the quality translation in any other way.\\\indent
+Although giving a detailed analysis of this viewpoint is beyond the scope of this paper, I would like to point out that something similar already seems to be happening for machine translation: since the release of Google Translate, various examples of websites, street signs, restaurant menus and other texts incorrectly translated by the service can be found on the internet.
+The authors of the text, apparently not skilled in the target language, have relied on machine translation only and did not check the quality translation in any other way.\\\indent
 And to make the problem even more complicated, there probably is no such thing as a 'correct' translation, in contrast to what the previous paragraphs may have suggested.

### 2013-01-07 Research paper.tex
RP: Fowlt corrections

@@
 A sentence like \emph{Mummy kisses daddy} for example entails that mummy is the agent, and daddy the patient of the sentence.
-After having identified these semantic roles, the algorithm would then look how this meaning is expressed in the target language, find the approriate syntactic structure for that, and fill it with the translated words.
+After having identified these semantic roles, the algorithm would then look how this meaning is expressed in the target language, find the appropriate syntactic structure for that, and fill it with the translated words.
 \item \emph{The interlingua approach}.
@@
 \item For trigrams and larger n-grams, taking into account the frequencies of the parts.
-If these parts do not exist or are very infrequent, there is no use in looking for the whole - if even the parts do not exist, the whole will not exist for sure and doesnt need to be investigated any further.
+If these parts do not exist or are very infrequent, there is no use in looking for the whole - if even the parts do not exist, the whole will not exist for sure and does not need to be investigated any further.
 \end{itemize}

### 2013-01-11 Paper doen.tex
Doen: Added text about dataset and included factors, some revisions

@@
 
-\title{Periphrastic \emph{doen} in Dutch}
+\title{Modelling periphrastic \emph{doen} in Dutch}
 \author{Wessel Stoop, s0808709}
@@
 
-\citet{g83} notes this construction with \emph{doen}, often called \emph{periphrastic doen} is not popular with language users - or at least the ones aware of the construction: 'language users consider [it] unacceptable, and such sentences are mostly associated with childish language.
+\citet{g83} notes this construction with \emph{doen}, often called \emph{periphrastic doen}, is not popular with language users - or at least the ones aware of the construction: 'language users consider [it] unacceptable, and such sentences are mostly associated with childish language.
 [It is a] frequent phenomenon in spoken language, contested on the basis of non-linguistic arguments or prescriptive norms' (\citealp[p.
@@
 Interestingly, the construction has also been claimed to be part of dialectal language use: it would be part of the dialects of Groningen \citep{tl53}, Twente \citep{n62} and Heerlen \citep{c94}, dialects that have in common that they are all spoken in the east of the Netherlands.\\\indent
-In this paper, I will show that \emph{doen}-support is not limited to dialectal language or the language of children or women,  but seems to be part of (the eastern variant of) Standard Dutch.
+In this paper, I will show that periphrastic \emph{doen} is not limited to dialectal language or the language of children or women,  but seems to be part of (the eastern variant of) Standard Dutch.
 Furthermore, I will try to explain what triggers its use.
@@
 
+%Isabel
+
 \section{The dataset} \label{data}
+
+The dataset was collected from the Corpus Gesproken Dutch ('Corpus Spoken Dutch'), henceforth CGN \citep{cgn}.
+The CGN contains different types of speech, including spontaneous face-to-face telephone conversations, interviews, debates, radio shows and read-aloud books.
+Part of the corpus was collected around Nijmegen, a city in the east of the Netherlands, making the CGN suitable for this research.\\\indent
+To obtain all examples of periphrastic \emph{doen}, the following query was used:
+
+\begin{exe}
+\ex lemma DOEN as finite verb + infinitive within 10 words
+\end{exe}
+
+This query is not perfect, both in terms of recall and precision: because of the limitation of 10 words, sentences with more material between the two verbs are missed, and this query returns a lot of examples which are not related to emphatic \emph{doen}.
+I present two examples of such false alarms:
+
+\begin{exe}
+\ex \gll wat doe jij hier eigenlijk als ik vragen mag? \\
+what do you here actually if I ask may?\\
+\trans 'What are you doing here, if I may ask?'
+
+\ex \gll en die jeugdboeken dat doe 'k {nog wel} …PAUSE… zo in een ruk uitlezen \\
+and that {children's books} that do I still ...Pause... so in one pull finish\\
+\trans 'For children's books I still do it, finishing them in one sitting'
+\end{exe}
+
+Although these examples indeed contain the lemma \emph{doen} as finite verb followed by an infinitive within ten positions, they are not examples of emphatic \emph{doen} because the finite \emph{doen} here does not replace the finite verb; instead, the two verbs are part of different clauses.\\\indent
+The resulting set of sentences was cleaned manually. A total of 66 useable examples of emphatic \emph{doen} remained. Three examples:
+
+\begin{exe} 
+\ex \gll doe jij morgen de hond uitlaten of moet ik het doen? \\ \label{hond}
+do you tomorrow the dog walk or must I it do?\\
+\trans 'Do you walk the dog tomorrow, or should I do it?'
+
+\ex \gll dan doe je twintig minuten zelf bakken \\
+then do you twenty minutes self bake\\
+\trans 'You then bake it for twenty minutes'
+
+\ex \gll dat doe ik {dan wel} kijken als ik dan toch voor Alex uh even kom kijken \\ \label{Alex}
+that do i then look when I then anyway for Alex er momentarily come look\\
+\trans 'I'll look at that when I come look for Alex'
+\end{exe}
+
+To these 66 examples, 100 randomly selected 'control' sentences were added.
+These are sentences with a finite verb but without emphatic \emph{doen}, which is almost every Dutch sentence.
+Picking a (sub)corpus to select these control sentences from turned out to be a rather difficult task.
+\\\indent
+For example, if 100 sentences selected from all Dutch corpora available would have been used, and then annotated for register, we would probably see a significant effect for register: the \emph{doen} examples are all spoken, informal language, while this would not true for all control sentences.
+We would thus find that there is no variation in formal languages.
+However, if we then wanted to use this same dataset to investigate other effects, these formal sentences might obscure the view; for example, while verb type might be a strong predictor for the use of emphatic \emph{doen} in informal language, it is not at all in formal language (as it never occurs there, regardless of verb type).
+Including formal sentences thus decreases the size of the effect.
+In other words, because the variation does not occur in formal language, it seems to make sense to exclude formal language, so we can zoom in on the place where the variation \emph{does} occur.
+\\\indent
+On the other hand, if it turned out that emphatic \emph{doen} is limited to the verb \emph{zijn} 'be', the researcher could decide to also only pick control sentences with \emph{zijn}.
+While doing this allows him or her to discover very specific effects about the variation, it is no longer possible to discover the general effect that \emph{doen} always co-occurs with \emph{zijn}, while this might be a very important predictor for emphatic \emph{doen} and part of the explanation why it is used.
+In other words, when being too picky in selecting the control sentences, one risks throwing away the baby with the bathwater.\\\indent
+Because the literature about emphatic \emph{doen} is so limited, I decided to focus on more general effects.
+There does not seem to be any consensus about possible predictors besides that it is limited to informal language, so that is the only place the control sentences were limited: all were taken from the \emph{face-to-face conversations} or \emph{telephone conversations} from the CGN, because that is also the place where all \emph{doen} sentences came from (with the exception of one example from sports commentaries).
+
+% Envelope of variation
+% Misschien iets doen met interacties?
 
@@
 
-\paragraph{Phrasal verbs.}
+\subsubsection{Phrasal verbs}
+In Dutch, several verbs consists of multiple parts.
+\emph{uitschelden} 'abuse, insult', for example, consists of \emph{uit} 'out' and \emph{schelden} 'scold'.
+It should be noted, however, that these verbs are relatively infrequent compared to normal verbs; in the frequency list based on the materials of the website \emph{http://www.opensubtitles.org}, the first phrasal verb encountered is \emph{voorstellen} 'imagine, introduce' at position 876, and others are even further down the list (for example, \emph{paardrijden} 'horse riding': 11341, \emph{opleggen} 'impose, lay on': 13577, \emph{uitschelden} 'abuse, insult': 24604).\\\indent
+When phrasal verbs are used as a finite verb in the main clause, only the verbal part appears on the second position.
+The rest of the verb appears in the sentence-final position, with the other verbs (if there are any).
 
-\paragraph{The given before new-principle.}
+\begin{exe}
+\ex \gll Ik \textbf{stel} haar aan je \textbf{voor}\\
+I introduce her to you\\
+'I'll introduce her to you'
+\end{exe}
 
-\paragraph{Person.}
+Some authors (ref) suggest emphatic \emph{doen} is a strategy to avoid the extra processing load this long-distance relationship causes, both for the speaker and the hearer.
+If that is true, we would expect to find much more phrasal verbs as the main verb in the \emph{doen} sentences than in the control sentences.
+This turns out to be the case:
 
-\paragraph{The imperative mood.}
+%ref
 
-\paragraph{Habitual aspect.}
+\begin{table}[h]
+\begin{tabular}{|l|ll|ll|}
+\hline
+&Phrasal verbs&&Non-phrasal verbs&\\
+\hline
+emphatic \emph{doen} sentences	&23 	& \emph{35\%}	&43		 &\emph{65\%}\\
+control sentences				&0		& \emph{0\%}	&100	 &\emph{100\%}\\
+\hline
+\end{tabular}
+\end{table}
+
+Because of their infrequency, phrasal verb do not even occur in the control sample, while one-third of the main verbs used with emphatic \emph{doen} is a phrasal verb.
+
+\subsubsection{The given before new-principle}
+
+Although there is a lot of disagreement over the exact details and the terminology, there is a wide consensus among linguists that languages often use the \emph{given before new}-principle (ref).
+This principle entails that speakers try to put information already known by the addressee in the beginning of the sentence, while the part of the sentence that contains the new information is put at the end as much as possible.
+Or, to put it differently, speakers first introduce their topic, and then say what they want to say about this topic.\\\indent
+%ref
+The Dutch V2-rule, however, limits the possibilities drastically.
+Unless there is an auxiliary present for other reasons, it is not possible to put the main verb at the end of the sentence in main clauses.
+Emphatic \emph{doen} could be a perfect tool to solve this problem, as it forces the main verb to go to the end of the sentence.
+If \emph{doen} is indeed used this way, we would expect that the main verb in \emph{doen} sentences indeed is new information.
+To investigate this, the conversations preceding all sentences were listened to.
+If the concept the main verb conveys was already mentioned in some form, the main verb was annotated as old information, whereas the it was annotated as new information if it was not mentioned somehow before.
+
+\begin{table}[h]
+\begin{tabular}{|l|ll|ll|}
+\hline
+&Old information&&New information&\\
+\hline
+emphatic \emph{doen} sentences	&15 	& \emph{23\%}	&51		 &\emph{77\%}\\
+control sentences				&24		& \emph{24\%}	&76		 &\emph{76\%}\\
+\hline
+\end{tabular}
+\end{table}
+
+%Vervolg
+
+\subsubsection{Habitual aspect}
+
+\citet{c94} discusses emphatic \emph{doen} as part of the Heerlen dialect, and claims it has a semantic function rather than being a syntactic strategy; emphatic \emph{doen} would be used to express habitual aspect.
+Although such a categorical explanation does not seem to work for all examples used in the current dataset, though (a habitual interpretation for sentences like \ref{hond} and \ref{Alex} would be very unlikely, for example), it might of course also play a role here.
+To investigate this, for every sentence it was annotated whether a habitual interpretation was possible.
+It should be added, that the possibility of such an interpretation does not necessarily mean the speaker also intended it.
+
+\begin{table}[h]
+\begin{tabular}{|l|ll|ll|}
+\hline
+&Possibly habitual&&Habitual impossible&\\
+\hline
+emphatic \emph{doen} sentences	&22 	& \emph{33\%}	&44		 &\emph{67\%}\\
+control sentences				&19		& \emph{19\%}	&81		 &\emph{81\%}\\
+\hline
+\end{tabular}
+\end{table}
+
+%Vervolg
+
+\subsubsection{Person}
+
+An observation done while annotating for the previously described factors is that emphatic \emph{doen} is only used in first and second person, but never in the third.
+While this can be linked to various more abstract explanations, as will be attempted in the next two sections, it was first investigated whether this really was the case, and whether this is expected on the basis of the control sentences.
+
+\begin{table}[h]
+\begin{tabular}{|l|ll|ll|ll|}
+\hline
+&First person&&Second person&&Third person&\\
+\hline
+emphatic \emph{doen} sentences	&26 	& \emph{39\%}	&40		&\emph{61\%} & 0 &\emph{0\%}\\
+control sentences				&55		& \emph{55\%}	&21	 	&\emph{21\%} & 24 &\emph{24\%}\\
+\hline
+\end{tabular}
+\end{table}
+
+For these data we not only see that the third person is absent from the \emph{doen} sentences, we also see that the second person is a lot more frequent than in the control sentences.
+This is related to the fact that the number of imperatives is exceptionally high among the \emph{doen} sentences.
+For this reason, mood was included as a factor as well.
+
+%stat
+
+\subsubsection{Mood}
+
+To investigate whether emphatic \emph{doen} is somehow related to the imperative, or any other mood, for each sentances the mood was annotated.
+
+\begin{table}[h]
+\begin{tabular}{|l|ll|ll|ll|}
+\hline
+&Indicative&&Imperative&&Interogative&\\
+\hline
+emphatic \emph{doen} sentences	&32 	& \emph{48.48\%}	&25		&\emph{37.88\%} & 9 &\emph{13.64\%}\\
+control sentences				&91		& \emph{91\%}	&0	 	&\emph{0\%} & 9 &\emph{9\%}\\
+\hline
+\end{tabular}
+\end{table}
+
+Here we see indeed that non-indicative moods are much more frequent for \emph{doen} sentences than they are for the control sentences.
+
+%stat
+
+\subsubsection{Inversion}
+
+\begin{table}[h]
+\begin{tabular}{|l|ll|ll|}
+\hline
+&Inverted&&Not inverted&\\
+\hline
+emphatic \emph{doen} sentences	&55 	& \emph{83.33\%}	&11		 &\emph{16.67\%}\\
+control sentences				&42		& \emph{42\%}	&54		 &\emph{54\%}\\
+\hline
+\end{tabular}
+\end{table}
+
+%pas op met causation
+%tekst
 
@@
 
+%perc met punt achter de komma
+
 \section{Conclusion} \label{conc}
@@
 \begin{thebibliography}{99}
+\bibitem[CGN(2006)]{cgn}
+\emph{Corpus Gesproken Nederlands, version 2.0.} Electronic resource. Available online at http://lands.let.ru.nl/cgn/home.htm.
 \bibitem[Duinhoven(1994)]{d94}

### 2013-01-15 Paper doen.tex
Doen: revised texts, added statistics

@@
 %Isabel
+%http://taaladvies.net/taal/advies/vraag/1438/
 
@@
 
-This query is not perfect, both in terms of recall and precision: because of the limitation of 10 words, sentences with more material between the two verbs are missed, and this query returns a lot of examples which are not related to emphatic \emph{doen}.
+This query is not perfect, both in terms of recall and precision: because of the limitation of 10 words, sentences with more material between the two verbs are missed, and this query returns a lot of examples which are not related to periprastic \emph{doen}.
 I present two examples of such false alarms:
@@
 
-Although these examples indeed contain the lemma \emph{doen} as finite verb followed by an infinitive within ten positions, they are not examples of emphatic \emph{doen} because the finite \emph{doen} here does not replace the finite verb; instead, the two verbs are part of different clauses.\\\indent
-The resulting set of sentences was cleaned manually. A total of 66 useable examples of emphatic \emph{doen} remained. Three examples:
+Although these examples indeed contain the lemma \emph{doen} as finite verb followed by an infinitive within ten positions, they are not examples of periprastic \emph{doen} because the finite \emph{doen} here does not replace the finite verb; instead, the two verbs are part of different clauses.\\\indent
+The resulting set of sentences was cleaned manually. A total of 66 useable examples of periprastic \emph{doen} remained. Three examples:
 
 \begin{exe} 
-\ex \gll doe jij morgen de hond uitlaten of moet ik het doen? \\ \label{hond}
-do you tomorrow the dog walk or must I it do?\\
+\ex \gll doe jij morgen de hond uitlaten of moet ik het doen? \\ 
+do you tomorrow the dog walk or must I it do?\\ \label{hond}
 \trans 'Do you walk the dog tomorrow, or should I do it?'
@@
 
-\ex \gll dat doe ik {dan wel} kijken als ik dan toch voor Alex uh even kom kijken \\ \label{Alex}
-that do i then look when I then anyway for Alex er momentarily come look\\
+\ex \gll dat doe ik {dan wel} kijken als ik dan toch voor Alex uh even kom kijken \\
+that do i then look when I then anyway for Alex er momentarily come look\\ \label{Alex}
 \trans 'I'll look at that when I come look for Alex'
@@
 To these 66 examples, 100 randomly selected 'control' sentences were added.
-These are sentences with a finite verb but without emphatic \emph{doen}, which is almost every Dutch sentence.
+These are sentences with a finite verb but without periprastic \emph{doen}, which is almost every Dutch sentence.
 Picking a (sub)corpus to select these control sentences from turned out to be a rather difficult task.
@@
 We would thus find that there is no variation in formal languages.
-However, if we then wanted to use this same dataset to investigate other effects, these formal sentences might obscure the view; for example, while verb type might be a strong predictor for the use of emphatic \emph{doen} in informal language, it is not at all in formal language (as it never occurs there, regardless of verb type).
+However, if we then wanted to use this same dataset to investigate other effects, these formal sentences might obscure the view; for example, while verb type might be a strong predictor for the use of periprastic \emph{doen} in informal language, it is not at all in formal language (as it never occurs there, regardless of verb type).
 Including formal sentences thus decreases the size of the effect.
@@
 \\\indent
-On the other hand, if it turned out that emphatic \emph{doen} is limited to the verb \emph{zijn} 'be', the researcher could decide to also only pick control sentences with \emph{zijn}.
-While doing this allows him or her to discover very specific effects about the variation, it is no longer possible to discover the general effect that \emph{doen} always co-occurs with \emph{zijn}, while this might be a very important predictor for emphatic \emph{doen} and part of the explanation why it is used.
+On the other hand, if it turned out that periprastic \emph{doen} is limited to the verb \emph{zijn} 'be', the researcher could decide to also only pick control sentences with \emph{zijn}.
+While doing this allows him or her to discover very specific effects about the variation, it is no longer possible to discover the general effect that \emph{doen} always co-occurs with \emph{zijn}, while this might be a very important predictor for periprastic \emph{doen} and part of the explanation why it is used.
 In other words, when being too picky in selecting the control sentences, one risks throwing away the baby with the bathwater.\\\indent
-Because the literature about emphatic \emph{doen} is so limited, I decided to focus on more general effects.
+Because the literature about periprastic \emph{doen} is so limited, I decided to focus on more general effects.
 There does not seem to be any consensus about possible predictors besides that it is limited to informal language, so that is the only place the control sentences were limited: all were taken from the \emph{face-to-face conversations} or \emph{telephone conversations} from the CGN, because that is also the place where all \emph{doen} sentences came from (with the exception of one example from sports commentaries).
@@
 \subsection{Factors included}
+
+In this section, I will elaborate on the factors included in the regression model, and use chi-square tests to see whether the concept investigated behaves differently for the \emph{doen} sentences.
+In section \ref{regressionresults} we will the predictive power of all these factors combined.
 
@@
 
-Some authors (ref) suggest emphatic \emph{doen} is a strategy to avoid the extra processing load this long-distance relationship causes, both for the speaker and the hearer.
+Some authors (ref) suggest periprastic \emph{doen} is a strategy to avoid the extra processing load this long-distance relationship causes, both for the speaker and the hearer.
 If that is true, we would expect to find much more phrasal verbs as the main verb in the \emph{doen} sentences than in the control sentences.
@@
 \hline
-emphatic \emph{doen} sentences	&23 	& \emph{35\%}	&43		 &\emph{65\%}\\
+periprastic \emph{doen} sentences	&23 	& \emph{34.85\%}	&43		 &\emph{65.15\%}\\
 control sentences					&0		& \emph{0\%}		&100	 &\emph{100\%}\\
@@
 
-Because of their infrequency, phrasal verb do not even occur in the control sample, while one-third of the main verbs used with emphatic \emph{doen} is a phrasal verb.
+Because of their infrequency, phrasal verb do not even occur in the control sample, while one-third of the main verbs used with periprastic \emph{doen} is a phrasal verb.
+A chi-square test showed this difference is indeed significant ($\chi^2(1, N=166) = 40.45, p > .001$).
 
@@
 Unless there is an auxiliary present for other reasons, it is not possible to put the main verb at the end of the sentence in main clauses.
-Emphatic \emph{doen} could be a perfect tool to solve this problem, as it forces the main verb to go to the end of the sentence.
+Periprastic \emph{doen} could be a perfect tool to solve this problem, as it forces the main verb to go to the end of the sentence.
 If \emph{doen} is indeed used this way, we would expect that the main verb in \emph{doen} sentences indeed is new information.
@@
 \hline
-emphatic \emph{doen} sentences	&15 	& \emph{23\%}	&51		 &\emph{77\%}\\
+periprastic \emph{doen} sentences	&15 	& \emph{22.73\%}	&51		 &\emph{77.27\%}\\
 control sentences					&24		& \emph{24\%}		&76		 &\emph{76\%}\\
@@
 
+A chi-square test showed this difference was not significant ($\chi^2(1, N=166) = 0.036, p = .85$).
+
 %Vervolg
@@
 
-\citet{c94} discusses emphatic \emph{doen} as part of the Heerlen dialect, and claims it has a semantic function rather than being a syntactic strategy; emphatic \emph{doen} would be used to express habitual aspect.
+\citet{c94} discusses periprastic \emph{doen} as part of the Heerlen dialect, and claims it has a semantic function rather than being a syntactic strategy; periprastic \emph{doen} would be used to express habitual aspect.
 Although such a categorical explanation does not seem to work for all examples used in the current dataset, though (a habitual interpretation for sentences like \ref{hond} and \ref{Alex} would be very unlikely, for example), it might of course also play a role here.
@@
 \hline
-emphatic \emph{doen} sentences	&22 	& \emph{33\%}	&44		 &\emph{67\%}\\
+periprastic \emph{doen} sentences	&22 	& \emph{33.33\%}	&44		 &\emph{66.67\%}\\
 control sentences					&19		& \emph{19\%}		&81		 &\emph{81\%}\\
@@
 
+A chi-square test showed this difference is indeed significant ($\chi^2(1, N=166) = 4.39, p = .036$).
+
 %Vervolg
@@
 
-An observation done while annotating for the previously described factors is that emphatic \emph{doen} is only used in first and second person, but never in the third.
+An observation done while annotating for the previously described factors is that periprastic \emph{doen} is used in first and second person, but never in the third.
 While this can be linked to various more abstract explanations, as will be attempted in the next two sections, it was first investigated whether this really was the case, and whether this is expected on the basis of the control sentences.
@@
 \hline
-&First person&&Second person&&Third person&\\
+&First&&Second&&Third&\\
 \hline
-emphatic \emph{doen} sentences	&26 	& \emph{39\%}	&40		&\emph{61\%} & 0 &\emph{0\%}\\
+periprastic \emph{doen} sentences	&26 	& \emph{39.39\%}	&40		&\emph{60.61\%} 	& 0 &\emph{0\%}\\
 control sentences					&55		& \emph{55\%}		&21	 	&\emph{21\%} 		& 24 &\emph{24\%}\\
@@
 
-For these data we not only see that the third person is absent from the \emph{doen} sentences, we also see that the second person is a lot more frequent than in the control sentences.
+Here we see indeed that the third person is absent from the \emph{doen} sentences, which suggests that periprastic \emph{doen} in third person is infrequent, and possibly even ungrammatical.
+It should be added that a counter-example to that last suggestion was found on \emph{taaladvies.net}, the language advice website discussed earlier, although it unclear to what extent this example is artificial:
+
+\begin{exe}
+\ex \gll Zij doen daar veel kaarten\\
+They do there much play.cardgames\\
+'They play cardgames a lot there'
+\end{exe}
+
+Another thing that strikes the eye is that second person is a lot more frequent in the \emph{doen} sentences than in the control sentences.
 This is related to the fact that the number of imperatives is exceptionally high among the \emph{doen} sentences.
-For this reason, mood was included as a factor as well.
-
+This will be investigated further in the next section.
+A chi-square test showed this difference is indeed significant ($\chi^2(2, N=166) = 34.8, p > .001$).
 %stat
@@
 
-To investigate whether emphatic \emph{doen} is somehow related to the imperative, or any other mood, for each sentances the mood was annotated.
+A possible interpretation of periprastic \emph{doen}'s preference for the first and second person is that \emph{doen} is somehow limited to interactional uses.
+The lack of emhatic \emph{doen} in third person could then be explained by the fact that the third person cannot be used for interaction.
+This interactional purpose also means that imperative and questions should occur more often in combination with \emph{doen}.
+To investigate whether periprastic \emph{doen} is somehow related to the imperative or the interrogative mood, for each sentence the mood was annotated.
 
@@
 \hline
-&Indicative&&Imperative&&Interogative&\\
+&Indicative&&Imperative&&Interrogative&\\
 \hline
-emphatic \emph{doen} sentences	&32 	& \emph{48.48\%}	&25		&\emph{37.88\%} & 9 &\emph{13.64\%}\\
+periprastic \emph{doen} sentences	&32 	& \emph{48.48\%}	&25		&\emph{37.88\%} & 9 &\emph{13.64\%}\\
 control sentences					&91		& \emph{91\%}		&0	 	&\emph{0\%} 	& 9 &\emph{9\%}\\
@@
 Here we see indeed that non-indicative moods are much more frequent for \emph{doen} sentences than they are for the control sentences.
-
+A chi-square test showed this difference is significant ($\chi^2(2, N=166) = 48.37, p > .001$).
 %stat
 
-\subsubsection{Inversion}
+\subsubsection{Word order}
+
+In the previous section we saw that the imperative and interrogative mood are more frequent for the \emph{doen} sentences than for the control sentences.
+Although this might be the result of periprastic \emph{doen}'s interactional character, it might also hint at a more syntactic interpretation: in Dutch, both questions and imperatives are formed by deviating from the standard SVO word order found in main clauses: questions are formed with \emph{inversion}, which means that the inflected verb is in first position and the subject in second position (VSO), imperatives are formed by dropping the subject (VO).
+
+\begin{exe}
+\ex \gll Wil je een koekje?\\
+Want you a cookie?\\
+'Do you want a cookie?'
+
+\ex \gll Eet je brood!\\
+Eat your bread!\\
+\end{exe}
+
+Possibly, periprastic \emph{doen} is used to indicate that an alternative word order is used.
 
@@
 \hline
-&Inverted&&Not inverted&\\
+&SVO&&V(S)O&\\
 \hline
-emphatic \emph{doen} sentences	&55 	& \emph{83.33\%}	&11		 &\emph{16.67\%}\\
-control sentences				&42		& \emph{42\%}	&54		 &\emph{54\%}\\
+periprastic \emph{doen} sentences	&55 	& \emph{83.33\%}	&11		 &\emph{16.67\%}\\
+control sentences				&42		& \emph{42\%}		&58		 &\emph{58\%}\\
 \hline
@@
 
+A chi-square test showed this difference is indeed significant ($\chi^2(1, N=166) = 27.97, p > .001$).
+
 %pas op met causation
-%tekst
+%mss ergens noemen dat die laatste drie niet echt verklaringen zijn. Of misschien is doen een mood-indicator? Over nadenken.
 
-\subsection{Results}
+\subsection{Results} \label{regressionresults}
 
@@
 The class that most of these training examples have is the class the MBL model predicts for the new example.
+An algorithm that predicts the class of new examples is called a \emph{classifier}.
+So to summarize, with this memory based learning model we are going to try to predict whether periphrastic \emph{doen} is used or not, only on the basis of lexical items.
 
 \subsection{Method}
+
+The classifier used is the IB1-IG in the TiMBL software package (ref).
+To make the dataset ready for TiMBL, two manipulations were done to make sure TiMBL can only use the lexical items for its predicts (and no other clues):
+
+%ref
+
+\begin{itemize}
+
+\item Of the control sentences, 34 sentences were randomly removed to create a balanced dataset.
+If we do not do this, 60.24\% of our data would be a control sentence, which would give the TiMBL an unfair advantage; by simply always predicting a control sentences, 60.24\% would be correct.
+This of course tells us nothing about the predictive power of the lexical items.
+\item All \emph{doen} sentences were rewritten into a non-\emph{doen} sentence.
+This is necessery, because TiMBL can simply use the presence of the lexical item \emph{doen} as an indicator whether this is an periprastic \emph{doen} sentence or not.
+
+\end{itemize}
+
+The resulting 132 sentences were split into a training set of 100 sentences and a test set of 32 sentences.
 

### 2013-01-22 research proposal ELPP.tex
Added ELPP research proposal

\documentclass[12pt]{article}

\title{Research proposal ELPP}
\author{Wessel Stoop, s0808709}

\usepackage{covington}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{float}

\renewcommand{\familydefault}{\sfdefault}

\let\stdsection\section
\renewcommand\section{\newpage\stdsection}

\begin{document}
\maketitle

\section{Introduction}
In a lot of cases, possibly even most cases, we do not say exactly what we want to convey. Sentence \ref{cold}, for example, can easily convey an order to close the window, and sentence \ref{hungry} could be a request for a sandwhich.

\begin{examples}
\item It's cold here. \label{cold}
\item Man, I'm hungry. \label{hungry}
\end{examples}

It is equally well possible, however, that sentence \ref{cold} is a request to leave, or that sentence \ref{hungry} is a suggestion to stop working and have a break. In pragmatic literature, where many similar examples can be found, these non-literal meanings are called the \emph{illocutionary force} \citep{a75}. \\\indent
But if we use indirect wordings like these all the time, how are we able to know what others mean? An answer many linguists (for example, \citet{c96,t08}) give to this question is the concept of \emph{common ground}. Common ground is the knowledge all conversation partners have, and know they all have. On the basis of this common ground, an addressee is able reconstruct the intention behind a particular utterance, and thus what the speaker really wanted to convey. This in turn allows the speaker to make his/her utterance shorter, or indirect (if one wants to be polite, for instance). Because of common ground, it is sometimes even possible to leave out language completely; a simple gesture is enough. This example, taken from \citet[pp. 3-5]{t08}, illustrates how powerful the common ground is:

\begin{quote}
Suppose that you and I are walking to the library, and out of the blue I point for you in the direction of some bicycles leaning against the library wall. [If] some days earlier you broke up with your boyfriend in a particularly nasty way, and we both know this mutually, and one the bicycles is his, which we also both know mutually, then [the pointing gesture might mean] "Your boyfriend's already at the library (so perhaps we should skip it)." On the other hand, if one of the bicycles is the one that we both know mutually was stolen from him recently, then the exact same pointing gesture will mean something completely different. [However, if] I am your friend from out of town and there is no way I could be familiar with your ex-boyfriend's bicycle, then you will not assume that I am indicating it for you. This is true even if, by some miracle, I do indeed know that this is his bicycle, but you do not know that I know this. In general, for smooth communication it is not enough that you and I each know separately and privately that this is his bicycle (and even that the other knows this); rather, this fact must be mutually known common ground between us. And in the case in which it is common ground between us that this is his bicycle, but not that the two of you have just broken up (even if we each know this privately), then you will probably think that I am indicating your boyfriend's bicycle as a way of encouraging our entrance into the library.
\end{quote}

According to Tomasello, if common grond is such an important part of language and communication in general, it must also be important to language acquisition. In fact, it might even be central to it: if, in the above example, the one who points at the bike would have uttered 'gavagai', the other person might be able to figure out that this could mean something like "let's not go in", solely on the basis of common ground.\\\indent
This does not necessarily mean infants and toddlers make complicated assumptions on the basis of shared experiences, as in the example above, however. Common ground can be as simple as both looking at the same thing, and both know that you are looking at this thing. This concept, common ground based on the immediate perceptual environment, is called \emph{joint attention}. So to translate the examples we started with into a language acquisition environment, a toddler hearing sentence \ref{hungry} might be able to figure out what 'hungry' means because he and his mother are both looking at a sandwhich. That is, joint attention can reduce referential uncertainty, and thus be a useful tool for language acquisition.\\\indent
Tomasello is not the first scholar to point out the importance of joint attention for language acquisition; the concept was first mentioned by \citet{b75}, and was investigated by various scholars afterwards. Much work related to the correlation between responsiveness to joint attention bids and language development has been done: there seems to be a positive correlation between joint attention abilities of 6-, 8-, 10-, 12-, and 18-month olds and later language abilities \citep{m00, c98} and infants who spent more time in joint attention activities with their mothers understood more language in the following months \citep{c98}. When approaching the matter from the opposite direction, it has been found that children with language development problems and autism often have joint attention impairments \citep{d02, m95, m86, od94, bc97}, and problems with joint attention can even be used as a predictor for autism \citep{m90}. Another popular topic is the difference between attention following and attention switching; it has been found that children who's mother had a tendency to follow the child's attention by talking about something the child was already looking at had a larger vocabulary than children who's mother tried to change the child's focus of attention \citep{c98,tt83}. \citet{ddc93} managed to replicate these findings in a controlled experiment: children who were taught the name of toys they were already looking at had a higher chance of remembering the name later than children who's attention was directed by the experimenter. \\\indent
An important focus in joint attention is the labeling of \emph{objects}: because joint attention reduces referential uncertainty, this is where its effects are expected to be the clearest. However, it is unclear whether findings concerning joint attention and objects can be generalized to language as a whole; the beginning of this introduction already indicated that language is much more complex than that, with its indirect utterances and assumptions one has to make about the intentions of the speaker. Do the advantages of joint attention for example carry over to learning verbs to refer to actions? \citet[pp. 312-313]{tk92} note:

\begin{quote}
'The ostensive context may not work in this same way in the case of action words or verbs. First of all, we know very little about how children determine that an adult is using a word to refer to an action or process in an ostensive context [...]. Assuming that the child knows that an adult is referring to an action, the central joint attentional problem is that, unlike objects, actions are not permanent. Thus, in ostensive presentations of object labels the object remains in view throughout, so that the child may alternate attention between adult and object as needed. In ostensive presentations of verbs, on the other hand, the action is often transient — which means that the child gets only one brief chance to identify the intended action-referent and connect it to the new word, with no chance of attention alternation.'
\end{quote}

The study proposed here will investigate both the exact nature of joint attention, and whether any discovered effects also hold for actions. In section \ref{rq}, the research questions will be discussed in detail, and in section \ref{exp} it will be explained how this will be tested. In section \ref{expectations}, finally, the expected results will be discussed.

\section{Research questions} \label{rq}

The main research question for this research proposal is the following:

\begin{examples}
\item To what extent does \emph{joint attention} facilitate language learning?
\end{examples}

One can try to answer this question in many different ways, looking at it from many different angles. For this project, the question is split up into two subquestions. These questions will discussed in detail in the next two sections:

\subsection{What exactly is joint attention?}
Whereas some define joint attention as 'coordinating visual attention on external objects with an adult during social interaction', or something very similar (e.g. \citet{ddc93} and \citet{dd04}), others go further and define it as the 'immediate perceptual environment'-part of the common ground (e.g. \citet{c96} and \citet{t08}). The difference between these two definitions, although the writers may not have been fully aware of it, is that for the second defition conversation partners both have to know they are attending to the same thing, while this is not necessarily the case for the first one. From the addressee's perspective it would of course make sense to know what the speaker is attending to, because otherwise it might be hard to figure what his / her utterances relate to, but does this also hold for the speaker? Learning new words from eavesdropping on conversations does not seem that counter-intuitive, for example. And if it is indeed possible to learn new words from speaker that is not aware of what you are attending to, is this harder, and to what extent?\\\indent For clarity, joint attention in which all conversation participants are attending to the same thing AND know they are all attending to the same thing will be called \emph{strong joint attention}, whereas joint attention in which the speaker does not know where all addressees are attending to will be called \emph{weak joint attention}.

\subsection{Is joint attention also useful when learning action labels?}
As stated in the introduction, the literature on joint attention and language acquisition mostly focused on labeling objects. However, language entails much more than labeling objects, and it is questionable to what extant joint attention is also beneficial for other parts of language. In this research project, the focus will be one such other part: describing actions.

\section{The experiment} \label{exp}

This experiment tries to mimic the methodology of \citet{ddc93}, one of the few studies investigating joint attention experimentally, as closely as possible. This is true in particular for the procedure, the materials and the setting.

\subsection{Experimental design}

The experiment is set up in a way that it can provide evidence for both questions at the same time. This is possible, because both questions correspond to a factor in the design. \\\indent
For the first question (\emph{What exactly is joint attention?}), the factor \emph{type of joint attention} will be used. This factor consists of three groups: the \textbf{strong joint attention} group, which will be taught words while the speaker is attending to something together with the participant, the \textbf{weak joint attention} group, which will be listening to a conversation without taking part in it, and the \textbf{control} group, which will be learning words from a television screen showing pictures, without communication. The relevant characteristics of these groups can be summarized in table \ref{jagroups}.

\begin{table}[h] 
\begin{tabular}{|l||l|l|} 
\hline
&Strong joint attention&Weak joint attention\\
\hline
\hline
Strong joint attention group&+&+\\
\hline
Weak joint attention group&-&+\\
\hline
Control group&-&-\\
\hline
\end{tabular}
\caption{Characteristics of the three groups in the factor \emph{type of joint attention}.}
\label{jagroups}
\end{table}

The control group is necessary to be able to compare any discovered differences between the two types of joint attention to a non-joint attention situation. If a difference is found between the results of the two joint attention groups, the results of the control group indicate whether this is a large difference or not. If the results for the control group are much worse than both joint attention groups, one can conclude that both types of joint attention are beneficial, whereas one can conclude that only strong joint attention is beneficial if the results of the weak joint attention group pair with the results of the control group.\\\indent
For the second question, the factor \emph{type of concept} will be used. This factor consists of two groups: in one group, the participants will be taught words for objects, in the other group, the participants will be taught words for actions. \\\indent
This gives us six groups of participants, as is exemplified in table \ref{overview}.

\begin{table}[h] 
\begin{tabular}{|l||l|l|}
\hline
&Objects&Actions\\
\hline
\hline
Strong joint attention group&Group 1&Group 2\\
\hline
Weak joint attention group&Group 3&Group 4\\
\hline
Control group&Group 5&Group 6\\
\hline
\end{tabular}
\caption{Overview of the experimental design.}
\label{overview}
\end{table}

The participants will be randomly assigned to one of these groups. To investigate whether the differences between groups are significant, multifactorial ANOVA for independent measures can be used.

\subsection{Procedure}

The procedure will be dependent on the group the participant is in, and will therefore be described separately.\\\\\indent
\textbf{Group 1 and 2}
\begin{enumerate}
\item Participants play with their mother, to get used to the environment.
\item Participants play with the experimenter, to get used to the experimenter.
\item The experimenter introduces a new word to the participant.
\item A 3 minute break, in which the participant plays alone.
\item The participants are brought to the eye tracking device, and do the eye-tracking experiment.
\end{enumerate}

\textbf{Group 3 and 4}
\begin{enumerate}
\item Participants play with their mother, to get used to the environment.
\item Participants play with the experimenter, to get used to the experimenter.
\item Another experiment comes in, kneels so that the child can see him / her. The experimenter already present introduces a new word to the new experimenter. Both experimenters do not make contact with the participant during this stage.
\item A 3 minute break, in which the participant plays alone.
\item The participants are brought to the eye tracking device, and do the eye-tracking experiment.
\end{enumerate}

\textbf{Group 5 and 6}
\begin{enumerate}
\item The participant watches a television screen, on which an object or an action is shown. Together with the object or action, a pre-recorded sentence is played.
\item The participants are brought to the eye tracking device, and do the eye-tracking experiment.
\end{enumerate}

\subsection{Materials}
\paragraph{Word.} The actual word taught to the participants can be any pseudoword that is not too long or hard to pronounce; it is more important that the same word is used consistently for all groups. The word \emph{dodo}, as used by \citet{ddc93}, could be used, for instance.
\paragraph{Action.} For the action (group 2, 4 and 6), it is important that participant has not yet associated another label with it, making things like jumping or sliding inappropriate. A possible action could be waving the toy in circles. It is important that the action is repeated with multiple toys, so the child does not associate the actual object with the label. 
\paragraph{Toys.} For the toys, items will be used for which the children cannot have a label already. To achieve this, very basic toys (for instance, balls and dolls) will be avoided, just as toys which a particular brand name associated with them (for instance, \emph{Duplo} or \emph{Barbie}).
\paragraph{Video.} The video watched by group 5 and 6 contains either one animation-sentence pair (for group 5), or multiple ones (for group 6). The animations for group 6 are all different objects doing the same action (moving in circles). The pre-recorded sentence played while showing the animations should be something to keep the participants focussed; for example 'Look! It's a dodo!' or 'Look! He's dodoing!'.
\paragraph{Eye tracking experiment. } The eye tracking experiment consists of multiple 'scenes', each consisting of two animations, and one on the right. While these scenes are shown, a pre-recorded sentence is played. The eye tracking device registers at which of the animations the child looks. One of these scenes will contain the object / action for which the participant was taught a label. The other scenes are simple fillers, and can be anything that will keep the participant interested.

\subsection{Participants}
As participants, infants are needed who (1) are not too old, so they not know an other word for the toy or action already but also (2) are not too young, because they have to be sensitive to joint attention. The exact time at which this sensitivity develops is subject to debate (see \citet{dd04}), but there seems to be a consensus that joint attention is fully developed around the first year (regardless of when it started), as claimed by \citet{t92}, so for this research children from that age group will be chosen. \\\indent
Following \citet{ddc93}, each group will contain 15 participants. This means a total of 90 participants will be needed.

\subsection{Setting}
\begin{itemize}
\item For the joint attention task, the setting of the experiment will be a bright room with a carpet in the middle. On this carpet will be various toys, which can be used by the participant to play with, or by the experimenters to teach the participant a new word. The entire room will be video-taped.
\item For the television task of the control group, participants will be in a dark room with a television screen. Participants can sit on the lap of their mothers while watching the screen and listening to the sounds.
\item For the eye-tracking task, all participants will be brought to a dark room with an eye-tracking device. Participants can sit on the lap of their mothers while watching the screen and listening to the sounds.
\end{itemize}

\section{Expectations} \label{expectations}

The eye-tracking device registers which part of the screen the participant is paying attention to. When the pre-recorded sentence is played, and the child pays most attention to the part of the screen corresponding to the sentence, it will be assumed the child has learned the noun or the verb correctly.\\\indent
It is expected that in all groups participants will be able to associate the correct label with the object or the action, but that the amount of participants able to make the correct association varies. For example, it is expected that groups with weak joint attention will do less good, and that groups with no joint attention will perform even worse. Furtermore, it expected that all groups in which the participants were taught a label for an action will perform worse than their object counterparts. These results could be visualized as follows:

\begin{figure}[htb]
\centering
\includegraphics[width=0.8\textwidth]{graph_elpp.png}
\caption{The expected results}
\label{fig:results}
\end{figure}

If these results are found indeed, than we could conclude that joint attention has more advantages for learning nouns than for learning verbs, and that strong joint attention works best. Other likely results are that there is no difference between nouns and verbs, in which case the blue and red line in the graph will overlap, that there is no difference between weak and strong joint attention, in which case both lines will not go down from the first to the second point, or that there is difference between weak and no joint attention, in which case both lines will not go down from the second to the third point.

\section{Summary and conclusion} \label{sum}
In this research proposal, a study about joint attention was described. Joint attention will be investigated from two viewpoints. Firstly, participants will be divided into either a verb or a noun group, to investigate whether joint attention also facilitates learning words for actions. Secondly, participants will be divided into a strong joint attention group, a weak joint attention group and a control group. The strong joint attention group will be taught a word while the speaker and the participant are both attending to the same thing, the weak joint attention group will be overhearing a conversation by two experimenters, and the control group will be taught words by looking at a screen while listening to pre-recorded sentences. Expextations are that joint attention, in particular the strong form, will facilitate language learning, but that this will go best for nouns. This is expected because joint attention is so central for language learning and communication in general.

\begin{thebibliography}{99}
\bibitem[Austin(1975)]{a75}
Austin, J. L. (1975). \emph{How to do things with words.} Oxford: Oxford University Press. 
\bibitem[Baron-Cohen et al.(1997)]{bc97}
Baron-Cohen, S., D. A. Baldwin \& M. Crowson (1997). Do children with autism use the speaker’s direction of gaze strategy to crack the code of language? \emph{Child Development}, 68(1), 48 - 57.
\bibitem[Bruner(1975)]{b75} 
Bruner, J. (1975). The ontogenesis of speech acts. \emph{Journal of Child Language}, 2, 1 - 20.
\bibitem[Carpenter et al(1998)]{c98}
Carpenter, M., K. Nagell, \& M. Tomasello (1998). Social cognition, joint attention and communicative competence from 9 to 15 months of age. \emph{Monographs of the Society for Research in Child Development}, 63, 1 - 14.
\bibitem[Clark(1996)]{c96}
Clark, H. (1996). \emph{Uses of language.} Cambridge: Cambridge University Press.
\bibitem[Dawson et al.(2002)]{d02}
Dawson, G., J. Munson, A. Estes, J. Osterling, J. McPartland, K. Toth, L. Carver, \& R. Abbot, (2002). Neurocognitive function and joint attention ability in young children with autism spectrum disorder versus developmental delay. \emph{Child Development}, 73(2), 345 - 358.
\bibitem[Dominey \& Dodane (2004)]{dd04}
Dominey, P. F. \& C. Dodane (2004). \emph{Journal of Neurolinguistics}, 17, 121 - 145.
\bibitem[Dunham, Dunham \& Curwin(1993)]{ddc93}
Dunham, P. J., F. Dunham, \& A. Curwin (1993). Joint-attentional states and lexical acquisition at 18 months. \emph{Developmental Psychology}, 29(5), 827 - 831.
\bibitem[Morales et al.(2000)]{m00}
Morales, M., P. Mundy, C. E. F. Delgado, M. Yale, D. Messinger, R. Neal \& H. K. Schwartz (2000). Responding to joint attention across the 6- through 24-month age period and early language acquisition. \emph{Journal of Applied Developmental Psychology}, 21(3), 283 - 298.
\bibitem[Mundy(1995)]{m95}
Mundy, P. (1995). Joint attention and social-emotional approach behavior in children with autism. \emph{Development and Psychopathology}, 7, 63 - 82.
\bibitem[Mundy et al.(1990)]{m90}
Mundy, P., M. Sigman, \& C. Kasari (1990). A longitudinal study of joint attention and language development in autistic children. \emph{Journal of Autism and Developmental Disorders}, 20(1), 115 - 128.
\bibitem[Mundy et al.(1986)]{m86}
Mundy, P., M. Sigman, J. Ungerer, \& T. Sherman (1986). Defining the social deficits of autism: the contribution of nonverbal communication measures. \emph{Journal of Child Psychology and Psychiatry}, 27, 657 - 669.
\bibitem[Osterling \& Dawson(1994)]{od94}
Osterling, J., \& G. Dawson (1994). Early recognition of children with autism: a study of first birthday home video-tapes. \emph{Journal of Autism and Developmental Disorders}, 24, 247 - 257.
\bibitem[Tomasello(1992)]{t92}
Tomasello, M. (1992). The social bases of language acquisition. \emph{Social Development}, 1, 67 - 87.
\bibitem[Tomasello(2008)]{t08}
Tomasello, M. (2008). \emph{Origins of Human Communication.} Cambridge: MIT Press.
\bibitem[Tomasello \& Kruger(1992)]{tk92}
Tomasello, M. \& A. Kruger (1992). Joint attention on actions: acquiring verbs in ostensive and nonostensive contexts. \emph{Journal of Child Language} 19, 311 - 333.
\bibitem[Tomasello \& Todd(1983)]{tt83}
Tomasello, M., \& J. Todd (1983). Joint attention and lexical acquisition style. \emph{First Language}, 4, 197 - 212.
\end{thebibliography}

\end{document}

### 2013-01-24 Paper doen.tex
Doen: revisions, added statistics, added conclusion

@@
 
-In standard Dutch, no such thing as do-support is known. For example, subject–auxiliary inversion is possible with all verbs to form questions.
+In standard Dutch, no such thing as do-support is known. For example, subject-auxiliary inversion is possible with all verbs to form questions.
 
@@
 [It is a] frequent phenomenon in spoken language, contested on the basis of non-linguistic arguments or prescriptive norms' (\citealp[p.
-58]{g83}, translation WS).\\\indent
-This dislike seems to be shared by some linguists, however.
+58]{g83}, translation WS).
+On \emph{taaladvies.net}, a Dutch website giving language-related advice, the use of periphrastic \emph{doen} is strongly discouraged because it 'does not add anything substantial' (translation WS).\\\indent
+This dislike seems to be shared by some linguists too.
 \citet[p.
@@
 Interestingly, the construction has also been claimed to be part of dialectal language use: it would be part of the dialects of Groningen \citep{tl53}, Twente \citep{n62} and Heerlen \citep{c94}, dialects that have in common that they are all spoken in the east of the Netherlands.\\\indent
-In this paper, I will show that periphrastic \emph{doen} is not limited to dialectal language or the language of children or women,  but seems to be part of (the eastern variant of) Standard Dutch.
+In this paper, I will show that periphrastic \emph{doen} is not limited to dialectal language or the language of children or women,  but seems to be part of (the eastern variant of) Dutch in general.
 Furthermore, I will try to explain what triggers its use.
@@
 Section \ref{regression} consists of an overview of factors that could possibly have an influence on whether periphrastic \emph{doen} is used or not, and to what extent these factors indeed can explain the data.
-In section \ref{mbl}, I will explain the working of Memory Based Learning software TiMBL, and discuss its results on predicting in which cases periphastic \emph{doen} is used.
+In section \ref{mbl}, I will explain the workings of Memory Based Learning software TiMBL, and discuss its results on predicting in which cases periphrastic \emph{doen} is used.
 The paper will be concluded in section \ref{conc}.
-
-%Isabel
-%http://taaladvies.net/taal/advies/vraag/1438/
 
@@
 
-This query is not perfect, both in terms of recall and precision: because of the limitation of 10 words, sentences with more material between the two verbs are missed, and this query returns a lot of examples which are not related to periprastic \emph{doen}.
+This query is not perfect, both in terms of recall and precision: because of the limitation of 10 words, sentences with more material between the two verbs are missed, and this query returns a lot of examples which are not related to periphrastic \emph{doen}.
 I present two examples of such false alarms:
@@
 
-Although these examples indeed contain the lemma \emph{doen} as finite verb followed by an infinitive within ten positions, they are not examples of periprastic \emph{doen} because the finite \emph{doen} here does not replace the finite verb; instead, the two verbs are part of different clauses.\\\indent
-The resulting set of sentences was cleaned manually. A total of 66 useable examples of periprastic \emph{doen} remained. Three examples:
+Although these examples indeed contain the lemma \emph{doen} as finite verb followed by an infinitive within ten positions, they are not examples of periphrastic \emph{doen} because the finite \emph{doen} here does not replace the finite verb; instead, the two verbs are part of different clauses.\\\indent
+The resulting set of sentences was cleaned manually. A total of 66 usable examples of periphrastic \emph{doen} remained. Three examples:
 
@@
 \trans 'Do you walk the dog tomorrow, or should I do it?'
-
-\ex \gll dan doe je twintig minuten zelf bakken \\
-then do you twenty minutes self bake\\
-\trans 'You then bake it for twenty minutes'
 
@@
 To these 66 examples, 100 randomly selected 'control' sentences were added.
-These are sentences with a finite verb but without periprastic \emph{doen}, which is almost every Dutch sentence.
+These are sentences with a finite verb but without periphrastic \emph{doen}, which is almost every Dutch sentence.
 Picking a (sub)corpus to select these control sentences from turned out to be a rather difficult task.
@@
 For example, if 100 sentences selected from all Dutch corpora available would have been used, and then annotated for register, we would probably see a significant effect for register: the \emph{doen} examples are all spoken, informal language, while this would not true for all control sentences.
-We would thus find that there is no variation in formal languages.
-However, if we then wanted to use this same dataset to investigate other effects, these formal sentences might obscure the view; for example, while verb type might be a strong predictor for the use of periprastic \emph{doen} in informal language, it is not at all in formal language (as it never occurs there, regardless of verb type).
+We would thus find that there is no variation in formal language.
+However, if we then wanted to use this same dataset to investigate other effects, these formal sentences might obscure the view; for example, while verb type might be a strong predictor for the use of periphrastic \emph{doen} in informal language, it is not at all in formal language (as it never occurs there, regardless of verb type).
 Including formal sentences thus decreases the size of the effect.
@@
 \\\indent
-On the other hand, if it turned out that periprastic \emph{doen} is limited to the verb \emph{zijn} 'be', the researcher could decide to also only pick control sentences with \emph{zijn}.
-While doing this allows him or her to discover very specific effects about the variation, it is no longer possible to discover the general effect that \emph{doen} always co-occurs with \emph{zijn}, while this might be a very important predictor for periprastic \emph{doen} and part of the explanation why it is used.
+On the other hand, if it turned out that periphrastic \emph{doen} is limited to the verb \emph{zijn} 'be', the researcher could decide to also only pick control sentences with \emph{zijn}.
+While doing this allows him or her to discover very specific effects about the variation, it is no longer possible to discover the general effect that \emph{doen} always co-occurs with \emph{zijn}, while this might be a very important predictor for periphrastic \emph{doen} and part of the explanation why it is used.
 In other words, when being too picky in selecting the control sentences, one risks throwing away the baby with the bathwater.\\\indent
-Because the literature about periprastic \emph{doen} is so limited, I decided to focus on more general effects.
+Because the literature about periphrastic \emph{doen} is so limited, I decided to focus on more general effects.
 There does not seem to be any consensus about possible predictors besides that it is limited to informal language, so that is the only place the control sentences were limited: all were taken from the \emph{face-to-face conversations} or \emph{telephone conversations} from the CGN, because that is also the place where all \emph{doen} sentences came from (with the exception of one example from sports commentaries).
@@
 
-In this section, I will elaborate on the factors included in the regression model, and use chi-square tests to see whether the concept investigated behaves differently for the \emph{doen} sentences.
-In section \ref{regressionresults} we will the predictive power of all these factors combined.
+In this section, I will elaborate on the factors included in the regression model, and use chi-square tests to see whether the \emph{doen} sentences behave differently from the control sentences with the regard to the concept investigated.
+In section \ref{regressionresults} the predictive power of all these factors combined will be discussed.
 
@@
 In Dutch, several verbs consists of multiple parts.
-\emph{uitschelden} 'abuse, insult', for example, consists of \emph{uit} 'out' and \emph{schelden} 'scold'.
+\emph{Uitschelden} 'abuse, insult', for example, consists of \emph{uit} 'out' and \emph{schelden} 'scold'.
 It should be noted, however, that these verbs are relatively infrequent compared to normal verbs; in the frequency list based on the materials of the website \emph{http://www.opensubtitles.org}, the first phrasal verb encountered is \emph{voorstellen} 'imagine, introduce' at position 876, and others are even further down the list (for example, \emph{paardrijden} 'horse riding': 11341, \emph{opleggen} 'impose, lay on': 13577, \emph{uitschelden} 'abuse, insult': 24604).\\\indent
-When phrasal verbs are used as a finite verb in the main clause, only the verbal part appears on the second position.
-The rest of the verb appears in the sentence-final position, with the other verbs (if there are any).
+When phrasal verbs are used as a finite verb in the main clause, only the verbal part appears on the second position. The rest of the verb appears in the sentence-final position.
 
@@
 
-Some authors (ref) suggest periprastic \emph{doen} is a strategy to avoid the extra processing load this long-distance relationship causes, both for the speaker and the hearer.
+\citet[pp.
+156]{n62} suggests periphrastic \emph{doen} is a strategy to avoid the extra processing load this long-distance relationship causes, both for the speaker and the hearer.
 If that is true, we would expect to find much more phrasal verbs as the main verb in the \emph{doen} sentences than in the control sentences.
@@
 
-%ref
-
 \begin{table}[h]
-\begin{tabular}{|l|ll|ll|}
+\begin{tabular}{|l||ll|ll|}
 \hline
@@
 \hline
-periprastic \emph{doen} sentences	&23 	& \emph{34.85\%}	&43		 &\emph{65.15\%}\\
+periphrastic \emph{doen} sentences	&23 	& \emph{34.85\%}	&43		 &\emph{65.15\%}\\
 control sentences					&0		& \emph{0\%}		&100	 &\emph{100\%}\\
@@
 
-Because of their infrequency, phrasal verb do not even occur in the control sample, while one-third of the main verbs used with periprastic \emph{doen} is a phrasal verb.
+Because of their infrequency, phrasal verb do not even occur in the control sample, while one-third of the main verbs used with periphrastic \emph{doen} is a phrasal verb.
 A chi-square test showed this difference is indeed significant ($\chi^2(1, N=166) = 40.45, p > .001$).
@@
 %ref
-The Dutch V2-rule, however, limits the possibilities drastically.
+The Dutch V2-rule, however, limits the possibilities to do this drastically.
 Unless there is an auxiliary present for other reasons, it is not possible to put the main verb at the end of the sentence in main clauses.
-Periprastic \emph{doen} could be a perfect tool to solve this problem, as it forces the main verb to go to the end of the sentence.
+periphrastic \emph{doen} could be a perfect tool to solve this problem, as it forces the main verb to go to the end of the sentence.
 If \emph{doen} is indeed used this way, we would expect that the main verb in \emph{doen} sentences indeed is new information.
 To investigate this, the conversations preceding all sentences were listened to.
-If the concept the main verb conveys was already mentioned in some form, the main verb was annotated as old information, whereas the it was annotated as new information if it was not mentioned somehow before.
+If the concept the main verb conveys was already mentioned in some form, the main verb was annotated as old information, whereas it was annotated as new information if it was not mentioned somehow before.
 
 \begin{table}[h]
-\begin{tabular}{|l|ll|ll|}
+\begin{tabular}{|l||ll|ll|}
 \hline
@@
 \hline
-periprastic \emph{doen} sentences	&15 	& \emph{22.73\%}	&51		 &\emph{77.27\%}\\
+periphrastic \emph{doen} sentences	&15 	& \emph{22.73\%}	&51		 &\emph{77.27\%}\\
 control sentences					&24		& \emph{24\%}		&76		 &\emph{76\%}\\
@@
 
+As can be seen from the table, there does not seem to be a clear difference between the \emph{doen} sentences and the control sentences; in both cases the main verb was new information most of the times, regardless of whether \emph{doen} was used.
 A chi-square test showed this difference was not significant ($\chi^2(1, N=166) = 0.036, p = .85$).
-
-%Vervolg
 
@@
 
-\citet{c94} discusses periprastic \emph{doen} as part of the Heerlen dialect, and claims it has a semantic function rather than being a syntactic strategy; periprastic \emph{doen} would be used to express habitual aspect.
-Although such a categorical explanation does not seem to work for all examples used in the current dataset, though (a habitual interpretation for sentences like \ref{hond} and \ref{Alex} would be very unlikely, for example), it might of course also play a role here.
+\citet{c94} discusses periphrastic \emph{doen} as part of the Heerlen dialect, and claims it has a semantic function rather than being a syntactic strategy; periphrastic \emph{doen} would be used to express habitual aspect.
+Although such a categorical explanation does not seem to work for all examples used in the current dataset, (a habitual interpretation for sentences like \ref{hond} and \ref{Alex} would be very unlikely, for example), it might also play a role here.
 To investigate this, for every sentence it was annotated whether a habitual interpretation was possible.
-It should be added, that the possibility of such an interpretation does not necessarily mean the speaker also intended it.
+It should be added that the possibility of such an interpretation does not necessarily mean the speaker also intended it.
 
 \begin{table}[h] 
-\begin{tabular}{|l|ll|ll|}
+\begin{tabular}{|l||ll|ll|} 
 \hline
@@
 \hline
-periprastic \emph{doen} sentences	&22 	& \emph{33.33\%}	&44		 &\emph{66.67\%}\\
+periphrastic \emph{doen} sentences	&22 	& \emph{33.33\%}	&44		 &\emph{66.67\%}\\
 control sentences					&19		& \emph{19\%}		&81		 &\emph{81\%}\\
@@
 
-A chi-square test showed this difference is indeed significant ($\chi^2(1, N=166) = 4.39, p = .036$).
+The data show a small difference between periphrastic \emph{doen} and the control sentences; a possible habitual doen interpretation is a little more likely when periphrastic \emph{doen} is present.
+A chi-square test showed this difference was significant ($\chi^2(1, N=166) = 4.39, p = .036$).
 
-%Vervolg
+\subsubsection{Variety}
+
+As discussed in the introduction, the literature suggests periphrastic \emph{doen} might be a phenomenon limited to eastern Dutch.
+Whereas the longitude of the speaker's origin cannot be checked easily for the CGN data, the latitude can: the data are annotated for being either Netherlandic or Belgian Dutch.
+One third of the CGN is recorded in Belgium.
+
+\begin{table}[h] 
+\begin{tabular}{|l||ll|ll|} 
+\hline
+&Netherlandic&&Belgian&\\
+\hline
+periphrastic \emph{doen} sentences	&63 	& \emph{95.45\%}	&3		 &\emph{4.55\%}\\
+control sentences					&67		& \emph{67\%}		&33		 &\emph{33\%}\\
+\hline
+\end{tabular}
+\end{table}
+
+These data reveal periphrastic \emph{doen} seems to be mostly limited to Netherlandic Dutch. A chi-square test showed this difference was significant ($\chi^2(1, N=166) = 18.954, p > .001$).
 
@@
 
-An observation done while annotating for the previously described factors is that periprastic \emph{doen} is used in first and second person, but never in the third.
+An observation done while annotating for the previously described factors is that periphrastic \emph{doen} is used in first and second person, but never in the third.
 While this can be linked to various more abstract explanations, as will be attempted in the next two sections, it was first investigated whether this really was the case, and whether this is expected on the basis of the control sentences.
@@
 \begin{table}[h]
-\begin{tabular}{|l|ll|ll|ll|}
+\begin{tabular}{|l||ll|ll|ll|}
 \hline
@@
 \hline
-periprastic \emph{doen} sentences	&26 	& \emph{39.39\%}	&40		&\emph{60.61\%} 	& 0 &\emph{0\%}\\
+periphrastic \emph{doen} sentences	&26 	& \emph{39.39\%}	&40		&\emph{60.61\%} 	& 0 &\emph{0\%}\\
 control sentences					&55		& \emph{55\%}		&21	 	&\emph{21\%} 		& 24 &\emph{24\%}\\
@@
 
-Here we see indeed that the third person is absent from the \emph{doen} sentences, which suggests that periprastic \emph{doen} in third person is infrequent, and possibly even ungrammatical.
+Here we see indeed that the third person is absent from the \emph{doen} sentences, which suggests that periphrastic \emph{doen} in third person is infrequent, and possibly even ungrammatical.
 It should be added that a counter-example to that last suggestion was found on \emph{taaladvies.net}, the language advice website discussed earlier, although it unclear to what extent this example is artificial:
@@
 This is related to the fact that the number of imperatives is exceptionally high among the \emph{doen} sentences.
-This will be investigated further in the next section.
-A chi-square test showed this difference is indeed significant ($\chi^2(2, N=166) = 34.8, p > .001$).
-%stat
+This will be investigated further in the next section.\\\indent
+A possible interpretation of periphrastic \emph{doen}'s preference for the first and second person is that \emph{doen} is somehow limited to interactional uses; the lack of periphrastic \emph{doen} in third person could then be explained by the fact that the third person cannot be used for interaction.
+In other words, it is possible that periphrastic \emph{doen} functions as something like an 'interaction marker'.\\\indent
+A chi-square test showed the difference between the \emph{doen} sentences and the control sentences was significant ($\chi^2(2, N=166) = 34.8, p > .001$).
 
@@
 
-A possible interpretation of periprastic \emph{doen}'s preference for the first and second person is that \emph{doen} is somehow limited to interactional uses.
-The lack of emhatic \emph{doen} in third person could then be explained by the fact that the third person cannot be used for interaction.
-This interactional purpose also means that imperative and questions should occur more often in combination with \emph{doen}.
-To investigate whether periprastic \emph{doen} is somehow related to the imperative or the interrogative mood, for each sentence the mood was annotated.
+If periphrastic \emph{doen} indeed has an interactional function, we would of course also expect it to occur more in interactional situations like imperatives and questions.
+In fact, it might even be the case that \emph{doen} is not so much an interaction marker, but is a marker for a less frequent mood - similar to how an accusative marker is a marker for a less frequent case.
+To investigate whether periphrastic \emph{doen} indeed is somehow related to the imperative or the interrogative mood, for each sentence the mood was annotated.
 
 \begin{table}[h]
-\begin{tabular}{|l|ll|ll|ll|}
+\begin{tabular}{|l||ll|ll|ll|}
 \hline
@@
 \hline
-periprastic \emph{doen} sentences	&32 	& \emph{48.48\%}	&25		&\emph{37.88\%} & 9 &\emph{13.64\%}\\
+periphrastic \emph{doen} sentences	&32 	& \emph{48.48\%}	&25		&\emph{37.88\%} & 9 &\emph{13.64\%}\\
 control sentences					&91		& \emph{91\%}		&0	 	&\emph{0\%} 	& 9 &\emph{9\%}\\
@@
 A chi-square test showed this difference is significant ($\chi^2(2, N=166) = 48.37, p > .001$).
-%stat
 
@@
 In the previous section we saw that the imperative and interrogative mood are more frequent for the \emph{doen} sentences than for the control sentences.
-Although this might be the result of periprastic \emph{doen}'s interactional character, it might also hint at a more syntactic interpretation: in Dutch, both questions and imperatives are formed by deviating from the standard SVO word order found in main clauses: questions are formed with \emph{inversion}, which means that the inflected verb is in first position and the subject in second position (VSO), imperatives are formed by dropping the subject (VO).
+Although this might be the result of periphrastic \emph{doen}'s interactional character, it might also hint at a more syntactic interpretation: in Dutch, both questions and imperatives are formed by deviating from the standard SVO word order found in main clauses: questions are formed with \emph{inversion}, which means that the inflected verb is in first position and the subject in second position (VSO), imperatives are formed by dropping the subject (VO).
 
@@
 
-Possibly, periprastic \emph{doen} is used to indicate that an alternative word order is used.
+Possibly, periphrastic \emph{doen} is used to indicate that a non-standard word order is used.
+To investigate this, all sentences where annotated for word order.
+Besides the reasons mentioned above, a frequent source of alternative word order turned out to be object fronting (OVS).
 
 \begin{table}[h]
-\begin{tabular}{|l|ll|ll|}
+\begin{tabular}{|l||ll|ll|}
 \hline
-&SVO&&V(S)O&\\
+&SVO&&V(S)O or OVS&\\
 \hline
-periprastic \emph{doen} sentences	&55 	& \emph{83.33\%}	&11		 &\emph{16.67\%}\\
-control sentences				&42		& \emph{42\%}		&58		 &\emph{58\%}\\
+periphrastic \emph{doen} sentences	&11		 &\emph{16.67\%}&55 	& \emph{83.33\%}	\\
+control sentences					&58		 &\emph{58\%}	&42		& \emph{42\%}		\\
 \hline
@@
 
-A chi-square test showed this difference is indeed significant ($\chi^2(1, N=166) = 27.97, p > .001$).
-
-%pas op met causation
-%mss ergens noemen dat die laatste drie niet echt verklaringen zijn. Of misschien is doen een mood-indicator? Over nadenken.
+As can be seen from these data, whereas most \emph{doen} examples indeed have an alternative word order, this is not the case for the control sentences.
+A chi-square test showed this difference is significant ($\chi^2(1, N=166) = 27.97, p > .001$).
+It should be added, however, that this correlation does not tell us anything about causation.
+For example, it has been suggested above that the presence of emphatic \emph{doen} is caused by alternative word orders, but it might as well be the case the presence of \emph{doen} actually causes alternative word orders.
 
 \subsection{Results} \label{regressionresults}
+
+When investigated seperately, only the factor \emph{given before new} did not have a relation to the \emph{doen} sentences.
+When all factors are investigated at the same time in a multiple regression, however, we might see something different: some of the factors might be correlated, which means that adding them both does not add more predictive power to the model.
+Another advantage of multiple regression is that we can calculate $R^2$, which indicates how much of the variation the model can explain.
+This is particularly important to this study, as the significant results of the $\chi^2$ tests might have camouflaged that the results never were really good predictors: no category had a majority of the \emph{doen} sentences in group A, but a majority of the control sentences in group B.
+
+\begin{table}[h] \centering
+\begin{tabular}{|l||l|l|l|}
+\hline
+Factor&B&SE B& $\beta$ \\
+\hline
+Constant&0.03&0.07&\\
+Phrasal verb&0.41&0.09&.29**\\
+Given before new&0.03&0.07&.03\\
+Habitual&0.21&0.07&.18*\\
+Variety&0.22&0.07&.18*\\
+Person&-0.27&0.09&-.20*\\
+Mood&0.32&0.08&.29**\\
+Inversion&0.06&0.07&.06\\
+\hline
+\end{tabular}
+\caption{Multiple regression results, * $ p > 0.005$, **  $p > 0.001$ }
+\end{table}
+The most important thing to note here is that inversion is no longer significant.
+As discussed above, it is strongly related to person and mood, which may make it useless to a regression model; investigations of correlation data showed that this was indeed the case.
+As expected, this model has a relatively low predictive value: $R^2 = .461$.\\\indent
 
@@
 
-The classifier used is the IB1-IG in the TiMBL software package (ref).
-To make the dataset ready for TiMBL, two manipulations were done to make sure TiMBL can only use the lexical items for its predicts (and no other clues):
-
-%ref
+The classifier used is IB1-IG in the TiMBL software package \citep{d10}.
+To make the dataset ready for TiMBL, two manipulations were done to make sure TiMBL can only use the lexical items for its predictions (and no other clues):
 
@@
 \item All \emph{doen} sentences were rewritten into a non-\emph{doen} sentence.
-This is necessery, because TiMBL can simply use the presence of the lexical item \emph{doen} as an indicator whether this is an periprastic \emph{doen} sentence or not.
+This is necessary, because TiMBL can simply use the presence of the lexical item \emph{doen} as an indicator whether this is an periphrastic \emph{doen} sentence or not.
 
@@
 
-%perc met punt achter de komma
+TiMBL was able to correctly predict the sentence type for 24 of the 32 examples, which means an prediction accuracy of \textbf{75\%}.
 
-\section{Conclusion} \label{conc}
+%tabel labels?
+%tabels too wide centeren
 
-\begin{thebibliography}{99}
+\section{Conclusion and discussion} \label{conc}
+In this paper we saw that the periphrastic \emph{doen} construction, although impopular, can be found in (the eastern variant of) spoken Dutch.
+Of the various factors investigated, phrasal verbs, a possible habitual interpretation, a Dutch speaker, first or second person and non-indicative mood turned out to be useful predictors of the construction, but together could explain slightly less than half of the variation.
+On the other hand, a Memory Based Learning model using only lexical information, achieved much better results.
+This can either mean not the right factors have been investigated in the regression model, or that periphrastic \emph{doen} is triggered by certain lexical items.
+
+\begin{thebibliography}{99}\footnotesize
 \bibitem[CGN(2006)]{cgn}
 \emph{Corpus Gesproken Nederlands, version 2.0.} Electronic resource. Available online at http://lands.let.ru.nl/cgn/home.htm.
+\bibitem[Daelemans et al.(2010)]{d10}
+Daelemans, W., J. Zavrel, K. Van der Sloot, \& A. Van den Bosch (2010). \emph{TiMBL: Tilburg Memory Based Learner, version 6.3, Reference Guide}. ILK Research Group Technical Report Series no. 10-01.
 \bibitem[Duinhoven(1994)]{d94}

### 2013-01-24 Research paper.tex
RP: Added suggestion by Maarten

@@
 A project overview therefore cannot be complete without an overview of the software's functionality.
-Colibri's functionality can be split in four parts: (1) identifying and extracting constructions in a language, (2) identifying aligned constructions between language-pairs, (3) memory-based classification using construction experts and (4) reassembling constructions into a coherent target sentence.
+Colibri's functionality can be split in four parts: (1) identifying and extracting constructions in a language, (2) identifying aligned constructions between language-pairs, (3) memory-based classification and (4) reassembling constructions into a coherent target sentence.
 I will discuss the four parts in more detail consecutively:
@@
 
-\subsubsection{Memory-based classification using construction experts}
+\subsubsection{Memory-based classification}
 As may have become clear from the previous sections, the Colibri project investigates the use of machine learning techniques to get from the source language construction to the target language constructions.
-More specifically, Colibri uses specialized classifiers, trained for only one construction, inspired by the effectivity of this approach in word sense ambiguation \citep{vangompel10}.
-The results of this approach are compared to the results of a combined classifier approach and statistical machine translation.
+Three approaches have been implemented:
+
+\begin{itemize}
+\item Monolithic classifier. All phrases go through one general classifier. This classifier thus has a very large number of classes to pick from.
+\item N-array classifiers. There is a separate classifier for each phrase length (n); the length of the source language construction thus decides which classifier is used.
+\item Construction experts.
+Each construction or phrase has its own classifier, which means that instead of one larger classifier, there are a lot of smaller classifier making a simpler decision.
+This is inspired by the effectivity of this approach in word sense ambiguation \citep{vangompel10}.
+\end{itemize}
 
@@
 However, as was explained in section \ref{Description}, Colibri is an attempt to improve the quality of machine translation by adding skipgrams and memory-based learning.
-Unfortunately it has turned out that both do not add much to the translation quality: memory-based learning only improves the results a little bit, and the skipgrams are mostly ignored by the decoder.
+Unfortunately it has turned out that both do not add much to the translation quality: all three memory-based learning approaches (construction experts, N-array classifiers and a monolithic classifier) only improve the results a little bit, and the skipgrams are mostly ignored by the decoder.
 Apparently, translation hypotheses without skipgrams are more likely to be good translations in most of the cases.

### 2013-01-29 Paper doen.tex
Added a few last missing things

@@
 \usepackage{gb4e}
+
+\addtolength{\textwidth}{0.4in}
+\addtolength{\hoffset}{-0.2in}
+\addtolength{\textheight}{0.4in}
+\addtolength{\voffset}{-0.2in}
 
@@
 These are sentences with a finite verb but without periphrastic \emph{doen}, which is almost every Dutch sentence.
-Picking a (sub)corpus to select these control sentences from turned out to be a rather difficult task.
+Picking a (sub)corpus from which to select these control sentences, which comes down to defining the envelope of variation, turned out to be a rather difficult task.
 \\\indent
@@
 There does not seem to be any consensus about possible predictors besides that it is limited to informal language, so that is the only place the control sentences were limited: all were taken from the \emph{face-to-face conversations} or \emph{telephone conversations} from the CGN, because that is also the place where all \emph{doen} sentences came from (with the exception of one example from sports commentaries).
-
-% Envelope of variation
-% Misschien iets doen met interacties?
 
@@
 
-\begin{table}[h]
+\begin{table}[h] \footnotesize
+
+\makebox[\textwidth][c]{
 \begin{tabular}{|l||ll|ll|} 
@@
 \hline
-\end{tabular}
+\end{tabular}}
+\caption{\footnotesize Distribution phrasal verbs versus non-phrasal verbs for the \emph{doen} sentences and the control sentences.}
 \end{table}
@@
 
-Although there is a lot of disagreement over the exact details and the terminology, there is a wide consensus among linguists that languages often use the \emph{given before new}-principle (ref).
+Although there is a lot of disagreement over the exact details and the terminology, there is a wide consensus among linguists that languages often use the \emph{given before new}-principle (see \citet{dsdh95} for an overview).
 This principle entails that speakers try to put information already known by the addressee in the beginning of the sentence, while the part of the sentence that contains the new information is put at the end as much as possible.
 Or, to put it differently, speakers first introduce their topic, and then say what they want to say about this topic.\\\indent
-%ref
 The Dutch V2-rule, however, limits the possibilities to do this drastically.
@@
 
-\begin{table}[h]
+\begin{table}[h] \footnotesize
+
+\makebox[\textwidth][c]{
 \begin{tabular}{|l||ll|ll|}
@@
 \hline
-\end{tabular}
+\end{tabular}}
+\caption{\footnotesize Distribution of whether the main verb contained new information for the \emph{doen} sentences and the control sentences.}
 \end{table}
@@
 
-\begin{table}[h] 
+\begin{table}[h] \footnotesize
+
+\makebox[\textwidth][c]{
 \begin{tabular}{|l||ll|ll|} 
@@
 \hline
-\end{tabular}
+\end{tabular}}
+\caption{\footnotesize Distribution of whether habitual interpration is possible for the \emph{doen} sentences and the control sentences.}
 \end{table}
@@
 \hline
-\end{tabular}
+\end{tabular}}
+\caption{\footnotesize Distribution of variety for the \emph{doen} sentences and the control sentences.}
 \end{table}
@@
 
-\begin{table}[h]
+\begin{table}[h] \footnotesize
+
+\makebox[\textwidth][c]{
 \begin{tabular}{|l||ll|ll|ll|}
@@
 \hline
-\end{tabular}
+\end{tabular}}
+\caption{\footnotesize Distribution person for the \emph{doen} sentences and the control sentences.}
 \end{table}
@@
 \hline
-\end{tabular}
+\end{tabular}}
+\caption{\footnotesize Distribution of mood for the \emph{doen} sentences and the control sentences.}
 \end{table}
@@
 
-\begin{table}[h]
+\begin{table}[h] \footnotesize
+
+\makebox[\textwidth][c]{
+
 \begin{tabular}{|l||ll|ll|}
@@
 \hline
-\end{tabular}
+\end{tabular}}
+\caption{\footnotesize Distribution of word order for the \emph{doen} sentences and the control sentences.}
 \end{table}
@@
 
-When investigated seperately, only the factor \emph{given before new} did not have a relation to the \emph{doen} sentences.
+When investigated separately, only the factor \emph{given before new} did not have a relation to the \emph{doen} sentences.
 When all factors are investigated at the same time in a multiple regression, however, we might see something different: some of the factors might be correlated, which means that adding them both does not add more predictive power to the model.
@@
 \caption{Multiple regression results, * $ p > 0.005$, **  $p > 0.001$ }
+\label{restable}
 \end{table}
-The most important thing to note here is that inversion is no longer significant.
+The most important thing to note from table \ref{restable} is that inversion is no longer significant.
 As discussed above, it is strongly related to person and mood, which may make it useless to a regression model; investigations of correlation data showed that this was indeed the case.
@@
 
-%tabel labels?
-%tabels too wide centeren
-
 \section{Conclusion and discussion} \label{conc}
-In this paper we saw that the periphrastic \emph{doen} construction, although impopular, can be found in (the eastern variant of) spoken Dutch.
-Of the various factors investigated, phrasal verbs, a possible habitual interpretation, a Dutch speaker, first or second person and non-indicative mood turned out to be useful predictors of the construction, but together could explain slightly less than half of the variation.
+In this paper we saw that the periphrastic \emph{doen} construction, although unpopular, can be found in (the eastern variant of) spoken Dutch.
+Of the various factors investigated, phrasal verbs, a possible habitual interpretation, a Dutch speaker, first or second person and non-indicative mood turned out to be useful predictors of the construction, but together could only explain slightly less than half of the variation.
 On the other hand, a Memory Based Learning model using only lexical information, achieved much better results.
-This can either mean not the right factors have been investigated in the regression model, or that periphrastic \emph{doen} is triggered by certain lexical items.
+This can either mean not enough factors have been included in the regression model, or that periphrastic \emph{doen} is triggered by certain lexical items.
 
@@
 Nuijtens, E. (1962). \emph{De tweetalige mens}. Assen: Van Gorcum \& Comp.
+\bibitem[de Swart \& de Hoop(1995)]{dsdh95}
+De Swart, H. \& H. De Hoop (1995). Topic and focus. \emph{Glot International} 1: 3 - 7.
 \end{thebibliography}

### 2013-01-30 Paper doen.tex
Doen: final revisions

@@
 
-Interestingly, however, in some non-standard, informal varieties of Dutch, something very similar to do-support can be found.
+Interestingly, however, in informal language, something very similar to do-support can be heard.
 Speakers replace the inflected verb in verb second position with an inflected form of \emph{doen} 'do', and place the meaningful verb in verb final position.
@@
 153]{n62} claims the construction is mainly limited to children and women.\\\indent
-Interestingly, the construction has also been claimed to be part of dialectal language use: it would be part of the dialects of Groningen \citep{tl53}, Twente \citep{n62} and Heerlen \citep{c94}, dialects that have in common that they are all spoken in the east of the Netherlands.\\\indent
+Interestingly, the construction has also been claimed to be part of dialectal language use: it would be part of the dialects of Groningen \citep{tl53}, Twente \citep{n62} and Heerlen \citep{c94}, dialects that have in common that they are all spoken in the eastern part of the Netherlands.\\\indent
 In this paper, I will show that periphrastic \emph{doen} is not limited to dialectal language or the language of children or women,  but seems to be part of (the eastern variant of) Dutch in general.
@@
 These are sentences with a finite verb but without periphrastic \emph{doen}, which is almost every Dutch sentence.
-Picking a (sub)corpus from which to select these control sentences, which comes down to defining the envelope of variation, turned out to be a rather difficult task.
+However, picking a (sub)corpus from which to select these control sentences, which comes down to defining the envelope of variation, turned out to be a rather difficult task.
 \\\indent
 For example, if 100 sentences selected from all Dutch corpora available would have been used, and then annotated for register, we would probably see a significant effect for register: the \emph{doen} examples are all spoken, informal language, while this would not true for all control sentences.
-We would thus find that there is no variation in formal language.
+We would thus find that there is no variation in formal language with respect to the presence perhiphrastic \emph{doen}.
 However, if we then wanted to use this same dataset to investigate other effects, these formal sentences might obscure the view; for example, while verb type might be a strong predictor for the use of periphrastic \emph{doen} in informal language, it is not at all in formal language (as it never occurs there, regardless of verb type).
@@
 
-\subsection{Factors included}
-
-In this section, I will elaborate on the factors included in the regression model, and use chi-square tests to see whether the \emph{doen} sentences behave differently from the control sentences with the regard to the concept investigated.
+In this section, I will try to find out in which cases periphrastic \emph{doen} is used by looking at various factors.
+In section \ref{factor} I will elaborate on the factors included in the regression model, and use chi-square tests to see whether the \emph{doen} sentences behave differently from the control sentences with the regard to the concept investigated.
 In section \ref{regressionresults} the predictive power of all these factors combined will be discussed.
 
+\subsection{Factors included} \label{factor}
+
 \subsubsection{Phrasal verbs}
-In Dutch, several verbs consists of multiple parts.
+In Dutch, several verbs consist of more than one part.
 \emph{Uitschelden} 'abuse, insult', for example, consists of \emph{uit} 'out' and \emph{schelden} 'scold'.
@@
 \citet[pp.
-156]{n62} suggests periphrastic \emph{doen} is a strategy to avoid the extra processing load this long-distance relationship causes, both for the speaker and the hearer.
+156]{n62} suggests that the extra processing load this long-distance relationship causes can be avoided by using periphrastic \emph{doen}, which makes the construction beneficial for both the speaker and the addressee.
 If that is true, we would expect to find much more phrasal verbs as the main verb in the \emph{doen} sentences than in the control sentences.
@@
 Unless there is an auxiliary present for other reasons, it is not possible to put the main verb at the end of the sentence in main clauses.
-periphrastic \emph{doen} could be a perfect tool to solve this problem, as it forces the main verb to go to the end of the sentence.
+Periphrastic \emph{doen} could be a perfect tool to solve this problem, as it forces the main verb to go to the end of the sentence.
 If \emph{doen} is indeed used this way, we would expect that the main verb in \emph{doen} sentences indeed is new information.
@@
 Another advantage of multiple regression is that we can calculate $R^2$, which indicates how much of the variation the model can explain.
-This is particularly important to this study, as the significant results of the $\chi^2$ tests might have camouflaged that the results never were really good predictors: no category had a majority of the \emph{doen} sentences in group A, but a majority of the control sentences in group B.
+This is particularly important to this study, as the significant results of the $\chi^2$ tests might have camouflaged that the results never were really good predictors: no factor had a majority of the \emph{doen} sentences in group A, but a majority of the control sentences in group B.
 
@@
 Mood&0.32&0.08&.29**\\
-Inversion&0.06&0.07&.06\\
+Word order&0.06&0.07&.06\\
 \hline
@@
 \end{table}
-The most important thing to note from table \ref{restable} is that inversion is no longer significant.
-As discussed above, it is strongly related to person and mood, which may make it useless to a regression model; investigations of correlation data showed that this was indeed the case.
+The most important thing to note from table \ref{restable} is that word order is no longer significant.
+As discussed above, it is strongly related to person and mood, which may make it useless to a regression model.
 As expected, this model has a relatively low predictive value: $R^2 = .461$.\\\indent
@@
 
-TiMBL was able to correctly predict the sentence type for 24 of the 32 examples, which means an prediction accuracy of \textbf{75\%}.
+TiMBL was able to correctly predict the sentence type for 24 of the 32 examples, which means a prediction accuracy of \textbf{75\%}.
 
@@
 Of the various factors investigated, phrasal verbs, a possible habitual interpretation, a Dutch speaker, first or second person and non-indicative mood turned out to be useful predictors of the construction, but together could only explain slightly less than half of the variation.
-On the other hand, a Memory Based Learning model using only lexical information, achieved much better results.
+On the other hand, a Memory Based Learning model using only lexical information achieved much better results.
 This can either mean not enough factors have been included in the regression model, or that periphrastic \emph{doen} is triggered by certain lexical items.

### 2013-05-07 research proposal MTS.tex
Started with research proposal MTS

\documentclass[12pt]{article}

\title{Research proposal: idiolects in word prediction}
\author{Wessel Stoop}

\usepackage{graphicx}
\usepackage{float}
\usepackage{apacite}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}

\section{Project title}
Idiolects in word prediction
\section{Summary}
[[]]
\section{Candidate}
Wessel Stoop
\section{Description of the proposed research}

\section{Summary in keywords}

\end{document}

### 2013-05-07 research proposal MTS.tex
Added introduction word completion

@@
 Idiolects in word prediction
+
 \section{Summary}
 [[]]
+
 \section{Candidate}
 Wessel Stoop
+
 \section{Description of the proposed research}
+
+This research is an attempt to link two academic fields: the field of \emph{word completion} and the field \emph{idiolects}.
+They will be introduced in section \ref{wordprediction} and \ref{idiolects} respectively.
+Section \ref{link} will show what the advantages of linking these two concepts are.
+
+\subsection{Word prediction} \label{wordprediction}
+
+\subsubsection{Introduction}
+
+Word prediction is used in many applications we use daily: Google completes our search queries, our webbrowser finishes the web addresses we type in, our email client finishes email addresses, Excel completes annotations based on earlier annotations, etc.
+It reduces the number of keystrokes we have to do, and thus saves us time and effort.
+However, when we type full sentences, whether they are part of an email, a scientific article, a tweet, or a research proposal, we often still key in every single character.
+This can be time consuming for regular users, but highly frustrating for people with physical disabilities; [[example]].
+Word prediction is not used for sentences because it is generally considered 'not good enough'; having to reject wrong predictions all the time apparently is more frustrating than having to key in every character.
+With the rise of smartphones during the last years word prediction system have become more widely known (because it takes more time to type on a phone, which makes autocompletion more attractive), although it is unclear what portion of the smartphone users also use some sort of autocompletion application.
+
+The goal of this research is to improve the quality of word prediction. We will build a word prediction system that incorperates the latest insights from the field, and ...
+
+% autocompletion
+X
+%N. Garay-Vitoria and J. Abascal. 2006. Text prediction systems: a survey. Universal Access in the Information Society, 4(3):188–203.
+
+\subsection{Idiolects} \label{idiolects}
+
+\subsection{Idiolects in word prediction} \label{link}
 

### 2013-05-08 research proposal MTS.tex
Revised first alinea

@@
 However, when we type full sentences, whether they are part of an email, a scientific article, a tweet, or a research proposal, we often still key in every single character.
-This can be time consuming for regular users, but highly frustrating for people with physical disabilities; [[example]].
-Word prediction is not used for sentences because it is generally considered 'not good enough'; having to reject wrong predictions all the time apparently is more frustrating than having to key in every character.
-With the rise of smartphones during the last years word prediction system have become more widely known (because it takes more time to type on a phone, which makes autocompletion more attractive), although it is unclear what portion of the smartphone users also use some sort of autocompletion application.
+
+Why don't we use word prediction here too?
+While, as far as I am aware, no quantitative research has been done to answer this question, my own experience with word prediction is that its prediction accuracy generally is too low to speed up typing.
+Keying in every character is often faster than checking and accepting the predictions.
+Once typing speed decreases, however, word prediction \emph{is} used.
+With the rise of smartphones,  on which one generally cannot type as fast as on a normal keyboard, word prediction has become more widely known - although it is unclear what proportion of the smartphone users also use some sort of word prediction application.
+Another example are  people with speech and motor disabilities, like cerebral palsy or hemiplexia.
+By using a device equiped with word prediction technology, they can increase their communication rate considerably \cite{Garay-Vitoria}.
+
+In the academic literature, several attempts have been made to improve word prediction technology, so that it can increase the communication rate for smartphone users, people with physical disabilities and typing in general.
+
+% WOORDENBOEK, CS, Antal & Suzan?
 
@@
 % autocompletion
-X
 %N. Garay-Vitoria and J. Abascal. 2006. Text prediction systems: a survey. Universal Access in the Information Society, 4(3):188–203.

### 2013-05-21 RS twitter.tex
Started with RS

\documentclass[12pt]{article}

\title{Twitter as a crystal ball: predicting future events with Twitter}
\author{Wessel Stoop, s0808709}

\usepackage{covington}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{float}

\renewcommand{\familydefault}{\sfdefault}

\begin{document}
\maketitle

\section{Introduction}

\section{Success stories}

%Eindingen met elections. Eerst Tumasjan, daarna Erik en nog wat anderen, om te laten zien dat die resultaten niet herhaald kunnen worden. Mag ook bij de volgende sectie.

\section{Problems}

With regard to elections, some authors have taken the opposite stance [[ref]]. Most fierce in this respect is [[x]], who says 'No, you cannot predict elections with Twitter'. He goes on to point out a number 'flaws' in research on doing predictions with Twitter. A summary of his most important points:

\begin{enumerate}
\item Scholars should actually predict something, instead of doing \emph{post-hoc} analyses. As [[x]] put it: 'if you are claiming you have a prediction method you should predict an election in the future!'
\item 'Not all tweets are trustworthy.' That is, some might be jokes or sarcasm and some might have been placed by the political leaders themselves.
\item 'Demographics are neglected.' Twitter users are likely to be higher-educated people interested in social media, which might not be a representative sample of the population . According to [[x]], all research on predicting elections with Twitter should take this into account.
\item 'There is not a commonly accepted way of “counting votes” in Twitter'. Whereas some authors use tweet volumes, others use tweet rate, and again other focus on the results of some kind of sentiment analyses.
\end{enumerate}

I am convinced all of these points are not really flaws in the research, but more general problems related to predicting with Twitter. As for the first point, you cannot evaluate a prediction system if you do not use past events; a paper saying 'sytem X predicts event Y', without the information where Y actually happened is meaningless. As long as the authors only used material available \emph{before} event Y, their result are perfectly valid. Arguments 2 and 3 claim there might be noise in the data, which I think is not only a general problem in Twitter research, but a problem in science in general: datasets which are 100\% representative of thing they were created to be representative of are scarce. The task of predicting events with Twitter is to do accurate predictions \emph{despite} these problems. If researchers can even report positive results by ignoring them, why bother? Finally, the last 'flaw' is actually the main goal of this research field: finding which way of 'counting votes' has the best correllation to the actual election results. Because it is not clear yet what technique consistently has the best results, various ideas will be proposed.

I therefore feel it would be more appropriate to rephrase the critique of [[x]] as problems common to this research field. Attempts to solve them might improve the result, and might even be research questions on their own. For example, [[x]] investigated Twitter demographics, and concluded that .

\end{document}

### 2013-05-22 research proposal MTS.tex
Finished first version MTS

@@
 \usepackage{apacite}
+\usepackage{covington}
 
@@
 \section{Project title}
-Idiolects in word prediction
+Defining idiolects for word prediction
 
 \section{Summary}
-[[]]
+Systems that predict what words we want to type as we are keying them in can save us a lot of time, in particular for users of smartphones and people with physical or cognitive disabilities.
+Recently, scholars have tried to implement knowledge of style register and topics into word prediction systems.
+This research will investigate a new approach to this task: using idiolects.
+
+The concept of idiolects, the language of only one person, has been well-known as a concept for a long time, but has rarely been investigated quantitively for practical and theoretical reasons.
+As a result of this, it is unclear how an idiolect should be defined (for example, in terms of style registers and changes over time).
+Now the linguistic landscape has changed (both practically and theoretically), these questions can finally be answered.
+
+The research proposed here consists of building a word prediction system based on idiolects.
+Various models, based on various definitions of idiolects, will be tested.
+Knowledge of which model (and thus definition) can be used to save the most keystrokes will be useful to both the field of word prediction and the field of idiolects.
 
@@
 
-This research is an attempt to link two academic fields: the field of \emph{word completion} and the field \emph{idiolects}.
+This research is an attempt to link two fields of research: the field of \emph{word completion} and the field \emph{idiolects}.
 They will be introduced in section \ref{wordprediction} and \ref{idiolects} respectively.
@@
 \subsection{Word prediction} \label{wordprediction}
-
-\subsubsection{Introduction}
 
@@
 Why don't we use word prediction here too?
-While, as far as I am aware, no quantitative research has been done to answer this question, my own experience with word prediction is that its prediction accuracy generally is too low to speed up typing.
+While no quantitative research has been done to answer this question (as far as we are aware), an obvious answer would be that prediction accuracy generally is too low to speed up typing.
 Keying in every character is often faster than checking and accepting the predictions.
 Once typing speed decreases, however, word prediction \emph{is} used.
-With the rise of smartphones,  on which one generally cannot type as fast as on a normal keyboard, word prediction has become more widely known - although it is unclear what proportion of the smartphone users also use some sort of word prediction application.
+With the rise of smartphones,  on which one  cannot type as fast as on a normal keyboard, word prediction has become more widely known - although it is unclear what proportion of the smartphone users also use some sort of word prediction application.
 Another example are  people with speech and motor disabilities, like cerebral palsy or hemiplexia.
-By using a device equiped with word prediction technology, they can increase their communication rate considerably \cite{Garay-Vitoria}.
+By using a device equiped with word prediction technology, they can increase their communication rate considerably \cite{Garay-Vitoria+06}.
+
+%In the academic literature, several attempts have been made to improve word prediction technology, so that it can increase the communication rate for smartphone users, people with physical disabilities and typing in general.
+One of the first techniques to achieve a useable word prediction system is to use frequency lists; although it is possible to wait until a word unicity point has been reached, much more keystrokes can be saved if the prediction can be done earlier.
+For example, if the user wants to type the word \emph{communication}, the system could wait until \emph{communicatio} has been keyed in to make sure not do an incorrect suggestion in case the user wanted to say \emph{communicative}, but it could also take the risk and guess one of the words much earlier.
+Which one of the two the system guesses should be the option that is more likely.
+This chance of occurence of a word is reflected by its frequency in texts.
+In the example, \emph{communication} will be more likely to be correct because it will probably be higher on a frequency list.
+One of the first word prediction systems using this technique, and also one of the first word predictions systems in general, was PAL \cite{swiffin+85}.
 
 In the academic literature, several attempts have been made to improve word prediction technology, so that it can increase the communication rate for smartphone users, people with physical disabilities and typing in general.
+One of the first techniques to achieve a useable word prediction system is to use frequency lists, because these lists reflect the chance of occurrence of a word.
+If a word has a high position on a frequency list, this means it has occurred a lot in previous text, and thus is (relatively) likely to occur again, which makes it an interesting choice for a word prediction system.
+One of the first word prediction systems using this technique, and also one of the first word predictions systems in general, was PAL \cite{swiffin+85}.
+Since then, numerous authors have shown that taking context into account improves prediction accuracy significantly \cite[among others]{Lesher+99,Garay-Vitoria+06,Tanaka-Ishii07,vandenbosch+08}.
+This entails that the system has some information on which words are likely to follow eachother, and uses this when giving predictions.
+For example, if the user has keyed in sentence \ref{iate}, a context-sensitive system should not predict the word \emph{communication}, even if it is high on the frequency lists, because \emph{communication} is unlikely to follow \emph{ate}.
 
-% WOORDENBOEK, CS, Antal & Suzan?
+\begin{examples}
+\item I ate co \label{iate}
+\end{examples}
 
-The goal of this research is to improve the quality of word prediction. We will build a word prediction system that incorperates the latest insights from the field, and ...
+%One of the earliest implementations of this concept, by \citeA{hunnicutt87}, used a 2D-matrix to store frequent collocations of words. A matrix for our example could be:
+
+%\begin{examples}
+
+%\item \begin{tabular}{l|ll}
+%&I&you\\
+%\hline
+%like&dancing&fishing\\
+%ate&cookies&cake\\
+%\end{tabular}
+
+%\end{examples}
+
+%However, when using millions of words and all possible combinations between them, and also more context than just two words, such a matrix quickly becomes impractical.
+
+Modern word predictions systems search through text corpora ('training material') and collect a number of similar contexts.
+The word that follows most of these contexts will be the word predicted \cite[among others]{vandenbosch+08}.
+
+% A key publication on context-sensitivity is that by \citeA{Lesher+99}, who show that the accuracy of a context-sensitive word prediction system is related to how much training material is provided.
+This is the case in particular when more words from the context are used, because then much more combinations are possible.
+In a hypothetical language with 10 words, 10 contexts are possible when using 1 word as the context, but $10*10*10 = 1000$ contexts are possible when using 3 words as the context.
+Thus, the larger the context the word prediction system uses, the more the system benefits from having a lot of training material.
+
+Context-sensitivity, in particular when using a lot of training material, is most powerful when it comes to collocations (for example, \emph{in combination with}, \emph{on the basis of}), i.e.
+when it comes to language.
+However, humans use more than knowledge of language alone to predict what somebody is going to say; we also take into account what style register the speaker is using, and what the speaker is talking about.
+For instance, the input \emph{wed} can be very likely to refer to \emph{Wednesday} when making an appointment for next week, whereas it can also be very likely to refer to \emph{wedding} when talking about a marriage.
+
+One way give some of this kind of knowledge to word prediction systems is to use training material from the same domain; \citeA{verberne+12} show that trying to predict Wikipedia text, tweets, transcriptions of conversational speech and Frequencty Asked Questions all worked best when using texts from the same type (i.e.
+training on Wikipedia text to predict Wikipedia text).
+Furthermore, they show that questions about neurological issues, for which little training material was available, could best be predicted with the corpus that was closest in terms of topics discussed: the medical pages from Wikipedia.
+
+\citeA{vandenbosch11} takes a completely different approach and tries to make a word prediction system learn about register and topic on the fly with a \emph{recency buffer}.
+This buffer stores the $n$ latest words; if a word the user is keying in matches one of the words in the recency buffer, this word is suggested.
+The idea behind this is that a word that has recently been used, it is likely that this word will be used again.
+
+An important thing to notice is that, although both approaches help to increase the number of keystrokes saved, they also have downsides: for the system by \citeA{verberne+12} training texts in the same genre are needed, which might not be available, whereas the system by \citeA{vandenbosch11} ignores context information that it should weigh more intelligently.
+For example, while a context-sensitive text-prediction system will probably be able to predict \emph{to} for the sentence \emph{Tommy was going t...}, the one with the recency buffer will predict \emph{Tommy}.
 
 % autocompletion
-%N. Garay-Vitoria and J. Abascal. 2006. Text prediction systems: a survey. Universal Access in the Information Society, 4(3):188–203.
 
 \subsection{Idiolects} \label{idiolects}
+Almost every claim done in the field of linguistics concerns language as a whole; whether the subject of investigation is a particular syntactic construction, phonological variable, or some other linguistic phenomenon, the results are always supposed to hold for an entire language variety.
+The concept of an idiolect, the language of only one person, is well-known, but rarely ever mentioned.
+According to \citeA{mollin09} it is a 'neglected area in corpus linguistics', and \citeA{barlow10} states that the term 'is distinguished by the fact that there is probably no other linguistic term in which there is such a large gap between the familiarity of the concept and lack of empirical data on the phenomenon.' This is remarkable, since 'idiolects are the only kind of language we can
+collect data on', as \citeA{Haugen72} points out; a language variety essentially is nothing more than a collection of idiolects.
 
-\subsection{Idiolects in word prediction} \label{link}
+Part of the of lack of research in this area can of course be explained by practical limitations (it is hard to collect a lot of data for one person, let alone doing this for multiple persons), but theoretical views have also played a role; for decades, most linguistists have focussed on \emph{langue} rather than \emph{parole}, and viewed deviations in actual usage as uninteresting exceptions to an otherwise perfect abstract system.
+As \citeA{labov89} put it: 'language is not a property of the individual, but of the community.
+Any description of a language must take the speech community as its object if it is to do justice to elegance and regularity of linguistic structure'.
+
+Now that usage-based theories of language become increasingly more accepted, and the progress in technology and in particular the internet has facilitated collecting, storing and working with large amount of data, these objections to studying idiolects slowly vanish.
+One of the first publications investigating idiolects quantitatively is that of \citeA{mollin09}.
+She created a corpus of speech by the former Prime Minister of the United Kingdom Tony Blair, and compared this the British National Corpus.
+Focussing on maximisers ('absolutely', 'totally') and the adjectives, verbs and adverbs they cooccur with, she has compiled a list of collocations that are used much more by Blair than by the general language community.
+\citeA{barlow10} investigated speech patterns of five White House Press Secretaries, using transcripts of spoken language available on the internet and dividing them in samples of 200.000 words.
+He compared the frequencies of bigrams and trigrams (like \emph{of the}, \emph{that we} and \emph{and that's}) and discovered that there are significant differences between individual speakers.
+Interestingly, however, when frequencies from one sample were compared to frequencies from another sample by the same speaker, they were very consistent.
+For example, Ari Fleischer rarely uses the bigram \emph{we are} in sample 1, and this is also the case for samples 2, 3 and 4, while the opposite is the true for Scott McClellan.
+To make sure these differences cannot be attributed to lexical preferences, Barlow repeated his research with POS $n$-grams ([article, noun], [preposition, verb], etc) and
+grammatical constructions ('\emph{noun} of the \emph{noun}',  'will' vs 'going to', usage of the passive, etc.), and got similar results.
+
+While the results of these studies are promising, there are some problems in the way they try to capture idiolects.
+For example, all of the language investigated here is that of people speaking to the press, but it unlikely that the speakers always speak like this; Tony Blair probably uses another style register when speaking to his friends and family.
+How should style registers be part of a theory about idiolects?
+Is a register a set of variables in an otherwise fixed system, are they completely seperate language varieties inside a person's mind, or something in between?
+Furthermore, \citeA{barlow10} points out that 'the samples from Mike McCurry's speeches seem to divide into two temporally distinct groups [...], suggesting a change in style over his long period of tenure as press secretary'.
+Apparently, idiolects are not time stable, but change over time.
+If idiolects are always changing, what should we call an idiolect?
+And if we have models for a particular person based on the idiolect of the last decade, last year, yesterday and a few minutes ago, to what extent do they overlap?
+And finally, both \citeA{barlow10} and \citeA{mollin09} try to capture idiolects by using the \emph{output} of persons.
+This of course makes sense (what comes out must have been inside), but this cannot be the whole story; wouldn't it be more logical to use the input for a particular person, in particular the things he or she hears very often?
+
+\subsection{Methodology: idiolects in word prediction} \label{link}
+As the previous two section have showed, in both academic fields discussed we have a problem: for word prediction, a technique is needed which is context-sensitive, but does not rely on many texts in the same genre, while for idiolects, we are looking for a good way to make concrete what an idiolect actually is.
+In this research project, these two concepts will be linked to help each other: a word prediction system will be built which will rely on text written by the user earlier (idiolects), while incorperating the latest insights on the field of word prediction, as described in \ref{wordprediction}.
+This way, the system will be able to capture a user's vocabulary, writing style, favorite topics, etc., which according to \citeA{vandenbosch11} is 'likely to have a major impact on the percentages of keystrokes saved'.
+
+As our dataset, we plan to use tweets, as Twitter essentially is nothing more than a collection of language ordered by author.
+The expectation thus is that a model based on earlier tweets is much better at predicting future tweets than a model based random tweets in the same language.
+Furthermore, tweets are annotated for time, which allows us to make models for a particular period, and in a lot of cases explicitly mention the addressee (with the \emph{@addressee} syntax).
+This not only allows us to make an approximation of register (different models could be made depending on to whom the author was talking), but also gives us some insight in social relations between the Twitter users.
+The expectation is that our word prediction will be able to do more accurate predictions for a particular twitter user when using a model based on the tweets produced by people that user communicates with a lot.
+
+The goal of this research thus is to improve word prediction and define idiolects at the same time.
+Several language models will be created by systematically manipulating a number of factors; for example, a model with the tweets of friends and a model without the tweets of friends, and a model with the tweets of this year, and a model with the tweets of last year.
+For each of these models, it will be calculated how much keystrokes can be saved when the algorithm uses it.
+Knowledge of which model works best can be used to improve word prediction technology, but also gives us useful insights on how we can best define an idiolect.
 
 \section{Summary in keywords}
+Word prediction; idiolects; language models; context-sensitivity
+
+\bibliography{thesisbib}{}
+\bibliographystyle{apacite}
 

### 2013-05-22 RS twitter.tex
Worked on Twitter paper

@@
 \usepackage{graphicx}
-\usepackage{natbib}
+\usepackage{apacite}
 \usepackage{float}
@@
 With regard to elections, some authors have taken the opposite stance [[ref]].
-Most fierce in this respect is [[x]], who says 'No, you cannot predict elections with Twitter'.
+Most fierce in this respect is \citeA{gayo12}, who claims 'No, you cannot predict elections with Twitter'.
 He goes on to point out a number 'flaws' in research on doing predictions with Twitter.
@@
 \item Scholars should actually predict something, instead of doing \emph{post-hoc} analyses.
-As [[x]] put it: 'if you are claiming you have a prediction method you should predict an election in the future!'
-\item 'Not all tweets are trustworthy.' That is, some might be jokes or sarcasm and some might have been placed by the political leaders themselves.
-\item 'Demographics are neglected.' Twitter users are likely to be higher-educated people interested in social media, which might not be a representative sample of the population .
-According to [[x]], all research on predicting elections with Twitter should take this into account.
+As \citeA{gayo12} put it: 'if you are claiming you have a prediction method you should predict an election in the future!'
+\item 'Not all tweets are trustworthy.' That is, some might be jokes or sarcasm and some might have been placed by the political leaders themselves. These tweets should not be taken into account.
+\item 'Demographics are neglected.' Twitter users are likely to be younger, higher-educated people interested in social media, which might not be a representative sample of the population .
+Researchers should correct for this bias somehow.
 \item 'There is not a commonly accepted way of “counting votes” in Twitter'.
@@
 
-I am convinced all of these points are not really flaws in the research, but more general problems related to predicting with Twitter.
-As for the first point, you cannot evaluate a prediction system if you do not use past events; a paper saying 'sytem X predicts event Y', without the information where Y actually happened is meaningless.
-As long as the authors only used material available \emph{before} event Y, their result are perfectly valid.
-Arguments 2 and 3 claim there might be noise in the data, which I think is not only a general problem in Twitter research, but a problem in science in general: datasets which are 100\% representative of thing they were created to be representative of are scarce.
+However, I am convinced all of these points are not really flaws in the research, but more general problems related to predicting the future with Twitter.
+As for the first point, you cannot evaluate a prediction system if you do not use past events; a paper claiming 'system X predicts event Y', without the information where Y actually happened, is meaningless.
+As long as the authors only used material available \emph{before} event Y, their results are perfectly valid.
+Arguments 2 and 3 are about noise in the data, which I think is not only a general problem in Twitter research, but a problem in science in general; datasets which are 100\% representative of thing they were created to be representative of are scarce.
 The task of predicting events with Twitter is to do accurate predictions \emph{despite} these problems.
-If researchers can even report positive results by ignoring them, why bother?
-Finally, the last 'flaw' is actually the main goal of this research field: finding which way of 'counting votes' has the best correllation to the actual election results.
+If researchers can even build useful systems and report positive results by ignoring them, why bother?
+Finally, the last 'flaw' is actually the main goal of this research field: finding which way of 'counting votes' has the best correlation to the actual election results.
 Because it is not clear yet what technique consistently has the best results, various ideas will be proposed.
+The fact that everybody uses another way to count votes thus is not a problem, but an attempt for a solution.
 
-I therefore feel it would be more appropriate to rephrase the critique of [[x]] as problems common to this research field.
-Attempts to solve them might improve the result, and might even be research questions on their own.
-For example, [[x]] investigated Twitter demographics, and concluded that .
+I therefore feel it would be more appropriate to rephrase the first three points of critique of \citeA{gayo12} as problems common to this research field:
+
+\begin{enumerate}
+\item To be able to tell whether a prediction system works, one is mostly limited to events of the past.
+If a prediction is done on the basis of very recent tweets, it might take some time before one can be sure this prediction was accurate.
+\item Some tweets might use figurative language or be misleading in another way, and this way put automated systems on the wrong track.
+\item Twitter users are probably not representative of the population.
+Although it is hard to retrieve exact numbers, \shortciteA{mislove+11} are able to do some estimates on the basis of the data available.
+Among other things, they (1) discovered that more populous areas are overrepresented on Twitter (so the percentage of the population that uses Twitter is much higher in these areas) and (2) estimated, on the basis of self-reported first names, that 72\% of the Twitter users are male, although this percentage seems to decrease over the years.
+\end{enumerate}
+
+\bibliography{twitbib}{}
+\bibliographystyle{apacite}
 

### 2013-05-29 RS twitter.tex
Almost finished twitter paper, added references

@@
 
-\title{Twitter as a crystal ball: predicting future events with Twitter}
+\title{Twitter as a crystal ball: detecting and predicting real-world events with Twitter}
 \author{Wessel Stoop, s0808709}
@@
 \section{Introduction}
+Over the last two decades, the way people exchange information has changed considerably.
+Social media have become a common way communicate for many people: at the time of writing, social networking service Facebook\footnote{\url{http://www.facebook.com}} claims to have 665 million daily active users, whereas micro-blogging service Twitter\footnote{\url{http://www.twitter.com}} claims to have 340 million Tweets per day.
+Interestingly, this communciation is not limited to peer-to-peer communication; communication via social media makes it possible for an individual to reach quite large audiences easily.
 
-\section{Success stories}
+This possibility to reach a lot of people quickly is important change in news reporting.
+Whereas people used to read about most events happening in the world in the newspaper, or hear about it on the television news, they can now hear it from their friends.
+On top of that, they can report events they encounter themselves.
+This essentially transforms social media users from news consumers to a giant group of news reporters, reporting their experiences to eachother.
+As \citeA{rui+12} put it, '[content] is produced by a large number of ordinary Internet users, rather than by only a few publishers and television networks, as in the traditional media.
+This ongoing shift to social media means that user-generated content is now playing an unprecedented role in people’s life and likely will completely alter the way information is generated in our society in the future.'
 
-%Eindingen met elections. Eerst Tumasjan, daarna Erik en nog wat anderen, om te laten zien dat die resultaten niet herhaald kunnen worden. Mag ook bij de volgende sectie.
+However, unlike professional news reporters, most people have no financial benefits of reporting the news well.
+How accurate, clear and useful is social media communication concerning recent, current or upcoming events?
+\citeA{rui+12} investigated the reasons why people provision public goods (like Tweets), and propose a theory where social media (like Twitter) are depicted as 'attention exchange platforms': since in our digital society, every hour more \emph{information} is produced than an individual can process in a lifetime, human \emph{attention} has become a scarce resource.
+Social media can help people to collect information that interests them (i.
+e.
+people with attention looking for information), whereas others use it spread information (i.
+e.
+people with information looking for attention).
+The most important reason for the desire to spread information seems to be social status; \citeA{rui+12} show that Twitter users with long biograpies - people that have written a lot about themselves for others to read, and thus are likely to be interested in improving their status - tweet more often, and that for most users the tweeting rate increases as the number of followers increases.
 
-\section{Problems}
+%abs. getallen?
+
+If social status is involved, at least part of the social media communication is very likely be meaningful to some people.
+But are they meaningful enough to when it comes to detecting events that are going to happen in the future?
+And more importantly, can this prediction be done automatically?
+Research on other forms of communication suggest it can.
+For example, \citeA{veronis07} shows that for the French presidential elections of 2007 simply counting the numbers of candidate mentions in the press was a better prediction of success than many polls.
+Similarly, \shortciteA{gruhl+05} were able to predict when there would be spikes in the number of book sales on amazon.com\footnote{\url{http://www.amazon.com}} with an accuracy of [[x]].
+By using the frequency of mentions of a particular book title on IBM's webfountain, which collects its content from a large number of weblogs, their algorithm could predict a rapid increase in the number of sales 1-2 days before it would happen.
+
+%abs
+
+The aim of this paper is to give an overview of studies that predict real-world events on the basis of Twitter data.
+Section \ref{success} will summarize various studies which succeeded in this task to a certain extent.
+Section \ref{problems} will address some critique and problems for the field.
+The paper will be concluded in section \ref{conclusion}.
+
+\section{Success stories} \label{success}
+
+Twitter is micro-blogging service.
+Each Twitter user submits status updates known as tweets, which consist 140 characters or less.
+These tweets typically consist of personal information about the users themselves, news or links to other pages on the internet.
+Tweets are displayed on the user’s profile page, as well as shown to his/her followers.
+In this section, four studies will be discussed which have succesfully detected or predicted events on the basis of the language in tweets alone.
+Although the terminology, 'detecting and predicting events', might suggest this paper is about two separate tasks, they are actually related in several ways.
+For example, if we detect a current event, this will almost always lead to assumptions about the future, and doing predictions about future event $x$ can only be done by detecting the current event that people are talking about $x$.
+For this reason, while discussing the literature, detecting and predicting events will be treated as a continuum.
+We will start on one end of this continuum (detecting ongoing events that have started in the past) and step by step move to the other (predicting events far into the future).
+
+\citeA{lampos+10} searched the Twitter API for tweets for which gps coordinates were provided, and which contained one or more of various keywords related to the flu (\emph{cough}, \emph{fever}, \emph{tired}, etc.).
+Using the coordinates, these tweets were linked to a particular area.
+For five areas in the United Kingdom, they then counted how much tweets contained one or more of these words, and compared these numbers to official data about the spread of the flu from the Health Protection Agency (HPA).
+Although there were some differences per area, on average a statistically significant linear correlation coefficient greater than 89\% could be achieved.
+For example, a lot more tweets containing health-related words were coming from areas were HPA discovered an increase in the number of people affected with the flu.
+Furthermore, by using correlation data, they automatically improved their set of queries.
+This resulted in an average correlation coefficent of 95\%.
+Interestingly, it takes HPA two weeks to gather and analyze the necessary data, while the results of \citeA{lampos+10} are available directly.
+That is, their algorithm can quite accurately detect where the flu prevails \emph{right now}.
+
+Similarly focussing on ongoing events, \shortciteA{sakaki+10} built an algorithm that detects earthquakes on the fly.
+According to \shortciteA{sakaki+10}, 'the government has allocated a considerable amount of its budget to mitigating earthquake damage', since earthquakes occur so regularly there.
+The system they have built uses the Twitter search API to search for queries like 'earthquake' and 'shaking', and then uses a classifier to find out whether this tweet really is about earthquakes (and not about, for instance, an earthquake conference).
+As features for this classifier, they use (1) statistical characteristics (the number of words and the position of the query words), (2) the other words in the tweet and (3) the words before and after the keyword.
+A model using all of these features turned out the most succesful.
+When a positive tweet has been detected with a confidence value higher than .99, the system reports an event.
+This way, they were able to detect 96\% of the earthquakes stronger than JMA seismic intensity scale 3 or more.
+On top of that, they try to estimate \emph{where} the event is happening, by using the gps coordinates given with the tweets (if any) and the twitter user's home location.
+They put all positive tweets in a particle filter, and are this way able to have a very close approximation of the earthquake center.
+
+ Importantly, earthquakes do not reach all places at the same time; instead, they start at one location and then gradually spread to other places.
+This means that if an event is detected for one location, it can function as a prediction as another.
+The algorithm by \citeA{sakaki+10} sends warnings to people who have subscribed to the service.
+According to them 'an earthquake is transmitted through the earth’s crust at about 3–7 km/s.
+Therefore, a person has about 20 seconds before its arrival at a point that is 100 km distant.
+[...] The delivery time is far faster than the rapid broadcast of announcements of JMA, which are widely broadcast on TV; on average, a JMA announcement is broadcast 6 min after an earthquake occurs.'
+
+\citeA{asur+10} concentrate on doing predictions a week into the future: their algorithm uses tweets to predict the box-office revenue of movies released next week.
+The gold standard for such predictions so far is the Hollywood Stock Exchange Index (HSX).
+The HSX\footnote{\url{http://www.hsx.com}} is a web-based multiplayer game in which players use simulated money to buy and sell "shares" of actors, directors and upcoming films.
+Since these shares can be exchanged for this simulated money [[two weeks]] after the movies have been released, it is the job of the players to buy them for less money than they produced (or to sell them for more than they costed or will produce).
+How much money these shares produce depends on the box-office revenue of the corresponding movies, which results in hundreds of people trying to guess it as accurately as possible.
+Combined with the number of theathers the movie is released in, a factor which of course has a very big influence on how much revenue can possibly be made, the HSX correlates with the actual revenue with an $R^2$ of .965.
+
+By using tweet rates, which is the number of tweets per hour, \citeA{asur+10} were able to improve this result.
+Their algorithm looks for tweets that contain the title of a movie.
+The number of tweets found is then divided by the number of hours, which results in the tweet rate.
+This rate will be high for movies which have either large burst of attention, or when there a lot of bursts, but low otherwise.
+Attention bursts are typically caused by the release of new promotional material \cite{asur+10}.
+When only using tweets, \citeA{asur+10} found a correlation of $R^2 = .92$ between tweet rate and the actual revenue.
+Adding the number of theaters resulted, as was done with HSX, resulted in the even higher $R^2$ of .973.
+The correlation between actual revenue and tweet rate was higher than the correlation between actual revenue and HSX for every movie in the study.
+\citeA{asur+10} also investigated to what extent adding sentiment analysis can improve the results.
+They used the DynamicLMClassifier with sentiment analyses produced by \emph{Amazon Mechanical Turk}.
+This improved results, but only for the second weekend (from .92 to .94 $R^2$), which according to the authors is caused by reviews and word of mouth.
+For example, the movie \emph{The Blind Side}, did not perform well in the opening weekend (34M), but got a positive response: it showed an enormous increase in positive sentiment from 5.02 to 9.65.
+It indeed performed better in its second weekend (40.1M).
+
+\shortciteA{Tumasjan+10} finally do predictions for an event several weeks beforehand: the elections.
+They have collected all tweets mentioning one of the six larger parties (CDU/CSU, SPD, FDP, B90/Die Grünen, and Die Linke) posted in the month before the German national elections of 2009.
+They also collected tweets mentioning several politicians linked to one of these parties.
+The percentage of tweets in this sample for each party was almost identical to the percentage of votes.
+\shortciteA{Tumasjan+10} report a \emph{Mean Absolute Error}, a measure to calculate to what extent the percentages are similar, of 1.65\%.
+Although traditional polls get even lower Mean Absolute Errors (ranging from for 0.80\% to 1.45\%), this is more than enough to establish the relative order of the parties in terms of number of votes.
+
+\section{Problems} \label{problems}
 
@@
 
+% GROVE DATA
+% VEEL MENSEN PRATEN NIET
+
+\section{Conclusion} \label{conclusion}
+
 \bibliography{twitbib}{}
@@
 \end{document}
+
+%TODO
+% Niet alleen predicting

### 2013-05-30 RS twitter.tex
Prefinal revision twitter paper

@@
 Over the last two decades, the way people exchange information has changed considerably.
-Social media have become a common way communicate for many people: at the time of writing, social networking service Facebook\footnote{\url{http://www.facebook.com}} claims to have 665 million daily active users, whereas micro-blogging service Twitter\footnote{\url{http://www.twitter.com}} claims to have 340 million Tweets per day.
+Social media have become a common way communicate for many people: at the time of writing, social networking service Facebook\footnote{\url{http://www.facebook.com}} claims to have 665 million daily active users, and micro-blogging service Twitter\footnote{\url{http://www.twitter.com}} claims to have 340 million Tweets per day.
 Interestingly, this communciation is not limited to peer-to-peer communication; communication via social media makes it possible for an individual to reach quite large audiences easily.
 
-This possibility to reach a lot of people quickly is important change in news reporting.
-Whereas people used to read about most events happening in the world in the newspaper, or hear about it on the television news, they can now hear it from their friends.
+This possibility to reach a lot of people quickly might be a very important change concerning the spread of knowledge about ongoing or upcoming events.
+Whereas people used to read about most events happening in the world in the newspaper, or hear about it on the television news, they can now hear it from their friends much faster.
 On top of that, they can report events they encounter themselves.
@@
 However, unlike professional news reporters, most people have no financial benefits of reporting the news well.
-How accurate, clear and useful is social media communication concerning recent, current or upcoming events?
-\citeA{rui+12} investigated the reasons why people provision public goods (like Tweets), and propose a theory where social media (like Twitter) are depicted as 'attention exchange platforms': since in our digital society, every hour more \emph{information} is produced than an individual can process in a lifetime, human \emph{attention} has become a scarce resource.
+How accurate, clear and useful is social media communication concerning ongoing or upcoming events?
+\citeA{rui+12} investigated the reasons why people provision public goods (like Tweets), and claim that the most important reason for the desire to spread information seems to be social status; \citeA{rui+12} show that Twitter users with long biograpies - people that have written a lot about themselves for others to read, and thus are likely to be interested in improving their status - tweet more often, and that for most users the tweeting rate increases as the number of followers increases.
+Apparently, when you can improve your social status for a lot of people at the same time, posting tweets becomes more interesting.
+
+%propose a theory where social media (like Twitter) are depicted as 'attention exchange platforms': since in our digital society, every hour more \emph{information} is produced than an individual can process in a lifetime, human \emph{attention} has become a scarce resource.
 Social media can help people to collect information that interests them (i.
@@
 people with information looking for attention).
-The most important reason for the desire to spread information seems to be social status; \citeA{rui+12} show that Twitter users with long biograpies - people that have written a lot about themselves for others to read, and thus are likely to be interested in improving their status - tweet more often, and that for most users the tweeting rate increases as the number of followers increases.
-
-%abs. getallen?
 
 If social status is involved, at least part of the social media communication is very likely be meaningful to some people.
-But are they meaningful enough to when it comes to detecting events that are going to happen in the future?
-And more importantly, can this prediction be done automatically?
+But are they meaningful enough to when it comes to detecting and predicting real-world events?
+And more importantly, can this be done automatically?
 Research on other forms of communication suggest it can.
 For example, \citeA{veronis07} shows that for the French presidential elections of 2007 simply counting the numbers of candidate mentions in the press was a better prediction of success than many polls.
-Similarly, \shortciteA{gruhl+05} were able to predict when there would be spikes in the number of book sales on amazon.com\footnote{\url{http://www.amazon.com}} with an accuracy of [[x]].
-By using the frequency of mentions of a particular book title on IBM's webfountain, which collects its content from a large number of weblogs, their algorithm could predict a rapid increase in the number of sales 1-2 days before it would happen.
-
-%abs
+Similarly, \shortciteA{gruhl+05} were able to predict when there would be spikes in the number of book sales on Amazon\footnote{\url{http://www.amazon.com}} by looking at the frequency of mentions of a particular book title on IBM's webfountain.
+IBM's webfountain collects its content from a large number of weblogs.
+The algorithm by \shortciteA{gruhl+05} could predict a rapid increase in the number of sales 1-2 days before it would happen.
 
 The aim of this paper is to give an overview of studies that predict real-world events on the basis of Twitter data.
-Section \ref{success} will summarize various studies which succeeded in this task to a certain extent.
+Twitter is micro-blogging service where each Twitter user can submit status updates known as tweets, which consist 140 characters or less.
+These tweets typically consist of personal information about the users themselves, news or links to other pages on the internet.
+Tweets are displayed on the user’s profile page, as well as shown to his/her followers.
+Probably because tweeting does not take a lot of effort (a lot less than, for example, writing a weblog), Twitter has a lot of users that tweet regularly.
+This makes Twitter a very interesting dataset for detecting and predicting real-world events.
+Section \ref{success} will summarize four studies which have succeeded in this task to a certain extent.
 Section \ref{problems} will address some critique and problems for the field.
@@
 
-Twitter is micro-blogging service.
-Each Twitter user submits status updates known as tweets, which consist 140 characters or less.
-These tweets typically consist of personal information about the users themselves, news or links to other pages on the internet.
-Tweets are displayed on the user’s profile page, as well as shown to his/her followers.
 In this section, four studies will be discussed which have succesfully detected or predicted events on the basis of the language in tweets alone.
@@
 For this reason, while discussing the literature, detecting and predicting events will be treated as a continuum.
-We will start on one end of this continuum (detecting ongoing events that have started in the past) and step by step move to the other (predicting events far into the future).
+We will start on one end of this continuum (detecting ongoing events that have started in the past) and step by step move to the other (predicting events several weeks into the future).
 
-\citeA{lampos+10} searched the Twitter API for tweets for which gps coordinates were provided, and which contained one or more of various keywords related to the flu (\emph{cough}, \emph{fever}, \emph{tired}, etc.).
+Starting with ongoing events, \citeA{lampos+10} used the Twitter search API to find tweets for which gps coordinates were provided, and which contained one or more of various keywords related to the flu (\emph{cough}, \emph{fever}, \emph{tired}, etc.).
 Using the coordinates, these tweets were linked to a particular area.
@@
 Although there were some differences per area, on average a statistically significant linear correlation coefficient greater than 89\% could be achieved.
-For example, a lot more tweets containing health-related words were coming from areas were HPA discovered an increase in the number of people affected with the flu.
+This means that for example a lot more tweets containing health-related words were coming from areas were HPA discovered an increase in the number of people affected with the flu.
 Furthermore, by using correlation data, they automatically improved their set of queries.
@@
 
-Similarly focussing on ongoing events, \shortciteA{sakaki+10} built an algorithm that detects earthquakes on the fly.
-According to \shortciteA{sakaki+10}, 'the government has allocated a considerable amount of its budget to mitigating earthquake damage', since earthquakes occur so regularly there.
+Similarly focussing on ongoing events, \shortciteA{sakaki+10} built an algorithm that detects earthquakes in Japan on the fly.
+According to \shortciteA{sakaki+10}, 'the government [of Japan] has allocated a considerable amount of its budget to mitigating earthquake damage', since earthquakes occur so regularly there.
 The system they have built uses the Twitter search API to search for queries like 'earthquake' and 'shaking', and then uses a classifier to find out whether this tweet really is about earthquakes (and not about, for instance, an earthquake conference).
 As features for this classifier, they use (1) statistical characteristics (the number of words and the position of the query words), (2) the other words in the tweet and (3) the words before and after the keyword.
-A model using all of these features turned out the most succesful.
-When a positive tweet has been detected with a confidence value higher than .99, the system reports an event.
-This way, they were able to detect 96\% of the earthquakes stronger than JMA seismic intensity scale 3 or more.
-On top of that, they try to estimate \emph{where} the event is happening, by using the gps coordinates given with the tweets (if any) and the twitter user's home location.
-They put all positive tweets in a particle filter, and are this way able to have a very close approximation of the earthquake center.
+A model using all of these features turned out to be the most succesful.
+When the classifier marks a tweet as positive, and is 99\% confident, the system reports an event.
+This way, the algorithm by \shortciteA{sakaki+10} is able to detect 96\% of the earthquakes stronger than JMA seismic intensity scale 3 or more.
+On top of that, the authors try to estimate \emph{where} the event is happening, by using the gps coordinates given with the tweets (if any) and the twitter user's home location.
+By putting all positive tweets in a particle filter, they were able to have a very close approximation of the earthquake center.
+%waar dan?
 
 Importantly, earthquakes do not reach all places at the same time; instead, they start at one location and then gradually spread to other places.
-This means that if an event is detected for one location, it can function as a prediction as another.
+This means that if an event is detected for one location, it can function as a prediction for another.
 The algorithm by \citeA{sakaki+10} sends warnings to people who have subscribed to the service.
@@
 The HSX\footnote{\url{http://www.hsx.com}} is a web-based multiplayer game in which players use simulated money to buy and sell "shares" of actors, directors and upcoming films.
-Since these shares can be exchanged for this simulated money [[two weeks]] after the movies have been released, it is the job of the players to buy them for less money than they produced (or to sell them for more than they costed or will produce).
-How much money these shares produce depends on the box-office revenue of the corresponding movies, which results in hundreds of people trying to guess it as accurately as possible.
-Combined with the number of theathers the movie is released in, a factor which of course has a very big influence on how much revenue can possibly be made, the HSX correlates with the actual revenue with an $R^2$ of .965.
+Players make money by buying shares of movies they think will do well, which makes HSX essentially a group of hundreds people trying to guess the box office revenue as accurately as possible.
+Data on how well shares for particular movies do, combined with the number of theathers the movie is released in (a factor which of course has a very big influence on how much revenue can possibly be made) therefore correlates with the actual revenue with an $R^2$ of .965.
 
 By using tweet rates, which is the number of tweets per hour, \citeA{asur+10} were able to improve this result.
-Their algorithm looks for tweets that contain the title of a movie.
-The number of tweets found is then divided by the number of hours, which results in the tweet rate.
-This rate will be high for movies which have either large burst of attention, or when there a lot of bursts, but low otherwise.
+This rate for tweets which mention a particular movie title will be high for movies which had either a large burst of attention, or a lot of smaller bursts, but low otherwise.
 Attention bursts are typically caused by the release of new promotional material \cite{asur+10}.
 When only using tweets, \citeA{asur+10} found a correlation of $R^2 = .92$ between tweet rate and the actual revenue.
-Adding the number of theaters resulted, as was done with HSX, resulted in the even higher $R^2$ of .973.
+Adding the number of theaters, as was done with HSX, resulted in the even higher $R^2$ of .973.
 The correlation between actual revenue and tweet rate was higher than the correlation between actual revenue and HSX for every movie in the study.
+
 \citeA{asur+10} also investigated to what extent adding sentiment analysis can improve the results.
-They used the DynamicLMClassifier with sentiment analyses produced by \emph{Amazon Mechanical Turk}.
-This improved results, but only for the second weekend (from .92 to .94 $R^2$), which according to the authors is caused by reviews and word of mouth.
+They used the DynamicLMClassifier with sentiment analyses produced by \emph{Amazon Mechanical Turk} as training material.
+This improved results, but only for the second weekend (the $R^2$ improved from .92 to .94), which according to the authors is caused by reviews and word of mouth.
 For example, the movie \emph{The Blind Side}, did not perform well in the opening weekend (34M), but got a positive response: it showed an enormous increase in positive sentiment from 5.02 to 9.65.
@@
 
-\shortciteA{Tumasjan+10} finally do predictions for an event several weeks beforehand: the elections.
-They have collected all tweets mentioning one of the six larger parties (CDU/CSU, SPD, FDP, B90/Die Grünen, and Die Linke) posted in the month before the German national elections of 2009.
+\shortciteA{Tumasjan+10} finally concentrate on predictions for an event several weeks beforehand: the elections.
+They collected all tweets mentioning one of the six larger parties (CDU/CSU, SPD, FDP, B90/Die Grunen, and Die Linke) posted in the month before the German national elections of 2009.
 They also collected tweets mentioning several politicians linked to one of these parties.
@@
 
-With regard to elections, some authors have taken the opposite stance [[ref]].
+With regard to elections, some authors have taken the opposite stance \shortcite{oconnor+10,jungherr+11,gayo12,skoric+12}.
 Most fierce in this respect is \citeA{gayo12}, who claims 'No, you cannot predict elections with Twitter'.
@@
 \item 'There is not a commonly accepted way of “counting votes” in Twitter'.
-Whereas some authors use tweet volumes, others use tweet rate, and again other focus on the results of some kind of sentiment analyses.
+Whereas some authors use tweet volumes, others use tweet rate, and again others focus on the results of some kind of sentiment analysis.
 \end{enumerate}
@@
 However, I am convinced all of these points are not really flaws in the research, but more general problems related to predicting the future with Twitter.
-As for the first point, you cannot evaluate a prediction system if you do not use past events; a paper claiming 'system X predicts event Y', without the information where Y actually happened, is meaningless.
-As long as the authors only used material available \emph{before} event Y, their results are perfectly valid.
+As for the first point, you cannot evaluate a prediction system if you do not use past events; a paper claiming 'system $X$ predicts event $Y$', without the information whether $Y$ actually happened, is meaningless.
+As long as the authors only used material available \emph{before} $Y$, their results are perfectly valid.
 Arguments 2 and 3 are about noise in the data, which I think is not only a general problem in Twitter research, but a problem in science in general; datasets which are 100\% representative of thing they were created to be representative of are scarce.
 The task of predicting events with Twitter is to do accurate predictions \emph{despite} these problems.
-If researchers can even build useful systems and report positive results by ignoring them, why bother?
+In the four studies described above researchers have been able to build useful systems and report positive results by mostly ignoring this noise, so calling this a 'flaw' might be too harsh.
 Finally, the last 'flaw' is actually the main goal of this research field: finding which way of 'counting votes' has the best correlation to the actual election results.
@@
 \item Some tweets might use figurative language or be misleading in another way, and this way put automated systems on the wrong track.
+In other words, there is a lot of noise in the data.
+I think the most important implication of this is that you cannot do very detailed analyses; indeed, a large part of the authors mentioned in this paper note that they were only able to very general detections and predictions \shortcite{sakaki+10,gruhl+05, Tumasjan+10,lampos+10}.
 \item Twitter users are probably not representative of the population.
-Although it is hard to retrieve exact numbers, \shortciteA{mislove+11} are able to do some estimates on the basis of the data available.
+Although it is hard to get exact numbers, \shortciteA{mislove+11} report some estimates on the basis of the data available.
 Among other things, they (1) discovered that more populous areas are overrepresented on Twitter (so the percentage of the population that uses Twitter is much higher in these areas) and (2) estimated, on the basis of self-reported first names, that 72\% of the Twitter users are male, although this percentage seems to decrease over the years.
@@
 
-% GROVE DATA
-% VEEL MENSEN PRATEN NIET
+As stated before, however, since several studies have been able to show positive results by ignoring these problems, I firmly believe they should be seen as areas where the results can be improved instead of insurmountable obstacles.
 
 \section{Conclusion} \label{conclusion}
+
+This paper discussed the way communication has changed over the last two decades, and the positive effects this can have on the automatical detection and prediction of real-world events.
+Very promising is microblogging service Twitter, because of its large userbase and frequent updates.
+We saw that Twitter can detect ongoing outbreaks of the flu and earthquakes.
+This earthquake detection can be used to predict earthquakes for other areas.
+Other predictions that can be done with Twitter are the box-office revenue of movies and the election results, although that last study is criticized by some authors.
+It was suggested to view this criticism as problems common to the field, instead of flaws of the research.
 

### 2013-06-06 RS twitter.tex
Finished Twitter paper

@@
 
-\title{Twitter as a crystal ball: detecting and predicting real-world events with Twitter}
+\title{Twitter as a crystal ball: detecting and predicting real-world events with tweets}
 \author{Wessel Stoop, s0808709}
@@
 This possibility to reach a lot of people quickly might be a very important change concerning the spread of knowledge about ongoing or upcoming events.
-Whereas people used to read about most events happening in the world in the newspaper, or hear about it on the television news, they can now hear it from their friends much faster.
-On top of that, they can report events they encounter themselves.
-This essentially transforms social media users from news consumers to a giant group of news reporters, reporting their experiences to eachother.
+Whereas people used to read about most events happening in the world in the newspaper, or hear about it on the television news, they can now hear it from their friends much earlier.
+On top of that, they can report the events they encounter themselves to others.
+This essentially transforms social media users from news consumers to a giant group of news reporters, reporting their experiences to each other.
 As \citeA{rui+12} put it, '[content] is produced by a large number of ordinary Internet users, rather than by only a few publishers and television networks, as in the traditional media.
@@
 How accurate, clear and useful is social media communication concerning ongoing or upcoming events?
-\citeA{rui+12} investigated the reasons why people provision public goods (like Tweets), and claim that the most important reason for the desire to spread information seems to be social status; \citeA{rui+12} show that Twitter users with long biograpies - people that have written a lot about themselves for others to read, and thus are likely to be interested in improving their status - tweet more often, and that for most users the tweeting rate increases as the number of followers increases.
+\citeA{rui+12} investigated the reasons why people provision public goods (like Tweets), and claim that the most important reason for the desire to spread information seems to be social status; \citeA{rui+12} show that Twitter users with long biographies - people that have written a lot about themselves for others to read, and thus are likely to be interested in improving their status - tweet more often, and that for most users the tweeting rate increases as the number of followers increases.
 Apparently, when you can improve your social status for a lot of people at the same time, posting tweets becomes more interesting.
@@
 And more importantly, can this be done automatically?
-Research on other forms of communication suggest it can.
+Research on other forms of communication suggests it can.
 For example, \citeA{veronis07} shows that for the French presidential elections of 2007 simply counting the numbers of candidate mentions in the press was a better prediction of success than many polls.
@@
 
-The aim of this paper is to give an overview of studies that predict real-world events on the basis of Twitter data.
-Twitter is micro-blogging service where each Twitter user can submit status updates known as tweets, which consist 140 characters or less.
+The aim of this paper is to give an overview of studies that detect and predict real-world events on the basis of Twitter data.
+Twitter is a micro-blogging service where each Twitter user can submit status updates known as tweets, which consist 140 characters or less.
 These tweets typically consist of personal information about the users themselves, news or links to other pages on the internet.
@@
 Probably because tweeting does not take a lot of effort (a lot less than, for example, writing a weblog), Twitter has a lot of users that tweet regularly.
-This makes Twitter a very interesting dataset for detecting and predicting real-world events.
-Section \ref{success} will summarize four studies which have succeeded in this task to a certain extent.
-Section \ref{problems} will address some critique and problems for the field.
+This makes Twitter a very interesting dataset for detecting and predicting events.
+Section \ref{success} will summarize four studies which have succeeded in this task to a certain extent, and section \ref{problems} will address some critique and problems for the field.
 The paper will be concluded in section \ref{conclusion}.
@@
 Although there were some differences per area, on average a statistically significant linear correlation coefficient greater than 89\% could be achieved.
-This means that for example a lot more tweets containing health-related words were coming from areas were HPA discovered an increase in the number of people affected with the flu.
+This means that, for example, a lot more tweets containing health-related words were coming from areas were HPA discovered an increase in the number of people affected with the flu.
 Furthermore, by using correlation data, they automatically improved their set of queries.
@@
 
-\citeA{asur+10} concentrate on doing predictions a week into the future: their algorithm uses tweets to predict the box-office revenue of movies released next week.
+\citeA{asur+10} aim at doing predictions a week into the future: their algorithm uses tweets to predict the box-office revenue of movies released next week.
 The gold standard for such predictions so far is the Hollywood Stock Exchange Index (HSX).
-The HSX\footnote{\url{http://www.hsx.com}} is a web-based multiplayer game in which players use simulated money to buy and sell "shares" of actors, directors and upcoming films.
+The HSX\footnote{\url{http://www.hsx.com}} is a web-based multiplayer game in which players use simulated money to buy and sell 'shares' of actors, directors and upcoming films.
 Players make money by buying shares of movies they think will do well, which makes HSX essentially a group of hundreds people trying to guess the box office revenue as accurately as possible.
-Data on how well shares for particular movies do, combined with the number of theathers the movie is released in (a factor which of course has a very big influence on how much revenue can possibly be made) therefore correlates with the actual revenue with an $R^2$ of .965.
+Data on how well shares for particular movies do, combined with the number of theaters the movie is released in (a factor which of course has a very big influence on how much revenue can possibly be made) therefore correlates with the actual revenue with an $R^2$ of .965.
 
 By using tweet rates, which is the number of tweets per hour, \citeA{asur+10} were able to improve this result.
-This rate for tweets which mention a particular movie title will be high for movies which had either a large burst of attention, or a lot of smaller bursts, but low otherwise.
+The rate for tweets which mention a particular movie title will be high for movies which had either a large burst of attention, or a lot of smaller bursts, but low otherwise.
 Attention bursts are typically caused by the release of new promotional material \cite{asur+10}.
 When only using tweets, \citeA{asur+10} found a correlation of $R^2 = .92$ between tweet rate and the actual revenue.
-Adding the number of theaters, as was done with HSX, resulted in the even higher $R^2$ of .973.
-The correlation between actual revenue and tweet rate was higher than the correlation between actual revenue and HSX for every movie in the study.
+Adding the number of theaters, as was done with the HSX, resulted in the even higher $R^2$ of .973.
+The correlation between actual revenue and tweet rate was higher than the correlation between actual revenue and the HSX for every movie in the study.
 
@@
 This improved results, but only for the second weekend (the $R^2$ improved from .92 to .94), which according to the authors is caused by reviews and word of mouth.
-For example, the movie \emph{The Blind Side}, did not perform well in the opening weekend (34M), but got a positive response: it showed an enormous increase in positive sentiment from 5.02 to 9.65.
+For example, the movie \emph{The Blind Side} did not perform well in the opening weekend (34M), but got a positive response: it showed an enormous increase in positive sentiment from 5.02 to 9.65.
 It indeed performed better in its second weekend (40.1M).
 
-\shortciteA{Tumasjan+10} finally concentrate on predictions for an event several weeks beforehand: the elections.
+\shortciteA{Tumasjan+10} finally try to do predictions for an event several weeks beforehand: the elections.
 They collected all tweets mentioning one of the six larger parties (CDU/CSU, SPD, FDP, B90/Die Grunen, and Die Linke) posted in the month before the German national elections of 2009.
@@
 The percentage of tweets in this sample for each party was almost identical to the percentage of votes.
-\shortciteA{Tumasjan+10} report a \emph{Mean Absolute Error}, a measure to calculate to what extent the percentages are similar, of 1.65\%.
-Although traditional polls get even lower Mean Absolute Errors (ranging from for 0.80\% to 1.45\%), this is more than enough to establish the relative order of the parties in terms of number of votes.
+The \emph{Mean Absolute Error} is a measure to calculate to what extent percentages are similar, and is often used to compare prediction polls to the actual result.
+\shortciteA{Tumasjan+10} report a MAE of 1.65\%.
+Although traditional polls get even lower MAEs (ranging from for 0.80\% to 1.45\%), this is more than enough to establish the relative order of the parties in terms of number of votes.
 
@@
 
-With regard to elections, some authors have taken the opposite stance \shortcite{oconnor+10,jungherr+11,gayo12,skoric+12}.
+With regard to predicting the election outcomes with Twitter, some authors have expressed doubts on whether this is really possible \shortcite{oconnor+10,jungherr+11,gayo12,skoric+12}.
 Most fierce in this respect is \citeA{gayo12}, who claims 'No, you cannot predict elections with Twitter'.
@@
 The task of predicting events with Twitter is to do accurate predictions \emph{despite} these problems.
-In the four studies described above researchers have been able to build useful systems and report positive results by mostly ignoring this noise, so calling this a 'flaw' might be too harsh.
+In the studies described above, researchers have been able to build useful systems and report positive results by mostly ignoring this noise, so calling this a 'flaw' might be too harsh.
 Finally, the last 'flaw' is actually the main goal of this research field: finding which way of 'counting votes' has the best correlation to the actual election results.
@@
 In other words, there is a lot of noise in the data.
-I think the most important implication of this is that you cannot do very detailed analyses; indeed, a large part of the authors mentioned in this paper note that they were only able to very general detections and predictions \shortcite{sakaki+10,gruhl+05, Tumasjan+10,lampos+10}.
+I think the most important implication of this is that you cannot do very detailed analyses; indeed, a large part of the authors mentioned in this paper note that they were only able to do very general detections and predictions \shortcite{sakaki+10,gruhl+05, Tumasjan+10,lampos+10}.
 \item Twitter users are probably not representative of the population.

### 2013-06-06 research proposal MTS.tex
Revised research proposal MTS

@@
 
-This research is an attempt to link two fields of research: the field of \emph{word completion} and the field \emph{idiolects}.
+This research is an attempt to link two fields of research: the field of \emph{word completion} (part of computational linguistics) and the field \emph{idiolects} (part of sociolinguistics).
 They will be introduced in section \ref{wordprediction} and \ref{idiolects} respectively.
@@
 Since then, numerous authors have shown that taking context into account improves prediction accuracy significantly \cite[among others]{Lesher+99,Garay-Vitoria+06,Tanaka-Ishii07,vandenbosch+08}.
-This entails that the system has some information on which words are likely to follow eachother, and uses this when giving predictions.
+This entails that the system has some information on which words are likely to follow each other, and uses this when giving predictions.
 For example, if the user has keyed in sentence \ref{iate}, a context-sensitive system should not predict the word \emph{communication}, even if it is high on the frequency lists, because \emph{communication} is unlikely to follow \emph{ate}.
@@
 
-One way give some of this kind of knowledge to word prediction systems is to use training material from the same domain; \citeA{verberne+12} show that trying to predict Wikipedia text, tweets, transcriptions of conversational speech and Frequencty Asked Questions all worked best when using texts from the same type (i.e.
+One way to give some of this kind of knowledge to word prediction systems is to use training material from the same domain; \citeA{verberne+12} show that trying to predict Wikipedia text, tweets, transcriptions of conversational speech and Frequently Asked Questions all worked best when using texts from the same type (i.e.
 training on Wikipedia text to predict Wikipedia text).
@@
 \citeA{vandenbosch11} takes a completely different approach and tries to make a word prediction system learn about register and topic on the fly with a \emph{recency buffer}.
-This buffer stores the $n$ latest words; if a word the user is keying in matches one of the words in the recency buffer, this word is suggested.
-The idea behind this is that a word that has recently been used, it is likely that this word will be used again.
+This buffer stores the $n$ latest words; if a word the user is keying in matches one of the words in the recency buffer, this word is suggested instead of what the system would actually have suggested.
+The idea behind this is that if the user is writing about, for example, traveling, words like 'go', 'hotel', and 'see  will always be in the buffer and thus be suggested.
+In other words, the systems remembers, in a simple way, what you are talking about.
+Similarly, if the user is writing a very formal letter in Dutch, the formal second pronoun \emph{u} is likely to be in the buffer, whereas this will not be the case (and thus is less likely to be suggested) when the user is writing to his/her friend.
+In other words, the system remembers, in a simple way, how you are talking.
 
@@
 
-Part of the of lack of research in this area can of course be explained by practical limitations (it is hard to collect a lot of data for one person, let alone doing this for multiple persons), but theoretical views have also played a role; for decades, most linguistists have focussed on \emph{langue} rather than \emph{parole}, and viewed deviations in actual usage as uninteresting exceptions to an otherwise perfect abstract system.
-As \citeA{labov89} put it: 'language is not a property of the individual, but of the community.
+Part of the lack of research in this area can of course be explained by practical limitations (it is hard to collect a lot of data for one person, let alone doing this for multiple persons), but theoretical views have also played a role; for decades, most linguistists have focussed on \emph{langue} rather than \emph{parole}, and viewed deviations in actual usage as uninteresting exceptions to an otherwise perfect abstract system.
+As \citeA{labov89} puts it: 'language is not a property of the individual, but of the community.
 Any description of a language must take the speech community as its object if it is to do justice to elegance and regularity of linguistic structure'.
@@
 One of the first publications investigating idiolects quantitatively is that of \citeA{mollin09}.
-She created a corpus of speech by the former Prime Minister of the United Kingdom Tony Blair, and compared this the British National Corpus.
+She created a corpus of speech by the former Prime Minister of the United Kingdom Tony Blair, and compared this to the British National Corpus.
 Focussing on maximisers ('absolutely', 'totally') and the adjectives, verbs and adverbs they cooccur with, she has compiled a list of collocations that are used much more by Blair than by the general language community.
@@
 
-While the results of these studies are promising, there are some problems in the way they try to capture idiolects.
-For example, all of the language investigated here is that of people speaking to the press, but it unlikely that the speakers always speak like this; Tony Blair probably uses another style register when speaking to his friends and family.
+While the results of these studies are promising, there are some problems in the way they try to capture idiolects. Three examples:
+
+\begin{itemize}
+\item All of the language investigated here is that of people speaking to the press, but it is unlikely that the speakers always speak like this; Tony Blair probably uses another style register when speaking to his friends and family, and even when speaking to one person there might be variation (happy, sad, angry).
 How should style registers be part of a theory about idiolects?
 Is a register a set of variables in an otherwise fixed system, are they completely seperate language varieties inside a person's mind, or something in between?
-Furthermore, \citeA{barlow10} points out that 'the samples from Mike McCurry's speeches seem to divide into two temporally distinct groups [...], suggesting a change in style over his long period of tenure as press secretary'.
+And if we have a language model for one register, to what extent does that overlap with the language model for another?
+\item \citeA{barlow10} points out that 'the samples from Mike McCurry's speeches seem to divide into two temporally distinct groups [...], suggesting a change in style over his long period of tenure as press secretary'.
 Apparently, idiolects are not time stable, but change over time.
@@
 And if we have models for a particular person based on the idiolect of the last decade, last year, yesterday and a few minutes ago, to what extent do they overlap?
-And finally, both \citeA{barlow10} and \citeA{mollin09} try to capture idiolects by using the \emph{output} of persons.
+\item Both \citeA{barlow10} and \citeA{mollin09} try to capture idiolects by using the \emph{output} of persons.
 This of course makes sense (what comes out must have been inside), but this cannot be the whole story; wouldn't it be more logical to use the input for a particular person, in particular the things he or she hears very often?
+\end{itemize}
 
 \subsection{Methodology: idiolects in word prediction} \label{link}
-As the previous two section have showed, in both academic fields discussed we have a problem: for word prediction, a technique is needed which is context-sensitive, but does not rely on many texts in the same genre, while for idiolects, we are looking for a good way to make concrete what an idiolect actually is.
+As the previous two section have shown, in both academic fields discussed we have a problem: for word prediction, a technique is needed which is context-sensitive, but does not rely on many texts in the same genre, while for idiolects, we are looking for a good way to make concrete what an idiolect actually is.
 In this research project, these two concepts will be linked to help each other: a word prediction system will be built which will rely on text written by the user earlier (idiolects), while incorperating the latest insights on the field of word prediction, as described in \ref{wordprediction}.
@@
 As our dataset, we plan to use tweets, as Twitter essentially is nothing more than a collection of language ordered by author.
+The quality of the language on Twitter might have a bad reputation, but we believe this is (at least partly) unjustified: Twitter is, for example, also used by politicians to talk about their opinions with their voters, companies to connect to their clients, developers to talk about their products, researchers to talk about their research, etc.
+
 The expectation thus is that a model based on earlier tweets is much better at predicting future tweets than a model based random tweets in the same language.
 Furthermore, tweets are annotated for time, which allows us to make models for a particular period, and in a lot of cases explicitly mention the addressee (with the \emph{@addressee} syntax).
-This not only allows us to make an approximation of register (different models could be made depending on to whom the author was talking), but also gives us some insight in social relations between the Twitter users.
+This not only allows us to make an approximation of register (different models could be made depending on to whom the author was talking to), but also gives us some insight in social relations between the Twitter users.
 The expectation is that our word prediction will be able to do more accurate predictions for a particular twitter user when using a model based on the tweets produced by people that user communicates with a lot.
